---
title: "Herbivory Plots"
author: "Kara Dobson"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

COLLABORATORS: Phoebe Zarnetske, Mark Hammond, Pat Bills, Moriah Young  
DATA INPUT: Clean & plot herb csv from the shared Google drive  
DATA OUTPUT: Code and Rmd are in the scripts folder in Github  
PROJECT: warmXtrophic  

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Clear all existing data
rm(list=ls())

# Load packages
library(tidyverse)
library(plotrix)

# Set working directory to Google Drive
# **** Update with the path to your Google drive on your computer
setwd("/Volumes/GoogleDrive/Shared drives/SpaCE_Lab_warmXtrophic/data/")

# Read in plant comp data & metadata
herb <- read.csv("L1/herbivory/final_herbivory_L1.csv")
#str(herb) # for some reason, date column converted back to character

# Fix date column & add column for the year and julian day
herb$date <- as.Date(herb$date,format="%Y-%m-%d")
#str(herb)

# Remove weird X column
herb <- subset(herb, select = -X)



#### Total herb by site and species ####
sum_herb <- herb %>%
  group_by(site, treatment_key, species, year) %>%
  summarize(avg_eaten = mean(p_eaten, na.rm = TRUE),
            se = std.error(p_eaten, na.rm = TRUE))

#### Total herb by site and species with insecticide treatment####
sum_herb_in <- herb %>%
  group_by(site, treatment_key, year) %>%
  summarize(avg_eaten = mean(p_eaten, na.rm = TRUE),
            se = std.error(p_eaten, na.rm = TRUE))

#### Total herb by site and species with separated insecticide and no insecticide####
sum_herb_in <- herb %>%
  group_by(site, treatment_key, year) %>%
  summarize(avg_eaten = mean(p_eaten, na.rm = TRUE),
            se = std.error(p_eaten, na.rm = TRUE))
```

### Makes a plot for each site with and without insecticide treatment
```{r}
herb_plot_in <- function(loc, key, key2) { 
  herb_spp <- subset(sum_herb_in, site == loc & treatment_key == key | site== loc & treatment_key == key2)
  return(ggplot(herb_spp, aes(x = treatment_key, y = avg_eaten, fill = treatment_key)) +
           facet_grid(.~year) +
           geom_bar(position = "identity", stat = "identity") +
           geom_errorbar(aes(ymin = avg_eaten - se, ymax = avg_eaten + se), width = 0.2,
                         position = "identity") +
           labs(x = "Treatment", y = "Average Percent of Leaf Eaten", title = loc) +
           scale_fill_manual(values = c("#a6bddb", "#fb6a4a")) +
           theme_classic())
}
herb_plot_in("umbs", "A0", "W0")
herb_plot_in("kbs", "A0", "W0")
herb_plot_in("umbs", "AI", "WI")
herb_plot_in("kbs", "AI", "WI")
```

### Makes a plot for a given site
```{r}
herb_plot_in <- function(loc) { 
  herb_spp <- subset(sum_herb_in, site == loc)
  return(ggplot(herb_spp, aes(x = treatment_key, y = avg_eaten, fill = treatment_key)) +
           facet_grid(.~year) +
           geom_bar(position = "identity", stat = "identity") +
           geom_errorbar(aes(ymin = avg_eaten - se, ymax = avg_eaten + se), width = 0.2,
                         position = "identity") +
           labs(x = "Treatment", y = "Average Percent of Leaf Eaten", title = loc) +
           scale_fill_manual(values = c("#abd9e9", "#2c7bb6", "#fdae61", "#d7191c")) +
           theme_classic())
}
herb_plot_in("umbs")
herb_plot_in("kbs")
```

### Makes a plot for a given species and site
### Ex: Popr at kbs, Eugr at kbs
```{r}
herb_plot <- function(spp, loc) { 
  herb_spp <- subset(sum_herb, species == spp & site == loc)
  return(ggplot(herb_spp, aes(x = treatment_key, y = avg_eaten, fill = treatment_key)) +
           facet_grid(.~year) +
           geom_bar(position = "identity", stat = "identity") +
           geom_errorbar(aes(ymin = avg_eaten - se, ymax = avg_eaten + se), width = 0.2,
                         position = "identity") +
           labs(x = "State", y = "Average Percent of Leaf Eaten", title = spp) +
           #scale_fill_manual(values = c("#a6bddb", "#fb6a4a")) +
           #scale_x_discrete(labels=c("ambient" = "A", "warmed" = "W")) +
           theme_classic())
}
herb_plot("Soca", "kbs")
```