---
title: "warmXtrophic Project: Greenup Analyses"
author: "Kara Dobson, Phoebe Zarnetske"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=80),tidy=TRUE)
```
COLLABORATORS:  Moriah Young, Mark Hammond, Pat Bills  
DATA INPUT:     Data imported as csv files from shared Google drive "SpaCE_Lab_warmXtrophic" plant comp folder  
DATA OUTPUT:    a brief description of the data output from through the script, including what format it’s in
OVERVIEW:       This script explores and analyses the greenup data from the KBS and UMBS stations. Greenup was recorded as percent cover of each species in each plot every 3 days in the beginning of the field season. Percent cover continued to be measured until the end of the field season. For calculating greenup dates, we use the date at which 50% of a species maximum percent cover was reached for each plot, to account for variation in litter depth affecting the early season's first greenup date.
REQUIRES:       Kara: elaborate on this to include any scripts or code sources that are required:                    Prior to running this script, run plant_comp_clean_L0.R, which creates greenup.csv ...
NOTES: 
The green_kbs and green_umbs dataframes contain data for greenup at each site.  
"half_cover_date" is the date at which 50% of a species max cover was reached (per plot, per year)  
"state" describes each treatment - warmed or ambient  

```{r, message = F}
# Clear all existing data
rm(list=ls())

#Load packages
library(tidyverse)
library(lme4)
library(olsrr)
library(predictmeans)
library(car)
library(fitdistrplus)
library(ggpubr)
library(rstatix)
library(vegan)

# Set working directory to Google Drive
# **** Update with the path to your Google drive on your computer
setwd("/Volumes/GoogleDrive/Shared drives/SpaCE_Lab_warmXtrophic/data/")

# Read in plant comp data - edit this in "plant_comp_clean_L0.R" so it's not an issue here with extra column
greenup <- read.csv("L1/greenup/final_greenup_L1.csv")
# check variable types
str(greenup)
# KARA: omit this once fixed in plant_comp_clean_L0.R
# remove first column 
greenup$X<-NULL

# adding sequential year variable starting at 1
greenup$year1<-greenup$year
greenup$year[greenup$year == 2015] <- 1
greenup$year[greenup$year == 2016] <- 2
greenup$year[greenup$year == 2017] <- 3
greenup$year[greenup$year == 2018] <- 4
greenup$year[greenup$year == 2019] <- 5
greenup$year[greenup$year == 2020] <- 6

# create dataframes for kbs and umbs
green_kbs <- subset(greenup, site == "kbs")
green_umbs <- subset(greenup, site == "umbs")
```

# Data Exploration: are there differences between warmed vs. ambient plots ?
# *Starting with KBS*

# First, checking for normality in raw data. 
```{r}
hist(green_kbs$half_cover_date)
qqnorm(green_kbs$half_cover_date)
shapiro.test(green_kbs$half_cover_date)

# histograms for each treatment separately - look almost identical
hist(green_kbs$half_cover_date[green_kbs$state == "ambient"])
hist(green_kbs$half_cover_date[green_kbs$state == "warmed"])

# histograms for each year
hist(green_kbs$half_cover_date[green_kbs$year == 1])
hist(green_kbs$half_cover_date[green_kbs$year == 2])
hist(green_kbs$half_cover_date[green_kbs$year == 3])
hist(green_kbs$half_cover_date[green_kbs$year == 4])
hist(green_kbs$half_cover_date[green_kbs$year == 5])
hist(green_kbs$half_cover_date[green_kbs$year == 6])
# looks like the 225 spike is from 2018 and 2020 
kbs_2018 <- subset(green_kbs, year == 4) # many records on 235
kbs_2020 <- subset(green_kbs, year == 6) # records from 227 & 228
```

\pagebreak

# Leverage plots? Not sure what these show exactly, but we discussed them previously
```{r}
# checking fit for date as a function of state
fit <- lm(half_cover_date~state, data = green_kbs)
hist(fit$residuals)
leveragePlots(fit)

# checking fit for date as a function of state and year
fit2 <- lm(half_cover_date~state+year, data = green_kbs)
hist(fit2$residuals)
leveragePlots(fit2)
```

\pagebreak

# Not normal, and previously attempted transformations don't help (in R script)

# Seeing what other distribution could fit
```{r}
descdist(green_kbs$half_cover_date, discrete = FALSE)
```
# While uniform looks the closest, I'll try poisson
```{r, fig.align='center'}
fit <- lm(half_cover_date~state, data = green_kbs)
residual <- fit$residuals
hist(residual, main="Raw residuals")  
pois <- glm(half_cover_date~state, data = green_kbs, family="poisson")
hist(pois$residuals, main="Poisson glm residuals")
```

\pagebreak

# Below I try a few different generalized linear models with poisson distribution:  

## An interaction between state and year, plus insecticide as a fixed effect and species and plot as random effects
```{r}
moda <- glmer(half_cover_date ~ state*year + insecticide + (1|species) + (1|plot),
              data=green_kbs, family = poisson)
summary(moda)
```

\pagebreak

## No interaction between state and year, but with state and insecticide as fixed effects and species and plot as random effects
```{r}
modb <- glmer(half_cover_date ~ state + year + insecticide + (1|species) + (1|plot),
              data=green_kbs, family = poisson)
summary(modb)
```

\pagebreak

## State and insecticide as fixed effects & year, species and plot as random effects
```{r}
modc <- glmer(half_cover_date ~ state + insecticide + (1|year) + (1|species) + (1|plot),
              data=green_kbs, family = poisson)
summary(modc)
```

\pagebreak

# Because no distributions seems to match well, I'll try a Friedman's test
```{r}
#friedman_kbs <- green_kbs %>% 
#  friedman_test(half_cover_date ~ state)
```
## Error: Must extract column with a single valid subscript. x Subscript `var` can't be `NA`
## Can't figure out what this means  

## If I include the blocks portion of the formula (from the documentation) I get this error
```{r}
#friedman_kbs <- green_kbs %>% 
#  friedman_test(half_cover_date ~ state | plot)
```
## Error in friedman.test.default(c(141L, 202L, 122L, 101L, 127L, 120L, 197L,  :  not an unreplicated complete block design

# Permanova?
```{r, cache = T}
per1 <- adonis2(green_kbs$half_cover_date ~ state*year + insecticide, data = green_kbs)
per1
per2 <- adonis(formula = green_kbs$half_cover_date ~ state*year + insecticide, strata = green_kbs$plot, data = green_kbs)
per2
```
# With per2, when controlling for "plot", there is a difference btwn treatments

\pagebreak

# *UMBS*

# Checking for normality

```{r}
hist(green_umbs$half_cover_date)
qqnorm(green_umbs$half_cover_date)
shapiro.test(green_umbs$half_cover_date)
hist(green_umbs$half_cover_date[green_kbs$state == "ambient"])
hist(green_umbs$half_cover_date[green_kbs$state == "warmed"])
```

## These look pretty good

\pagebreak

# Trying log transformation
```{r}
green_umbs$date_log <- log(green_umbs$half_cover_date)
hist(green_umbs$date_log)
shapiro.test(green_umbs$date_log)
```
## I think this looks good but shapiro-wilk is lower than 0.05

\pagebreak

# Trying inverse tranformation
```{r}
green_umbs$date_inv <- 1/(green_umbs$half_cover_date)
hist(green_umbs$date_inv)
shapiro.test(green_umbs$date_inv)
```
## This also looks good but is also still low for shapiro-wilk