---
title: "Greenup Plots"
author: "Kara Dobson"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
---

COLLABORATORS: Phoebe Zarnetske, Mark Hammond, Pat Bills, Moriah Young  
DATA INPUT: Clean & plot plant comp csv from the shared Google drive  
DATA OUTPUT: Code and Rmd are in the scripts folder in Github  
PROJECT: warmXtrophic 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = F, message=FALSE, warning=FALSE}
# Clear all existing data
rm(list=ls())

# Load packages
library(tidyverse)
library(plotrix)

# Set working directory to Google Drive
# **** Update with the path to your Google drive on your computer
setwd("/Volumes/GoogleDrive/Shared drives/SpaCE_Lab_warmXtrophic/data/")

# Read in plant comp data
comp <- read.csv("L1/plant_composition/final_plantcomp_L1.csv")
#str(comp)

# remove uneeded X column
comp$X <- NULL



# filter data to contain the averages and std error for each site
sum_comp_site <- comp %>%
  group_by(site, state, year) %>%
  summarize(avg_cover = mean(cover, na.rm = TRUE),
            se = std.error(cover, na.rm = TRUE))

# by plant origin (native/exotic) - no grouping by year
sum_comp_org <- comp %>%
  group_by(site, state, origin) %>%
  summarize(avg_comp = mean(cover, na.rm = TRUE),
            se = std.error(cover, na.rm = TRUE))
sum_comp_org <- subset(sum_comp_org, origin == "Exotic" | origin == "Native")

# by plant origin (native/exotic) - grouped by year
sum_compyear_org <- comp %>%
  group_by(site, state, origin, year) %>%
  summarize(avg_comp = mean(cover, na.rm = TRUE),
            se = std.error(cover, na.rm = TRUE))
sum_compyear_org <- subset(sum_compyear_org, origin == "Exotic" | origin == "Native")

# by plant growth habit (forb, graminoid, shrub)
sum_comp_habit <- comp %>%
  group_by(site, growth_habit, state) %>%
  summarize(avg_cover = mean(cover, na.rm = TRUE),
            se = std.error(cover, na.rm = TRUE))
sum_comp_habit <- subset(sum_comp_habit, growth_habit == "Forb" | growth_habit == "Graminoid")
```

## Plot for plant cover between warmed and ambient treatments
```{r, echo = F}
# Plot for all species between warmed and ambient
comp_plot_all <- function(loc) { 
  comp_spp <- subset(sum_comp_site, site == loc)
  return(ggplot(comp_spp, aes(x = state, y = avg_cover, fill = state)) +
           facet_grid(.~year) +
           geom_bar(position = "identity", stat = "identity", color = "black") +
           geom_errorbar(aes(ymin = avg_cover - se, ymax = avg_cover + se), width = 0.2,
                         position = "identity") +
           labs(x = "State", y = "Percent Cover", title = loc) +
           scale_fill_manual(values = c("#a6bddb", "#fb6a4a")) +
           scale_x_discrete(labels=c("ambient" = "A", "warmed" = "W")) +
           theme_classic())
}
comp_plot_all("kbs")
comp_plot_all("umbs")
```

## Plot for percent cover between native/exotic for warmed and ambient
```{r, echo = F}
comp_plot_org <- function(loc) { 
  comp_spp <- subset(sum_comp_org, site == loc)
  return(ggplot(comp_spp, aes(x = origin, y = avg_comp, fill = state)) +
           #facet_grid(.~year) +
           geom_bar(position = "dodge", stat = "identity", color = "black") +
           geom_errorbar(aes(ymin = avg_comp - se, ymax = avg_comp + se), width = 0.2,
                         position = position_dodge(0.9)) +
           labs(x = "State", y = "Percent Cover", title = loc) +
           scale_fill_manual(values = c("#a6bddb", "#fb6a4a")) +
           scale_x_discrete(labels=c("ambient" = "A", "warmed" = "W")) +
           theme_classic())
}
comp_plot_org("kbs")
comp_plot_org("umbs")
```

## Plot for percent cover between native/exotic for warmed and ambient, separated by year
```{r, echo = F}
compyear_plot_org <- function(loc) { 
  comp_spp <- subset(sum_compyear_org, site == loc)
  return(ggplot(comp_spp, aes(x = origin, y = avg_comp, fill = state)) +
           facet_grid(.~year) +
           geom_bar(position = "dodge", stat = "identity", color = "black") +
           geom_errorbar(aes(ymin = avg_comp - se, ymax = avg_comp + se), width = 0.2,
                         position = position_dodge(0.9)) +
           labs(x = "State", y = "Percent Cover", title = loc) +
           scale_fill_manual(values = c("#a6bddb", "#fb6a4a")) +
           scale_x_discrete(labels=c("Exotic" = "E", "Native" = "N")) +
           theme_classic())
}
compyear_plot_org("kbs")
compyear_plot_org("umbs")
```

## Plot for percent cover between forb/graminoid for warmed and ambient
```{r, echo = F}
comp_plot_habit <- function(loc) { 
  comp_spp <- subset(sum_comp_habit, site == loc)
  return(ggplot(comp_spp, aes(x = growth_habit, y = avg_cover, fill = state)) +
           #facet_grid(.~year) +
           geom_bar(position = "dodge", stat = "identity", color = "black") +
           geom_errorbar(aes(ymin = avg_cover - se, ymax = avg_cover + se), width = 0.2,
                         position = position_dodge(0.9)) +
           labs(x = "State", y = "Percent Cover", title = loc) +
           scale_fill_manual(values = c("#a6bddb", "#fb6a4a")) +
           theme_classic())
}
comp_plot_habit("kbs")
comp_plot_habit("umbs")  
```