---
title: "warmXtrophic: Phenology Analysis"
author: "Moriah Young"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
---

COLLABORATORS:  Kara Dobson, Phoebe Zarnetske, Mark Hammond, Pat Bills  
DATA INPUT:     Data imported as csv files from shared Google drive "SpaCE_Lab_warmXtrophic" plant comp folder  
DATA OUTPUT:    a brief description of the data output from through the script, including what format it’s in
OVERVIEW:       This script ... 
REQUIRES:       any scripts or code sources that are required
NOTES: 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy.opts = list(width.cutoff=80), tidy=TRUE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Clear existing data
rm(list=ls())

#Load packages
library(tidyverse)
library(lme4)
library(lmerTest)
library(emmeans)
library(vegan)
library(car)
library(rstatix)
library(scales)
library(fitdistrplus)
library(moments)# for calculating skewness of data
library(ggpubr)
library(jtools) # summ() function
library(predictmeans)

# Set working directory to Google Drive
setwd("/Volumes/GoogleDrive/Shared drives/SpaCE_Lab_warmXtrophic/data/")

# Read in data
# cleaned phenology data from L1
phen_data <- read.csv("/Volumes/GoogleDrive/Shared drives/SpaCE_Lab_warmXtrophic/data/L1/phenology/final_flw_sd_L1.csv", stringsAsFactors=FALSE)
phen_data$X <- NULL # get rid of "X" column that shows up
View(phen_data) # take a look at the data to see if looks good

# make a column that breaks down years as 1, 2, 3, 4, 5, 6 and into factors
phen_data$year_factor <- 
        ifelse(phen_data$year == 2015, "1",
               ifelse(phen_data$year == 2016, "2",
                      ifelse(phen_data$year == 2017, "3",
                             ifelse(phen_data$year == 2018, "4",
                                    ifelse(phen_data$year == 2019, "5",
                                           ifelse(phen_data$year == 2020, "6", NA))))))

# Create separate data frames for flowering and seeding
phen_flwr <- subset(phen_data, action == "flower")
phen_sd <- subset(phen_data, action == "seed")

# Separate data frame by site for flower
kbs_flwr <- subset(phen_flwr, site == "kbs")
umbs_flwr <- subset(phen_flwr, site == "umbs")

# Separate data frame by site for seed set
kbs_sd <- subset(phen_sd, site == "kbs")
umbs_sd <- subset(phen_sd, site == "umbs")

# Filter data to contain the date of first flower for each species at each plot
FirstFlower_all <- phen_flwr %>%
        group_by(plot, year, species, state, site, action, origin, insecticide, year_factor) %>%
        summarize(julian = min(julian, na.rm=T))

MedianFlwr_all <- phen_flwr %>%
        group_by(plot, year, species, state, site, action, origin, insecticide, year_factor, treatment_key) %>%
        summarize(julian = median(julian, na.rm=T))
```

## UMBS Data

First Flower
```{r}
umbs_firstflwr <- subset(FirstFlower_all, site == "umbs") # pull out umbs only data

# first flower
hist(umbs_firstflwr$julian) # poisson?
qqnorm(umbs_firstflwr$julian)
fit1 <- lm(julian ~ state, data = umbs_firstflwr)
qqPlot(fit1)
shapiro.test(umbs_firstflwr$julian) 

descdist(umbs_firstflwr$julian, discrete = FALSE) # uniform? normal?

skewness(umbs_firstflwr$julian)

umbs_firstflwr$sqrt_julian <- sqrt(umbs_firstflwr$julian)
hist(umbs_firstflwr$sqrt_julian)
skewness(umbs_firstflwr$sqrt_julian, na.rm = TRUE)
```

Median Flower 
```{r}
umbs_flwr_med <- subset(MedianFlwr_all, site == "umbs") # pull out umbs only data

hist(umbs_flwr_med$julian)
#explore what's going on around julian date 190

umbs_flwr_explore <- subset(umbs_flwr_med, julian > 190)
View(umbs_flwr_explore) 
# I'm not really seeing anything glaring here when I look at the data. Cest is the most dominate species it
# appears. I'm going to pull out all the Cest data and look at it separately. 

umbs_flwr_cest <- subset(umbs_flwr_med, species == "Cest") #just Cest median flower date
View(umbs_flwr_cest)
hist(umbs_flwr_cest$julian) # this looks pretty normal to me - slight left skew in the data

# Looking into making early and late flowering categories for species (make new column)
umbs_median_species <- umbs_flwr %>% #median date of flower for each species across EVERYTHING at umbs
        group_by(species) %>%
        summarize(julian = median(julian, na.rm=T))
View(umbs_median_species)
summary(umbs_median_species) # Mean = 183.2 

# Not sure if this is the best way to do this, but I calculated the median date of flower for each species 
# across all years and plots then took the summary of those median dates, took the mean (183.2) and made a
# new column for a flower group broken into "early" (species with a median date less than 183) and "late" 
# (species with a median date greater than 183) median flower date. 
umbs_median_species$flwr_group <- with(umbs_median_species, 
                                       ifelse(julian >= 183, "late", "early"))
View(umbs_median_species)
# "early" species: Acru, Ansp, Cape, Dasp, Frve, Hiau, Hica, Hipi, Hisp, Poco, Popr, Ruac, Rual, Spsp, Trdu, Unknown, Vaan, Vear
# "late" species: Anma, Apan, Assy, Asun, Cest, Elre, Gnul, Hype, Sogi, Sone, Sohi, Syla

umbs_flwr_med$flwr_group <- with(umbs_flwr_med, 
                                       ifelse(species == c("Acru", "Ansp", "Cape", "Dasp", "Frve", "Hiau", "Hica", "Hipi", "Hisp", "Poco", "Popr", "Ruac", "Rual", "Spsp", "Trdu", "Unknown", "Vaan", "Vear"), "early", "late"))

# histograms for both "early" and "late" flower group separately 
hist(umbs_flwr_med$julian[umbs_flwr_med$flwr_group == "early"])
hist(umbs_flwr_med$julian[umbs_flwr_med$flwr_group == "late"])
# I don't think this did what I wanted it too..... I need a break.



qqnorm(umbs_flwr_med$julian)
fit2 <- lm(julian ~ state, data = umbs_flwr_med)
qqPlot(fit2)
shapiro.test(umbs_flwr_med$julian)

descdist(umbs_flwr_med$julian, discrete = FALSE) # uniform?

skewness(umbs_flwr_med$julian, na.rm = TRUE)  # slight positive skew

# Because of slight positive skew I'm transforming the data by taking the square root.
umbs_flwr_med$sqrt_julian <- sqrt(umbs_flwr_med$julian) 
hist(umbs_flwr_med$sqrt_julian)
skewness(umbs_flwr_med$sqrt_julian, na.rm = TRUE)

ggdensity(umbs_flwr_med, x = "sqrt_julian", fill = "lightgray", title = "CONT") +
        stat_overlay_normal_density(color = "red", linetype = "dashed") 
# there's that weird dip around 14 on the x-axis
```

Models!
What is the effect of warming (state) on the median date of flower?
```{r}
# Models using median flowering julian date

# Null model?
model0 <- lmer(sqrt_julian ~ 1 # This means median flower date (julian) is predicted by the intercept
               + (1|plot), data = umbs_flwr_med, REML=FALSE) # each plot gets its own intercept?
summary(model0) 

# The effects of warming/ambient (state) on median date of flowering (julian) with species (species) as a random effect.
model1 <- lmer(sqrt_julian ~ state + (1|species), data = umbs_flwr_med, REML=FALSE)
summary(model1)

# The effects of warming/ambient on median date of flowering (julian - response variable) with an interaction between warming and native vs exotic species (origin) with species and year as a random effect.
model2 <- lmer(sqrt_julian ~ state*origin + (1|species) + (1|year_factor), data = umbs_flwr_med, REML=FALSE)
summary(model2)

# The effects of warming/ambient on median date of flowering with origin and insecticide as fixed effects and species and year_factor as random effects
model3 <- lmer(sqrt_julian ~ state + origin + insecticide + (1|species) + (1|year_factor), data = umbs_flwr_med, REML=FALSE)
summary(model3)
summ(model3)
```

Permanova
```{r}
#perma1 <- adonis2(julian ~ state + (1|species), data = umbs_flwr_med, permutations = 999, by = "terms")
#summary(perma1)

permanova.lmer(model1)
```

Updated from Kileigh's script
```{r, warning=FALSE}
#should plot be a random effect?
moda <- lmer(sqrt_julian ~ state*origin + insecticide + (1|species) + (1|plot), umbs_flwr_med, REML=FALSE)
mod1 <- lmer(sqrt_julian ~ state*origin + insecticide + (1|species), umbs_flwr_med, REML=FALSE)
anova(mod1,moda)
AIC(mod1, moda)

#Do we need insects?
mod2 <- lmer(sqrt_julian ~ state*origin + (1|species), umbs_flwr_med, REML=FALSE)
anova(mod2,mod1)

#Do we need interaction term?
mod3 <- lmer(sqrt_julian ~ state + origin + (1|species), umbs_flwr_med, REML=FALSE)
anova(mod3, mod2)
anova(mod3)
summary(mod3)
confint(mod3, method="boot", nsim=999)
difflsmeans(mod3, test.effs=NULL, ddf="Satterthwaite")
```

## KBS Data

```{r}
kbs_firstflwr <- subset(FirstFlower_all, site == "kbs") # pull out kbs only data

kbs_flwr_med <- subset(MedianFlwr_all, site == "kbs") # pull out kbs only data

hist(kbs_flwr_med$julian)
qqnorm(kbs_flwr_med$julian)
fit2 <- lm(julian ~ state, data = umbs_flwr_med)
qqPlot(fit2)
shapiro.test(kbs_flwr_med$julian) #not normal

descdist(kbs_flwr_med$julian, discrete = FALSE) # beta?

skewness(kbs_flwr_med$julian, na.rm = TRUE)  # slight positive skew

ggdensity(kbs_flwr_med, x = "julian", fill = "lightgray", title = "CONT") +
        stat_overlay_normal_density(color = "red", linetype = "dashed") 

# Because of positive skew I'm transforming the data by taking the square root
kbs_flwr_med$sqrt_julian <- sqrt(kbs_flwr_med$julian) 
hist(kbs_flwr_med$sqrt_julian)
skewness(kbs_flwr_med$sqrt_julian, na.rm = TRUE)

# Now try the log of julian dates
kbs_flwr_med$log_julian <- log10(kbs_flwr_med$julian)
hist(kbs_flwr_med$log_julian)
skewness(kbs_flwr_med$log_julian, na.rm = TRUE)

# Ok now try the inverse
kbs_flwr_med$inv_julian <- 1/(kbs_flwr_med$julian)
hist(kbs_flwr_med$inv_julian)
skewness(kbs_flwr_med$inv_julian, na.rm = TRUE) 
#taking the inverse of the julian dates made the data the least skewed
qqnorm(kbs_flwr_med$inv_julian)
shapiro.test(kbs_flwr_med$inv_julian) #still not normal

#pois <- glm(julian ~ state, data = kbs_flwr_med, family="poisson")
#hist(pois$residuals)
```

