---
title: "warmXtrophic Project: Flowering Duration Phenology Analyses"
author: "Moriah Young & Kara Dobson"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load in packages & data
```{r, message = F}
# clear all existing data
rm(list=ls())

#Load packages
library(tidyverse)
library(ggplot2)
library(lme4)
library(lmerTest)
library(emmeans)
library(vegan)
library(car)
library(rstatix)
library(scales)
library(fitdistrplus)
library(moments)# for calculating skewness of data
library(ggpubr)
library(jtools) # summ() function
library(predictmeans)
library(olsrr)
library(car)
library(fitdistrplus)
library(ggpubr)
library(interactions)
library(sjPlot)
library(effects)
library(glmmTMB)
library(GGally) # ggpairs() function
library(bbmle) # AICtab() function

# Set working directory
Sys.getenv("L1DIR")
L0_dir <- Sys.getenv("L0DIR")
L1_dir <- Sys.getenv("L1DIR")
L2_dir <- Sys.getenv("L2DIR")

# Read in data
flwr_plot <- read.csv(file.path(L2_dir, "phenology/final_flwr_plot_L2.csv")) # plot level data

# species level flower duration
# note: the species data frame is only used to demonstrate that flwr duration differs between species (for the supplement)
flwr_spp <- read.csv(file.path(L2_dir, "phenology/final_flwr_species_L2.csv")) 
flwr_spp$julian_min <- as.numeric(as.integer(flwr_spp$julian_min))

# Order warmed and ambient so that warmed shows up first in plotting (and is default is red = warmed; blue = ambient). First make it a factor
#flwr_plot$state <- as.factor(flwr_plot$state)
#levels(flwr_plot$state)
# [1] "ambient" "warmed" 
#flwr_plot$state <- factor(flwr_plot$state, levels(flwr_plot$state)[c(2,1)])
#levels(flwr_plot$state)
# [1] "warmed"  "ambient"

flwr_plot$flwr_duration <- as.numeric(as.character(flwr_plot$flwr_duration))
flwr_spp$flwr_duration <- as.numeric(as.character(flwr_spp$flwr_duration))

# delete 2021 data from umbs dataframe - doesn't make sense to have bc we can't get min flower/sd, median flower, or 
# flower duration from 2021 UMBS since data was not collected at the same frequency as previous years
#flwr_plot1 <- flwr_plot[-which(flwr_plot$year == "2021" & flwr_plot$site == "umbs"),]
#flwr_spp <- flwr_spp[-which(flwr_spp$year == "2021" & flwr_plot$site == "umbs"),]

# also delete date 2015 data since some species may have started flowering before chambers were placed
flwr_plot <- flwr_plot[-which(flwr_plot$year == "2015" & flwr_plot$site == "kbs"),]
flwr_spp <- flwr_spp[-which(flwr_spp$year == "2015" & flwr_plot$site == "kbs"),]

# Kara's edits:
# adding 1 to each occurrence in the flwr_duration column so everything is scaled up the same & removes 0's
flwr_plot$flwr_duration_scaled <- flwr_plot$flwr_duration+1
flwr_spp$flwr_duration_scaled <- flwr_spp$flwr_duration+1

umbs_flwr_plot <- subset(flwr_plot, site == "umbs") # pull out umbs only data at plot level
kbs_flwr_plot <- subset(flwr_plot, site == "kbs") # pull out kbs only data at plot level

# species specific dataframes for kbs and umbs
flwr_kbs_spp <- subset(flwr_spp, site == "kbs")
flwr_umbs_spp <- subset(flwr_spp, site == "umbs")
```

# KBS PLOT LEVEL - Looking at DURATION of flowering
```{r}
### KBS ###
hist(kbs_flwr_plot$flwr_duration_scaled)
qqnorm(kbs_flwr_plot$flwr_duration_scaled)
shapiro.test(kbs_flwr_plot$flwr_duration_scaled) # not normally distributed but close

# Exploring distributions for these data:
descdist(kbs_flwr_plot$flwr_duration_scaled, discrete = FALSE)

# Gamma distribution 
fit.gamma <- fitdist(kbs_flwr_plot$flwr_duration_scaled, "gamma")
plot(fit.gamma)

# Weibull distribution
fit.weibull  <- fitdist(kbs_flwr_plot$flwr_duration_scaled, "weibull")
plot(fit.weibull)

# Lognormal distribution 
fit.ln <- fitdist(kbs_flwr_plot$flwr_duration_scaled, "lnorm")
plot(fit.ln)

par(mfrow=c(2,2))
plot.legend <- c("Gamma", "Weibull", "Log Normal")
denscomp(list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
cdfcomp (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
qqcomp  (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
ppcomp  (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
dev.off()

# Goodness of fit comparisons across fits (can't include the log normal bc it becomes diff response values)
gofstat(list(fit.gamma, fit.weibull, fit.ln), fitnames = c("Gamma", "Weibull", "Log Normal"))
# weibull, probably going with log transformation
```

# UMBS PLOT LEVEL - Looking at DURATION of flowering
```{r, warning=F, error=TRUE}
### UMBS ###
hist(umbs_flwr_plot$flwr_duration_scaled)
qqnorm(umbs_flwr_plot$flwr_duration_scaled)
shapiro.test(umbs_flwr_plot$flwr_duration_scaled) # pvalue is > 0.05 so we accept the null hypothesis that the data is normally distributed! WOO

# Exploring distributions for these data:
descdist(umbs_flwr_plot$flwr_duration_scaled, discrete = FALSE)

# Gamma distribution 
fit.gamma <- fitdist(umbs_flwr_plot$flwr_duration_scaled, "gamma")
plot(fit.gamma)

# Weibull distribution
fit.weibull  <- fitdist(umbs_flwr_plot$flwr_duration_scaled, "weibull")
plot(fit.weibull)

# Lognormal distribution 
fit.ln <- fitdist(umbs_flwr_plot$flwr_duration_scaled, "lnorm")
plot(fit.ln)

par(mfrow=c(2,2))
plot.legend <- c("Gamma", "Weibull", "Log Normal")
denscomp(list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
cdfcomp (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
qqcomp  (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
ppcomp  (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
dev.off()

# Goodness of fit comparisons across fits (can't include the log normal bc it becomes diff response values)
gofstat(list(fit.gamma, fit.weibull, fit.ln), fitnames = c("Gamma", "Weibull", "Log Normal"))
# Weibull, probably going with log transformation
```

# Leverage plots and detecting Outliers. https://www.statmethods.net/stats/rdiagnostics.html
# These illustrate whether certain data points have more leverage (more influence), and thus could be outliers. It's a way of detecting outliers. Leverage plots can help identify whether a point has high or low influence, based on its leverage and residual and determining model fit with and without the point in question. Ultimately you decide whether the points are outliers or not, based on the knowledge of the system and how much it changes the model when included vs. excluded from the data used to fit the model. Here is a good overview of the combination of leverage and residual: scroll down to sections beginning at "13.3 Unusual Observations": https://daviddalpiaz.github.io/appliedstats/model-diagnostics.html

```{r}
# Plot level data
# KBS State-only model
fit_plot_state_kbs <- lm(log(flwr_duration_scaled) ~ state, data = kbs_flwr_plot)
outlierTest(fit_plot_state_kbs) # line 52 is an outlier
qqPlot(fit_plot_state_kbs, main="QQ Plot") 
hist(fit_plot_state_kbs$residuals)
leveragePlots(fit_plot_state_kbs)
ols_test_normality(fit_plot_state_kbs) # looks ok besides Kolmogorov-Smirnov test

# KBS State and year model
fit_plot_stateyear_kbs <- lm(log(flwr_duration_scaled) ~ state + year_factor, data = kbs_flwr_plot)
outlierTest(fit_plot_stateyear_kbs) # outliers - 142, 155, 220, 116, 207, 272
qqPlot(fit_plot_stateyear_kbs, main="QQ Plot") 
hist(fit_plot_stateyear_kbs$residuals)
leveragePlots(fit_plot_stateyear_kbs)
ols_test_normality(fit_plot_stateyear_kbs) # a couple tests say not normal 

# Interaction plot (ignore for now the repeated measures with species); see: https://cran.r-project.org/web/packages/interactions/vignettes/interactions.html and: https://interactions.jacob-long.com/

fit3 <- lm(log(flwr_duration_scaled) ~ state * year, data = kbs_flwr_plot)
interact_plot(fit3, pred = year, modx = state) # this looks very strange to me
```

```{r}
# Plot level data
# UMBS State-only model
fit_plot_state_umbs <- lm(flwr_duration_scaled ~ state, data = umbs_flwr_plot)
outlierTest(fit_plot_state_umbs) # no outliers
qqPlot(fit_plot_state_umbs, main="QQ Plot") 
hist(fit_plot_state_umbs$residuals)
leveragePlots(fit_plot_state_umbs)
ols_test_normality(fit_plot_state_umbs) # these look good

# UMBS State and year model
fit_plot_stateyear_umbs <- lm(flwr_duration_scaled ~ state + year_factor, data = umbs_flwr_plot)
outlierTest(fit_plot_stateyear_umbs) # no outliers
qqPlot(fit_plot_stateyear_umbs, main="QQ Plot") 
hist(fit_plot_stateyear_umbs$residuals)
leveragePlots(fit_plot_stateyear_umbs)
ols_test_normality(fit_plot_stateyear_umbs) # these look good except Kolmogorov-Smirnov which is > 0.05

# Interaction plot (ignore for now the repeated measures with species); see: https://cran.r-project.org/web/packages/interactions/vignettes/interactions.html and: https://interactions.jacob-long.com/

fit3 <- lm(flwr_duration_scaled ~ state * year, data = umbs_flwr_plot)
interact_plot(fit3, pred = year, modx = state)
```

# KBS Plot-level Mixed Effects Models:
```{r, error=TRUE}
mod1p <- lmer(flwr_duration_scaled ~ state + (1|plot), kbs_flwr_plot, REML=FALSE)
mod2p <- lmer(flwr_duration_scaled ~ insecticide + (1|plot), kbs_flwr_plot, REML=FALSE)
mod3p <- lmer(flwr_duration_scaled ~ insecticide + state + (1|plot), kbs_flwr_plot, REML=FALSE)
mod4p <- lmer(flwr_duration_scaled ~ insecticide * state + (1|plot), kbs_flwr_plot, REML=FALSE)
mod5p <- lmer(flwr_duration_scaled ~ state + year_factor + (1|plot), kbs_flwr_plot, REML=FALSE)
mod6p <- lmer(flwr_duration_scaled ~ state + year_factor + insecticide + (1|plot), kbs_flwr_plot, REML=FALSE)
mod7p <- lmer(flwr_duration_scaled ~ state * year_factor + (1|plot), kbs_flwr_plot, REML=FALSE)
mod8p <- lmer(flwr_duration_scaled ~ state * year_factor + insecticide + (1|plot), kbs_flwr_plot, REML=FALSE)
mod9p <- lmer(flwr_duration_scaled ~ state * insecticide + year_factor + (1|plot), kbs_flwr_plot, REML=FALSE)
mod10p <- lmer(flwr_duration_scaled ~ state + insecticide * year_factor + (1|plot), kbs_flwr_plot, REML=FALSE)
mod11p <- lmer(flwr_duration_scaled ~ state * year_factor * insecticide + (1|plot), kbs_flwr_plot, REML=FALSE)
AICctab(mod1p, mod2p, mod3p, mod4p, mod5p, mod6p, mod7p, mod8p, mod9p, mod10p, mod11p, weights=T) # model 5p and 7p are the top models
anova(mod1p, mod2p) # go with 1p although they are nearly identical
AICctab(mod1p, mod2p, weights=T)
anova(mod1p)
summary(mod1p)

# Our hypothesized model is  mod9p
summary(mod9p)
anova(mod9p)

# comparisons
emmeans(mod9p, list(pairwise ~ state*insecticide), adjust = "tukey")
mod9p.emm <- emmeans(mod9p, ~ state*insecticide)
contrast(mod9p.emm, "consec", simple = "each", combine = F, adjust = "mvt")
emmip(mod9p, insecticide~state)

# making a table
kable(anova(mod9p)) %>% kableExtra::kable_styling()

# adding in our temp data into some models
# note: including state, year, and temp data into a model leads to rank deficiency
# so below, we test for date of first flower as a function of just temp to see how real temp data affects date of first flower
kbs_flwr_plot_amb <- kbs_flwr_plot %>%
        filter(state == "ambient")
modtest1u <- lmer(flwr_duration_scaled ~ mean_temp + (1|plot), kbs_flwr_plot, REML=FALSE)
modtest2u <- lmer(flwr_duration_scaled ~ GDD_cumulative + (1|plot), kbs_flwr_plot, REML=FALSE)

AICtab(modtest1u, modtest2u, weights=T)
# growing degree days model is better but we are going to use the mean temp for all models

anova(modtest1u)
summary(modtest1u)

# making a table
kable(anova(modtest1u)) %>% kableExtra::kable_styling()
```

# UMBS Plot-level Mixed Effects Models:
```{r}
mod1up <- lmer(flwr_duration_scaled ~ state + (1|plot), umbs_flwr_plot, REML=FALSE)
mod2up <- lmer(flwr_duration_scaled ~ insecticide + (1|plot), umbs_flwr_plot, REML=FALSE)
mod3up <- lmer(flwr_duration_scaled ~ insecticide + state + (1|plot), umbs_flwr_plot, REML=FALSE)
mod4up <- lmer(flwr_duration_scaled ~ insecticide * state + (1|plot), umbs_flwr_plot, REML=FALSE)
mod5up <- lmer(flwr_duration_scaled ~ state + year_factor + (1|plot), umbs_flwr_plot, REML=FALSE)
mod6up <- lmer(flwr_duration_scaled ~ state + year_factor + insecticide + (1|plot), umbs_flwr_plot, REML=FALSE)
mod7up <- lmer(flwr_duration_scaled ~ state * year_factor + (1|plot), umbs_flwr_plot, REML=FALSE)
mod8up <- lmer(flwr_duration_scaled ~ state * year_factor + insecticide + (1|plot), umbs_flwr_plot, REML=FALSE)
mod9up <- lmer(flwr_duration_scaled ~ state * insecticide + year_factor + (1|plot), umbs_flwr_plot, REML=FALSE)
mod10up <- lmer(flwr_duration_scaled ~ state + insecticide * year_factor + (1|plot), umbs_flwr_plot, REML=FALSE)
mod11up <- lmer(flwr_duration_scaled ~ state * year_factor * insecticide + (1|plot), umbs_flwr_plot, REML=FALSE)
AICctab(mod1up, mod2up, mod3up, mod4up, mod5up, mod6up, mod7up, mod8up, mod9up, mod10up, mod11up, weights=T)
anova(mod1up, mod2up) # go with simpler model 1up
AICctab(mod2up, mod1up, weights=T) #mod1up
anova(mod1up)
summary(mod1up)

# with herb reduction included in model
anova(mod3up)
# with herb reduction as interactive term
anova(mod9up)

#Our hypothesized model is  mod9p
summary(mod9up)
anova(mod9up)

# making a table
kable(anova(mod9up)) %>% kableExtra::kable_styling()

# comparisons
emmeans(mod9up, list(pairwise ~ state*insecticide), adjust = "tukey")
mod9up.emm <- emmeans(mod9up, ~ state*insecticide)
contrast(mod9up.emm, "consec", simple = "each", combine = F, adjust = "mvt")
emmip(mod9up, insecticide~state)

# adding in our temp data into some models
# note: including state, year, and temp data into a model leads to rank deficiency
# so below, we test for date of first flower as a function of just temp to see how real temp data affects date of first flower
umbs_flwr_plot_amb <- umbs_flwr_plot %>%
        filter(state == "ambient")
modtest1u <- lmer(flwr_duration_scaled ~ mean_temp + (1|plot), umbs_flwr_plot, REML=FALSE)
modtest2u <- lmer(flwr_duration_scaled ~ GDD_cumulative + (1|plot), umbs_flwr_plot, REML=FALSE)

AICtab(modtest1u, modtest2u, weights=T)
# growing degree days model is better but we are going to use the mean temp for all models

anova(modtest1u)
summary(modtest1u)

# making a table
kable(anova(modtest1u)) %>% kableExtra::kable_styling()
```

```{r}
# UMBS
# species model for supp
mod_spp_k <- lmer(flwr_duration_scaled ~ state * insecticide + species + as.factor(year_factor) + (1|plot), flwr_kbs_spp, REML=FALSE)
anova(mod_spp_k)
# making a table
kable(anova(mod_spp_k)) %>% kableExtra::kable_styling()

# UMBS
# species model for supp
mod_spp_u <- lmer(flwr_duration_scaled ~ state * insecticide + species + as.factor(year_factor) + (1|plot), flwr_umbs_spp, REML=FALSE)
anova(mod_spp_u)
# making a table
kable(anova(mod_spp_u)) %>% kableExtra::kable_styling()
```
