---
title: "warmXtrophic Project: Phenology - Flowering Median Day Analyses"
author: "Moriah Young"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=80),tidy=TRUE)
```

# Load in packages & data
```{r, message = F}
# clear all existing data
rm(list=ls())

#Load packages
library(tidyverse)
library(ggplot2)
library(lme4)
library(lmerTest)
library(emmeans)
library(vegan)
library(car)
library(rstatix)
library(scales)
library(fitdistrplus)
library(moments)# for calculating skewness of data
library(ggpubr)
library(jtools) # summ() function
library(predictmeans)
library(olsrr)
library(car)
library(fitdistrplus)
library(ggpubr)
library(interactions)
library(sjPlot)
library(effects)
library(glmmTMB)
library(GGally) # ggpairs() function
library(bbmle) # AICtab() function

# Set working directory
Sys.getenv("L1DIR")
L0_dir <- Sys.getenv("L0DIR")
L1_dir <- Sys.getenv("L1DIR")
L2_dir <- Sys.getenv("L2DIR")

# Set ggplot2 plotting
# This code for ggplot2 sets the theme to mostly black and white 
# (Arial font, and large font, base size=24)
theme_set(theme_bw(14))
theme_update(axis.text.x = element_text(size = 12, angle = 90),
             axis.text.y = element_text(size = 12))

# Read in data
flwr_species <- read.csv(file.path(L2_dir, "phenology/final_flwr_species_L2.csv")) # species level data
flwr_plot <- read.csv(file.path(L2_dir, "phenology/final_flwr_plot_L2.csv")) # plot level data
flwr_plot_origin <- read.csv(file.path(L2_dir, "phenology/final_flwr_plot_origin_L2.csv")) # plot level data for origin
flwr_plot_growthhabit <- read.csv(file.path(L2_dir, "phenology/final_flwr_plot_growthhabit_L2.csv")) # plot level data for growth habit

# get rid of "X" column that shows up
flwr_species$X <- NULL
flwr_plot$X <- NULL
flwr_plot_origin$X <- NULL
flwr_plot_growthhabit$X <- NULL
# take a look at the data to see if looks good
View(flwr_species) 
View(flwr_plot) 

# Order warmed and ambient so that warmed shows up first in plotting (and is default is red = warmed; blue = ambient). First make it a factor
flwr_species$state <- as.factor(flwr_species$state)
levels(flwr_species$state)
# [1] "ambient" "warmed" 
flwr_species$state <- factor(flwr_species$state, levels(flwr_species$state)[c(2,1)])
levels(flwr_species$state)
# [1] "warmed"  "ambient"

# again for plot level data
flwr_plot$state <- as.factor(flwr_plot$state)
levels(flwr_plot$state)
# [1] "ambient" "warmed" 
flwr_plot$state <- factor(flwr_plot$state, levels(flwr_plot$state)[c(2,1)])
levels(flwr_plot$state)
# [1] "warmed"  "ambient"

umbs_flwr_plot <- subset(flwr_plot, site == "umbs") # pull out umbs only data at plot level
# delete 2021 data from umbs dataframe (doesn't make sense to have bc we can't get min flower/sd, median flower, or flower duration from 2021 UMBS)
umbs_flwr_plot <- subset(umbs_flwr_plot, year!="2021")
umbs_flwr_spp <- subset(flwr_species, site == "umbs") # pull out umbs only data at species level
umbs_flwr_spp <- subset(umbs_flwr_spp, year!="2021") # delete 2021 data
kbs_flwr_plot <- subset(flwr_plot, site == "kbs") # pull out kbs only data at plot level
kbs_flwr_spp <- subset(flwr_species, site == "kbs") # pull out kbs only data at species level
kbs_flwr_plot_origin <- subset(flwr_plot_origin, site == "kbs")
kbs_flwr_plot_growthhabit <- subset(flwr_plot_growthhabit, site == "kbs")
umbs_flwr_plot_origin <- subset(flwr_plot_origin, site == "umbs")
umbs_flwr_plot_origin <- subset(flwr_plot_origin, year != "2021") # delete 2021 data
umbs_flwr_plot_growthhabit <- subset(flwr_plot_growthhabit, site == "umbs")
umbs_flwr_plot_growthhabit <- subset(flwr_plot_growthhabit, year != "2021") # delete 2021 data
```

# Phoebe
# UMBS SPECIES LEVEL - Looking at MEDIAN JULIAN DAY
```{r, warning=F}
### UMBS ###
hist(umbs_flwr_spp$julian_median)
qqnorm(umbs_flwr_spp$julian_median) # this looks ok
shapiro.test(umbs_flwr_spp$julian_median) # pvalue is < 0.05 so we reject the null hypothesis that the data is normal (aka not normally distributed)

# Visualizing median Julian date for umbs at the PLOT LEVEL
ggplot(umbs_flwr_spp, aes(julian_median, fill = plot)) + 
        geom_histogram(binwidth = 0.5) + 
        facet_grid(year ~ site, margins = TRUE, scales = "free")

ggplot(umbs_flwr_spp, aes(julian_median, fill = as.factor(plot))) + geom_histogram(binwidth = 0.5) + 
        facet_grid(state~year, margins = TRUE, scales = "free")

ggplot(umbs_flwr_spp, aes(julian_median, fill = plot, color=plot)) +
        geom_density(alpha = 0.1)

ggplot(umbs_flwr_spp, aes(julian_median, fill = plot, color=plot)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year)

ggplot(umbs_flwr_spp, aes(julian_median, fill = plot, color=plot)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year + plot)

# Exploring distributions for these right-skewed data:
descdist(umbs_flwr_spp$julian_median, discrete = FALSE)

# Gamma distribution 
fit.gamma <- fitdist(umbs_flwr_spp$julian_median, "gamma")
plot(fit.gamma)

# Weibull distribution
fit.weibull  <- fitdist(umbs_flwr_spp$julian_median, "weibull")
plot(fit.weibull)

# Lognormal distribution 
fit.ln <- fitdist(umbs_flwr_spp$julian_median, "lnorm")
plot(fit.ln)

par(mfrow=c(2,2))
plot.legend <- c("Gamma", "Weibull", "Log Normal")
denscomp(list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
cdfcomp (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
qqcomp  (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
ppcomp  (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)

# Goodness of fit comparisons across fits (can't include the sqrt normal bc it becomes diff response values)
gofstat(list(fit.gamma, fit.weibull, fit.ln), fitnames = c("Gamma", "Weibull", "Log Normal"))
# Lognormal and gamma are very close, log normal slightly better, so going with log transforming
```

# Phoebe
# UMBS PLOT LEVEL - Looking at MEDIAN JULIAN DAY
```{r, warning=F}
### UMBS ###
hist(umbs_flwr_plot$julian_median)
qqnorm(umbs_flwr_plot$julian_median)
shapiro.test(umbs_flwr_plot$julian_median) # pvalue is > 0.05 so we accept the null hypothesis that the data is normal (aka normally distributed)

# Visualizing median Julian date for umbs at the PLOT LEVEL
ggplot(umbs_flwr_plot, aes(julian_median, fill = plot)) + 
        geom_histogram(binwidth = 0.5) + 
        facet_grid(year ~ site, margins = TRUE, scales = "free")

ggplot(umbs_flwr_plot, aes(julian_median, fill = as.factor(plot))) + geom_histogram(binwidth = 0.5) + 
        facet_grid(state~year, margins = TRUE, scales = "free")

ggplot(umbs_flwr_plot, aes(julian_median, fill = plot, color=plot)) +
        geom_density(alpha = 0.1)

ggplot(umbs_flwr_plot, aes(julian_median, fill = plot, color=plot)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year)

ggplot(umbs_flwr_plot, aes(julian_median, fill = plot, color=plot)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year + plot)

# Exploring distributions for these right-skewed data:
descdist(umbs_flwr_plot$julian_median, discrete = FALSE)

# Gamma distribution 
fit.gamma <- fitdist(umbs_flwr_plot$julian_median, "gamma")
plot(fit.gamma)

# Weibull distribution
fit.weibull  <- fitdist(umbs_flwr_plot$julian_median, "weibull")
plot(fit.weibull)

# Lognormal distribution 
fit.ln <- fitdist(umbs_flwr_plot$julian_median, "lnorm")
plot(fit.ln)

par(mfrow=c(2,2))
plot.legend <- c("Gamma", "Weibull", "Log Normal")
denscomp(list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
cdfcomp (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
qqcomp  (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
ppcomp  (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)

# Goodness of fit comparisons across fits (can't include the sqrt normal bc it becomes diff response values)
gofstat(list(fit.gamma, fit.weibull, fit.ln), fitnames = c("Gamma", "Weibull", "Log Normal"))
# Gamma is the best fit based on this function with log normal really close or we could go with normal distribution
```

# UMBS PLOT LEVEL ORIGIN - Looking at MEDIAN JULIAN DAY
```{r, warning=F, error=TRUE}
### UMBS ###
hist(umbs_flwr_plot_origin$julian_median)
qqnorm(umbs_flwr_plot_origin$julian_median)
shapiro.test(umbs_flwr_plot_origin$julian_median) # pvalue is < 0.05 so we reject the null hypothesis that the data is normal (aka not normally distributed) 

# Visualizing median Julian date for umbs at the PLOT LEVEL
ggplot(umbs_flwr_plot_origin, aes(julian_median, fill = plot)) + 
        geom_histogram(binwidth = 0.5) + 
        facet_grid(year ~ site, margins = TRUE, scales = "free")

ggplot(umbs_flwr_plot_origin, aes(julian_median, fill = as.factor(plot))) + geom_histogram(binwidth = 0.5) + 
        facet_grid(state~year, margins = TRUE, scales = "free")

ggplot(umbs_flwr_plot_origin, aes(julian_median, fill = plot, color=plot)) +
        geom_density(alpha = 0.1)

ggplot(umbs_flwr_plot_origin, aes(julian_median, fill = plot, color=plot)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year)

ggplot(umbs_flwr_plot_origin, aes(julian_median, fill = plot, color=plot)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year + plot)

# Exploring distributions for these data:
descdist(umbs_flwr_plot_origin$julian_median, discrete = FALSE)

# Gamma distribution 
fit.gamma <- fitdist(umbs_flwr_plot_origin$julian_median, "gamma")
plot(fit.gamma)
#
# Weibull distribution
fit.weibull  <- fitdist(umbs_flwr_plot_origin$julian_median, "weibull")
plot(fit.weibull)

# Lognormal distribution 
fit.ln <- fitdist(umbs_flwr_plot_origin$julian_median, "lnorm")
plot(fit.ln)

par(mfrow=c(2,2))
plot.legend <- c("Gamma", "Weibull", "Log Normal")
denscomp(list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
cdfcomp (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
qqcomp  (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
ppcomp  (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)

# Goodness of fit comparisons across fits (can't include the sqrt normal bc it becomes diff response values)
gofstat(list(fit.gamma, fit.weibull, fit.ln), fitnames = c("Gamma", "Weibull", "Log Normal"))
# log normal looks like it's the best fit
```

# UMBS PLOT LEVEL GROWTH HABIT - Looking at MEDIAN JULIAN DAY
```{r, warning=F, error=TRUE}
### UMBS ###
hist(umbs_flwr_plot_growthhabit$julian_median)
qqnorm(umbs_flwr_plot_growthhabit$julian_median)
shapiro.test(umbs_flwr_plot_growthhabit$julian_median) # pvalue is < 0.05 so we reject the null hypothesis that the data is normal (aka not normally distributed) 

# Visualizing median Julian date for umbs at the PLOT LEVEL
ggplot(umbs_flwr_plot_growthhabit, aes(julian_median, fill = plot)) + 
        geom_histogram(binwidth = 0.5) + 
        facet_grid(year ~ site, margins = TRUE, scales = "free")

ggplot(umbs_flwr_plot_growthhabit, aes(julian_median, fill = as.factor(plot))) + geom_histogram(binwidth = 0.5) + 
        facet_grid(state~year, margins = TRUE, scales = "free")

ggplot(umbs_flwr_plot_growthhabit, aes(julian_median, fill = plot, color=plot)) +
        geom_density(alpha = 0.1)

ggplot(umbs_flwr_plot_growthhabit, aes(julian_median, fill = plot, color=plot)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year)

ggplot(umbs_flwr_plot_growthhabit, aes(julian_median, fill = plot, color=plot)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year + plot)

# Exploring distributions for these data:
descdist(umbs_flwr_plot_growthhabit$julian_median, discrete = FALSE)

# Gamma distribution 
fit.gamma <- fitdist(umbs_flwr_plot_growthhabit$julian_median, "gamma")
plot(fit.gamma)
#
# Weibull distribution
fit.weibull  <- fitdist(umbs_flwr_plot_growthhabit$julian_median, "weibull")
plot(fit.weibull)

# Lognormal distribution 
fit.ln <- fitdist(umbs_flwr_plot_growthhabit$julian_median, "lnorm")
plot(fit.ln)

par(mfrow=c(2,2))
plot.legend <- c("Gamma", "Weibull", "Log Normal")
denscomp(list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
cdfcomp (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
qqcomp  (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
ppcomp  (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)

# Goodness of fit comparisons across fits (can't include the sqrt normal bc it becomes diff response values)
gofstat(list(fit.gamma, fit.weibull, fit.ln), fitnames = c("Gamma", "Weibull", "Log Normal"))
# Log normal looks like it's the best fit
```

# KBS SPECIES LEVEL - Looking at MEDIAN JULIAN DAY
```{r, warning=F}
### KBS ###
hist(kbs_flwr_spp$julian_median)
qqnorm(kbs_flwr_spp$julian_median)
shapiro.test(kbs_flwr_spp$julian_median) # pvalue is < 0.05 so we reject the null hypothesis that the data is normal (aka not normally distributed)

# Visualizing median Julian date for kbs at the SPECIES LEVEL
ggplot(kbs_flwr_spp, aes(julian_median, fill = plot)) + 
        geom_histogram(binwidth = 0.5) + 
        facet_grid(year ~ site, margins = TRUE, scales = "free")

ggplot(kbs_flwr_spp, aes(julian_median, fill = as.factor(plot))) + geom_histogram(binwidth = 0.5) + 
        facet_grid(state~year, margins = TRUE, scales = "free")

ggplot(kbs_flwr_spp, aes(julian_median, fill = plot, color=plot)) +
        geom_density(alpha = 0.1)

ggplot(kbs_flwr_spp, aes(julian_median, fill = plot, color=plot)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year)

ggplot(kbs_flwr_spp, aes(julian_median, fill = plot, color=plot)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year + plot)

# Exploring distributions for these right-skewed data:
descdist(kbs_flwr_spp$julian_median, discrete = FALSE)

# Gamma distribution 
fit.gamma <- fitdist(kbs_flwr_spp$julian_median, "gamma")
plot(fit.gamma)

# Weibull distribution
fit.weibull  <- fitdist(kbs_flwr_spp$julian_median, "weibull")
plot(fit.weibull)

# Lognormal distribution 
fit.ln <- fitdist(kbs_flwr_spp$julian_median, "lnorm")
plot(fit.ln)

par(mfrow=c(2,2))
plot.legend <- c("Gamma", "Weibull", "Log Normal")
denscomp(list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
cdfcomp (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
qqcomp  (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
ppcomp  (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)

# Goodness of fit comparisons across fits (can't include the sqrt normal bc it becomes diff response values)
gofstat(list(fit.gamma, fit.weibull, fit.ln), fitnames = c("Gamma", "Weibull", "Log Normal"))
# Lognormal is the best so going to log transform response variable julian_median
```

# KBS PLOT LEVEL - Looking at MEDIAN JULIAN DAY
```{r, warning=F}
### UMBS ###
hist(kbs_flwr_plot$julian_median)
qqnorm(kbs_flwr_plot$julian_median)
shapiro.test(kbs_flwr_plot$julian_median) # pvalue is < 0.05 so we reject the null hypothesis that the data is normal (aka not normally distributed)

# Visualizing median Julian date for kbs at the PLOT LEVEL
ggplot(kbs_flwr_plot, aes(julian_median, fill = plot)) + 
        geom_histogram(binwidth = 0.5) + 
        facet_grid(year ~ site, margins = TRUE, scales = "free")

ggplot(kbs_flwr_plot, aes(julian_median, fill = as.factor(plot))) + geom_histogram(binwidth = 0.5) + 
        facet_grid(state~year, margins = TRUE, scales = "free")

ggplot(kbs_flwr_plot, aes(julian_median, fill = plot, color=plot)) +
        geom_density(alpha = 0.1)

#ggplot(umbs_flwr_plot, aes(julian_median, fill = plot, color=plot)) +
#        geom_density(alpha = 0.1) +
#        facet_wrap(~year)
#
#ggplot(umbs_flwr_plot, aes(julian_median, fill = plot, color=plot)) +
#        geom_density(alpha = 0.1) +
#        facet_wrap(~year + plot)

# Exploring distributions for these right-skewed data:
descdist(kbs_flwr_plot$julian_median, discrete = FALSE)

# Gamma distribution 
fit.gamma <- fitdist(kbs_flwr_plot$julian_median, "gamma")
plot(fit.gamma)

# Weibull distribution
fit.weibull  <- fitdist(kbs_flwr_plot$julian_median, "weibull")
plot(fit.weibull)

# Lognormal distribution 
fit.ln <- fitdist(kbs_flwr_plot$julian_median, "lnorm")
plot(fit.ln)

par(mfrow=c(2,2))
plot.legend <- c("Gamma", "Weibull", "Log Normal")
denscomp(list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
cdfcomp (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
qqcomp  (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
ppcomp  (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)

# Goodness of fit comparisons across fits (can't include the sqrt normal bc it becomes diff response values)
gofstat(list(fit.gamma, fit.weibull, fit.ln), fitnames = c("Gamma", "Weibull", "Log Normal"))
# Log normal is the best first so going to log transform response variable julian_median
```
# KBS PLOT LEVEL ORIGIN - Looking at MEDIAN JULIAN DAY
```{r, warning=F, error=TRUE}
### KBS ###
hist(kbs_flwr_plot_origin$julian_median)
qqnorm(kbs_flwr_plot_origin$julian_median)
shapiro.test(umbs_flwr_plot_origin$julian_median) # pvalue is < 0.05 so we reject the null hypothesis that the data is normal (aka not normally distributed) 

# Visualizing median Julian date for kbs at the PLOT LEVEL
ggplot(kbs_flwr_plot_origin, aes(julian_median, fill = plot)) + 
        geom_histogram(binwidth = 0.5) + 
        facet_grid(year ~ site, margins = TRUE, scales = "free")

ggplot(kbs_flwr_plot_origin, aes(julian_median, fill = as.factor(plot))) + geom_histogram(binwidth = 0.5) + 
        facet_grid(state~year, margins = TRUE, scales = "free")

ggplot(kbs_flwr_plot_origin, aes(julian_median, fill = plot, color=plot)) +
        geom_density(alpha = 0.1)

ggplot(kbs_flwr_plot_origin, aes(julian_median, fill = plot, color=plot)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year)

ggplot(kbs_flwr_plot_origin, aes(julian_median, fill = plot, color=plot)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year + plot)

# Exploring distributions for these data:
descdist(kbs_flwr_plot_origin$julian_median, discrete = FALSE)

# Gamma distribution 
fit.gamma <- fitdist(kbs_flwr_plot_origin$julian_median, "gamma")
plot(fit.gamma)
#
# Weibull distribution
fit.weibull  <- fitdist(kbs_flwr_plot_origin$julian_median, "weibull")
plot(fit.weibull)

# Lognormal distribution 
fit.ln <- fitdist(kbs_flwr_plot_origin$julian_median, "lnorm")
plot(fit.ln)

par(mfrow=c(2,2))
plot.legend <- c("Gamma", "Weibull", "Log Normal")
denscomp(list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
cdfcomp (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
qqcomp  (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
ppcomp  (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)

# Goodness of fit comparisons across fits (can't include the sqrt normal bc it becomes diff response values)
gofstat(list(fit.gamma, fit.weibull, fit.ln), fitnames = c("Gamma", "Weibull", "Log Normal"))
# log normal looks like it's the best fit
```
# KBS PLOT LEVEL GROWTH HABIT - Looking at MEDIAN JULIAN DAY
```{r, warning=F, error=TRUE}
### KBS ###
hist(kbs_flwr_plot_growthhabit$julian_median)
qqnorm(kbs_flwr_plot_growthhabit$julian_median)
shapiro.test(kbs_flwr_plot_growthhabit$julian_median) # pvalue is < 0.05 so we reject the null hypothesis that the data is normal (aka not normally distributed) 

# Visualizing median Julian date for kbs at the PLOT LEVEL
ggplot(kbs_flwr_plot_growthhabit, aes(julian_median, fill = plot)) + 
        geom_histogram(binwidth = 0.5) + 
        facet_grid(year ~ site, margins = TRUE, scales = "free")

ggplot(kbs_flwr_plot_growthhabit, aes(julian_median, fill = as.factor(plot))) + geom_histogram(binwidth = 0.5) + 
        facet_grid(state~year, margins = TRUE, scales = "free")

ggplot(kbs_flwr_plot_growthhabit, aes(julian_median, fill = plot, color=plot)) +
        geom_density(alpha = 0.1)

ggplot(kbs_flwr_plot_growthhabit, aes(julian_median, fill = plot, color=plot)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year)

ggplot(kbs_flwr_plot_growthhabit, aes(julian_median, fill = plot, color=plot)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year + plot)

# Exploring distributions for these data:
descdist(kbs_flwr_plot_growthhabit$julian_median, discrete = FALSE)

# Gamma distribution 
fit.gamma <- fitdist(umbs_flwr_plot_growthhabit$julian_median, "gamma")
plot(fit.gamma)
#
# Weibull distribution
fit.weibull  <- fitdist(umbs_flwr_plot_growthhabit$julian_median, "weibull")
plot(fit.weibull)

# Lognormal distribution 
fit.ln <- fitdist(umbs_flwr_plot_growthhabit$julian_median, "lnorm")
plot(fit.ln)

par(mfrow=c(2,2))
plot.legend <- c("Gamma", "Weibull", "Log Normal")
denscomp(list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
cdfcomp (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
qqcomp  (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
ppcomp  (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)

# Goodness of fit comparisons across fits (can't include the sqrt normal bc it becomes diff response values)
gofstat(list(fit.gamma, fit.weibull, fit.ln), fitnames = c("Gamma", "Weibull", "Log Normal"))
# Log normal looks like it's the best fit
```

# Leverage plots and detecting Outliers. https://www.statmethods.net/stats/rdiagnostics.html
# These illustrate whether certain data points have more leverage (more influence), and thus could be outliers. It's a way of detecting outliers. Leverage plots can help identify whether a point has high or low influence, based on its leverage and residual and determining model fit with and without the point in question. Ultimately you decide whether the points are outliers or not, based on the knowledge of the system and how much it changes the model when included vs. excluded from the data used to fit the model. Here is a good overview of the combination of leverage and residual: scroll down to sections beginning at "13.3 Unusual Observations": https://daviddalpiaz.github.io/appliedstats/model-diagnostics.html

# Phoebe
```{r}
# species level data
# UMBS State-only model
fit_spp_state <- lm(log(julian_median) ~ state, data = umbs_flwr_spp)
outlierTest(fit_spp_state) # no outliers
qqPlot(fit_spp_state, main="QQ Plot") 
hist(fit_spp_state$residuals)
leveragePlots(fit_spp_state)
ols_test_normality(fit_spp_state)

# UMBS State and year model
fit_spp_stateyear <- lm(log(julian_median) ~ state + year_factor, data = umbs_flwr_spp)
outlierTest(fit_spp_stateyear) # no outliers
qqPlot(fit_spp_stateyear, main="QQ Plot") 
hist(fit_spp_stateyear$residuals)
leveragePlots(fit_spp_stateyear)
ols_test_normality(fit_spp_stateyear)

# Interaction plot (ignore for now the repeated measures with species); see: https://cran.r-project.org/web/packages/interactions/vignettes/interactions.html and: https://interactions.jacob-long.com/

fit3 <- lm(log(julian_median) ~ state + year_factor + species, data = umbs_flwr_spp)
interact_plot(fit3, pred = year_factor, modx = state)
fit4 <- lm(log(julian_median) ~ state * year_factor + species, data = umbs_flwr_spp)
interact_plot(fit4, pred = year_factor, modx = state, mod2 = species)
```
# Phoebe
```{r}
# Plot level data
# UMBS State-only model
fit_plot_state <- lm(log(julian_median) ~ state, data = umbs_flwr_plot)
outlierTest(fit_plot_state) # no outliers
qqPlot(fit_plot_state, main="QQ Plot") 
hist(fit_plot_state$residuals)
leveragePlots(fit_plot_state)
ols_test_normality(fit_plot_state) # looks ok besides Kolmogorov-Smirnov test

# UMBS State and year model
fit_plot_stateyear <- lm(log(julian_median) ~ state + year_factor, data = umbs_flwr_plot)
outlierTest(fit_plot_stateyear) # no outliers
qqPlot(fit_plot_stateyear, main="QQ Plot") 
hist(fit_plot_stateyear$residuals)
leveragePlots(fit_plot_stateyear)
ols_test_normality(fit_plot_stateyear) # not good

# Interaction plot (ignore for now the repeated measures with species); see: https://cran.r-project.org/web/packages/interactions/vignettes/interactions.html and: https://interactions.jacob-long.com/

fit3 <- lm(log(julian_median) ~ state * year_factor, data = umbs_flwr_plot)
interact_plot(fit3, pred = year_factor, modx = state)
```

```{r}
# species level data
# KBS State-only model
fit_spp_state_kbs <- lm(log(julian_median) ~ state, data = kbs_flwr_spp)
outlierTest(fit_spp_state_kbs) # no outliers
qqPlot(fit_spp_state_kbs, main="QQ Plot") 
hist(fit_spp_state_kbs$residuals)
leveragePlots(fit_spp_state_kbs)
ols_test_normality(fit_spp_state_kbs)

# KBS State and year model
fit_spp_stateyear_kbs <- lm(log(julian_median) ~ state + year_factor, data = kbs_flwr_spp)
outlierTest(fit_spp_stateyear_kbs) # no outliers
qqPlot(fit_spp_stateyear_kbs, main="QQ Plot") 
hist(fit_spp_stateyear_kbs$residuals)
leveragePlots(fit_spp_stateyear_kbs)
ols_test_normality(fit_spp_stateyear_kbs)

# Interaction plot (ignore for now the repeated measures with species); see: https://cran.r-project.org/web/packages/interactions/vignettes/interactions.html and: https://interactions.jacob-long.com/

fit3_spp_kbs <- lm(log(julian_median) ~ state + year_factor + species, data = kbs_flwr_spp)
interact_plot(fit3_spp_kbs, pred = year_factor, modx = state, data = kbs_flwr_spp)
fit4_spp_kbs <- lm(log(julian_median) ~ state * year_factor + species, data = kbs_flwr_spp)
interact_plot(fit4_spp_kbs, pred = year_factor, modx = state, mod2 = species, data = kbs_flwr_spp)
```

```{r}
# Plot level data
# KBS State-only model
fit_plot_state_kbs <- lm(log(julian_median) ~ state, data = kbs_flwr_plot)
outlierTest(fit_plot_state_kbs) # no outliers
qqPlot(fit_plot_state_kbs, main="QQ Plot") 
hist(fit_plot_state_kbs$residuals)
leveragePlots(fit_plot_state_kbs)
ols_test_normality(fit_plot_state_kbs)

# KBS State and year model
fit_plot_stateyear_kbs <- lm(log(julian_median) ~ state + year_factor, data = kbs_flwr_plot)
outlierTest(fit_plot_stateyear_kbs) # outliers
qqPlot(fit_plot_stateyear_kbs, main="QQ Plot") 
hist(fit_plot_stateyear_kbs$residuals)
leveragePlots(fit_plot_stateyear_kbs)
ols_test_normality(fit_plot_stateyear_kbs)

# Interaction plot (ignore for now the repeated measures with species); see: https://cran.r-project.org/web/packages/interactions/vignettes/interactions.html and: https://interactions.jacob-long.com/

fit3_plot_kbs <- lm(log(julian_median) ~ state * year_factor, data = kbs_flwr_plot)
interact_plot(fit3_plot_kbs, pred = year_factor, modx = state)
```
MIXED EFFECT MODELS

# Phoebe
UMBS Species-level Mixed Effects Models
```{r, error=TRUE}
# KBS SPECIES LEVEL - Looking at DURATION of flowering
mod1 <- lmer(log(julian_median) ~ state*year_factor + insecticide*year_factor + (1|species) + (1|plot), umbs_flwr_spp, REML = FALSE)
ggplot(umbs_flwr_spp,aes(x=plot,y=julian_median,col=species)) + geom_jitter() + geom_boxplot(alpha=0.2) + facet_wrap(~plot)

# Check Assumptions:
# (1) Linearity: if covariates are not categorical (year isn't)
# (2) Homogeneity: Need to Check by plotting residuals vs predicted values.
par(mfrow=c(1,2))
plot(mod1) # thus looks ok?
# Homogeneity of variance looks weird here 
# Check for homogeneity of variances (true if p>0.05). If the result is not significant, the assumption of equal variances (homoscedasticity) is met (no significant difference between the group variances).
# *****Levene’s Test - tests whether or not the variance among two or more groups is equal - If the p-value is less than our chosen significance level, we can reject the null hypothesis and conclude that we have enough evidence to state that the variance among the groups is not equal (which we want).

leveneTest(residuals(mod1) ~ umbs_flwr_spp$state)
# Assumption met
leveneTest(residuals(mod1) ~ umbs_flwr_spp$species)
# Assumption met
leveneTest(residuals(mod1) ~ umbs_flwr_spp$insecticide) 
# Assumption not met
leveneTest(residuals(mod1) ~ umbs_flwr_spp$plot)
# Assumption not met

# (3) Normality of error term: need to check by histogram, QQplot of residuals, could do Kolmogorov-Smirnov test.
# Check for normal residuals
qqPlot(resid(mod1))
hist(residuals(mod1))
shapiro.test(resid(mod1)) # not normally distributed resids bc p<0.05
outlierTest(mod1) # 1676 outlier

# (4) Normality of random effect: Get the estimate of random effect (e.g., random intercepts), and check them as you would check the residual. 
require(lme4)
r_int<- ranef(mod1)$plot$`(Intercept)`
qqnorm(r_int)
# qqline(r_int) doesn't work
# shapiro.test(r_int) # this doens't run
```
# Phoebe
```{r, error=TRUE}
# Do we need to include plot as a random effect with the UMBS models?   
mod1 <- lmer(log(julian_median) ~ state*year_factor + insecticide*year_factor + (1|species) + (1|plot), umbs_flwr_spp, REML=FALSE)
mod2 <- lmer(log(julian_median) ~ state*year_factor + insecticide*year_factor + (1|species), umbs_flwr_spp, REML=FALSE)
# Run analysis of variance on each model (see this for more explanation on how anova on a linear mixed effects model is similar to an anove on a regular linear model: https://m-clark.github.io/docs/mixedModels/anovamixed.html)
anova(mod1)
anova(mod2)

# Run an ANOVA to test the 2 models to test whether the more complex model is significantly better at capturing the data than the simpler model. If the resulting p-value is sufficiently low (usually less than 0.05), we conclude that the more complex model is significantly better than the simpler model, and thus favor the more complex model. If the p-value is not sufficiently low (usually greater than 0.05), we should favor the simpler model. https://bookdown.org/ndphillips/YaRrr/comparing-regression-models-with-anova.html
anova(mod1, mod2) # Go with model 2 since pvalue >0.05, aka more complex model does not have something in it that matters (the two models are essentially the same so go w model with fewer parameters)
summ(mod1) 
summ(mod2)
AICctab(mod1, mod2, weights=T) # yup, model 2

# Next, plot the model. There are multiple variables but here's one way to do it based on this package sjPlot:
# https://strengejacke.github.io/sjPlot/articles/plot_model_estimates.html
# Annoyingly, this package somehow overwrites the factor order in its plotting so we will have to modify the code to get warmed = red. I haven't figured this out yet. It does seem to work on some of the plots. hmm.
#?plot_model
# Plot the fixed effects estimates for different models
# these are the fixed effects estimates from summary(mod1)
plot_model(mod2, sort.est = TRUE)
# these are the fixed predicted values:
plot_model(mod2, type = "pred", terms = c("year_factor", "state", "insecticide"))
# these are the random effects estimates
plot_model(mod2, type = "re", terms = c("species"))

# Do we need to include insecticide? (dropping insecticide from the model)
mod3 <- lmer(log(julian_median) ~ state*year_factor + (1|species), umbs_flwr_spp, REML=FALSE)
anova(mod2, mod3) # Looks like no pvalue > 0.05, so insecticide does not improve model fit so we will go with mod 3 the simpler model

AICctab(mod2, mod3, weights=T)

# Does year need to be interactive with insecticide? - already removed insecticide
# mod4 <- lmer(log(julian_median) ~ state*year_factor + insecticide + (1|species) + (1|plot), umbs_flwr_spp, REML=FALSE)

# Does year need to be interactive with state?
mod5 <- lmer(log(julian_median) ~ state + year_factor + (1|species), umbs_flwr_spp, REML=FALSE)
anova(mod3, mod5)
# No, P>0.05 so state*year_factor doesn't strongly improve model fit so we will shift to mod5, but AIC values are super close!!

# Plot the fixed effects estimates for different models
# these are the fixed effects estimates from summary(mod5)
plot_model(mod5, sort.est = TRUE)
# these are the fixed predicted values:
plot_model(mod5, type = "pred", terms = c("year_factor", "state"))
# these are the random effects estimates
plot_model(mod5, type = "re", terms = c("species"))

# If we wanted to include plots nested within year it would look like this:
mod6 <- lmer(log(julian_median) ~ state + year_factor + (1|species) + (1 + year|plot), umbs_flwr_spp, REML=FALSE)
# model failed to converge...
anova(mod5, mod6)
anova(mod5)

# mod4 (and mod5) are pretty complex in terms of interpretation (they actually don't have many parameters though). We could consider an alternative model that's simpler to understand and also one that provides more insight about the species. That would be something like this:
mod7 <- lmer(log(julian_median) ~ state + species + (1+factor(year_factor)|plot), umbs_flwr_spp, REML=FALSE)
mod7a <- lmer(log(julian_median)  ~ state + species + factor(year_factor) + (1|plot), umbs_flwr_spp, REML=FALSE)
mod7b <- lmer(log(julian_median)  ~ state * factor(year_factor) + species + (1|plot), umbs_flwr_spp, REML=FALSE)
mod7c <- lmer(log(julian_median)  ~ state + species + factor(year_factor) + insecticide + (1|plot), umbs_flwr_spp, REML=FALSE)
anova(mod5, mod7) # model 7 is a better fit to data
anova(mod7, mod7a) #mod 7a
anova(mod7a, mod7b) #mod 7a - interaction between state and year does not improve model, but pvalue = 0.08 so going to go with more complex model
anova(mod7a, mod7c) # model 7a
summ(mod7a)
anova(mod7a) # investigates whether at least one of the levels within each factor is significantly different.
# Yes, at least one of the species is different (they do not all have the median first flwr dates). 
emmeans(mod7a, list(pairwise ~ state + year_factor), adjust = "tukey")
emmeans(mod7a, list(pairwise ~ year_factor), adjust = "tukey")
emmeans(mod7a, list(pairwise ~ species), adjust = "tukey")
emmeans(mod7a, list(pairwise ~ insecticide), adjust = "tukey")

# Take a look at the estimates for each fixed effect. These are the estimates from summary(mod7a). You'll see that species vary a lot - and many of them are different from zero (meaning their half cover date is significantly different from zero).
plot_model(mod7a, sort.est = TRUE)
# if you want to standardize the estimates:
plot_model(mod7a, sort.est = TRUE, type="std") 
# these are the fixed predicted values: - note this is a new plot
plot_model(mod7a, type = "pred", terms = c("species", "state", "year_factor"))
# these are the random effects estimates
plot_model(mod7a, type = "re")

# including native vs. exotic
umbs_flwr_spp <- within(umbs_flwr_spp, origin <- relevel(factor(origin), ref = "Native")) # releveling so native is the reference
mod8 <- lmer(log(julian_median) ~ state * origin + (1+factor(year_factor)|plot), umbs_flwr_spp, REML=FALSE)
mod9 <- lmer(log(julian_median) ~ state + origin + (1+factor(year_factor)|plot), umbs_flwr_spp, REML=FALSE)
mod9a <- lmer(log(julian_median) ~ state + origin + factor(year_factor) + (1|plot), umbs_flwr_spp, REML=FALSE)
mod9b <- lmer(log(julian_median) ~ state + origin + insecticide + factor(year_factor) + (1|plot), umbs_flwr_spp, REML=FALSE)
anova(mod8, mod9) # model 9 is a better fit to data
anova(mod9, mod9a) # mod 9a
anova(mod9a, mod9b) # mod 9a
summary(mod9a)
anova(mod9a)
emmeans(mod9a, list(pairwise ~ state * origin), adjust = "tukey")

# including growth form - first with interaction term
umbs_flwr_spp <- within(umbs_flwr_spp, growth_habit <- relevel(factor(growth_habit), ref = "Forb")) # releveling so forb is the reference
mod10 <- lmer(log(julian_median) ~ state * growth_habit + (1+factor(year_factor)|plot), umbs_flwr_spp, REML=FALSE)
mod11 <- lmer(log(julian_median) ~ state + growth_habit + (1+factor(year_factor)|plot), umbs_flwr_spp, REML=FALSE)
mod11a <- lmer(log(julian_median) ~ state + growth_habit + factor(year_factor) + (1|plot), umbs_flwr_spp, REML=FALSE)
mod11b <- lmer(log(julian_median) ~ state + growth_habit + factor(year_factor) + insecticide + (1|plot), umbs_flwr_spp, REML=FALSE)
anova(mod10, mod11) # model 11 is a better fit to data
anova(mod11, mod11a) # model 11a is still a better fit to data
anova(mod11a, mod11b) # model 11a still a better fit to data
summary(mod11a)
anova(mod11a)
emmeans(mod11a, list(pairwise ~ factor(year_factor) * growth_habit), adjust = "tukey")

# New version of our model incorporating interaction term and species within year so that there is a separate intercept and slope for each species. The issue here is that there are some species that are not found each year. Easiest to remove those from another version of this dataframe before running below. Otherwise, it's not a balanced design.
# updated mod4
mod12 <- lmer(log(julian_median) ~ state * factor(year_factor) + (1+factor(year_factor)|species), umbs_flwr_spp)

# So another version of this model would include the interaction but not include the nesting (and thus would assume that species aren't observed ea yr)
# updated mod5
mod13 <- lmer(log(julian_median) ~ state * factor(year_factor) + (1|species), umbs_flwr_spp)

# All the models ran:
mod1 <- lmer(log(julian_median) ~ state*year_factor + insecticide*year_factor + (1|species) + (1|plot), umbs_flwr_spp, REML=FALSE)
mod2 <- lmer(log(julian_median) ~ state*year_factor + insecticide*year_factor + (1|species), umbs_flwr_spp, REML=FALSE)
mod3 <- lmer(log(julian_median) ~ state*year_factor + (1|species), umbs_flwr_spp, REML=FALSE)
mod4 <- lmer(log(julian_median) ~ state*year_factor + insecticide + (1|species) + (1|plot), umbs_flwr_spp, REML=FALSE)
mod5 <- lmer(log(julian_median) ~ state + year_factor + (1|species), umbs_flwr_spp, REML=FALSE)
mod6 <- lmer(log(julian_median) ~ state + year_factor + (1|species) + (1 + year|plot), umbs_flwr_spp, REML=FALSE)
mod7 <- lmer(log(julian_median) ~ state + species + (1+factor(year_factor)|plot), umbs_flwr_spp, REML=FALSE)
mod7a <- lmer(log(julian_median) ~ state + species + factor(year_factor) + (1|plot), umbs_flwr_spp, REML=FALSE)
mod7b <- lmer(log(julian_median) ~ state * factor(year_factor) + species + (1|plot), umbs_flwr_spp, REML=FALSE)
mod7c <- lmer(log(julian_median) ~ state + species + factor(year_factor) + insecticide + (1|plot), umbs_flwr_spp, REML=FALSE)
mod8 <- lmer(log(julian_median) ~ state * origin + (1+factor(year_factor)|plot), umbs_flwr_spp, REML=FALSE)
mod9 <- lmer(log(julian_median) ~ state + origin + (1+factor(year_factor)|plot), umbs_flwr_spp, REML=FALSE)
mod9a <- lmer(log(julian_median) ~ state + origin + factor(year_factor) + (1|plot), umbs_flwr_spp, REML=FALSE)
mod10 <- lmer(log(julian_median) ~ state * growth_habit + (1+factor(year_factor)|plot), umbs_flwr_spp, REML=FALSE)
mod11 <- lmer(log(julian_median) ~ state + growth_habit + (1+factor(year_factor)|plot), umbs_flwr_spp, REML=FALSE)
mod11a <- lmer(log(julian_median) ~ state + growth_habit + factor(year_factor) + (1|plot), umbs_flwr_spp, REML=FALSE)
mod12 <- lmer(log(julian_median) ~ state * factor(year_factor) + (1+factor(year_factor)|species), umbs_flwr_spp)
mod13 <- lmer(log(julian_median) ~ state * factor(year_factor) + (1|species), umbs_flwr_spp)
AICctab(mod1, mod2, mod3, mod5, mod6, mod7, mod7a, mod7b, mod7c, mod8, mod9, mod9a, mod10, mod11, mod11a, mod13, weights=T) #mod7a is the best fitting model (took out mod12 bc it wouldn't run)
summary(mod7a)
anova(mod7a)
```
# Phoebe
# UMBS Plot-level Mixed Effects Models:
```{r}
mod1p <- lmer(log(julian_median) ~ state + (1|plot), umbs_flwr_plot, REML=FALSE)
mod2p <- lmer(log(julian_median) ~ insecticide + (1|plot), umbs_flwr_plot, REML=FALSE)
mod3p <- lmer(log(julian_median) ~ insecticide + state + (1|plot), umbs_flwr_plot, REML=FALSE)
mod4p <- lmer(log(julian_median) ~ insecticide * state + (1|plot), umbs_flwr_plot, REML=FALSE)
mod5p <- lmer(log(julian_median) ~ state + year_factor + (1|plot), umbs_flwr_plot, REML=FALSE)
mod6p <- lmer(log(julian_median) ~ state + year_factor + insecticide + (1|plot), umbs_flwr_plot, REML=FALSE)
mod7p <- lmer(log(julian_median) ~ state * year_factor + (1|plot), umbs_flwr_plot, REML=FALSE)
mod8p <- lmer(log(julian_median) ~ state * year_factor + insecticide + (1|plot), umbs_flwr_plot, REML=FALSE)
mod9p <- lmer(log(julian_median) ~ state * insecticide + year_factor + (1|plot), umbs_flwr_plot, REML=FALSE)
mod10p <- lmer(log(julian_median) ~ state + insecticide * year_factor + (1|plot), umbs_flwr_plot, REML=FALSE)
mod11p <- lmer(log(julian_median) ~ state * year_factor * insecticide + (1|plot), umbs_flwr_plot, REML=FALSE)
AICctab(mod1p, mod2p, mod3p, mod4p, mod5p, mod6p, mod7p, mod8p, mod9p, mod10p, mod11p, weights=T) # model 5p and 7p are the top models
anova(mod5p, mod7p) # go with simpler model 5p
AICctab(mod5p, mod7p, weights=T)
summ(mod5p)
summary(mod5p)
emmeans(mod5p, list(pairwise ~ state+year_factor), adjust = "tukey")

# including native vs. exotic
umbs_flwr_plot_origin <- within(umbs_flwr_plot_origin, origin <- relevel(factor(origin), ref = "Native")) # releveling so native is the reference
mod12p <- lmer(log(julian_median) ~ state * origin + (1+year_factor|plot), umbs_flwr_plot_origin, REML=FALSE)
mod13p <- lmer(log(julian_median) ~ state + origin + (1+year_factor|plot), umbs_flwr_plot_origin, REML=FALSE)
mod14p <- lmer(log(julian_median) ~ state + origin + year_factor + (1|plot), umbs_flwr_plot_origin, REML=FALSE)
anova(mod12p, mod13p) # go with model 13p
anova(mod13p, mod14p) # mod 14p
summary(mod14p)
anova(mod14p)
emmeans(mod14p, list(pairwise ~ state + origin + year_factor), adjust = "tukey")

# Phoebe ignore
# including growth form - first with interaction term
umbs_flwr_plot_growthhabit <- within(umbs_flwr_plot_growthhabit, growth_habit <- relevel(factor(growth_habit), ref = "Forb")) # releveling so forb is the reference
mod15p <- lmer(log(julian_median) ~ state * growth_habit + (1+year_factor|plot), umbs_flwr_plot_growthhabit, REML=FALSE)
mod16p <- lmer(log(julian_median) ~ state + growth_habit + (1+year_factor|plot), umbs_flwr_plot_growthhabit, REML=FALSE)
mod17p <- lmer(log(julian_median) ~ state + growth_habit + year_factor + (1|plot), umbs_flwr_plot_growthhabit, REML=FALSE)
anova(mod15p, mod16p) # go with model 16pu 
anova(mod16p, mod17p) # mod 17pu
summary(mod17p)
anova(mod17p)
emmeans(mod17p, list(pairwise ~ state + growth_habit + year_factor), adjust = "tukey")
```

KBS Species-level Mixed Effects Models
```{r, error=TRUE}
# Do we need to include plot as a random effect with the KBS models?   
mod1k <- lmer(log(julian_median) ~ state*year_factor + insecticide*year_factor + (1|species) + (1|plot), kbs_flwr_spp, REML=FALSE)
mod2k <- lmer(log(julian_median) ~ state*year_factor + insecticide*year_factor + (1|species), kbs_flwr_spp, REML=FALSE)
# Run analysis of variance on each model (see this for more explanation on how anova on a linear mixed effects model is similar to an anova on a regular linear model: https://m-clark.github.io/docs/mixedModels/anovamixed.html)
anova(mod1k)
anova(mod2k)

# Run an ANOVA to test the 2 models to test whether the more complex model is significantly better at capturing the data than the simpler model. If the resulting p-value is sufficiently low (usually less than 0.05), we conclude that the more complex model is significantly better than the simpler model, and thus favor the more complex model. If the p-value is not sufficiently low (usually greater than 0.05), we should favor the simpler model. https://bookdown.org/ndphillips/YaRrr/comparing-regression-models-with-anova.html
anova(mod1k, mod2k) # models are not that different from each other so go with simpler model = model 2
summary(mod1k) 
summary(mod2k)

# Next, plot the model. There are multiple variables but here's one way to do it based on this package sjPlot:
# https://strengejacke.github.io/sjPlot/articles/plot_model_estimates.html
# Annoyingly, this package somehow overwrites the factor order in its plotting so we will have to modify the code to get warmed = red. I haven't figured this out yet. It does seem to work on some of the plots. hmm.
#?plot_model
# Plot the fixed effects estimates for different models
# these are the fixed effects estimates 
plot_model(mod2k, sort.est = TRUE)
# these are the fixed predicted values:
plot_model(mod2k, type = "pred", terms = c("year_factor", "state", "insecticide"))
# these are the random effects estimates
plot_model(mod2k, type = "re", terms = c("species"))

# Do we need to include insecticide? 
mod3k <- lmer(log(julian_median) ~ state*year_factor + (1|species), kbs_flwr_spp, REML=FALSE)
anova(mod2k, mod3k) # Don't need insecticide, continue with mod3 - the simpler model

# Does year need to be interactive with insecticide? - already removed insecticide
# mod4k <- lmer(log(julian_median) ~ state*year_factor + insecticide + (1|species) + (1|plot), kbs_flwr_spp, REML=FALSE)

# Does year need to be interactive with state?
mod5k <- lmer(log(julian_median) ~ state + year_factor + (1|species) + (1|plot), kbs_flwr_spp, REML=FALSE)
anova(mod3k, mod5k) # models are pretty similar
AICctab(mod3k, mod5k, weights=T) # go with mod3k

# Plot the fixed effects estimates for different models
# these are the fixed effects estimates from summary(mod3k)
plot_model(mod3k, sort.est = TRUE)
# these are the fixed predicted values:
plot_model(mod3k, type = "pred", terms = c("year_factor", "state"))
# these are the random effects estimates
plot_model(mod3k, type = "re", terms = c("species"))

# If we wanted to include plots nested within year it would look like this:
mod6k <- lmer(log(julian_median) ~ state + year_factor + (1|species) + (1 + year|plot), kbs_flwr_spp, REML=FALSE)
anova(mod3k, mod6k) #stick with mod3k
anova(mod3k)

# We could consider an alternative model that's simpler to understand and also one that provides more insight about the species. That would be something like this:
mod7k <- lmer(log(julian_median) ~ state + species + (1+year_factor|plot), kbs_flwr_spp, REML=FALSE)
mod7ak <- lmer(log(julian_median)  ~ state + species + year_factor + (1|plot), kbs_flwr_spp, REML=FALSE)
mod7bk <- lmer(log(julian_median)  ~ state * year_factor + species + (1|plot), kbs_flwr_spp, REML=FALSE)
mod7ck <- lmer(log(julian_median)  ~ state + species + year_factor + insecticide + (1|plot), kbs_flwr_spp, REML=FALSE)
anova(mod5k, mod7k) # model 7k is a better fit to data
anova(mod7k, mod7ak) #mod 7ak
anova(mod7ak, mod7bk) #mod 7ak
anova(mod7ak, mod7ck) #mod7ak
summary(mod7ak)
anova(mod7ak) # investigates whether at least one of the levels within each factor is significantly different.
# Yes, at least one of the species is different (they do not all have the min first seed dates). 
emmeans(mod7ak, list(pairwise ~ state + year_factor), adjust = "tukey")
emmeans(mod7ak, list(pairwise ~ year_factor), adjust = "tukey")
emmeans(mod7ak, list(pairwise ~ species), adjust = "tukey")
emmeans(mod7ak, list(pairwise ~ insecticide), adjust = "tukey")

# Take a look at the estimates for each fixed effect. These are the estimates from summary(mod7a). You'll see that species vary a lot - and many of them are different from zero (meaning their half cover date is significantly different from zero).
plot_model(mod7ak, sort.est = TRUE)
# if you want to standardize the estimates:
plot_model(mod7ak, sort.est = TRUE, type="std") 
# these are the fixed predicted values: - note this is a new plot
plot_model(mod7ak, type = "pred", terms = c("species", "state", "year_factor"))
# these are the random effects estimates
plot_model(mod7ak, type = "re")

AICctab(mod3k, mod7ak, weights=T)

# including native vs. exotic
kbs_flwr_spp <- within(kbs_flwr_spp, origin <- relevel(factor(origin), ref = "Native")) # releveling so native is the reference
mod8k <- lmer(log(julian_median) ~ state * origin + (1+year_factor|plot), kbs_flwr_spp, REML=FALSE)
mod9k <- lmer(log(julian_median) ~ state + origin + (1+year_factor|plot), kbs_flwr_spp, REML=FALSE)
mod9ak <- lmer(log(julian_median) ~ state + origin + year_factor + (1|plot), kbs_flwr_spp, REML=FALSE)
anova(mod8k, mod9k) # model 9k is a better fit to data
anova(mod9k, mod9ak) # mod 9ak
AICctab(mod8k, mod9k, mod9ak, weights=T)
summary(mod9ak)
anova(mod9ak)
emmeans(mod9ak, list(pairwise ~ state * origin), adjust = "tukey")

# including growth form - first with interaction term
kbs_flwr_spp <- within(kbs_flwr_spp, growth_habit <- relevel(factor(growth_habit), ref = "Forb")) # releveling so forb is the reference
mod10k <- lmer(log(julian_median) ~ state * growth_habit + (1+year_factor|plot), kbs_flwr_spp, REML=FALSE)
mod11k <- lmer(log(julian_median) ~ state + growth_habit + (1+year_factor|plot), kbs_flwr_spp, REML=FALSE)
mod11ak <- lmer(log(julian_median) ~ state + growth_habit + year_factor + (1|plot), kbs_flwr_spp, REML=FALSE)
anova(mod10k, mod11k) # model 11k is a better fit to data
anova(mod11k, mod11ak) # model 11ak is a better fit to data
AICctab(mod10k, mod11k, mod11ak, weights=T)
summary(mod11ak)
anova(mod11ak)
emmeans(mod11ak, list(pairwise ~ factor(year_factor) * growth_habit), adjust = "tukey")

# New version of our model incorporating interaction term and species within year so that there is a separate intercept and slope for each species. The issue here is that there are some species that are not found each year. Easiest to remove those from another version of this dataframe before running below. Otherwise, it's not a balanced design.
# updated mod4
mod12k <- lmer(log(julian_median) ~ state * factor(year_factor) + (1+factor(year_factor)|species), kbs_flwr_spp)

# So another version of this model would include the interaction but not include the nesting (and thus would assume that species aren't observed ea yr)
# updated mod5
mod13k <- lmer(log(julian_median) ~ state * factor(year_factor) + (1|species), kbs_flwr_spp)

# All the models ran:
mod1k <- lmer(log(julian_median) ~ state*year_factor + insecticide*year_factor + (1|species) + (1|plot), kbs_flwr_spp, REML=FALSE)
mod2k <- lmer(log(julian_median) ~ state*year_factor + insecticide*year_factor + (1|species), kbs_flwr_spp, REML=FALSE)
mod3k <- lmer(log(julian_median) ~ state*year_factor + (1|species), kbs_flwr_spp, REML=FALSE)
mod4k <- lmer(log(julian_median) ~ state*year_factor + insecticide + (1|species) + (1|plot), kbs_flwr_spp, REML=FALSE)
mod5k <- lmer(log(julian_median) ~ state + year_factor + (1|species), kbs_flwr_spp, REML=FALSE)
mod6k <- lmer(log(julian_median) ~ state + year_factor + (1|species) + (1 + year|plot), kbs_flwr_spp, REML=FALSE)
mod7k <- lmer(log(julian_median) ~ state + species + (1+factor(year_factor)|plot), kbs_flwr_spp, REML=FALSE)
mod7ak <- lmer(log(julian_median) ~ state + species + factor(year_factor) + (1|plot), kbs_flwr_spp, REML=FALSE)
mod7bk <- lmer(log(julian_median) ~ state * factor(year_factor) + species + (1|plot), kbs_flwr_spp, REML=FALSE)
mod7ck <- lmer(log(julian_median) ~ state + species + factor(year_factor) + insecticide + (1|plot), kbs_flwr_spp, REML=FALSE)
mod8k <- lmer(log(julian_median) ~ state * origin + (1+factor(year_factor)|plot), kbs_flwr_spp, REML=FALSE)
mod9k <- lmer(log(julian_median) ~ state + origin + (1+factor(year_factor)|plot), kbs_flwr_spp, REML=FALSE)
mod9ak <- lmer(log(julian_median) ~ state + origin + factor(year_factor) + (1|plot), kbs_flwr_spp, REML=FALSE)
mod10k <- lmer(log(julian_median) ~ state * growth_habit + (1+factor(year_factor)|plot), kbs_flwr_spp, REML=FALSE)
mod11k <- lmer(log(julian_median) ~ state + growth_habit + (1+factor(year_factor)|plot), kbs_flwr_spp, REML=FALSE)
mod11ak <- lmer(log(julian_median) ~ state + growth_habit + factor(year_factor) + (1|plot), kbs_flwr_spp, REML=FALSE)
mod12k <- lmer(log(julian_median) ~ state * factor(year_factor) + (1+factor(year_factor)|species), kbs_flwr_spp)
mod13k <- lmer(log(julian_median) ~ state * factor(year_factor) + (1|species), kbs_flwr_spp)
AICctab(mod1k, mod2k, mod3k, mod5k, mod6k, mod7k, mod7ak, mod7bk, mod7ck, mod8k, mod9k, mod9ak, mod10k, mod11k, mod11ak, mod13k, weights=T) #mod7a is the best fitting model (took out mod12 bc it wouldn't run)
summary(mod7ak)
summ(mod7ak)
anova(mod7ak)
```

# KBS Plot-level Mixed Effects Models:
```{r, error=TRUE}
mod1pk <- lmer(log(julian_median) ~ state + (1|plot), kbs_flwr_plot, REML=FALSE)
mod2pk <- lmer(log(julian_median) ~ insecticide + (1|plot), kbs_flwr_plot, REML=FALSE)
mod3pk <- lmer(log(julian_median) ~ insecticide + state + (1|plot), kbs_flwr_plot, REML=FALSE)
mod4pk <- lmer(log(julian_median) ~ insecticide * state + (1|plot), kbs_flwr_plot, REML=FALSE)
mod5pk <- lmer(log(julian_median) ~ state + year_factor + (1|plot), kbs_flwr_plot, REML=FALSE)
mod6pk <- lmer(log(julian_median) ~ state + year_factor + insecticide + (1|plot), kbs_flwr_plot, REML=FALSE)
mod7pk <- lmer(log(julian_median) ~ state * year_factor + (1|plot), kbs_flwr_plot, REML=FALSE)
mod8pk <- lmer(log(julian_median) ~ state * year_factor + insecticide + (1|plot), kbs_flwr_plot, REML=FALSE)
mod9pk <- lmer(log(julian_median) ~ state * insecticide + year_factor + (1|plot), kbs_flwr_plot, REML=FALSE)
mod10pk <- lmer(log(julian_median) ~ state + insecticide * year_factor + (1|plot), kbs_flwr_plot, REML=FALSE)
mod11pk <- lmer(log(julian_median) ~ state * year_factor * insecticide + (1|plot), kbs_flwr_plot, REML=FALSE)
AICctab(mod1pk, mod2pk, mod3pk, mod4pk, mod5pk, mod6pk, mod7pk, mod8pk, mod9pk, mod10pk, mod11pk, weights=T)
anova(mod5pk, mod7pk) # go with more complex model mod7pk
AICctab(mod5pk, mod7pk, weights=T)
summ(mod5pk)
summary(mod5pk)
emmeans(mod5pk, list(pairwise ~ state+year_factor), adjust = "tukey")

# including native vs. exotic
kbs_flwr_plot_origin <- within(kbs_flwr_plot_origin, origin <- relevel(factor(origin), ref = "Native")) # releveling so native is the reference
mod12pk <- lmer(log(julian_median) ~ state * origin + (1+year_factor|plot), kbs_flwr_plot_origin, REML=FALSE)
mod13pk <- lmer(log(julian_median) ~ state + origin + (1+year_factor|plot), kbs_flwr_plot_origin, REML=FALSE)
mod14pk <- lmer(log(julian_median) ~ state + origin + year_factor + (1|plot), kbs_flwr_plot_origin, REML=FALSE)
anova(mod12pk, mod13pk) # go with model 12pk
anova(mod12pk, mod14pk) # mod 14pk
AICctab(mod12pk, mod13pk, mod14pk, weights=T)
summary(mod14pk)
summ(mod14pk)
anova(mod14pk)
emmeans(mod14pk, list(pairwise ~ state + origin + year_factor), adjust = "tukey")

# including growth form - first with interaction term
kbs_flwr_plot_growthhabit <- within(kbs_flwr_plot_growthhabit, growth_habit <- relevel(factor(growth_habit), ref = "Forb")) # releveling so forb is the reference
mod15pk <- lmer(log(julian_median) ~ state * growth_habit + (1+year_factor|plot), kbs_flwr_plot_growthhabit, REML=FALSE)
mod16pk <- lmer(log(julian_median) ~ state + growth_habit + (1+year_factor|plot), kbs_flwr_plot_growthhabit, REML=FALSE)
mod17pk <- lmer(log(julian_median) ~ state + growth_habit + year_factor + (1|plot), kbs_flwr_plot_growthhabit, REML=FALSE)
anova(mod15pk, mod16pk) # go with model 16pk
anova(mod16pk, mod17pk) # mod 17pk
AICctab(mod15pk, mod16pk, mod17pk, weights=T)
summary(mod17pk)
summ(mod17pk)
anova(mod17pk)
emmeans(mod17pk, list(pairwise ~ state + growth_habit + year_factor), adjust = "tukey")
```

