---
title: "warmXtrophic: OTC Data Plots"
author: "Kara Dobson"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```
COLLABORATORS: Phoebe Zarnetske, Nina Lany, Kathryn Schmidt, Mark Hammond, Pat Bills, Kileigh Welshofer, Moriah Young  
DATA INPUT: CSV files are located in the L1 hobo folder in the shared Google drive   
DATA OUTPUT: Relevant plots for the HOBO data, other plots located in HOBO_plots_L2.R script
NOTES: Figures marked with (older) are plots generated in the past and may not be relevant/useful for the paper


```{r, include = FALSE, message = FALSE}
library(tidyverse)
library(plotrix)
library(ggpubr)
# Set working directory
#Sys.getenv("L1DIR")
L1_dir<-Sys.getenv("L1DIR")
#list.files(L1_dir)
KBS <- read.csv(file.path(L1_dir,"HOBO_data/HOBO_paired_sensor_data/KBS/KBS_pairedsensors_L1.csv"))
UMBS <- read.csv(file.path(L1_dir,"HOBO_data/HOBO_paired_sensor_data/UMBS/UMBS_pairedsensors_L1.csv"))
KBS_1 <- read.csv(file.path(L1_dir,"HOBO_data/HOBO_paired_sensor_data/KBS/KBS_pair1_L1.csv"))
KBS_2 <- read.csv(file.path(L1_dir,"HOBO_data/HOBO_paired_sensor_data/KBS/KBS_pair2_L1.csv"))
KBS_3 <- read.csv(file.path(L1_dir,"HOBO_data/HOBO_paired_sensor_data/KBS/KBS_pair3_L1.csv"))
UMBS_1 <- read.csv(file.path(L1_dir,"HOBO_data/HOBO_paired_sensor_data/UMBS/UMBS_pair1_L1.csv"))
UMBS_2 <- read.csv(file.path(L1_dir,"HOBO_data/HOBO_paired_sensor_data/UMBS/UMBS_pair2_L1.csv"))
UMBS_3 <- read.csv(file.path(L1_dir,"HOBO_data/HOBO_paired_sensor_data/UMBS/UMBS_pair3_L1.csv"))
KBS_par <- read.csv(file.path(L1_dir,"PAR/KBS_PAR_L1.csv"))
UMBS_par <- read.csv(file.path(L1_dir,"PAR/UMBS_PAR_L1.csv"))
KBS_pend <- read.csv(file.path(L1_dir,"HOBO_data/HOBO_pendant_data/KBS/KBS_HOBOpendant_L1.csv"))
UMBS_pend <- read.csv(file.path(L1_dir,"HOBO_data/HOBO_pendant_data/UMBS/UMBS_HOBOpendant_L1.csv"))

# date is a character column - convert to date format
KBS$Date_Time <- as.POSIXct(KBS$Date_Time, format = "%Y-%m-%d %H:%M")
UMBS$Date_Time <- as.POSIXct(UMBS$Date_Time, format = "%Y-%m-%d %H:%M")
KBS_pend$Date_Time <- as.POSIXct(KBS_pend$Date_Time, format = "%Y-%m-%d %H:%M")
UMBS_pend$Date_Time <- as.POSIXct(UMBS_pend$Date_Time, format = "%Y-%m-%d %H:%M")
KBS_par$Date_Time <- as.POSIXct(KBS_par$Date_Time, format = "%Y-%m-%d")
UMBS_par$Date_Time <- as.POSIXct(UMBS_par$Date_Time, format = "%Y-%m-%d")
KBS_1$Date_Time <- as.POSIXct(KBS_1$Date_Time, format = "%Y-%m-%d %H:%M")
KBS_2$Date_Time <- as.POSIXct(KBS_2$Date_Time, format = "%Y-%m-%d %H:%M")
KBS_3$Date_Time <- as.POSIXct(KBS_3$Date_Time, format = "%Y-%m-%d %H:%M")
UMBS_1$Date_Time <- as.POSIXct(UMBS_1$Date_Time, format = "%Y-%m-%d %H:%M")
UMBS_2$Date_Time <- as.POSIXct(UMBS_2$Date_Time, format = "%Y-%m-%d %H:%M")
UMBS_3$Date_Time <- as.POSIXct(UMBS_3$Date_Time, format = "%Y-%m-%d %H:%M")

test_KBS_1 <- KBS_1
test_KBS_1$month <- format(test_KBS_1$Date_Time,format="%m")
test_KBS_1$year <- format(test_KBS_1$Date_Time,format="%y")
mean_monthly_warmk1 <- aggregate( XH_warmed_air_1m ~ month + year , test_KBS_1 , mean )
mean_monthly_ambk1 <- aggregate( XH_ambient_air_1m ~ month + year, test_KBS_1, mean)
mean_monthlyk1 <- merge(mean_monthly_ambk1, mean_monthly_warmk1, by=c("month", "year"))
mean_monthly_gatherk1 <- mean_monthlyk1 %>%
  gather(key = "treatment",value="temp", -month, -year)

test_KBS_2 <- KBS_2
test_KBS_2$month <- format(test_KBS_2$Date_Time,format="%m")
test_KBS_2$year <- format(test_KBS_2$Date_Time,format="%y")
mean_monthly_warmk2 <- aggregate( XH_warmed_air_1m ~ month + year , test_KBS_2 , mean )
mean_monthly_ambk2 <- aggregate( XH_ambient_air_1m ~ month + year, test_KBS_2, mean)
mean_monthlyk2 <- merge(mean_monthly_ambk2, mean_monthly_warmk2, by=c("month", "year"))
mean_monthly_gatherk2 <- mean_monthlyk2 %>%
  gather(key = "treatment",value="temp", -month, -year)

test_KBS_3 <- KBS_3
test_KBS_3$month <- format(test_KBS_3$Date_Time,format="%m")
test_KBS_3$year <- format(test_KBS_3$Date_Time,format="%y")
mean_monthly_warmk3 <- aggregate( XH_warmed_air_1m ~ month + year , test_KBS_3 , mean )
mean_monthly_ambk3 <- aggregate( XH_ambient_air_1m ~ month + year, test_KBS_3, mean)
mean_monthlyk3 <- merge(mean_monthly_ambk3, mean_monthly_warmk3, by=c("month", "year"))
mean_monthly_gatherk3 <- mean_monthlyk3 %>%
  gather(key = "treatment",value="temp", -month, -year)

KBS_season <- KBS
KBS_season$month <- format(KBS_season$Date_Time,format="%m")
KBS_season$year <- format(KBS_season$Date_Time,format="%Y")
KBS_season$hour <- format(KBS_season$Date_Time, format="%H")
KBS_season_air <- KBS_season %>%
  filter(month > "03") %>%
  filter(month < "09") %>%
  filter(hour > "06") %>%
  filter(hour < "20") %>%
  dplyr::select(Date_Time, year, month, hour, XH_warmed_air_1m, XH_ambient_air_1m)
KBS_season_air2 <- KBS_season %>%
        filter(month > "03") %>%
        filter(month < "09") %>%
        filter(hour > "06") %>%
        filter(hour < "20") %>%
        dplyr::select(Date_Time, year, month, hour, XH_warmed_air_1m, XH_ambient_air_1m,XU_warmed_air_10cm, XU_ambient_air_10cm)
KBS_season_soil <- KBS_season %>%
        filter(month > "03") %>%
        filter(month < "09") %>%
        filter(hour > "06") %>%
        filter(hour < "20") %>%
        dplyr::select(Date_Time, year, month, hour, XU_warmed_soil_temp_5cm, XU_ambient_soil_temp_5cm)
KBS_season_soilmo <- KBS_season %>%
        filter(month > "03") %>%
        filter(month < "09") %>%
        filter(hour > "06") %>%
        filter(hour < "20") %>%
        dplyr::select(Date_Time, year, month, hour, XH_warmed_soil_moisture_5cm, XH_ambient_soil_moisture_5cm)
KBS_avg_month <- KBS_season_air %>%  # by month and year
  gather(key = "treatment", value = "temp", -year, -month, -hour, -Date_Time) %>%
  group_by(month, year, treatment) %>%
  summarize(average_temp = mean(temp, na.rm = TRUE),
            se = std.error(temp, na.rm = TRUE))
KBS_avg_year <- KBS_season_air %>%  # by year
  gather(key = "treatment", value = "temp", -year, -month, -hour, -Date_Time) %>%
  group_by(year, treatment) %>%
  summarize(count = n(),
            average_temp = mean(temp, na.rm = TRUE),
            se = std.error(temp, na.rm = TRUE))
KBS_avg_year_air <- KBS_season_air2 %>%  # by year
        gather(key = "treatment", value = "temp", -year, -month, -hour, -Date_Time) %>%
        group_by(year, treatment) %>%
        summarize(count = n(),
                  average_temp = mean(temp, na.rm = TRUE),
                  se = std.error(temp, na.rm = TRUE))
KBS_avg_year_soil <- KBS_season_soil %>%  # by year
        gather(key = "treatment", value = "temp", -year, -month, -hour, -Date_Time) %>%
        group_by(year, treatment) %>%
        summarize(count = n(),
                  average_temp = mean(temp, na.rm = TRUE),
                  se = std.error(temp, na.rm = TRUE))
KBS_avg_year_soilmo <- KBS_season_soilmo %>%  # by year
        gather(key = "treatment", value = "moisture", -year, -month, -hour, -Date_Time) %>%
        group_by(year, treatment) %>%
        summarize(count = n(),
                  average_moist = mean(moisture, na.rm = TRUE),
                  se = std.error(moisture, na.rm = TRUE))
KBS_soil_merged <- rbind(KBS_avg_year_soil, KBS_avg_year_soilmo)

UMBS_season <- UMBS
UMBS_season$month <- format(UMBS_season$Date_Time,format="%m")
UMBS_season$year <- format(UMBS_season$Date_Time,format="%Y")
UMBS_season$hour <- format(UMBS_season$Date_Time, format="%H")
UMBS_season_air <- UMBS_season %>%
  filter(month > "03") %>%
  filter(month < "09") %>%
  filter(hour > "06") %>%
  filter(hour < "20") %>%
  dplyr::select(Date_Time, year, month, hour, XH_warmed_air_1m, XH_ambient_air_1m)
UMBS_season_air2 <- UMBS_season %>%
        filter(month > "03") %>%
        filter(month < "09") %>%
        filter(hour > "06") %>%
        filter(hour < "20") %>%
        dplyr::select(Date_Time, year, month, hour, XH_warmed_air_1m, XH_ambient_air_1m, XU_warmed_air_10cm, XU_ambient_air_10cm)
UMBS_season_soil <- UMBS_season %>%
        filter(month > "03") %>%
        filter(month < "09") %>%
        filter(hour > "06") %>%
        filter(hour < "20") %>%
        dplyr::select(Date_Time, year, month, hour, XU_warmed_soil_temp_5cm, XU_ambient_soil_temp_5cm)
UMBS_season_soilmo <- UMBS_season %>%
        filter(month > "03") %>%
        filter(month < "09") %>%
        filter(hour > "06") %>%
        filter(hour < "20") %>%
        dplyr::select(Date_Time, year, month, hour, XH_warmed_soil_moisture_5cm, XH_ambient_soil_moisture_5cm)
UMBS_avg_month <- UMBS_season_air %>%
  gather(key = "treatment", value = "temp", -year, -month, -hour, -Date_Time) %>%
  group_by(month, year, treatment) %>%
  summarize(average_temp = mean(temp, na.rm = TRUE),
            se = std.error(temp, na.rm = TRUE))
UMBS_avg_year <- UMBS_season_air %>%
  gather(key = "treatment", value = "temp", -year, -month, -hour, -Date_Time) %>%
  group_by(year, treatment) %>%
  summarize(average_temp = mean(temp, na.rm = TRUE),
            se = std.error(temp, na.rm = TRUE))
UMBS_avg_year_air <- UMBS_season_air2 %>%
        gather(key = "treatment", value = "temp", -year, -month, -hour, -Date_Time) %>%
        group_by(year, treatment) %>%
        summarize(count = n(),
                  average_temp = mean(temp, na.rm = TRUE),
                  se = std.error(temp, na.rm = TRUE))
UMBS_avg_year_soil <- UMBS_season_soil %>%
        gather(key = "treatment", value = "temp", -year, -month, -hour, -Date_Time) %>%
        group_by(year, treatment) %>%
        summarize(count = n(),
                  average_temp = mean(temp, na.rm = TRUE),
                  se = std.error(temp, na.rm = TRUE))
UMBS_avg_year_soilmo <- UMBS_season_soilmo %>%  # by year
        gather(key = "treatment", value = "moisture", -year, -month, -hour, -Date_Time) %>%
        group_by(year, treatment) %>%
        summarize(count = n(),
                  average_moist = mean(moisture, na.rm = TRUE),
                  se = std.error(moisture, na.rm = TRUE))
UMBS_soil_merged <- rbind(UMBS_avg_year_soil, UMBS_avg_year_soilmo)

KBS_hobo <- KBS
KBS_hobo$month <- format(KBS_hobo$Date_Time,format="%m")
KBS_hobo$year <- format(KBS_hobo$Date_Time,format="%Y")
KBS_hobo$day <- format(KBS_hobo$Date_Time,format="%d")
KBS_hobo1 <- KBS_hobo %>%
  dplyr::select(Date_Time, year, month, day, XH_warmed_air_1m, XH_ambient_air_1m) %>%
  gather(key = "treatment", value = "temp", -year, -month, -day, -Date_Time) %>%
  group_by(month, year, day, treatment) %>%
  summarize(average_temp = mean(temp, na.rm = TRUE),
            se_temp = std.error(temp, na.rm = TRUE)) %>%
  filter(year == "2017" & month == "07" & day == "04"|
           year == '2018' & month == '08' & day == '23'|
           year == '2019' & month == '08' & day == '02')
KBS_hobo2 <- KBS_hobo %>%
  dplyr::select(Date_Time, year, month, day, XH_warmed_air_1m, XH_ambient_air_1m) %>%
  gather(key = "treatment", value = "temp", -year, -month, -day, -Date_Time) %>%
  group_by(month, year, day, treatment) %>%
  summarize(average_temp = mean(temp, na.rm = TRUE),
            se_temp = std.error(temp, na.rm = TRUE)) %>%
  filter(year == "2019" & month == "05" & day == "15"|
           year == '2019' & month == '07' & day == '08'|
           year == '2019' & month == '08' & day == '02'|
           year == '2019' & month == '08' & day == '29'|
           year == '2019' & month == '09' & day == '18')
KBS_par$month <- format(KBS_par$Date_Time,format="%m")
KBS_par$year <- format(KBS_par$Date_Time,format="%Y")
KBS_par$day <- format(KBS_par$Date_Time,format="%d")
KBS_par1 <- KBS_par %>%
  dplyr::select(Date_Time, Plot, year, month, day, Percent_Sunlight) %>%
  filter(year == "2017" & month == "07" & day == "04"|
           year == '2018' & month == '08' & day == '23'|
           year == '2019' & month == '08' & day == '02') %>%
  filter(Plot == "B6"|Plot == "A6"|Plot == "B3"|Plot == "B4"|Plot == "C1"|Plot == "B1")
KBS_par2 <- KBS_par %>%
  dplyr::select(Date_Time, Plot, year, month, day, Percent_Sunlight) %>%
  filter(year == "2019" & month == "05" & day == "15"|
           year == '2019' & month == '07' & day == '08'|
           year == '2019' & month == '08' & day == '02'|
           year == '2019' & month == '08' & day == '29'|
           year == '2019' & month == '09' & day == '18') %>%
  filter(Plot == "B6"|Plot == "A6"|Plot == "B3"|Plot == "B4"|Plot == "C1"|Plot == "B1")
par_names <- c(B6="XH_warmed_air_1m", B3="XH_warmed_air_1m", C1="XH_warmed_air_1m", 
                A6="XH_ambient_air_1m", B4="XH_ambient_air_1m", B1="XH_ambient_air_1m")
KBS_par1$treatment <- as.character(par_names[KBS_par1$Plot])
KBS_par2$treatment <- as.character(par_names[KBS_par2$Plot])
KBS_par1 <- KBS_par1 %>%
  group_by(month, year, day, treatment) %>%
  summarize(average_par = mean(Percent_Sunlight, na.rm = TRUE),
            se_par = std.error(Percent_Sunlight, na.rm = TRUE))
KBS_par2 <- KBS_par2 %>%
  group_by(month, year, day, treatment) %>%
  summarize(average_par = mean(Percent_Sunlight, na.rm = TRUE),
            se_par = std.error(Percent_Sunlight, na.rm = TRUE))
KBS_comb <- merge(KBS_hobo1,KBS_par1,by=c("year","month","day","treatment"))
KBS_comb2 <- merge(KBS_hobo2,KBS_par2,by=c("year","month","day","treatment"))
ym <- c(paste0(KBS_comb$year,'_',KBS_comb$month,'_',KBS_comb$day))
KBS_comb <- data.frame(KBS_comb,ym)
ym2 <- c(paste0(KBS_comb2$year,'_',KBS_comb2$month,'_',KBS_comb2$day))
KBS_comb_2019 <- data.frame(KBS_comb2,ym2)
UMBS_hobo <- UMBS
UMBS_hobo$month <- format(UMBS_hobo$Date_Time,format="%m")
UMBS_hobo$year <- format(UMBS_hobo$Date_Time,format="%Y")
UMBS_hobo$day <- format(UMBS_hobo$Date_Time,format="%d")
UMBS_hobo1 <- UMBS_hobo %>%
  dplyr::select(Date_Time, year, month, day, XH_warmed_air_1m, XH_ambient_air_1m) %>%
  gather(key = "treatment", value = "temp", -year, -month, -day, -Date_Time) %>%
  group_by(month, year, day, treatment) %>%
  summarize(average_temp = mean(temp, na.rm = TRUE),
            se_temp = std.error(temp, na.rm = TRUE)) %>%
  filter(year == "2017" & month == "07" & day == "03"|
           year == '2017' & month == '07' & day == '10'|
           year == '2017' & month == '07' & day == '17'|
           year == '2017' & month == '07' & day == '24'|
           year == '2017' & month == '08' & day == '09'|
           year == '2017' & month == '08' & day == '16'|
           year == '2018' & month == '08' & day == '03'|
           year == '2018' & month == '08' & day == '13'|
           year == '2018' & month == '08' & day == '22'|
           year == '2018' & month == '08' & day == '31'|
           year == '2018' & month == '09' & day == '07'|
           year == '2018' & month == '09' & day == '13'|
           year == '2019' & month == '06' & day == '03'|
           year == '2019' & month == '06' & day == '18'|
           year == '2019' & month == '07' & day == '28')
UMBS_par$month <- format(UMBS_par$Date_Time,format="%m")
UMBS_par$year <- format(UMBS_par$Date_Time,format="%Y")
UMBS_par$day <- format(UMBS_par$Date_Time,format="%d")
UMBS_par1 <- UMBS_par %>%
  dplyr::select(Date_Time, Plot, year, month, day, Percent_Sunlight) %>%
  filter(Plot == "A5"|Plot == "A6"|Plot == "B4"|Plot == "C4"|Plot == "C2"|Plot == "D2")
par_names2 <- c(A5="XH_warmed_air_1m", C4="XH_warmed_air_1m", D2="XH_warmed_air_1m", 
               A6="XH_ambient_air_1m", B4="XH_ambient_air_1m", C2="XH_ambient_air_1m")
UMBS_par1$treatment <- as.character(par_names2[UMBS_par1$Plot])
UMBS_par1 <- UMBS_par1 %>%
  group_by(month, year, day, treatment) %>%
  summarize(average_par = mean(Percent_Sunlight, na.rm = TRUE),
            se_par = std.error(Percent_Sunlight, na.rm = TRUE))
UMBS_comb <- merge(UMBS_hobo1,UMBS_par1,by=c("year","month","day","treatment")) # over the years
ym3 <- c(paste0(UMBS_comb$year,'_',UMBS_comb$month,'_',UMBS_comb$day))
UMBS_comb <- data.frame(UMBS_comb,ym3)
UMBS_pend_season <- UMBS_pend
UMBS_pend_season$month <- format(UMBS_pend_season$Date_Time,format="%m")
UMBS_pend_season$year <- format(UMBS_pend_season$Date_Time,format="%Y")
UMBS_pend_season$hour <- format(UMBS_pend_season$Date_Time, format="%H")
UMBS_pend_season <- UMBS_pend_season %>%
  filter(month > "03") %>%
  filter(month < "09") %>%
  filter(hour > "06") %>%
  filter(hour < "20") %>%
  na.omit(Temp_F_XP_air_1m) %>%
  dplyr::select(Date_Time, year, month, hour, Pendant_ID, Temp_F_XP_air_1m)
KBS_pend_season <- KBS_pend
KBS_pend_season$month <- format(KBS_pend_season$Date_Time,format="%m")
KBS_pend_season$year <- format(KBS_pend_season$Date_Time,format="%Y")
KBS_pend_season$hour <- format(KBS_pend_season$Date_Time, format="%H")
KBS_pend_season <- KBS_pend_season %>%
  filter(month > "03") %>%
  filter(month < "09") %>%
  filter(hour > "06") %>%
  filter(hour < "20") %>%
  dplyr::select(Date_Time, year, month, hour, Pendant_ID, Temp_F_XP_air_1m)
```

## Air temperatures at KBS and UMBS
```{r, echo = FALSE, message=FALSE}
Fig1_kbs <- ggplot(KBS_avg_year_air, aes(x=year, y=average_temp, fill=treatment, shape=treatment)) +
  geom_pointrange(aes(ymin = average_temp - se, ymax = average_temp + se), size=1, color="black") +
  scale_fill_manual(labels = c("Ambient air (1m)", "Warmed air (1m)", "Ambient air (10cm)", "Warmed air (10cm)"), values=c('steelblue3','#fb6a4a','steelblue3','#fb6a4a'))+
  scale_shape_manual(labels = c("Ambient air (1m)", "Warmed air (1m)", "Ambient air (10cm)", "Warmed air (10cm)"), values=c(22, 22, 21, 21))+
  labs(title="KBS",y=NULL, x=NULL, fill="Treatment", shape="Treatment") +
  theme(legend.position="bottom") +
  theme_classic()

Fig1_umbs <- ggplot(UMBS_avg_year_air, aes(x=year, y=average_temp, fill=treatment, shape=treatment)) +
        geom_pointrange(aes(ymin = average_temp - se, ymax = average_temp + se), size=1, color="black") +
        scale_fill_manual(labels = c("Ambient air (1m)", "Warmed air (1m)", "Ambient air (10cm)", "Warmed air (10cm)"), values=c('steelblue3','#fb6a4a','steelblue3','#fb6a4a'))+
        scale_shape_manual(labels = c("Ambient air (1m)", "Warmed air (1m)", "Ambient air (10cm)", "Warmed air (10cm)"), values=c(22, 22, 21, 21))+
        labs(title="UMBS",y=NULL, x=NULL, fill="Treatment", shape="Treatment") +
        theme(legend.position="bottom") +
        theme_classic()

Fig1 <- ggarrange(Fig1_kbs, Fig1_umbs, ncol = 2, common.legend = T, legend = "right")
annotate_figure(Fig1,
                left = text_grob("Average Temperature (°C)", color = "black", rot = 90),
                bottom = text_grob("Year", color = "black"))
```

## Soil temperatures at KBS and UMBS
```{r, echo=F, message=F}
Fig1_soil_kbs <- ggplot(KBS_avg_year_soil, aes(x=year, y=average_temp, fill=treatment, shape=treatment)) +
        geom_pointrange(aes(ymin = average_temp - se, ymax = average_temp + se), size=1, color="black") +
        scale_fill_manual(labels = c("Ambient soil (5cm)", "Warmed soil (5cm)"), values=c('steelblue3','#fb6a4a'))+
        scale_shape_manual(labels = c("Ambient soil (5cm)", "Warmed soil (5cm)"), values=c(24, 24))+
        labs(title="KBS",y=NULL, x=NULL, fill="Treatment", shape="Treatment") +
        theme(legend.position="bottom") +
        theme_classic()

Fig1_soil_umbs <- ggplot(UMBS_avg_year_soil, aes(x=year, y=average_temp, fill=treatment, shape=treatment)) +
        geom_pointrange(aes(ymin = average_temp - se, ymax = average_temp + se), size=1, color="black") +
        scale_fill_manual(labels = c("Ambient soil (5cm)", "Warmed soil (5cm)"), values=c('steelblue3','#fb6a4a'))+
        scale_shape_manual(labels = c("Ambient soil (5cm)", "Warmed soil (5cm)"), values=c(24, 24))+
        labs(title="UMBS",y=NULL, x=NULL, fill="Treatment", shape="Treatment") +
        theme(legend.position="bottom") +
        theme_classic()

Fig1.1 <- ggarrange(Fig1_soil_kbs, Fig1_soil_umbs, ncol = 2, common.legend = T, legend = "right")
annotate_figure(Fig1.1,
                left = text_grob("Average Temperature (°C)", color = "black", rot = 90),
                bottom = text_grob("Year", color = "black"))
```

## KBS and UMBS soil temperature and moisture - dual y axis plot
```{r, echo=F, message=F}
Fig1_soil_kbs_dualy <- ggplot(KBS_soil_merged, aes(x=year, fill=treatment, shape=treatment)) +
        geom_pointrange(aes(y=average_temp, ymin = average_temp - se, ymax = average_temp + se), size=1, color="black") +
        geom_pointrange(aes(y=average_moist*100, ymin = average_moist*100 - se*100, ymax = average_moist*100 + se*100), size=1, color="black") +
        scale_y_continuous(
                name = NULL,
                sec.axis = sec_axis(~./100, name=NULL)) +
        scale_fill_manual(labels = c("Ambient soil moisture (5cm)", "Warmed soil moisture (5cm)","Ambient soil temperature (5cm)","Warmed soil temperature (5cm)"), values=c('steelblue3','#fb6a4a','steelblue3','#fb6a4a'))+
        scale_shape_manual(labels = c("Ambient soil moisture (5cm)", "Warmed soil moisture (5cm)","Ambient soil temperature (5cm)","Warmed soil temperature (5cm)"), values=c(22,22,21,21))+
        labs(title="KBS",y=NULL, x="Year", fill="Treatment", shape="Treatment") +
        theme(legend.position="bottom") +
        theme_classic()

Fig1_soil_umbs_dualy <- ggplot(UMBS_soil_merged, aes(x=year, fill=treatment, shape=treatment)) +
        geom_pointrange(aes(y=average_temp, ymin = average_temp - se, ymax = average_temp + se), size=1, color="black") +
        geom_pointrange(aes(y=average_moist*100, ymin = average_moist*100 - se*100, ymax = average_moist*100 + se*100), size=1, color="black") +
        scale_y_continuous(
                name = NULL,
                sec.axis = sec_axis(~./100, name=NULL)) +
        scale_fill_manual(labels = c("Ambient soil moisture (5cm)", "Warmed soil moisture (5cm)","Ambient soil temperature (5cm)","Warmed soil temperature (5cm)"), values=c('steelblue3','#fb6a4a','steelblue3','#fb6a4a'))+
        scale_shape_manual(labels = c("Ambient soil moisture (5cm)", "Warmed soil moisture (5cm)","Ambient soil temperature (5cm)","Warmed soil temperature (5cm)"), values=c(22,22,21,21))+
        labs(title="UMBS",y=NULL, x="Year", fill="Treatment", shape="Treatment") +
        theme(legend.position="bottom") +
        theme_classic()

Fig1.2 <- ggarrange(Fig1_soil_kbs_dualy, Fig1_soil_umbs_dualy, ncol = 2, common.legend = T, legend = "bottom")
annotate_figure(Fig1.2,
                left = text_grob("Soil Temperature (°C)", color = "black", rot = 90),
                right = text_grob("Soil Moisture", color = "black", rot = 270))
```

## KBS and UMBS air temperature pendants - looking to see if warming is consistent btwn chambers
```{r, echo = FALSE, message=FALSE}
ggplot(KBS_pend_season, aes(x = Pendant_ID, y = Temp_F_XP_air_1m)) +
  facet_grid(.~year) +
  geom_boxplot() +
  theme_classic() +
  labs(title = "KBS")
ggplot(UMBS_pend_season, aes(x = Pendant_ID, y = Temp_F_XP_air_1m)) +
  facet_grid(.~year) +
  geom_boxplot() +
  theme_classic() +
  labs(title = "UMBS")
```

## Other air temperature plot by year (older)
```{r, echo = FALSE, message=FALSE}
k_year <- ggplot(KBS_avg_year, aes(x = year, y = average_temp, fill = treatment)) + 
  geom_bar(position = "identity", alpha = 0.5, stat = "identity", color = 'black') +
  geom_errorbar(aes(ymin = average_temp - se, ymax = average_temp + se), width = 0.2,
                position = "identity") +
  ylim(0, 30) +
  scale_fill_manual(labels = c("Ambient", "Warmed"), values=c('darkblue','lightblue'))+
  theme_classic() +
  labs(title = "KBS", x=NULL, y = NULL, fill = "Treatment")
u_year <- ggplot(UMBS_avg_year, aes(x = year, y = average_temp, fill = treatment)) + 
  geom_bar(position = "identity", alpha = 0.5, stat = "identity", color = 'black') +
  geom_errorbar(aes(ymin = average_temp - se, ymax = average_temp + se), width = 0.2,
                position = "identity") +
  ylim(0, 30) +
  scale_fill_manual(labels = c("Ambient", "Warmed"), values=c('darkblue','lightblue'))+
  theme_classic() +
  labs(title = "UMBS", x="Year", y = NULL, fill = "Treatment")
final_year <- ggarrange(k_year, u_year, nrow = 2, common.legend = TRUE, legend = "bottom")
annotate_figure(final_year,
                left = text_grob("Average Temperature (°C)", color = "black", rot = 90))
```

## Other air temperature plot by month (older)
```{r, echo = FALSE, message=FALSE}
k_month <- ggplot(KBS_avg_month, aes(x = month, y = average_temp, fill = treatment)) + 
  facet_grid(.~year) +
  geom_bar(position = "identity", alpha = 0.5, stat = "identity", color = 'black') +
  geom_errorbar(aes(ymin = average_temp - se, ymax = average_temp + se), width = 0.2,
                position = "identity") +
  ylim(0, 30) +
  scale_fill_manual(labels = c("Ambient", "Warmed"), values=c('darkblue','lightblue'))+
  theme_classic() +
  labs(x = NULL, y = NULL, fill = "Treatment")
u_month <- ggplot(UMBS_avg_month, aes(x = month, y = average_temp, fill = treatment)) + 
  facet_grid(.~year) +
  geom_bar(position = "identity", alpha = 0.5, stat = "identity", color = 'black') +
  geom_errorbar(aes(ymin = average_temp - se, ymax = average_temp + se), width = 0.2,
                position = "identity") +
  ylim(0, 30) +
  scale_fill_manual(labels = c("Ambient", "Warmed"), values=c('darkblue','lightblue'))+
  theme_classic() +
  labs(x="Month", y = NULL, fill = "Treatment")
final_month <- ggarrange(k_month, u_month, nrow = 2, common.legend = TRUE, legend = "bottom")
annotate_figure(final_month,
                left = text_grob("Average Temperature (°C)", color = "black", rot = 90),
                top = text_grob("Year"))
```

## KBS - Comparing PAR to HOBO from 2017-2019; only one day is shown for each year because PAR measurements were only taken on one day at KBS for 2017 and 2018 (older)
```{r, echo = FALSE, message=FALSE}
ggplot(KBS_comb, aes(x=ym, fill=treatment, color = treatment)) +
  geom_bar( aes(y=average_temp), position = "identity",alpha = 0.5, stat = "identity", color = 'black') + 
  geom_errorbar(aes(ymin = average_temp - se_temp, ymax = average_temp + se_temp), width = 0.05,
                position = "identity", color = 'black') +
  geom_point( aes(y=average_par * 100), size=2) +
  geom_line(aes(y = average_par * 100, group = treatment)) +
  scale_y_continuous(
    name = "Air Temperature (°C)",
    sec.axis = sec_axis(~./100, name="PAR")) + 
  scale_fill_manual(labels = c("Ambient", "Warmed"), values=c('darkblue','lightblue')) +
  scale_color_manual(labels = c("Ambient", "Warmed"), values=c('black','white'))+
  theme_classic()
```

## KBS - Again, comparing PAR to HOBO, this time in line format - PAR in dashed lines (older)
```{r, echo = FALSE, message=FALSE}
ggplot(KBS_comb, aes(x = ym, color = treatment, shape = treatment)) +
  geom_point(aes(y = average_temp)) +
  geom_point(aes(y = average_par * 100)) +
  geom_line(aes(y = average_temp, group = treatment)) +
  geom_line(aes(y = average_par * 100, group = treatment), linetype = "dashed") +
  scale_y_continuous(
    name = "Air Temperature (°C)",
    sec.axis = sec_axis(~./100, name="PAR")) + 
  labs(x = "Year, month, day") +
  scale_color_manual(labels = c("Ambient", "Warmed"), name = "Treatment", values=c('midnightblue','coral')) +
  scale_shape_manual(labels = c("Ambient", "Warmed"), name = "Treatment", values=c(17,16)) +
  theme_classic() +
  theme(legend.position="bottom")
```

## KBS- Comparing PAR to HOBO for only 2019 because multiple PAR measurements were taken that year - PAR in dashed lines (older)
```{r, echo = FALSE, message=FALSE}
ggplot(KBS_comb_2019, aes(x = ym2, color = treatment, shape = treatment)) +
  geom_point(aes(y = average_temp)) +
  geom_point(aes(y = average_par * 100)) +
  geom_line(aes(y = average_temp, group = treatment)) +
  geom_line(aes(y = average_par * 100, group = treatment), linetype = 'dashed') +
  scale_y_continuous(
    name = "Air Temperature (°C)",
    sec.axis = sec_axis(~./100, name="PAR")) + 
  labs(x = "Year, month, day") +
  scale_color_manual(labels = c("Ambient", "Warmed"), name = "Treatment", values=c('midnightblue','coral')) +
  scale_shape_manual(labels = c("Ambient", "Warmed"), name = "Treatment", values=c(17,16)) +
  theme_classic()+
  theme(legend.position="bottom")
```

## KBS - Simple linear regression between temp and par: F(1,3) = 32.21, p-value = 0.011 (older)
```{r, echo = FALSE, message=FALSE}
KBS_comb_lm <- KBS_comb_2019 %>%
  filter(treatment == "XH_warmed_air_1m")
lm_kbs <- lm(average_temp ~ average_par, data = KBS_comb_lm)
plot(average_temp ~ average_par, data = KBS_comb_lm)
abline(lm_kbs)
```

## UMBS - Simple linear regression between temp and par: F(1,13) = 1.45, p-value = 0.25 (older)
```{r, echo = FALSE, message=FALSE}
UMBS_comb_lm <- UMBS_comb %>%
  filter(treatment == "XH_warmed_air_1m")
lm_umbs <- lm(average_temp ~ average_par, data = UMBS_comb_lm)
plot(average_temp ~ average_par, data = UMBS_comb_lm)
abline(lm_umbs)
```

## Soil temperature over time - varies between ambient or warmed treatments
## no sig difference for 2018 (p-value = 0.87) (older)
```{r, echo = FALSE, message=FALSE}
ggplot(KBS_avg_year_soil, aes(x = year, y = average_temp, fill = treatment)) + 
  geom_bar(stat = "identity", position = position_dodge(), color = 'black') +
  geom_errorbar(aes(ymin = average_temp - se, ymax = average_temp + se), width = 0.2,
                position = position_dodge(0.9)) +
  scale_fill_manual(labels = c("Ambient", "Warmed"), values=c('#a3b4d3','#d8ecf2'))+
  theme_classic() +
  labs(title = "KBS", x="Year", y = "Average Soil Temperature (°C)", fill = "Treatment") +
  theme(legend.position = "bottom")
```

## Soil moisture over time - ambient retains more moisture (p-value < 0.001 for all) (older)
```{r, echo = FALSE, message=FALSE}
ggplot(KBS_avg_year_soilmo, aes(x = year, y = average_moist, fill = treatment)) + 
  geom_bar(position = "identity", stat = "identity", color = 'black', alpha = 1) +
  geom_errorbar(aes(ymin = average_moist - se, ymax = average_moist + se), width = 0.2,
                position = "identity") +
  scale_fill_manual(labels = c("Ambient", "Warmed"), values=c('#a3b4d3','#d8ecf2'))+
  theme_classic() +
  labs(title = "KBS", x="Year", y = "Soil Moisture", fill = "Treatment")
```

### 1H sensor
```{r, echo = FALSE, message=FALSE}
ggplot(mean_monthly_gatherk1, aes(x = as.numeric(month), y = temp)) +
  geom_point(aes(col = treatment, shape = year), size = 1) +
  geom_smooth(method = "loess", aes(linetype = treatment, col = treatment), size = 1.2,
              fill = "grey90") +
  labs(x = "Month",y = "Temperature °C") +
  theme_minimal() + 
  theme(legend.position = "bottom") +
  scale_x_discrete(limits=c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12"))
```

### 2H sensor
```{r, echo = FALSE, message=FALSE}
ggplot(mean_monthly_gatherk2, aes(x = as.numeric(month), y = temp)) +
  geom_point(aes(col = treatment, shape = year), size = 1) +
  geom_smooth(method = "loess", aes(linetype = treatment, col = treatment), size = 1.2,
              fill = "grey90") +
  labs(x = "Month",y = "Temperature °C") +
  theme_minimal() + 
  theme(legend.position = "bottom") +
  scale_x_discrete(limits=c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12"))
```

### 3H sensor
```{r, echo = FALSE, message=FALSE}
ggplot(mean_monthly_gatherk3, aes(x = as.numeric(month), y = temp)) +
  geom_point(aes(col = treatment, shape = year), size = 1) +
  geom_smooth(method = "loess", aes(linetype = treatment, col = treatment), size = 1.2,
              fill = "grey90") +
  labs(x = "Month",y = "Temperature °C") +
  theme_minimal() + 
  theme(legend.position = "bottom") +
  scale_x_discrete(limits=c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12"))
```
