---
title: 'warmXtrophic: HOBO analyses'
author: "Kara Dobson"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## COLLABORATORS: Phoebe Zarnetske, Nina Lany, Kathryn Schmidt, Mark Hammond, Pat Bills, Kileigh Welshofer, Moriah Young
## DATA INPUT: CSV files are located in the HOBO_data folder in the shared Google drive
## DATA OUTPUT: Computes different averages + error for hobo variables measured at each site

```{r, message=F}
# clear all existing data
rm(list=ls())

# load in packages and set working directory
library(tidyverse)
library(plotrix)
library(ggpubr)
library(rstatix)
library(sjPlot)

# Set working directory
L1_dir<-Sys.getenv("L1DIR")

# load in the data
KBS <- read.csv(file.path(L1_dir,"HOBO_data/HOBO_paired_sensor_data/KBS/KBS_pairedsensors_dataremoved_L1.csv"))
UMBS <- read.csv(file.path(L1_dir,"HOBO_data/HOBO_paired_sensor_data/UMBS/UMBS_pairedsensors_dataremoved_L1.csv"))

KBS_pend <- read.csv(file.path(L1_dir,"HOBO_data/HOBO_pendant_data/KBS/KBS_HOBOpendant_L1.csv"))
UMBS_pend <- read.csv(file.path(L1_dir,"HOBO_data/HOBO_pendant_data/UMBS/UMBS_HOBOpendant_L1.csv"))

# date is a character column - convert to date format
KBS$Date_Time <- as.POSIXct(KBS$Date_Time, format = "%Y-%m-%d %H:%M")
UMBS$Date_Time <- as.POSIXct(UMBS$Date_Time, format = "%Y-%m-%d %H:%M")
```

## KBS air temperatures
```{r}
# merge the data + filter data for only the daytime during the growing season
KBS_season1 <- KBS
KBS_season1$month <- format(KBS_season1$Date_Time,format="%m")
KBS_season1$year <- format(KBS_season1$Date_Time,format="%Y")
KBS_season1$hour <- format(KBS_season1$Date_Time, format="%H")

# 1m
KBS_season <- KBS_season1 %>%
  filter(month > "03") %>%
  filter(month < "09") %>%
  filter(hour > "06") %>%
  filter(hour < "20") %>%
  dplyr::select(Date_Time, year, month, hour, sensor, XH_warmed_air_1m, XH_ambient_air_1m)

# 10cm
KBS_season_10cm <- KBS_season1 %>%
  filter(month > "03") %>%
  filter(month < "09") %>%
  filter(hour > "06") %>%
  filter(hour < "20") %>%
  filter(sensor == 2) %>% # only sensor 2 has 10cm data, 1 and 3 removed (see HOBO_data_removal_L1.R)
  dplyr::select(Date_Time, year, month, hour, sensor, XU_warmed_air_10cm, XU_ambient_air_10cm)

# create new dataframes for temperatures averaged by year & averaged by month and year
KBS_avg_year <- KBS_season %>%
  gather(key = "treatment", value = "temp", -year, -month, -hour, -Date_Time, -sensor)
KBS_avg_year_10cm <- KBS_season_10cm %>%
  gather(key = "treatment", value = "temp", -year, -month, -hour, -Date_Time, -sensor)

# test for outliers - none extreme
outliers <- KBS_avg_year %>%
  group_by(treatment, year, sensor) %>%
  identify_outliers(temp)
view(outliers)
outliers <- KBS_avg_year_10cm %>%
  group_by(treatment, year, sensor) %>%
  identify_outliers(temp)
view(outliers)

# t-test
avgyear_t.test <- t.test(temp~treatment,data=KBS_avg_year)
avgyear_t.test
avgyear10cm_t.test <- t.test(temp~treatment,data=KBS_avg_year_10cm)
avgyear10cm_t.test

# avg temps in the chambers during the daytime (over all years)
KBS_avg_temp <- KBS_avg_year %>%
        group_by(treatment, sensor) %>%
        summarize(mean_temp = mean(temp, na.rm = T)) %>%
        group_by(treatment) %>%
        summarize(mean = mean(mean_temp, na.rm = T),
                  sd = sd(mean_temp, na.rm = T))
print(KBS_avg_temp)
KBS_avg_temp_10cm <- KBS_avg_year_10cm %>%
        group_by(treatment, sensor) %>%
        summarize(mean_temp = mean(temp, na.rm = T)) %>%
        group_by(treatment) %>%
        summarize(mean = mean(mean_temp, na.rm = T),
                  sd = sd(mean_temp, na.rm = T))
print(KBS_avg_temp_10cm) # no sd because there is only one sensor


# avg temps in the chambers during the daytime for each year + the difference between warmed and ambient
KBS_avg_temp_year <- KBS_avg_year %>%
        group_by(year, treatment, sensor) %>%
        summarize(mean_temp = mean(temp, na.rm = T)) %>%
        group_by(year, treatment) %>%
        summarize(mean = mean(mean_temp, na.rm = T),
                  sd = sd(mean_temp, na.rm = T))
kbs_avg_temp_year2 <- KBS_avg_temp_year %>%
        select(-sd) %>%
        pivot_wider(names_from = treatment, values_from = mean) %>%
        group_by(year) %>%
        summarize(diff = XH_warmed_air_1m-XH_ambient_air_1m)
print(kbs_avg_temp_year2)
tab_df(KBS_avg_temp_year)


# avg temps in the chambers on hot days
KBS_avg_hot_day <- KBS_season %>%
        filter(XH_ambient_air_1m > 27) %>%
        gather(key = "treatment", value = "temp", -year, -month, -hour, -Date_Time, -sensor) %>%
        group_by(treatment, sensor) %>%
        summarize(mean_temp = mean(temp, na.rm = T)) %>%
        group_by(treatment) %>%
        summarize(mean = mean(mean_temp, na.rm = T),
                  sd = sd(mean_temp, na.rm = T))
print(KBS_avg_hot_day)


# avg temps from march-april and july-august (early season vs late season)
KBS_early <- KBS_season %>%
        filter(month == "03" | month == "04") %>%
        gather(key = "treatment", value = "temp", -year, -month, -hour, -Date_Time, -sensor) %>%
        group_by(year, sensor, treatment) %>%
        summarize(mean_temp = mean(temp, na.rm = T)) %>%
        group_by(year, treatment) %>%
        summarize(mean = mean(mean_temp, na.rm = T),
                  sd = sd(mean_temp, na.rm = T))
KBS_early2 <- KBS_early %>%
        select(-sd) %>%
        pivot_wider(names_from = treatment, values_from = mean) %>%
        group_by(year) %>%
        summarize(diff = XH_warmed_air_1m-XH_ambient_air_1m)
print(KBS_early2) # averaged these values to get early season amount
tab_df(KBS_early)
KBS_late <- KBS_season %>%
        filter(month == "07" | month == "08") %>%
        gather(key = "treatment", value = "temp", -year, -month, -hour, -Date_Time, -sensor) %>%
        group_by(year, sensor, treatment) %>%
        summarize(mean_temp = mean(temp, na.rm = T)) %>%
        group_by(year, treatment) %>%
        summarize(mean = mean(mean_temp, na.rm = T),
                  sd = sd(mean_temp, na.rm = T))
KBS_late2 <- KBS_late %>%
        select(-sd) %>%
        pivot_wider(names_from = treatment, values_from = mean) %>%
        group_by(year) %>%
        summarize(diff = XH_warmed_air_1m-XH_ambient_air_1m)
print(KBS_late2) # averaged these values to get late season amount
tab_df(KBS_late)


# winter warming?
KBS_winter <- KBS
KBS_winter$month <- format(KBS_winter$Date_Time,format="%m")
KBS_winter$year <- format(KBS_winter$Date_Time,format="%Y")
KBS_winter$hour <- format(KBS_winter$Date_Time, format="%H")

KBS_winter <- KBS_winter %>%
        filter(month > "10" | month < "03") %>%
        filter(hour > "06") %>%
        filter(hour < "20") %>%
        dplyr::select(Date_Time, year, month, hour, XH_warmed_air_1m, XH_ambient_air_1m, sensor)

KBS_winter <- KBS_winter %>%
        gather(key = "treatment", value = "temp", -year, -month, -hour, -Date_Time, -sensor)

# t-test
winter_t.test <- t.test(temp~treatment,data=KBS_winter)
winter_t.test

KBS_avg_winter <- KBS_winter %>%
        group_by(sensor, treatment) %>%
        summarize(mean_temp = mean(temp, na.rm = T)) %>%
        group_by(treatment) %>%
        summarize(mean = mean(mean_temp, na.rm = T),
                  sd = sd(mean_temp, na.rm = T))
#KBS_avg_winter2 <- KBS_avg_winter %>%
#        select(-sd) %>%
#        pivot_wider(names_from = treatment, values_from = mean) %>%
#        group_by(year) %>%
#        summarize(diff = XH_warmed_air_1m-XH_ambient_air_1m)
print(KBS_avg_winter) 


#nighttime warming?
KBS_night <- KBS
KBS_night$month <- format(KBS_night$Date_Time,format="%m")
KBS_night$year <- format(KBS_night$Date_Time,format="%Y")
KBS_night$hour <- format(KBS_night$Date_Time, format="%H")
KBS_night <- KBS_night %>%
        filter(month > "03" | month < "09") %>%
        filter(hour > "20" | hour < "06") %>%
        select(Date_Time, year, month, hour, XH_warmed_air_1m, XH_ambient_air_1m, sensor)

KBS_night <- KBS_night %>%
        gather(key = "treatment", value = "temp", -year, -month, -hour, -Date_Time, -sensor)

# t-test
night_t.test <- t.test(temp~treatment,data=KBS_night)
night_t.test

KBS_avg_night <- KBS_night %>%
        group_by(sensor, treatment) %>%
        summarize(mean_temp = mean(temp, na.rm = T)) %>%
        group_by(treatment) %>%
        summarize(mean = mean(mean_temp, na.rm = T),
                  sd = sd(mean_temp, na.rm = T))
#KBS_avg_night2 <- KBS_avg_night %>%
#        select(-sd) %>%
#        pivot_wider(names_from = treatment, values_from = mean) %>%
#        group_by(year) %>%
#        summarize(diff = XH_warmed_air_1m-XH_ambient_air_1m)
print(KBS_avg_night)
```

## UMBS air temperatures
```{r}
UMBS_season1 <- UMBS
UMBS_season1$month <- format(UMBS_season1$Date_Time,format="%m")
UMBS_season1$year <- format(UMBS_season1$Date_Time,format="%Y")
UMBS_season1$hour <- format(UMBS_season1$Date_Time, format="%H")

# 1m
UMBS_season <- UMBS_season1 %>%
  filter(month > "03") %>%
  filter(month < "09") %>%
  filter(hour > "06") %>%
  filter(hour < "20") %>%
  select(Date_Time, year, month, hour, XH_warmed_air_1m, XH_ambient_air_1m, sensor)

# 10cm
UMBS_season_10cm <- UMBS_season1 %>%
  filter(month > "03") %>%
  filter(month < "09") %>%
  filter(hour > "06") %>%
  filter(hour < "20") %>%
  dplyr::select(Date_Time, year, month, hour, XU_warmed_air_10cm, XU_ambient_air_10cm, sensor)

# create new dataframes for temperatures averaged by year & averaged by month and year
UMBS_avg_year <- UMBS_season %>%
  gather(key = "treatment", value = "temp", -year, -month, -hour, -Date_Time, -sensor)
UMBS_avg_year_10cm <- UMBS_season_10cm %>%
  gather(key = "treatment", value = "temp", -year, -month, -hour, -Date_Time, -sensor)

# test for outliers - none extreme
outliers <- UMBS_avg_year %>%
  group_by(treatment, year) %>%
  identify_outliers(temp)
View(outliers)
outliers <- UMBS_avg_year_10cm %>%
  group_by(treatment, year) %>%
  identify_outliers(temp)
View(outliers)

# t-test
avgyear_umbs_t.test <- t.test(temp~treatment,data=UMBS_avg_year)
avgyear_umbs_t.test
avgyear10cm_umbs_t.test <- t.test(temp~treatment,data=UMBS_avg_year_10cm)
avgyear10cm_umbs_t.test


# avg temps in the chambers during the daytime
UMBS_avg_temp <- UMBS_avg_year %>%
        group_by(sensor, treatment) %>%
        summarize(mean_temp = mean(temp, na.rm = T))  %>%
        group_by(treatment) %>%
        summarize(mean = mean(mean_temp, na.rm = T),
                  sd = sd(mean_temp, na.rm = T))
print(UMBS_avg_temp)
UMBS_avg_temp_10cm <- UMBS_avg_year_10cm %>%
        group_by(sensor, treatment) %>%
        summarize(mean_temp = mean(temp, na.rm = T)) %>%
        group_by(treatment) %>%
        summarize(mean = mean(mean_temp, na.rm = T),
                  sd = sd(mean_temp, na.rm = T))
print(UMBS_avg_temp_10cm)


# avg temps in the chambers during the daytime for each year
UMBS_avg_temp_year <- UMBS_avg_year %>%
        group_by(year, sensor, treatment) %>%
        summarize(mean_temp = mean(temp, na.rm = T)) %>%
        group_by(year, treatment) %>%
        summarize(mean = mean(mean_temp, na.rm = T),
                  sd = sd(mean_temp, na.rm = T))
UMBS_avg_temp_year2 <- UMBS_avg_temp_year %>%
        select(-sd) %>%
        pivot_wider(names_from = treatment, values_from = mean) %>%
        group_by(year) %>%
        summarize(diff = XH_warmed_air_1m-XH_ambient_air_1m)
print(UMBS_avg_temp_year2)
tab_df(UMBS_avg_temp_year)


# avg temps in the chambers on hot days
UMBS_avg_hot_day <- UMBS_season %>%
        filter(XH_ambient_air_1m > 27) %>%
        gather(key = "treatment", value = "temp", -year, -month, -hour, -Date_Time, -sensor) %>%
        group_by(sensor, treatment) %>%
        summarize(mean_temp = mean(temp, na.rm = T))  %>%
        group_by(treatment) %>%
        summarize(mean = mean(mean_temp, na.rm = T),
                  sd = sd(mean_temp, na.rm = T))
print(UMBS_avg_hot_day)


# avg temps from march-april and july-august (early season vs late season)
UMBS_early <- UMBS_season %>%
        filter(month == "03" | month == "04") %>%
        gather(key = "treatment", value = "temp", -year, -month, -hour, -Date_Time, -sensor) %>%
        group_by(year, sensor, treatment) %>%
        summarize(mean_temp = mean(temp, na.rm = T))  %>%
        group_by(year, treatment) %>%
        summarize(mean = mean(mean_temp, na.rm = T),
                  sd = sd(mean_temp, na.rm = T))
UMBS_early2 <- UMBS_early %>%
        select(-sd) %>%
        pivot_wider(names_from = treatment, values_from = mean) %>%
        group_by(year) %>%
        summarize(diff = XH_warmed_air_1m-XH_ambient_air_1m)
print(UMBS_early2) #averaged these
tab_df(UMBS_early)

UMBS_late <- UMBS_season %>%
        filter(month == "07" | month == "08") %>%
        gather(key = "treatment", value = "temp", -year, -month, -hour, -Date_Time, -sensor) %>%
        group_by(year, sensor, treatment) %>%
        summarize(mean_temp = mean(temp, na.rm = T))  %>%
        group_by(year, treatment) %>%
        summarize(mean = mean(mean_temp, na.rm = T),
                  sd = sd(mean_temp, na.rm = T))
UMBS_late2 <- UMBS_late %>%
        select(-sd) %>%
        pivot_wider(names_from = treatment, values_from = mean) %>%
        group_by(year) %>%
        summarize(diff = XH_warmed_air_1m-XH_ambient_air_1m)
print(UMBS_late2) #averaged these
tab_df(UMBS_late)


# winter warming?
UMBS_winter <- UMBS
UMBS_winter$month <- format(UMBS_winter$Date_Time,format="%m")
UMBS_winter$year <- format(UMBS_winter$Date_Time,format="%Y")
UMBS_winter$hour <- format(UMBS_winter$Date_Time, format="%H")
UMBS_winter <- UMBS_winter %>%
        filter(month > "10" | month < "03") %>%
        filter(hour > "06") %>%
        filter(hour < "20") %>%
        select(Date_Time, year, month, hour, XH_warmed_air_1m, XH_ambient_air_1m, sensor)
UMBS_winter <- UMBS_winter %>%
        gather(key = "treatment", value = "temp", -year, -month, -hour, -Date_Time, -sensor)

winter_umbs_t.test <- t.test(temp~treatment,data=UMBS_winter)
winter_umbs_t.test

UMBS_avg_winter <- UMBS_winter %>%
        group_by(sensor, treatment) %>%
        summarize(mean_temp = mean(temp, na.rm = T))  %>%
        group_by(treatment) %>%
        summarize(mean = mean(mean_temp, na.rm = T),
                  sd = sd(mean_temp, na.rm = T))
#UMBS_avg_winter2 <- UMBS_avg_winter %>%
#        select(-sd) %>%
#        pivot_wider(names_from = treatment, values_from = mean) %>%
#        group_by(year) %>%
#        summarize(diff = XH_warmed_air_1m-XH_ambient_air_1m)
print(UMBS_avg_winter)


#nighttime warming?
UMBS_night <- UMBS
UMBS_night$month <- format(UMBS_night$Date_Time,format="%m")
UMBS_night$year <- format(UMBS_night$Date_Time,format="%Y")
UMBS_night$hour <- format(UMBS_night$Date_Time, format="%H")
UMBS_night <- UMBS_night %>%
        filter(month > "03" | month < "09") %>%
        filter(hour > "20" | hour < "06") %>%
        select(Date_Time, year, month, hour, XH_warmed_air_1m, XH_ambient_air_1m, sensor)
UMBS_night <- UMBS_night %>%
        gather(key = "treatment", value = "temp", -year, -month, -hour, -Date_Time, -sensor)

night_umbs_t.test <- t.test(temp~treatment,data=UMBS_night)
night_umbs_t.test

UMBS_avg_night <- UMBS_night %>%
        group_by(sensor, treatment) %>%
        summarize(mean_temp = mean(temp, na.rm = T))  %>%
        group_by(treatment) %>%
        summarize(mean = mean(mean_temp, na.rm = T),
                  sd = sd(mean_temp, na.rm = T))
#UMBS_avg_night2 <- UMBS_avg_night %>%
#        select(-sd) %>%
#        pivot_wider(names_from = treatment, values_from = mean) %>%
#        group_by(year) %>%
#        summarize(diff = XH_warmed_air_1m-XH_ambient_air_1m)
print(UMBS_avg_night)
```

## KBS soil temp & moisture
```{r}
# merge the data + filter data for only the daytime during the growing season
KBS_season_soil <- KBS
KBS_season_soil$month <- format(KBS_season_soil$Date_Time,format="%m")
KBS_season_soil$year <- format(KBS_season_soil$Date_Time,format="%y")
KBS_season_soil$hour <- format(KBS_season_soil$Date_Time, format="%H")

KBS_season_soil <- KBS_season_soil %>%
  filter(month > "03") %>%
  filter(month < "09") %>%
  filter(hour > "06") %>%
  filter(hour < "20") %>%
  dplyr::select(Date_Time, year, month, hour, XU_warmed_soil_temp_5cm, XU_ambient_soil_temp_5cm, sensor) %>%
        filter(!is.na(XU_warmed_soil_temp_5cm))

# create new dataframes for temperatures averaged by year & averaged by month and year
KBS_avg_soil <- KBS_season_soil %>%
  gather(key = "treatment", value = "temp", -year, -month, -hour, -Date_Time, -sensor)

soil_t.test <- t.test(temp~treatment,data=KBS_avg_soil)
soil_t.test

# avg soil temps in the chambers during the daytime
KBS_avg_soil_temp <- KBS_avg_soil %>%
        group_by(sensor, treatment) %>%
        summarize(mean_temp = mean(temp, na.rm = T)) %>%
        group_by(treatment) %>%
        summarize(mean = mean(mean_temp, na.rm = T),
                  sd = sd(mean_temp, na.rm = T))
print(KBS_avg_soil_temp)


# merge the data + filter data for only the daytime during the growing season
KBS_season_moist <- KBS
KBS_season_moist$month <- format(KBS_season_moist$Date_Time,format="%m")
KBS_season_moist$year <- format(KBS_season_moist$Date_Time,format="%y")
KBS_season_moist$hour <- format(KBS_season_moist$Date_Time, format="%H")

KBS_season_moist <- KBS_season_moist %>%
  filter(month > "03") %>%
  filter(month < "09") %>%
  filter(hour > "06") %>%
  filter(hour < "20") %>%
  dplyr::select(Date_Time, year, month, hour, XH_warmed_soil_moisture_5cm, XH_ambient_soil_moisture_5cm, sensor) %>%
        filter(!is.na(XH_warmed_soil_moisture_5cm))

# create new dataframes for temperatures averaged by year & averaged by month and year
# note temp here is actually moisture level
KBS_avg_moist <- KBS_season_moist %>%
  gather(key = "treatment", value = "temp", -year, -month, -hour, -Date_Time, -sensor)

moist_t.test <- t.test(temp~treatment,data=KBS_avg_moist)
moist_t.test

# avg soil temps in the chambers during the daytime
KBS_avg_soil_moist <- KBS_avg_moist %>%
        group_by(sensor, treatment) %>%
        summarize(mean_temp = mean(temp, na.rm = T))  %>%
        group_by(treatment) %>%
        summarize(mean = mean(mean_temp, na.rm = T),
                  sd = sd(mean_temp, na.rm = T))
print(KBS_avg_soil_moist)
```

## UMBS soil temp & moisture
```{r}
# merge the data + filter data for only the daytime during the growing season
UMBS_season_soil1 <- UMBS
UMBS_season_soil1$month <- format(UMBS_season_soil1$Date_Time,format="%m")
UMBS_season_soil1$year <- format(UMBS_season_soil1$Date_Time,format="%y")
UMBS_season_soil1$hour <- format(UMBS_season_soil1$Date_Time, format="%H")

UMBS_season_soil <- UMBS_season_soil1 %>%
        filter(month > "03") %>%
        filter(month < "09") %>%
        filter(hour > "06") %>%
        filter(hour < "20") %>%
        select(Date_Time, year, month, hour, XU_warmed_soil_temp_5cm, XU_ambient_soil_temp_5cm, sensor)

# create new dataframes for temperatures averaged by year & averaged by month and year
UMBS_avg_soil <- UMBS_season_soil %>%
        gather(key = "treatment", value = "temp", -year, -month, -hour, -Date_Time, -sensor)

soil_umbs_t.test <- t.test(temp~treatment,data=UMBS_avg_soil)
soil_umbs_t.test

# avg soil temps in the chambers during the daytime
UMBS_avg_soil_temp <- UMBS_avg_soil %>%
        group_by(sensor, treatment) %>%
        summarize(mean_temp = mean(temp, na.rm = T)) %>%
        group_by(treatment) %>%
        summarize(mean = mean(mean_temp, na.rm = T),
                  sd = sd(mean_temp, na.rm = T))
print(UMBS_avg_soil_temp)


# merge the data + filter data for only the daytime during the growing season
UMBS_season_moist <- UMBS
UMBS_season_moist$month <- format(UMBS_season_moist$Date_Time,format="%m")
UMBS_season_moist$year <- format(UMBS_season_moist$Date_Time,format="%y")
UMBS_season_moist$hour <- format(UMBS_season_moist$Date_Time, format="%H")

UMBS_season_moist <- UMBS_season_moist %>%
        filter(month > "03") %>%
        filter(month < "09") %>%
        filter(hour > "06") %>%
        filter(hour < "20") %>%
        select(Date_Time, year, month, hour, XH_warmed_soil_moisture_5cm, XH_ambient_soil_moisture_5cm, sensor)

# create new dataframes for temperatures averaged by year & averaged by month and year
UMBS_avg_moist <- UMBS_season_moist %>%
        gather(key = "treatment", value = "temp", -year, -month, -hour, -Date_Time, -sensor)

moist_umbs_t.test <- t.test(temp~treatment,data=UMBS_avg_moist)
moist_umbs_t.test

# avg soil temps in the chambers during the daytime
UMBS_avg_soil_moist <- UMBS_avg_moist %>%
        group_by(sensor, treatment) %>%
        summarize(mean_temp = mean(temp, na.rm = T)) %>%
        group_by(treatment) %>%
        summarize(mean = mean(mean_temp, na.rm = T),
                  sd = sd(mean_temp, na.rm = T))
print(UMBS_avg_soil_moist)
```