---
title: "warmXtrophic Project: Plant Composition Diversity Data Analyses"
author: "Moriah Young"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=80),tidy=TRUE)
```

# Load in packages & data
```{r, message = F}
# Clear all existing data
rm(list=ls())

#Load packages
library(tidyverse)
library(ggplot2)
library(lme4)
library(olsrr)
library(predictmeans)
library(car)
library(fitdistrplus)
library(ggpubr)
library(rstatix)
library(vegan)
library(interactions)
library(emmeans)
library(sjPlot)
library(effects)
library(glmmTMB)
library(labdsv) # used with Vegan package, the matrify() and matrify2() functions
library(agricolae) # HSD.test() function
library(bbmle)
library(jtools) # summ() function
library(knitr)

# Set working directory
Sys.getenv("L1DIR")
L0_dir <- Sys.getenv("L0DIR")
L1_dir <- Sys.getenv("L1DIR")
L2_dir <- Sys.getenv("L2DIR")
list.files(L1_dir)

# read in plant comp data
diver <- read.csv(file.path(L2_dir, "plant_composition/final_plant_comp_diversity_L2.csv"))
diver$year_factor <- as.factor(diver$year_factor)

# Order warmed and ambient so that warmed shows up first in plotting (and is default is red = warmed; blue = ambient). First make it a factor
diver$state <- as.factor(diver$state)
levels(diver$state)
# [1] "ambient" "warmed" 
diver$state <- factor(diver$state, levels(diver$state)[c(2,1)])
levels(diver$state)
# [1] "warmed"  "ambient"

# create separate data frames for kbs and umbs sites
kbs_diversity <- subset(diver, site == "kbs")
umbs_diversity <- subset(diver, site == "umbs")
```

Simpson's Index
KBS
```{r, warning=F}
### KBS ###
hist(kbs_diversity$simpson) # skewed left
qqnorm(kbs_diversity$simpson)
shapiro.test(kbs_diversity$simpson) # pvalue is < 0.05 so we reject the null hypothesis that the data is normal (aka not normally distributed)

# Exploring distributions for these data:
descdist(kbs_diversity$simpson, discrete = FALSE)

# Gamma distribution 
fit.gamma <- fitdist(kbs_diversity$simpson, "gamma")
plot(fit.gamma)

# Weibull distribution
fit.weibull  <- fitdist(kbs_diversity$simpson, "weibull")
plot(fit.weibull)

# Lognormal distribution 
fit.ln <- fitdist(kbs_diversity$simpson, "lnorm")
plot(fit.ln)

# Exponential distribution is another option
fit.exp <- fitdist(kbs_diversity$simpson, "exp")
plot(fit.exp)

par(mfrow=c(2,2))
plot.legend <- c("Gamma", "Weibull", "Log Normal", "Exponential")
denscomp(list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)
cdfcomp (list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)
qqcomp  (list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)
ppcomp  (list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)
dev.off()

# Goodness of fit comparisons across fits
gofstat(list(fit.gamma, fit.weibull, fit.ln, fit.exp), fitnames = c("Gamma", "Weibull", "Log Normal", "Exp"))
# log normal distribution looks to be the best based on AIC and BIC values or would it be gamma? (closest to zero?)
```

UMBS
```{r, warning=F}
### UMBS ###
hist(umbs_diversity$simpson) #skewed to the left
qqnorm(umbs_diversity$simpson)
shapiro.test(umbs_diversity$simpson) # pvalue is < 0.05 so we reject the null hypothesis that the data is normal (aka not normally distributed)

# Exploring distributions for these data:
descdist(umbs_diversity$simpson, discrete = FALSE)

# Gamma distribution 
fit.gamma <- fitdist(umbs_diversity$simpson, "gamma")
plot(fit.gamma)

# Weibull distribution
fit.weibull  <- fitdist(umbs_diversity$simpson, "weibull")
plot(fit.weibull)

# Lognormal distribution 
fit.ln <- fitdist(umbs_diversity$simpson, "lnorm")
plot(fit.ln)

# Exponential distribution is another option
fit.exp <- fitdist(umbs_diversity$simpson, "exp")
plot(fit.exp)

par(mfrow=c(2,2))
plot.legend <- c("Gamma", "Weibull", "Log Normal", "Exponential")
denscomp(list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)
cdfcomp (list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)
qqcomp  (list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)
ppcomp  (list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)
dev.off()

# Goodness of fit comparisons across fits
gofstat(list(fit.gamma, fit.weibull, fit.ln, fit.exp), fitnames = c("Gamma", "Weibull", "Log Normal", "Exp"))
# log normal distribution looks to be the best based on AIC and BIC values
```

Shannon Index
KBS
```{r, warning=F}
### KBS ###
hist(kbs_diversity$shannon) 
qqnorm(kbs_diversity$shannon)
shapiro.test(kbs_diversity$shannon) # pvalue is > 0.05 - normally distributed
```

UMBS
```{r, warning=F}
### UMBS ###
hist(umbs_diversity$shannon)
qqnorm(umbs_diversity$shannon)
shapiro.test(umbs_diversity$shannon) # pvalue is close to > 0.05 so we do not reject the null hypothesis that the data is normal (aka normally distributed)

# Exploring distributions for these data:
descdist(umbs_diversity$shannon, discrete = FALSE)
```

Species Richness
KBS
```{r, warning=F}
### KBS ###
hist(kbs_diversity$richness) # looks pretty good
qqnorm(kbs_diversity$richness)
shapiro.test(kbs_diversity$richness) # pvalue is < 0.05 so we reject the null hypothesis that the data is normal (aka not normally distributed) but it's pretty close to normal

# Exploring distributions for these data:
descdist(kbs_diversity$richness, discrete = FALSE) # close to normal

# Gamma distribution 
fit.gamma <- fitdist(kbs_diversity$richness, "gamma")
plot(fit.gamma)

# Weibull distribution
fit.weibull  <- fitdist(kbs_diversity$richness, "weibull")
plot(fit.weibull)

# Lognormal distribution 
fit.ln <- fitdist(kbs_diversity$richness, "lnorm")
plot(fit.ln)

# Exponential distribution is another option
fit.exp <- fitdist(kbs_diversity$richness, "exp")
plot(fit.exp)

par(mfrow=c(2,2))
plot.legend <- c("Gamma", "Weibull", "Log Normal", "Exponential")
denscomp(list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)
cdfcomp (list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)
qqcomp  (list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)
ppcomp  (list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)
dev.off()

# Goodness of fit comparisons across fits
gofstat(list(fit.gamma, fit.weibull, fit.ln, fit.exp), fitnames = c("Gamma", "Weibull", "Log Normal", "Exp"))
# weibull distribution looks to be the best based on AIC and BIC values
```

UMBS
```{r, warning=F}
### UMBS ###
hist(umbs_diversity$richness) 
qqnorm(umbs_diversity$richness)
shapiro.test(umbs_diversity$richness) # pvalue is < 0.05 so we reject the null hypothesis that the data is normal (aka not normally distributed)

# Exploring distributions for these data:
descdist(umbs_diversity$richness, discrete = FALSE)

# Gamma distribution 
fit.gamma <- fitdist(umbs_diversity$richness, "gamma")
plot(fit.gamma)

# Weibull distribution
fit.weibull  <- fitdist(umbs_diversity$richness, "weibull")
plot(fit.weibull)

# Lognormal distribution 
fit.ln <- fitdist(umbs_diversity$richness, "lnorm")
plot(fit.ln)

# Exponential distribution is another option
fit.exp <- fitdist(umbs_diversity$richness, "exp")
plot(fit.exp)

par(mfrow=c(2,2))
plot.legend <- c("Gamma", "Weibull", "Log Normal", "Exponential")
denscomp(list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)
cdfcomp (list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)
qqcomp  (list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)
ppcomp  (list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)
dev.off()

# Goodness of fit comparisons across fits
gofstat(list(fit.gamma, fit.weibull, fit.ln, fit.exp), fitnames = c("Gamma", "Weibull", "Log Normal", "Exp"))
# log normal and gamma are essentially tied 
```

# Leverage plots and detecting Outliers. https://www.statmethods.net/stats/rdiagnostics.html
# These illustrate whether certain data points have more leverage (more influence), and thus could be outliers. It's a way of detecting outliers. Leverage plots can help identify whether a point has high or low influence, based on its leverage and residual and determining model fit with and without the point in question. Ultimately you decide whether the points are outliers or not, based on the knowledge of the system and how much it changes the model when included vs. excluded from the data used to fit the model. Here is a good overview of the combination of leverage and residual: scroll down to sections beginning at "13.3 Unusual Observations": https://daviddalpiaz.github.io/appliedstats/model-diagnostics.html

SIMPSON
```{r}
# KBS State-only model
fit_state_kbs <- lm(log(simpson) ~ state, data = kbs_diversity)
outlierTest(fit_state_kbs) # yes row 60 (2017), 145 (2021)
qqPlot(fit_state_kbs, main="QQ Plot") 
hist(fit_state_kbs$residuals)
leveragePlots(fit_state_kbs)
ols_test_normality(fit_state_kbs)

# KBS State and year model
fit_stateyear_kbs <- lm(log(simpson) ~ state + year_factor, data = kbs_diversity)
outlierTest(fit_stateyear_kbs) # yes
qqPlot(fit_stateyear_kbs, main="QQ Plot") 
hist(fit_stateyear_kbs$residuals)
leveragePlots(fit_stateyear_kbs)
ols_test_normality(fit_stateyear_kbs)

# Interaction plot (ignore for now the repeated measures with species); see: https://cran.r-project.org/web/packages/interactions/vignettes/interactions.html and: https://interactions.jacob-long.com/

fit3 <- lm(log(simpson) ~ state*year, data = kbs_diversity)
interact_plot(fit3, pred = year, modx = state)
```

```{r}
# UMBS State-only model
fit_state_umbs <- lm(log(simpson) ~ state, data = umbs_diversity)
outlierTest(fit_state_umbs) # yes, row 186 and 198
qqPlot(fit_state_umbs, main="QQ Plot") 
hist(fit_state_umbs$residuals)
leveragePlots(fit_state_umbs)
ols_test_normality(fit_state_umbs)

# UMBS State and year model
fit_stateyear_umbs <- lm(log(simpson) ~ state + year_factor, data = umbs_diversity)
outlierTest(fit_stateyear_kbs) # rows 60, 145
qqPlot(fit_stateyear_umbs, main="QQ Plot") 
hist(fit_stateyear_umbs$residuals)
leveragePlots(fit_stateyear_umbs)
ols_test_normality(fit_stateyear_umbs)

# Interaction plot (ignore for now the repeated measures with species); see: https://cran.r-project.org/web/packages/interactions/vignettes/interactions.html and: https://interactions.jacob-long.com/

# I can't get these to work
fit3 <- lm(log(simpson) ~ state*year, data = umbs_diversity)
interact_plot(fit3, pred = year, modx = state)
```

SHANNON
```{r}
# KBS State-only model
fit_state_kbs_shannon <- lm(shannon ~ state, data = kbs_diversity)
outlierTest(fit_state_kbs_shannon) # no
qqPlot(fit_state_kbs_shannon, main="QQ Plot") 
hist(fit_state_kbs_shannon$residuals)
leveragePlots(fit_state_kbs_shannon)
ols_test_normality(fit_state_kbs_shannon)

# KBS State and year model
fit_stateyear_kbs_shannon <- lm(shannon ~ state + year_factor, data = kbs_diversity)
outlierTest(fit_stateyear_kbs_shannon) # 140
qqPlot(fit_stateyear_kbs_shannon, main="QQ Plot") 
hist(fit_stateyear_kbs_shannon$residuals)
leveragePlots(fit_stateyear_kbs_shannon)
ols_test_normality(fit_stateyear_kbs_shannon)

# Interaction plot (ignore for now the repeated measures with species); see: https://cran.r-project.org/web/packages/interactions/vignettes/interactions.html and: https://interactions.jacob-long.com/

# I can't get these to work
fit3 <- lm(log(shannon) ~ state*year, data = kbs_diversity)
interact_plot(fit3, pred = year, modx = state)
```

```{r}
# UMBS State-only model
fit_state_umbs_shannon <- lm(shannon ~ state, data = umbs_diversity)
outlierTest(fit_state_umbs_shannon) # no outliers
qqPlot(fit_state_umbs_shannon, main="QQ Plot") 
hist(fit_state_umbs_shannon$residuals)
leveragePlots(fit_state_umbs_shannon)
ols_test_normality(fit_state_umbs_shannon)

# UMBS State and year model
fit_stateyear_umbs_shannon <- lm(shannon ~ state + year_factor, data = umbs_diversity)
outlierTest(fit_stateyear_umbs_shannon) # no outliers
qqPlot(fit_stateyear_umbs_shannon, main="QQ Plot") 
hist(fit_stateyear_umbs_shannon$residuals)
leveragePlots(fit_stateyear_umbs_shannon)
ols_test_normality(fit_stateyear_umbs_shannon)

# Interaction plot (ignore for now the repeated measures with species); see: https://cran.r-project.org/web/packages/interactions/vignettes/interactions.html and: https://interactions.jacob-long.com/

# I can't get these to work
fit3 <- lm(shannon ~ state*year, data = umbs_diversity)
interact_plot(fit3, pred = year, modx = state)
```

RICHNESS
```{r}
# KBS State-only model
fit_state_kbs_richness <- lm(log(richness) ~ state, data = kbs_diversity)
outlierTest(fit_state_kbs_richness) # no
qqPlot(fit_state_kbs_richness, main="QQ Plot") 
hist(fit_state_kbs_richness$residuals)
leveragePlots(fit_state_kbs_richness)
ols_test_normality(fit_state_kbs_richness)

# KBS State and year model
fit_stateyear_kbs_richness <- lm(log(richness) ~ state + year_factor, data = kbs_diversity)
outlierTest(fit_stateyear_kbs_richness) # no outliers
qqPlot(fit_stateyear_kbs_richness, main="QQ Plot") 
hist(fit_stateyear_kbs_richness$residuals)
leveragePlots(fit_stateyear_kbs_richness)
ols_test_normality(fit_stateyear_kbs_richness)

# Interaction plot (ignore for now the repeated measures with species); see: https://cran.r-project.org/web/packages/interactions/vignettes/interactions.html and: https://interactions.jacob-long.com/

fit3 <- lm(log(richness) ~ state*year, data = kbs_diversity)
interact_plot(fit3, pred = year, modx = state)
```

```{r}
# UMBS State-only model
fit_state_umbs_richness <- lm(log(richness) ~ state, data = umbs_diversity)
outlierTest(fit_state_umbs_richness) # no outliers
qqPlot(fit_state_umbs_richness, main="QQ Plot") 
hist(fit_state_umbs_richness$residuals)
leveragePlots(fit_state_umbs_richness)
ols_test_normality(fit_state_umbs_richness)

# UMBS State and year model
fit_stateyear_umbs_richness <- lm(log(richness) ~ state + year_factor, data = umbs_diversity)
outlierTest(fit_stateyear_umbs_richness) # no outliers
qqPlot(fit_stateyear_umbs_richness, main="QQ Plot") 
hist(fit_stateyear_umbs_richness$residuals)
leveragePlots(fit_stateyear_umbs_richness)
ols_test_normality(fit_stateyear_umbs_richness)

# Interaction plot (ignore for now the repeated measures with species); see: https://cran.r-project.org/web/packages/interactions/vignettes/interactions.html and: https://interactions.jacob-long.com/

# I can't get these to work
fit3 <- lm(log(richness) ~ state*year, data = umbs_diversity)
interact_plot(fit3, pred = year, modx = state)
```

MIXED EFFECT MODELS
SIMPSON
KBS 
```{r}
modfull <- lmer(log(simpson) ~ state*year_factor + insecticide*year_factor + (1|plot), kbs_diversity, REML = FALSE)

# Check Assumptions:
# (1) Linearity: if covariates are not categorical (year isn't)
# (2) Homogeneity: Need to Check by plotting residuals vs predicted values.
par(mfrow=c(1,2))
plot(modfull)
# Homogeneity of variance is ok here (increasing variance in resids is not increasing with fitted values)
# Check for homogeneity of variances (true if p>0.05). If the result is not significant, the assumption of equal variances (homoscedasticity) is met (no significant difference between the group variances).
# *****Levene’s Test - tests whether or not the variance among two or more groups is equal - If the p-value is less than our chosen significance level, we can reject the null hypothesis and conclude that we have enough evidence to state that the variance among the groups is not equal (which we want).

leveneTest(residuals(modfull) ~ kbs_diversity$state)
# Assumption not met
leveneTest(residuals(modfull) ~ kbs_diversity$insecticide) 
# Assumption not met 
leveneTest(residuals(modfull) ~ kbs_diversity$plot)
# Assumption not met

# (3) Normality of error term: need to check by histogram, QQplot of residuals, could do Kolmogorov-Smirnov test.
# Check for normal residuals
qqPlot(resid(modfull))
hist(residuals(modfull))
shapiro.test(resid(modfull)) # not normally distributed resids bc p<0.05
outlierTest(modfull) # row 60 and 152

# (4) Normality of random effect: Get the estimate of random effect (e.g., random intercepts), and check them as you would check the residual. 
require(lme4)
r_int<- ranef(modfull)$plot$`(Intercept)`
qqnorm(r_int)
qqline(r_int)
shapiro.test(r_int) 
# Normally distributed random effect pvalue > 0.05
```

```{r}
# testing if year as continuous or year as a factor makes more sense (test models)
year.mod.test1 <- lmer(log(simpson) ~ state + year + (1|plot), kbs_diversity, REML=FALSE)
year.mod.test2 <- lmer(log(simpson) ~ state + year_factor + (1|plot), kbs_diversity, REML=FALSE)
anova(year.mod.test1, year.mod.test2) #year.mod.test2

mod1 <- lmer(log(simpson) ~ state + (1|plot), kbs_diversity, REML=FALSE)
mod2 <- lmer(log(simpson) ~ insecticide + (1|plot), kbs_diversity, REML=FALSE)
mod3 <- lmer(log(simpson) ~ insecticide + state + (1|plot), kbs_diversity, REML=FALSE)
mod4 <- lmer(log(simpson) ~ insecticide * state + (1|plot), kbs_diversity, REML=FALSE)
mod5 <- lmer(log(simpson) ~ state + year_factor + (1|plot), kbs_diversity, REML=FALSE)
mod6 <- lmer(log(simpson) ~ state + year_factor + insecticide + (1|plot), kbs_diversity, REML=FALSE)
mod7 <- lmer(log(simpson) ~ state * year_factor + (1|plot), kbs_diversity, REML=FALSE)
mod8 <- lmer(log(simpson) ~ state * year_factor + insecticide + (1|plot), kbs_diversity, REML=FALSE)
mod9 <- lmer(log(simpson) ~ state * insecticide + year_factor + (1|plot), kbs_diversity, REML=FALSE)
mod10 <- lmer(log(simpson) ~ state + insecticide * year_factor + (1|plot), kbs_diversity, REML=FALSE)
mod11 <- lmer(log(simpson) ~ state * year_factor * insecticide + (1|plot), kbs_diversity, REML=FALSE)
AICctab(mod1, mod2, mod3, mod4, mod5, mod6, mod7, mod8, mod9, mod10, mod11, weights=T) 
anova(mod10, mod5)
AICctab(mod10, mod5, weights=T) #10p
summary(mod10)
summ(mod10)
anova(mod10)
emmeans(mod10, list(pairwise ~ state + year_factor), adjust = "tukey")
emmeans(mod10, list(pairwise ~ insecticide*year_factor), adjust = "tukey")
emmeans(mod10, list(pairwise ~ year_factor), adjust = "tukey")

# with herb reduction as interactive term
anova(mod9)

#Our hypothesized model is  mod9p
summary(mod9)
anova(mod9)

# comparisons
emmeans(mod9, list(pairwise ~ state*insecticide), adjust = "tukey")
mod9.emm <- emmeans(mod9, ~ state*insecticide)
contrast(mod9.emm, "consec", simple = "each", combine = F, adjust = "mvt")
emmip(mod9, insecticide~state)

```

UMBS
```{r}
modfull_u <- lmer(log(simpson) ~ state*year_factor + insecticide*year_factor + (1|plot), umbs_diversity, REML = FALSE)

# Check Assumptions:
# (1) Linearity: if covariates are not categorical (year isn't)
# (2) Homogeneity: Need to Check by plotting residuals vs predicted values.
par(mfrow=c(1,2))
plot(modfull_u )
# Homogeneity of variance is ok here (increasing variance in resids is not increasing with fitted values)
# Check for homogeneity of variances (true if p>0.05). If the result is not significant, the assumption of equal variances (homoscedasticity) is met (no significant difference between the group variances).
# *****Levene’s Test - tests whether or not the variance among two or more groups is equal - If the p-value is less than our chosen significance level, we can reject the null hypothesis and conclude that we have enough evidence to state that the variance among the groups is not equal (which we want).

leveneTest(residuals(modfull_u) ~ umbs_diversity$state)
# Assumption not met
leveneTest(residuals(modfull_u) ~ umbs_diversity$insecticide) 
# Assumption not met 
leveneTest(residuals(modfull_u) ~ umbs_diversity$plot)
# Assumption not met

# (3) Normality of error term: need to check by histogram, QQplot of residuals, could do Kolmogorov-Smirnov test.
# Check for normal residuals
qqPlot(resid(modfull_u))
hist(residuals(modfull_u))
shapiro.test(resid(modfull_u)) # not normally distributed resids bc p<0.05
outlierTest(modfull_u) # yes outliers

# (4) Normality of random effect: Get the estimate of random effect (e.g., random intercepts), and check them as you would check the residual. 
require(lme4)
r_int_u <- ranef(modfull_u )$plot$`(Intercept)`
qqnorm(r_int_u)
qqline(r_int_u)
shapiro.test(r_int_u) 
# Normally distributed random effect pvalue > 0.05
```

UMBS
```{r}
# hypothesized model
mod1u <- lmer(log(simpson) ~ state * insecticide + year_factor + (1|plot), umbs_diversity, REML=FALSE)

summary(mod1u)
anova(mod1u)

# comparisons
emmeans(mod1u, list(pairwise ~ state*insecticide), adjust = "tukey")
mod1u.emm <- emmeans(mod1u, ~ state*insecticide)
contrast(mod1u.emm, "consec", simple = "each", combine = F, adjust = "mvt")
emmip(mod1u, insecticide~state)
```

SHANNON
KBS 
```{r}
modfull_k_sh <- lmer(shannon ~ state*year_factor + insecticide*year_factor + (1|plot), kbs_diversity, REML = FALSE)

# Check Assumptions:
# (1) Linearity: if covariates are not categorical (year isn't)
# (2) Homogeneity: Need to Check by plotting residuals vs predicted values.
par(mfrow=c(1,2))
plot(modfull_k_sh)
# Homogeneity of variance is ok here (increasing variance in resids is not increasing with fitted values)
# Check for homogeneity of variances (true if p>0.05). If the result is not significant, the assumption of equal variances (homoscedasticity) is met (no significant difference between the group variances).
# *****Levene’s Test - tests whether or not the variance among two or more groups is equal - If the p-value is less than our chosen significance level, we can reject the null hypothesis and conclude that we have enough evidence to state that the variance among the groups is not equal (which we want).

leveneTest(residuals(modfull_k_sh) ~ kbs_diversity$state)
# Assumption met
leveneTest(residuals(modfull_k_sh) ~ kbs_diversity$insecticide) 
# Assumption not met 
leveneTest(residuals(modfull_k_sh) ~ kbs_diversity$plot)
# Assumption not met

# (3) Normality of error term: need to check by histogram, QQplot of residuals, could do Kolmogorov-Smirnov test.
# Check for normal residuals
qqPlot(resid(modfull_k_sh))
hist(residuals(modfull_k_sh))
shapiro.test(resid(modfull_k_sh)) # not normally distributed resids bc p<0.05
outlierTest(modfull_k_sh) # no outliers

# (4) Normality of random effect: Get the estimate of random effect (e.g., random intercepts), and check them as you would check the residual. 
require(lme4)
r_int_k_sh <- ranef(modfull_k_sh)$plot$`(Intercept)`
qqnorm(r_int_k_sh)
qqline(r_int_k_sh)
shapiro.test(r_int_k_sh) 
# Normally distributed random effect pvalue > 0.05

# testing if year as continuous or year as a factor makes more sense (test models)
year.mod.testk_sh1 <- lmer(shannon ~ state + year + (1|plot), kbs_diversity, REML=FALSE)
year.mod.testk_sh2 <- lmer(shannon ~ state + year_factor + (1|plot), kbs_diversity, REML=FALSE)
anova(year.mod.testk_sh1, year.mod.testk_sh2) #year.mod.test2

# hypothesized model
mod_k_sh <- lmer(shannon ~ state * insecticide + year_factor + (1|plot), kbs_diversity, REML=FALSE)

summary(mod_k_sh)
anova(mod_k_sh)

# comparisons
emmeans(mod_k_sh, list(pairwise ~ state*insecticide), adjust = "tukey")
mod_k_sh.emm <- emmeans(mod_k_sh, ~ state*insecticide)
contrast(mod_k_sh.emm, "consec", simple = "each", combine = F, adjust = "mvt")
emmip(mod_k_sh, insecticide~state)
```

UMBS
```{r}
modfull_u_sh <- lmer(shannon ~ state*year_factor + insecticide*year_factor + (1|plot), umbs_diversity, REML = FALSE)

# Check Assumptions:
# (1) Linearity: if covariates are not categorical (year isn't)
# (2) Homogeneity: Need to Check by plotting residuals vs predicted values.
par(mfrow=c(1,2))
plot(modfull_u_sh)
# Homogeneity of variance is ok here (increasing variance in resids is not increasing with fitted values)
# Check for homogeneity of variances (true if p>0.05). If the result is not significant, the assumption of equal variances (homoscedasticity) is met (no significant difference between the group variances).
# *****Levene’s Test - tests whether or not the variance among two or more groups is equal - If the p-value is less than our chosen significance level, we can reject the null hypothesis and conclude that we have enough evidence to state that the variance among the groups is not equal (which we want).

leveneTest(residuals(modfull_u_sh) ~ umbs_diversity$state)
# Assumption not met
leveneTest(residuals(modfull_u_sh) ~ umbs_diversity$insecticide) 
# Assumption not met 
leveneTest(residuals(modfull_u_sh) ~ umbs_diversity$plot)
# Assumption not met

# (3) Normality of error term: need to check by histogram, QQplot of residuals, could do Kolmogorov-Smirnov test.
# Check for normal residuals
qqPlot(resid(modfull_u_sh))
hist(residuals(modfull_u_sh))
shapiro.test(resid(modfull_u_sh)) # not normally distributed resids bc p<0.05
outlierTest(modfull_u_sh) # no

# (4) Normality of random effect: Get the estimate of random effect (e.g., random intercepts), and check them as you would check the residual. 
require(lme4)
r_int_u_sh <- ranef(modfull_u_sh)$plot$`(Intercept)`
qqnorm(r_int_u_sh)
qqline(r_int_u_sh)
shapiro.test(r_int_u_sh) 
# Normally distributed random effect pvalue > 0.05

# testing if year as continuous or year as a factor makes more sense (test models)
year.mod.testu_sh1 <- lmer(shannon ~ state + year + (1|plot), umbs_diversity, REML=FALSE)
year.mod.testu_sh2 <- lmer(shannon ~ state + year_factor + (1|plot), umbs_diversity, REML=FALSE)
anova(year.mod.testu_sh1, year.mod.testu_sh2) #year.mod.test2

# hypothesized model
mod_u_sh <- lmer(shannon ~ state * insecticide + year_factor + (1|plot), umbs_diversity, REML=FALSE)

summary(mod_u_sh)
anova(mod_u_sh)

# comparisons
emmeans(mod_u_sh, list(pairwise ~ state*insecticide), adjust = "tukey")
mod_u_sh.emm <- emmeans(mod_u_sh, ~ state*insecticide)
contrast(mod_u_sh.emm, "consec", simple = "each", combine = F, adjust = "mvt")
emmip(mod_u_sh, insecticide~state)
```

RICHNESS
KBS 
```{r}
modfull_kr <- lmer(log(richness) ~ state*year_factor + insecticide*year_factor + (1|plot), kbs_diversity, REML = FALSE)

# Check Assumptions:
# (1) Linearity: if covariates are not categorical (year isn't)
# (2) Homogeneity: Need to Check by plotting residuals vs predicted values.
par(mfrow=c(1,2))
plot(modfull_kr)
# Homogeneity of variance is ok here (increasing variance in resids is not increasing with fitted values)
# Check for homogeneity of variances (true if p>0.05). If the result is not significant, the assumption of equal variances (homoscedasticity) is met (no significant difference between the group variances).
# *****Levene’s Test - tests whether or not the variance among two or more groups is equal - If the p-value is less than our chosen significance level, we can reject the null hypothesis and conclude that we have enough evidence to state that the variance among the groups is not equal (which we want).

leveneTest(residuals(modfull_kr) ~ kbs_diversity$state)
# Assumption not met
leveneTest(residuals(modfull_kr) ~ kbs_diversity$insecticide) 
# Assumption not met 
leveneTest(residuals(modfull_kr) ~ kbs_diversity$plot)
# Assumption not met

# (3) Normality of error term: need to check by histogram, QQplot of residuals, could do Kolmogorov-Smirnov test.
# Check for normal residuals
qqPlot(resid(modfull_kr))
hist(residuals(modfull_kr))
shapiro.test(resid(modfull_kr)) # normally distributed resids bc p>0.05
outlierTest(modfull_kr) # no

# (4) Normality of random effect: Get the estimate of random effect (e.g., random intercepts), and check them as you would check the residual. 
require(lme4)
r_int_kr <- ranef(modfull_kr)$plot$`(Intercept)`
qqnorm(r_int_kr)
qqline(r_int_kr)
shapiro.test(r_int_kr) 
# Normally distributed random effect pvalue > 0.05

# testing if year as continuous or year as a factor makes more sense (test models)
year.mod.testkr1 <- lmer(log(richness) ~ state + year + (1|plot), kbs_diversity, REML=FALSE)
year.mod.testkr2 <- lmer(log(richness) ~ state + year_factor + (1|plot), kbs_diversity, REML=FALSE)
anova(year.mod.testkr1, year.mod.testkr2) #year.mod.test2

# hypothesized model
mod_kr <- lmer(log(richness) ~ state * insecticide + year_factor + (1|plot), kbs_diversity, REML=FALSE)

summary(mod_kr)
anova(mod_kr)

# comparisons
emmeans(mod_kr, list(pairwise ~ state*insecticide), adjust = "tukey")
mod_kr.emm <- emmeans(mod_kr, ~ state*insecticide)
contrast(mod_kr.emm, "consec", simple = "each", combine = F, adjust = "mvt")
emmip(mod_kr, insecticide~state)

# calculating effect size - accounting for log transformation
# overall
exp((1.96971) +  (0.23710)*0) # 7.168597 
exp((1.96971) +  (0.23710)*1) # 9.086684
# effect:
9.086684 - 7.168597 # effect of warming - warmed plots had 1.918087 less species than ambient plots

# herbivores present
exp((1.96971) +  (0.237)*0) # 7.168597 
exp((1.96971) +  (0.237)*1) # 9.085775
# effect:
9.085775 - 7.168597 # effect of warming - warmed plots had 1.917178 less species than ambient plots

# reduced herbivory
exp((1.96971) +  (0.207)*0) # 7.168597
exp((1.96971) +  (0.207)*1) # 8.81725
# effect:
8.81725 - 7.168597 # effect of warming - warmed plots had 1.648653 less species than ambient plots
```

UMBS 
```{r}
modfull_ur <- lmer(log(richness) ~ state*year_factor + insecticide*year_factor + (1|plot), umbs_diversity, REML = FALSE)

# Check Assumptions:
# (1) Linearity: if covariates are not categorical (year isn't)
# (2) Homogeneity: Need to Check by plotting residuals vs predicted values.
par(mfrow=c(1,2))
plot(modfull_ur)
# Homogeneity of variance is ok here (increasing variance in resids is not increasing with fitted values)
# Check for homogeneity of variances (true if p>0.05). If the result is not significant, the assumption of equal variances (homoscedasticity) is met (no significant difference between the group variances).
# *****Levene’s Test - tests whether or not the variance among two or more groups is equal - If the p-value is less than our chosen significance level, we can reject the null hypothesis and conclude that we have enough evidence to state that the variance among the groups is not equal (which we want).

leveneTest(residuals(modfull_ur) ~ umbs_diversity$state)
# Assumption not met
leveneTest(residuals(modfull_ur) ~ umbs_diversity$insecticide) 
# Assumption not met 
leveneTest(residuals(modfull_ur) ~ umbs_diversity$plot)
# Assumption not met

# (3) Normality of error term: need to check by histogram, QQplot of residuals, could do Kolmogorov-Smirnov test.
# Check for normal residuals
qqPlot(resid(modfull_ur))
hist(residuals(modfull_ur))
shapiro.test(resid(modfull_ur)) # not normally distributed resids bc p<0.05
outlierTest(modfull_ur) # no

# (4) Normality of random effect: Get the estimate of random effect (e.g., random intercepts), and check them as you would check the residual. 
require(lme4)
r_int_ur <- ranef(modfull_ur)$plot$`(Intercept)`
qqnorm(r_int_ur)
qqline(r_int_ur)
shapiro.test(r_int_ur) 
# Normally distributed random effect pvalue > 0.05

# testing if year as continuous or year as a factor makes more sense (test models)
year.mod.testur1 <- lmer(log(richness) ~ state + year + (1|plot), umbs_diversity, REML=FALSE)
year.mod.testur2 <- lmer(log(richness) ~ state + year_factor + (1|plot), umbs_diversity, REML=FALSE)
anova(year.mod.testur1, year.mod.testur2) #year.mod.test2

# hypothesized model
mod_ur <- lmer(log(richness) ~ state * insecticide + year_factor + (1|plot), umbs_diversity, REML=FALSE)

summary(mod_ur)
anova(mod_ur)

# comparisons
emmeans(mod_ur, list(pairwise ~ state*insecticide), adjust = "tukey")
mod_ur.emm <- emmeans(mod_ur, ~ state*insecticide)
contrast(mod_ur.emm, "consec", simple = "each", combine = F, adjust = "mvt")
emmip(mod_ur, insecticide~state)

```

