---
title: "Plant Comp Plots"
author: "Moriah Young"
date: "December 18th, 2023"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

COLLABORATORS: Kara Dobson, Phoebe Zarnetske, Mark Hammond, Pat Bills 
DATA INPUT: Clean plot plant comp csv and plant comp diversity csv files from the shared Google drive  
DATA OUTPUT: Code and Rmd are in the scripts folder in Github  
PROJECT: warmXtrophic 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=F}
# Clear all existing data
rm(list=ls())

# Load packages
library(tidyverse)
library(plotrix)
library(ggpubr)

# Set working directory
Sys.getenv("L0DIR")
L0_dir <- Sys.getenv("L0DIR")
L1_dir <- Sys.getenv("L1DIR")
L2_dir <- Sys.getenv("L2DIR")

# Read in plant comp data and meta-data - plot level
comp <- read.csv(file.path(L2_dir, "plant_composition/final_plant_comp_plot_L2.csv"))
diversity <- read.csv(file.path(L2_dir, "plant_composition/final_plant_comp_diversity_L2.csv"))

# making site names capital for cleaner plots - from Kara
change_site <- function(df){
  df$site[df$site == "umbs"] <- "UMBS"
  df$site[df$site == "kbs"] <- "KBS"
  return(df)
}

diversity <- change_site(diversity) # using function from above
comp <- change_site(comp)

# creating a vector with cleaned up insecticide labels for plotting - from Kara
insect_labels <- c("insects" = "Herbivory", "no_insects" = "Reduced Herbivory")

```

Averages + std error for plot level data + insecticide treatment across all years (what is used in the paper)
```{r}
sum_comp_plot_i <- comp %>%
        group_by(site, state, insecticide) %>%
        summarize(avg_pc_plot = mean(avg_pc, na.rm = TRUE),
                  se = std.error(avg_pc, na.rm = TRUE))

sum_richness_all <- diversity %>%
  group_by(site, state, insecticide) %>%
  summarize(avg_richness = mean(richness, na.rm = TRUE),
            se = std.error(richness, na.rm = TRUE))

```


Plant Comp and Species Richness combined dot plot
```{r}
# plant composition
sum_comp_plot_i$full_treat <- paste(sum_comp_plot_i$state, sum_comp_plot_i$insecticide, sep="_")
comp_dot_i <- function(loc) { 
       comp_plot <- subset(sum_comp_plot_i, site == loc)
        return(ggplot(comp_plot, aes(x = state, y = avg_pc_plot, fill=insecticide)) +
                       #geom_point(stat = "identity",position=position_dodge(0.2),size=5) +
                       geom_pointrange(aes(ymin=avg_pc_plot-se, ymax=avg_pc_plot+se), pch=21,size=1,position=position_dodge(0.3)) +
                       #scale_shape_manual(name="Treatment",
                       #                   values = c(1, 19),
                       #                   labels=c("Herbivory","Reduced Herbivory")) +
                       scale_fill_manual(name="Treatment",
                                         values = c("#FFB451", "#0b0055"),
                                         labels=c("Herbivory","Reduced Herbivory")) +
                               labs(x = NULL, y = NULL, title = loc) +
                       scale_x_discrete(labels=c("ambient" = "Ambient", "warmed" = "Warmed")) +
                       coord_cartesian(ylim=c(14,24)) +
                       theme_bw(14))
}
comp_dot_i_kbs <- comp_dot_i("KBS")
comp_dot_i_kbs <- comp_dot_i_kbs + labs(y = "Cover (%)") +
        annotate("text", x = 0.5, y=24, label = "A", size=5) +
        theme(plot.title = element_text(size=20),
                                     axis.text.y = element_text(size=17),
                                     axis.text.x = element_blank(),
                                     axis.title.y = element_text(size=17),
                                     legend.text = element_text(size=17),
                                     legend.title = element_text(size=17))
comp_dot_i_umbs <- comp_dot_i("UMBS")
comp_dot_i_umbs <- comp_dot_i_umbs + labs(y=NULL) + 
        annotate("text", x = 0.5, y=24, label = "B", size=5) + 
        theme(axis.title.y=element_blank(),
              plot.title = element_text(size=20),
              axis.text.y = element_blank(),
              axis.text.x = element_blank(),
              legend.text = element_text(size=17),
              legend.title = element_text(size=17))

# Species Richness
sum_richness_all$full_treat <- paste(sum_richness_all$state, sum_richness_all$insecticide, sep="_")
richness_dot_i <- function(loc) { 
       richness_plot <- subset(sum_richness_all, site == loc)
        return(ggplot(richness_plot, aes(x = state, y = avg_richness, fill=insecticide)) +
                       #geom_point(stat = "identity",position=position_dodge(0.2),size=5) +
                       geom_pointrange(aes(ymin=avg_richness-se, ymax=avg_richness+se), pch=21,size=1,position=position_dodge(0.3)) +
                       #scale_shape_manual(name="Treatment",
                       #                   values = c(1, 19),
                       #                   labels=c("Herbivory","Reduced Herbivory")) +
                       scale_fill_manual(name="Treatment",
                                         values = c("#FFB451", "#0b0055"),
                                         labels=c("Herbivory","Reduced Herbivory")) +
                               labs(x = NULL, y = NULL, title = NULL) +
                       scale_x_discrete(labels=c("ambient" = "Ambient", "warmed" = "Warmed")) +
                       coord_cartesian(ylim=c(4.0,8.5)) +
                       theme_bw(14))
}
richness_dot_i_kbs <- richness_dot_i("KBS")
richness_dot_i_kbs <- richness_dot_i_kbs + labs(y = "Species Richness") +
        annotate("text", x = 0.5, y=8.5, label = "C", size=5) +
        theme(plot.title = element_text(size=20),
                                     axis.text.y = element_text(size=17),
                                     axis.text.x = element_text(size=17),
                                     axis.title.y = element_text(size=17),
                                     legend.text = element_text(size=17),
                                     legend.title = element_text(size=17)) +
        scale_y_continuous(breaks = c(4, 5, 6, 7, 8), 
                           labels = c("  4", "  5", "  6", "  7", "  8"))

richness_dot_i_umbs <- richness_dot_i("UMBS")
richness_dot_i_umbs <- richness_dot_i_umbs + labs(y=NULL) + 
        annotate("text", x = 0.5, y=8.5, label = "D", size=5) +
        theme(axis.title.y=element_blank(),
              axis.text.y = element_blank(),
              axis.text.x = element_text(size=17),
              legend.text = element_text(size=17),
              legend.title = element_text(size=17)) +
        scale_y_continuous(breaks = c(4, 5, 6, 7, 8), 
                           labels = c("4", "5", "6", "7", "8"))

pc_richness_all_dot_i <- ggpubr::ggarrange(comp_dot_i_kbs, comp_dot_i_umbs, 
                                        richness_dot_i_kbs, richness_dot_i_umbs,
                                nrow = 2, ncol = 2, common.legend = T, legend="right",
                                widths = c(1, 0.85))

png("plant_comp_and_richness_L2_all_dot_insect.png", units="in", width=10, height=8, res=300)
pc_richness_all_dot_i
dev.off()

```
