---
title: "warmXtrophic Project: Plant Composition Data Analyses"
author: "Kara Dobson, Moriah Young, Pat Bills"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=80),tidy=TRUE)
```

# Load in packages & data
```{r, message = F}
# Clear all existing data
rm(list=ls())

#Load packages
library(tidyverse)
library(ggplot2)
library(lme4)
library(lmerTest)
library(olsrr)
library(predictmeans)
library(car)
library(fitdistrplus)
library(ggpubr)
library(rstatix)
library(vegan)
library(interactions)
library(sjPlot)
library(effects)
library(glmmTMB)
library(labdsv) # used with Vegan package, the matrify() and matrify2() functions
library(agricolae) # HSD.test() function
library(emmeans)
library(scales)
library(fitdistrplus)
library(moments)# for calculating skewness of data
library(jtools) # summ() function
library(GGally) # ggpairs() function
library(bbmle) # AICtab() function

# Set working directory
Sys.getenv("L1DIR")
L0_dir <- Sys.getenv("L0DIR")
L1_dir <- Sys.getenv("L1DIR")
L2_dir <- Sys.getenv("L2DIR")
list.files(L1_dir)

# Set ggplot2 plotting
# This code for ggplot2 sets the theme to mostly black and white 
# (Arial font, and large font, base size=24)
theme_set(theme_bw(14))
theme_update(axis.text.x = element_text(size = 12, angle = 90),
             axis.text.y = element_text(size = 12))

# read in plant comp data
comp_species <- read.csv(file.path(L2_dir, "plant_composition/final_plant_comp_species_L2.csv"))
comp_plot <- read.csv(file.path(L2_dir, "plant_composition/final_plant_comp_plot_L2.csv"))
comp_plot_origin <- read.csv(file.path(L2_dir, "plant_composition/final_plant_comp_plot_origin_L2.csv")) # plot level data for origin
comp_plot_growthhabit <- read.csv(file.path(L2_dir, "plant_composition/final_plant_comp_plot_growthhabit_L2.csv")) # plot level data for growth habit

comp_species <- comp_species %>% select(-X) # get rid of "X" column that shows up
comp_plot <- comp_plot %>% select(-X) # get rid of "X" column that shows up
comp_plot_origin <- comp_plot_origin %>% select(-X)
comp_plot_growthhabit <- comp_plot_growthhabit %>% select(-X)

# Order warmed and ambient so that warmed shows up first in plotting (and is default is red = warmed; blue = ambient). First make it a factor
comp_species$state <- as.factor(comp_species$state)
levels(comp_species$state)
# [1] "ambient" "warmed" 
comp_species$state <- factor(comp_species$state, levels(comp_species$state)[c(2,1)])
levels(comp_species$state)
# [1] "warmed"  "ambient"

comp_plot$state <- as.factor(comp_plot$state)
levels(comp_plot$state)
# [1] "ambient" "warmed" 
comp_plot$state <- factor(comp_plot$state, levels(comp_plot$state)[c(2,1)])
levels(comp_plot$state)
# [1] "warmed"  "ambient"

# create dataframes for kbs and umbs - remember that these contain species within plots
comp_kbs_spp <- subset(comp_species, site == "kbs")
comp_umbs_spp <- subset(comp_species, site == "umbs")
comp_umbs_spp <- subset(comp_umbs_spp, year_factor!="7") # delete 2021 data from umbs dataframe (doesn't make sense to have bc data was not collect at same frequency as previous years)


# create dataframes for kbs and umbs
comp_kbs_plot <- subset(comp_plot, site == "kbs")
comp_umbs_plot <- subset(comp_plot, site == "umbs")
comp_umbs_plot <- subset(comp_umbs_plot, year_factor!="7") # delete 2021 data from umbs dataframe (doesn't make sense to have bc data was not collect at same frequency as previous years)

kbs_comp_plot_origin <- subset(comp_plot_origin , site == "kbs")
kbs_comp_plot_growthhabit <- subset(comp_plot_growthhabit, site == "kbs")
umbs_comp_plot_origin <- subset(comp_plot_origin, site == "umbs")
umbs_comp_plot_origin <- subset(umbs_comp_plot_origin, year_factor!="7") # delete 2021 data
umbs_comp_plot_growthhabit <- subset(comp_plot_growthhabit, site == "umbs")
umbs_comp_plot_growthhabit <- subset(umbs_comp_plot_growthhabit, year_factor!="7") # delete 2021 data

```

# Data exploration - some different ways of visualizing these data
# KBS SPECIES LEVEL - Looking at RELATIVE ABUNDANCE
```{r, warning=F}
### KBS ###
hist(comp_kbs_spp$relabun) # very skewed to the right
qqnorm(comp_kbs_spp$relabun)
shapiro.test(comp_kbs_spp$relabun) # pvalue is < 0.05 so we reject the null hypothesis that the data is normal (aka not normally distributed)

# Visualizing relative abundance for kbs at the PLOT LEVEL
ggplot(comp_kbs_spp, aes(relabun, fill = plot)) + 
        geom_histogram(binwidth = 0.5) + 
        facet_grid(year_factor ~ site, margins = TRUE, scales = "free")

ggplot(comp_kbs_spp, aes(relabun, fill = as.factor(plot))) + geom_histogram(binwidth = 0.5) + 
        facet_grid(state~year_factor, margins = TRUE, scales = "free")

ggplot(comp_kbs_spp, aes(relabun, fill = plot, color=plot)) +
        geom_density(alpha = 0.1)

ggplot(comp_kbs_spp, aes(relabun, fill = plot, color=plot)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year_factor)

ggplot(comp_kbs_spp, aes(relabun, fill = plot, color=plot)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year_factor + plot)

# Exploring distributions for these right-skewed data:
descdist(comp_kbs_spp$relabun, discrete = FALSE)

# Gamma distribution 
fit.gamma <- fitdist(comp_kbs_spp$relabun, "gamma")
plot(fit.gamma)

# Weibull distribution
fit.weibull  <- fitdist(comp_kbs_spp$relabun, "weibull")
plot(fit.weibull)

# Beta distribution
fit.beta <- fitdist(comp_kbs_spp$relabun, "beta")
plot(fit.beta)

# Lognormal distribution 
fit.ln <- fitdist(comp_kbs_spp$relabun, "lnorm")
plot(fit.ln)

# Exponential distribution is another option
fit.exp <- fitdist(comp_kbs_spp$relabun, "exp")
plot(fit.exp)

par(mfrow=c(2,2))
plot.legend <- c("Gamma", "Weibull", "Log Normal", "Exponential", "Beta")
denscomp(list(fit.gamma, fit.weibull, fit.ln, fit.exp, fit.beta), legendtext = plot.legend)
cdfcomp (list(fit.gamma, fit.weibull, fit.ln, fit.exp, fit.beta), legendtext = plot.legend)
qqcomp  (list(fit.gamma, fit.weibull, fit.ln, fit.exp, fit.beta), legendtext = plot.legend)
ppcomp  (list(fit.gamma, fit.weibull, fit.ln, fit.exp, fit.beta), legendtext = plot.legend)

# Goodness of fit comparisons across fits (can't include the sqrt normal bc it becomes diff response values)
gofstat(list(fit.gamma, fit.weibull, fit.ln, fit.exp, fit.beta), fitnames = c("Gamma", "Weibull", "Log Normal", "Exp", "Beta"))
# Beta and exp distributions look to be the best based on AIC and BIC values
```

# KBS PLOT LEVEL - Looking at PLOT AVG TOTALS 
```{r, warning=F}
### KBS ###
hist(comp_kbs_plot$plot_cover_avg) # skewed to the right
qqnorm(comp_kbs_plot$plot_cover_avg)
shapiro.test(comp_kbs_plot$plot_cover_avg) # pvalue is < 0.05 so we reject the null hypothesis that the data is normal (aka not normally distributed)

# Visualizing plot average totals for kbs at the PLOT LEVEL
ggplot(comp_kbs_plot, aes(plot_cover_avg, fill = plot)) + 
        geom_histogram(binwidth = 0.5) + 
        facet_grid(year_factor ~ site, margins = TRUE, scales = "free")

ggplot(comp_kbs_plot, aes(plot_cover_avg, fill = as.factor(plot))) + geom_histogram(binwidth = 0.5) + 
        facet_grid(state~year_factor, margins = TRUE, scales = "free")

ggplot(comp_kbs_plot, aes(plot_cover_avg, fill = plot, color=plot)) +
        geom_density(alpha = 0.1)

ggplot(comp_kbs_plot, aes(plot_cover_avg, fill = plot, color=plot)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year_factor)

ggplot(comp_kbs_plot, aes(plot_cover_avg, fill = plot, color=plot)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year_factor + plot)

# Exploring distributions for these right-skewed data:
descdist(comp_kbs_plot$plot_cover_avg, discrete = FALSE)

# Gamma distribution 
fit.gamma <- fitdist(comp_kbs_plot$plot_cover_avg, "gamma")
plot(fit.gamma)

# Weibull distribution
fit.weibull  <- fitdist(comp_kbs_plot$plot_cover_avg, "weibull")
plot(fit.weibull)

# Lognormal distribution 
fit.ln <- fitdist(comp_kbs_plot$plot_cover_avg, "lnorm")
plot(fit.ln)

# Exponential distribution is another option
fit.exp <- fitdist(comp_kbs_plot$plot_cover_avg, "exp")
plot(fit.exp)

par(mfrow=c(2,2))
plot.legend <- c("Gamma", "Weibull", "Log Normal", "Exponential")
denscomp(list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)
cdfcomp (list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)
qqcomp  (list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)
ppcomp  (list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)

# Goodness of fit comparisons across fits
gofstat(list(fit.gamma, fit.weibull, fit.ln, fit.exp), fitnames = c("Gamma", "Weibull", "Log Normal", "Exp"))
# log normal distribution looks to be the best based on AIC and BIC values
```
# KBS PLOT LEVEL ORIGIN - Looking at PLOT AVG TOTALS
```{r, warning=F, error=TRUE}
### KBS ###
hist(kbs_comp_plot_origin$plot_cover_avg) # skewed to the right
qqnorm(kbs_comp_plot_origin$plot_cover_avg)
shapiro.test(kbs_comp_plot_origin$plot_cover_avg) # pvalue is < 0.05 so we reject the null hypothesis that the data is normal (aka not normally distributed) 

# Visualizing plot average totals for kbs at the PLOT LEVEL
ggplot(kbs_comp_plot_origin, aes(plot_cover_avg, fill = plot)) + 
        geom_histogram(binwidth = 0.5) + 
        facet_grid(year_factor ~ site, margins = TRUE, scales = "free")

ggplot(kbs_comp_plot_origin, aes(plot_cover_avg, fill = as.factor(plot))) + geom_histogram(binwidth = 0.5) + 
        facet_grid(state~year_factor, margins = TRUE, scales = "free")

ggplot(kbs_comp_plot_origin, aes(plot_cover_avg, fill = plot, color=plot)) +
        geom_density(alpha = 0.1)

ggplot(kbs_comp_plot_origin, aes(plot_cover_avg, fill = plot, color=plot)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year_factor)

ggplot(kbs_comp_plot_origin, aes(plot_cover_avg, fill = plot, color=plot)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year_factor + plot)

# Exploring distributions for these data:
descdist(kbs_comp_plot_origin$plot_cover_avg, discrete = FALSE)

# Gamma distribution 
fit.gamma <- fitdist(kbs_comp_plot_origin$plot_cover_avg, "gamma")
plot(fit.gamma)

# Weibull distribution
fit.weibull  <- fitdist(kbs_comp_plot_origin$plot_cover_avg, "weibull")
plot(fit.weibull)

# Lognormal distribution 
fit.ln <- fitdist(kbs_comp_plot_origin$plot_cover_avg, "lnorm")
plot(fit.ln)

par(mfrow=c(2,2))
plot.legend <- c("Gamma", "Weibull", "Log Normal")
denscomp(list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
cdfcomp (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
qqcomp  (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
ppcomp  (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)

# Goodness of fit comparisons across fits
gofstat(list(fit.gamma, fit.weibull, fit.ln), fitnames = c("Gamma", "Weibull", "Log Normal"))
# log normal looks like it's the best fit
```

# KBS PLOT LEVEL GROWTH HABIT - Looking at PLOT AVG TOTALS
```{r, warning=F, error=TRUE}
### KBS ###
hist(kbs_comp_plot_growthhabit$plot_cover_avg) # skewed to the right
qqnorm(kbs_comp_plot_growthhabit$plot_cover_avg)
shapiro.test(kbs_comp_plot_growthhabit$plot_cover_avg) # pvalue is < 0.05 so we reject the null hypothesis that the data is normal (aka not normally distributed) 

# Visualizing plot average totals for kbs at the PLOT LEVEL
ggplot(kbs_comp_plot_growthhabit, aes(plot_cover_avg, fill = plot)) + 
        geom_histogram(binwidth = 0.5) + 
        facet_grid(year_factor ~ site, margins = TRUE, scales = "free")

ggplot(kbs_comp_plot_origin, aes(plot_cover_avg, fill = as.factor(plot))) + geom_histogram(binwidth = 0.5) + 
        facet_grid(state~year_factor, margins = TRUE, scales = "free")

ggplot(kbs_comp_plot_origin, aes(plot_cover_avg, fill = plot, color=plot)) +
        geom_density(alpha = 0.1)

ggplot(kbs_comp_plot_origin, aes(plot_cover_avg, fill = plot, color=plot)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year_factor)

ggplot(kbs_comp_plot_origin, aes(plot_cover_avg, fill = plot, color=plot)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year_factor + plot)

# Exploring distributions for these data:
descdist(kbs_comp_plot_origin$plot_cover_avg, discrete = FALSE)

# Gamma distribution 
fit.gamma <- fitdist(kbs_comp_plot_origin$plot_cover_avg, "gamma")
plot(fit.gamma)
#
# Weibull distribution
fit.weibull  <- fitdist(kbs_comp_plot_origin$plot_cover_avg, "weibull")
plot(fit.weibull)

# Lognormal distribution 
fit.ln <- fitdist(kbs_comp_plot_origin$plot_cover_avg, "lnorm")
plot(fit.ln)

par(mfrow=c(2,2))
plot.legend <- c("Gamma", "Weibull", "Log Normal")
denscomp(list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
cdfcomp (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
qqcomp  (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
ppcomp  (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)

# Goodness of fit comparisons across fits (can't include the sqrt normal bc it becomes diff response values)
gofstat(list(fit.gamma, fit.weibull, fit.ln), fitnames = c("Gamma", "Weibull", "Log Normal"))
# log normal looks like it's the best fit
```

# UMBS SPECIES LEVEL - Looking at RELATIVE ABUNDANCE
```{r, warning=F}
### UMBS ###
hist(comp_umbs_spp$relabun) # skewed to the right
qqnorm(comp_umbs_spp$relabun)
shapiro.test(comp_umbs_spp$relabun) # pvalue is < 0.05 so we reject the null hypothesis that the data is normal (aka not normally distributed)

# Visualizing median Julian date for umbs at the PLOT LEVEL
ggplot(comp_umbs_spp, aes(relabun, fill = plot)) + 
        geom_histogram(binwidth = 0.5) + 
        facet_grid(year_factor ~ site, margins = TRUE, scales = "free")

ggplot(comp_umbs_spp, aes(relabun, fill = as.factor(plot))) + geom_histogram(binwidth = 0.5) + 
        facet_grid(state~year_factor, margins = TRUE, scales = "free")

ggplot(comp_umbs_spp, aes(relabun, fill = plot, color=plot)) +
        geom_density(alpha = 0.1)

ggplot(comp_umbs_spp, aes(relabun, fill = plot, color=plot)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year_factor)

ggplot(comp_umbs_spp, aes(relabun, fill = plot, color=plot)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year_factor + plot)

# Exploring distributions for these right-skewed data:
descdist(comp_umbs_spp$relabun, discrete = FALSE)

# Gamma distribution 
fit.gamma <- fitdist(comp_umbs_spp$relabun, "gamma")
plot(fit.gamma)

# Weibull distribution
fit.weibull  <- fitdist(comp_umbs_spp$relabun, "weibull")
plot(fit.weibull)

# Beta
fit.beta <- fitdist(comp_umbs_spp$relabun, "beta")
plot(fit.beta)

# Lognormal distribution 
fit.ln <- fitdist(comp_umbs_spp$relabun, "lnorm")
plot(fit.ln)

# Exponential distribution is another option
fit.exp <- fitdist(comp_umbs_spp$relabun, "exp")
plot(fit.exp)

par(mfrow=c(2,2))
plot.legend <- c("Gamma", "Weibull", "Log Normal", "Exponential", "Beta")
denscomp(list(fit.gamma, fit.weibull, fit.ln, fit.exp, fit.beta), legendtext = plot.legend)
cdfcomp (list(fit.gamma, fit.weibull, fit.ln, fit.exp, fit.beta), legendtext = plot.legend)
qqcomp  (list(fit.gamma, fit.weibull, fit.ln, fit.exp, fit.beta), legendtext = plot.legend)
ppcomp  (list(fit.gamma, fit.weibull, fit.ln, fit.exp, fit.beta), legendtext = plot.legend)

# Goodness of fit comparisons across fits (can't include the sqrt normal bc it becomes diff response values)
gofstat(list(fit.gamma, fit.weibull, fit.ln, fit.exp, fit.beta), fitnames = c("Gamma", "Weibull", "Log Normal", "Exp", "Beta"))
# Beta and exp distributions look to be the best based on AIC and BIC values
```

# UMBS PLOT LEVEL - Looking at PLOT AVG TOTALS 
```{r, warning=F}
### UMBS ###
hist(comp_umbs_plot$plot_cover_avg) # skewed to the right
qqnorm(comp_umbs_plot$plot_cover_avg)
shapiro.test(comp_umbs_plot$plot_cover_avg) # pvalue is < 0.05 so we reject the null hypothesis that the data is normal (aka not normally distributed)

# Visualizing median Julian date for umbs at the PLOT LEVEL
ggplot(comp_umbs_plot, aes(plot_cover_avg, fill = plot)) + 
        geom_histogram(binwidth = 0.5) + 
        facet_grid(year_factor ~ site, margins = TRUE, scales = "free")

ggplot(comp_umbs_plot, aes(plot_cover_avg, fill = as.factor(plot))) + geom_histogram(binwidth = 0.5) + 
        facet_grid(state~year_factor, margins = TRUE, scales = "free")

ggplot(comp_umbs_plot, aes(plot_cover_avg, fill = plot, color=plot)) +
        geom_density(alpha = 0.1)


ggplot(comp_umbs_plot, aes(plot_cover_avg, fill = plot, color=plot)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year_factor)

ggplot(comp_umbs_plot, aes(plot_cover_avg, fill = plot, color=plot)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year_factor + plot)

# Exploring distributions for these right-skewed data:
descdist(comp_umbs_plot$plot_cover_avg, discrete = FALSE)

# Gamma distribution 
fit.gamma <- fitdist(comp_umbs_plot$plot_cover_avg, "gamma")
plot(fit.gamma)

# Weibull distribution
fit.weibull  <- fitdist(comp_umbs_plot$plot_cover_avg, "weibull")
plot(fit.weibull)

# Lognormal distribution 
fit.ln <- fitdist(comp_umbs_plot$plot_cover_avg, "lnorm")
plot(fit.ln)

# Exponential distribution is another option
fit.exp <- fitdist(comp_umbs_plot$plot_cover_avg, "exp")
plot(fit.exp)

par(mfrow=c(2,2))
plot.legend <- c("Gamma", "Weibull", "Log Normal", "Exponential")
denscomp(list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)
cdfcomp (list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)
qqcomp  (list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)
ppcomp  (list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)

# Goodness of fit comparisons across fits (can't include the sqrt normal bc it becomes diff response values)
gofstat(list(fit.gamma, fit.weibull, fit.ln, fit.exp), fitnames = c("Gamma", "Weibull", "Log Normal", "Exp"))
# log normal distribution looks to be the best based on AIC and BIC values
```

# UMBS PLOT LEVEL ORIGIN - Looking at PLOT AVG TOTALS
```{r, warning=F, error=TRUE}
### UMBS ###
hist(umbs_comp_plot_origin$plot_cover_avg) # skewed to the right
qqnorm(umbs_comp_plot_origin$plot_cover_avg)
shapiro.test(umbs_comp_plot_origin$plot_cover_avg) # pvalue is < 0.05 so we reject the null hypothesis that the data is normal (aka not normally distributed) 

# Visualizing plot average totals for umbs at the PLOT LEVEL
ggplot(umbs_comp_plot_origin, aes(plot_cover_avg, fill = plot)) + 
        geom_histogram(binwidth = 0.5) + 
        facet_grid(year_factor ~ site, margins = TRUE, scales = "free")

ggplot(umbs_comp_plot_origin, aes(plot_cover_avg, fill = as.factor(plot))) + geom_histogram(binwidth = 0.5) + 
        facet_grid(state~year_factor, margins = TRUE, scales = "free")

ggplot(umbs_comp_plot_origin, aes(plot_cover_avg, fill = plot, color=plot)) +
        geom_density(alpha = 0.1)

ggplot(umbs_comp_plot_origin, aes(plot_cover_avg, fill = plot, color=plot)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year_factor)

ggplot(umbs_comp_plot_origin, aes(plot_cover_avg, fill = plot, color=plot)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year_factor + plot)

# Exploring distributions for these data:
descdist(umbs_comp_plot_origin$plot_cover_avg, discrete = FALSE)

# Gamma distribution 
fit.gamma <- fitdist(umbs_comp_plot_origin$plot_cover_avg, "gamma")
plot(fit.gamma)
#
# Weibull distribution
fit.weibull  <- fitdist(umbs_comp_plot_origin$plot_cover_avg, "weibull")
plot(fit.weibull)

# Lognormal distribution 
fit.ln <- fitdist(umbs_comp_plot_origin$plot_cover_avg, "lnorm")
plot(fit.ln)

par(mfrow=c(2,2))
plot.legend <- c("Gamma", "Weibull", "Log Normal")
denscomp(list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
cdfcomp (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
qqcomp  (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
ppcomp  (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)

# Goodness of fit comparisons across fits
gofstat(list(fit.gamma, fit.weibull, fit.ln), fitnames = c("Gamma", "Weibull", "Log Normal"))
# log normal looks like it's the best fit
```

# UMBS PLOT LEVEL GROWTH HABIT - Looking at PLOT AVG TOTALS
```{r, warning=F, error=TRUE}
### UMBS ###
hist(umbs_comp_plot_growthhabit$plot_cover_avg)
qqnorm(umbs_comp_plot_growthhabit$plot_cover_avg)
shapiro.test(umbs_comp_plot_growthhabit$plot_cover_avg) # pvalue is < 0.05 so we reject the null hypothesis that the data is normal (aka not normally distributed) 

# Visualizing plot average totals for umbs at the PLOT LEVEL
ggplot(umbs_comp_plot_growthhabit, aes(plot_cover_avg, fill = plot)) + 
        geom_histogram(binwidth = 0.5) + 
        facet_grid(year ~ site, margins = TRUE, scales = "free")

ggplot(umbs_comp_plot_origin, aes(plot_cover_avg, fill = as.factor(plot))) + geom_histogram(binwidth = 0.5) + 
        facet_grid(state~year, margins = TRUE, scales = "free")

ggplot(umbs_comp_plot_origin, aes(plot_cover_avg, fill = plot, color=plot)) +
        geom_density(alpha = 0.1)

ggplot(umbs_comp_plot_origin, aes(plot_cover_avg, fill = plot, color=plot)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year)

ggplot(umbs_comp_plot_origin, aes(plot_cover_avg, fill = plot, color=plot)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year + plot)

# Exploring distributions for these data:
descdist(umbs_comp_plot_origin$plot_cover_avg, discrete = FALSE)

# Gamma distribution 
fit.gamma <- fitdist(umbs_comp_plot_origin$plot_cover_avg, "gamma")
plot(fit.gamma)
#
# Weibull distribution
fit.weibull  <- fitdist(umbs_comp_plot_origin$plot_cover_avg, "weibull")
plot(fit.weibull)

# Lognormal distribution 
fit.ln <- fitdist(umbs_comp_plot_origin$plot_cover_avg, "lnorm")
plot(fit.ln)

par(mfrow=c(2,2))
plot.legend <- c("Gamma", "Weibull", "Log Normal")
denscomp(list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
cdfcomp (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
qqcomp  (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
ppcomp  (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)

# Goodness of fit comparisons across fits (can't include the sqrt normal bc it becomes diff response values)
gofstat(list(fit.gamma, fit.weibull, fit.ln), fitnames = c("Gamma", "Weibull", "Log Normal"))
# log normal looks like it's the best fit
```


# Leverage plots and detecting Outliers. https://www.statmethods.net/stats/rdiagnostics.html
# These illustrate whether certain data points have more leverage (more influence), and thus could be outliers. It's a way of detecting outliers. Leverage plots can help identify whether a point has high or low influence, based on its leverage and residual and determining model fit with and without the point in question. Ultimately you decide whether the points are outliers or not, based on the knowledge of the system and how much it changes the model when included vs. excluded from the data used to fit the model. Here is a good overview of the combination of leverage and residual: scroll down to sections beginning at "13.3 Unusual Observations": https://daviddalpiaz.github.io/appliedstats/model-diagnostics.html

KBS
```{r}
# species level data
# KBS State-only model
fit_spp_state_kbs <- lm(log(relabun) ~ state, data = comp_kbs_spp)
outlierTest(fit_spp_state_kbs) # outliers - row 449
qqPlot(fit_spp_state_kbs, main="QQ Plot") 
hist(fit_spp_state_kbs$residuals)
leveragePlots(fit_spp_state_kbs)
ols_test_normality(fit_spp_state_kbs)

comp_kbs_spp_no_outliers <- comp_kbs_spp[-449,] # remove outliers
fit_spp_state_kbs1 <- lm(log(relabun) ~ state, data = comp_kbs_spp_no_outliers)
outlierTest(fit_spp_state_kbs1) # outliers - row 449
qqPlot(fit_spp_state_kbs1, main="QQ Plot") 
hist(fit_spp_state_kbs1$residuals)
leveragePlots(fit_spp_state_kbs1)
ols_test_normality(fit_spp_state_kbs1)

# KBS State and year model
fit_spp_stateyear_kbs <- lm(log(relabun) ~ state + year_factor, data = comp_kbs_spp)
outlierTest(fit_spp_stateyear_kbs) # no outliers
qqPlot(fit_spp_stateyear_kbs, main="QQ Plot") 
hist(fit_spp_stateyear_kbs$residuals)
leveragePlots(fit_spp_stateyear_kbs)
ols_test_normality(fit_spp_stateyear_kbs)

# Interaction plot (ignore for now the repeated measures with species); see: https://cran.r-project.org/web/packages/interactions/vignettes/interactions.html and: https://interactions.jacob-long.com/

fit3 <- lm(log(relabun) ~ state + year_factor + species, data = comp_kbs_spp)
interact_plot(fit3, pred = year_factor, modx = state)
fit4 <- lm(log(relabun) ~ state * year_factor + species, data = comp_kbs_spp)
interact_plot(fit4, pred = year_factor, modx = state, mod2 = species)
```
```{r}
# Plot level data
# KBS State-only model
fit_plot_state_kbs <- lm(log(plot_cover_avg) ~ state, data = comp_kbs_plot)
outlierTest(fit_plot_state_kbs) # outlier - row 252
comp_kbs_plot_no_outliers <- comp_kbs_plot[-252,] # remove outliers
qqPlot(fit_plot_state_kbs, main="QQ Plot") 
hist(fit_plot_state_kbs$residuals)
leveragePlots(fit_plot_state_kbs)
ols_test_normality(fit_plot_state_kbs)

# KBS State and year model
fit_plot_stateyear_kbs <- lm(log(plot_cover_avg) ~ state + year_factor, data = comp_kbs_plot)
outlierTest(fit_plot_stateyear_kbs) # outlier - row 190
comp_kbs_plot_no_outliers <- comp_kbs_plot[-190,] # remove outliers
qqPlot(fit_plot_stateyear_kbs, main="QQ Plot") 
hist(fit_plot_stateyear_kbs$residuals)
leveragePlots(fit_plot_stateyear_kbs) 
ols_test_normality(fit_plot_stateyear_kbs) # not all are under 0.05 but I think we can move forward with assumptions of normality here after log transformation

# Interaction plot (ignore for now the repeated measures with species); see: https://cran.r-project.org/web/packages/interactions/vignettes/interactions.html and: https://interactions.jacob-long.com/

fit3 <- lm(log(plot_cover_avg) ~ state * year_factor, data = comp_kbs_plot)
interact_plot(fit3, pred = year_factor, modx = state)
```
UMBS
```{r}
# species level data
# UMBS State-only model
fit_spp_state_umbs <- lm(log(relabun) ~ state, data = comp_umbs_spp)
outlierTest(fit_spp_state_umbs) # outliers - row 449
qqPlot(fit_spp_state_umbs, main="QQ Plot") 
hist(fit_spp_state_umbs$residuals)
leveragePlots(fit_spp_state_umbs)
ols_test_normality(fit_spp_state_umbs)

# UMBS State and year model
fit_spp_stateyear_umbs <- lm(log(relabun) ~ state + year_factor, data = comp_umbs_spp)
outlierTest(fit_spp_stateyear_umbs) # no outliers
qqPlot(fit_spp_stateyear_umbs, main="QQ Plot") 
hist(fit_spp_stateyear_umbs$residuals)
leveragePlots(fit_spp_stateyear_umbs)
ols_test_normality(fit_spp_stateyear_umbs)

# Interaction plot (ignore for now the repeated measures with species); see: https://cran.r-project.org/web/packages/interactions/vignettes/interactions.html and: https://interactions.jacob-long.com/

fit3 <- lm(log(relabun) ~ state + year_factor + species, data = comp_umbs_spp)
interact_plot(fit3, pred = year_factor, modx = state)
fit4 <- lm(log(relabun) ~ state * year_factor + species, data = comp_umbs_spp)
interact_plot(fit4, pred = year_factor, modx = state, mod2 = species)
```
```{r}
# Plot level data
# UMBS State-only model
fit_plot_state_umbs <- lm(log(plot_cover_avg) ~ state, data = comp_umbs_plot)
outlierTest(fit_plot_state_umbs) # outlier - row 252
qqPlot(fit_plot_state_umbs, main="QQ Plot") 
hist(fit_plot_state_umbs$residuals)
leveragePlots(fit_plot_state_umbs)
ols_test_normality(fit_plot_state_umbs)

comp_umbs_plot_no_outliers <- comp_umbs_plot[-298,] # remove outliers

# KBS State and year model
fit_plot_stateyear_umbs <- lm(log(plot_cover_avg) ~ state + year_factor, data = comp_umbs_plot)
outlierTest(fit_plot_stateyear_umbs) # outlier - row 190
comp_umbs_plot_no_outliers <- comp_umbs_plot[-190,] # remove outliers
qqPlot(fit_plot_stateyear_umbs, main="QQ Plot") 
hist(fit_plot_stateyear_umbs$residuals)
leveragePlots(fit_plot_stateyear_umbs) 
ols_test_normality(fit_plot_stateyear_umbs) # not all are under 0.05 but I think we can move forward with assumptions of normality here after log transformation

# Interaction plot (ignore for now the repeated measures with species); see: https://cran.r-project.org/web/packages/interactions/vignettes/interactions.html and: https://interactions.jacob-long.com/

fit3 <- lm(log(plot_cover_avg) ~ state * year_factor, data = comp_umbs_plot)
interact_plot(fit3, pred = year_factor, modx = state)
```

MIXED EFFECT MODELS

KBS species level data
```{r}
# KBS SPECIES LEVEL - Looking at RELATIVE ABUNDANCE
mod1 <- lmer(log(relabun) ~ state*year_factor + insecticide*year_factor + (1|species) + (1|plot), comp_kbs_spp, REML = FALSE)
ggplot(comp_kbs_spp,aes(x=plot,y=relabun,col=species)) + geom_jitter() + geom_boxplot(alpha=0.2) + facet_wrap(~plot)

# Check Assumptions:
# (1) Linearity: if covariates are not categorical (year isn't)
# (2) Homogeneity: Need to Check by plotting residuals vs predicted values.
par(mfrow=c(1,2))
plot(mod1)
# Homogeneity of variance is ok here (increasing variance in resids is not increasing with fitted values)
# Check for homogeneity of variances (true if p>0.05). If the result is not significant, the assumption of equal variances (homoscedasticity) is met (no significant difference between the group variances).
# *****Levene’s Test - tests whether or not the variance among two or more groups is equal - If the p-value is less than our chosen significance level, we can reject the null hypothesis and conclude that we have enough evidence to state that the variance among the groups is not equal (which we want).

leveneTest(residuals(mod1) ~ comp_kbs_spp$state)
# Assumption met
leveneTest(residuals(mod1) ~ comp_kbs_spp$species)
# Assumption met
leveneTest(residuals(mod1) ~ comp_kbs_spp$insecticide) 
# Assumption not met - this seems weird
leveneTest(residuals(mod1) ~ comp_kbs_spp$plot)
# Assumption met

# (3) Normality of error term: need to check by histogram, QQplot of residuals, could do Kolmogorov-Smirnov test.
# Check for normal residuals
qqPlot(resid(mod1))
hist(residuals(mod1))
shapiro.test(resid(mod1)) # not normally distributed resids bc p<0.05
outlierTest(mod1) # no outliers

# (4) Normality of random effect: Get the estimate of random effect (e.g., random intercepts), and check them as you would check the residual. 
require(lme4)
r_int<- ranef(mod1)$plot$`(Intercept)`
qqnorm(r_int)
# qqline(r_int) doesn't work
shapiro.test(r_int) 
# Not normally distributed random effect - p-value = 0.0015
```

```{r, error=TRUE}
# Do we need to include plot as a random effect with the KBS models?   
mod1 <- lmer(log(relabun) ~ state*year_factor + insecticide*year_factor + (1|species) + (1|plot), comp_kbs_spp, REML = FALSE)
mod2 <- lmer(log(relabun) ~ state*year_factor + insecticide*year_factor + (1|species), comp_kbs_spp, REML=FALSE)
# Run analysis of variance on each model (see this for more explanation on how anova on a linear mixed effects model is similar to an anove on a regular linear model: https://m-clark.github.io/docs/mixedModels/anovamixed.html)
anova(mod1)
anova(mod2)

anova(mod1, mod2) # Go with model 2 since pvalue >0.05, aka more complex model does not have something in it that matters (the two models are essentially the same so go w model with fewer parameters)
summ(mod1) 
summ(mod2)
AICctab(mod1, mod2, weights=T)

# Plot the fixed effects estimates for different models
# these are the fixed effects estimates from summary(mod1)
plot_model(mod2, sort.est = TRUE)
# these are the fixed predicted values:
plot_model(mod2, type = "pred", terms = c("year_factor", "state", "insecticide"))
# these are the random effects estimates
plot_model(mod2, type = "re", terms = c("species"))

# Do we need to include insecticide? (dropping insecticide from the model)
# mod2 <- lmer(log(relabun) ~ state*year_factor + insecticide*year_factor + (1|species), comp_kbs_spp, REML=FALSE)
mod3 <- lmer(log(relabun) ~ state*year_factor + (1|species), comp_kbs_spp, REML=FALSE)
anova(mod2, mod3)

AICctab(mod2, mod3, weights=T)
# Looks like no, pvalue > 0.05, so insecticide does not improve model fit so we will go with mod 3

# Does year need to be interactive with insecticide? - already removed insecticide
# mod4 <- lmer(log(relabun) ~ state*year_factor + insecticide + (1|species) + (1|plot), comp_kbs_spp, REML=FALSE)

# Does year need to be interactive with state?
mod5 <- lmer(log(relabun) ~ state + year_factor + (1|species), comp_kbs_spp, REML=FALSE)
anova(mod3, mod5)
# No, P>0.05 so state*year_factor doesn't strongly improve model fit so we will shift to mod5, but AIC values are super close!!

# Plot the fixed effects estimates for different models
# these are the fixed effects estimates from summary(mod5)
plot_model(mod5, sort.est = TRUE)
# these are the fixed predicted values:
plot_model(mod5, type = "pred", terms = c("year_factor", "state"))
# these are the random effects estimates
plot_model(mod5, type = "re", terms = c("species"))

# If we wanted to include plots nested within year it would look like this:
mod6 <- lmer(log(relabun) ~ state + year_factor + (1|species) + (1 + year_factor|plot), comp_kbs_spp, REML=FALSE)
# Model failed to converge with 1 negative eigenvalue: -1.9e+00
anova(mod5, mod6)
anova(mod5)

# mod4 (and mod5) are pretty complex in terms of interpretation (they actually don't have many parameters though). We could consider an alternative model that's simpler to understand and also one that provides more insight about the species. That would be something like this:
mod7 <- lmer(log(relabun) ~ state + species + (1+factor(year_factor)|plot), comp_kbs_spp, REML=FALSE)
mod7a <- lmer(log(relabun)  ~ state + species + factor(year_factor) + (1|plot), comp_kbs_spp, REML=FALSE)
mod7b <- lmer(log(relabun)  ~ state * factor(year_factor) + species + (1|plot), comp_kbs_spp, REML=FALSE)
mod7c <- lmer(log(relabun)  ~ state + species + factor(year_factor) + insecticide + (1|plot), comp_kbs_spp, REML=FALSE)
anova(mod5, mod7) # model 7 is a better fit to data
anova(mod7, mod7a) #mod 7a
anova(mod7a, mod7b) #mod 7a - interaction between state and year does not improve model go with simpler model (mod7a)
anova(mod7a, mod7c) #models are not different than one another, go with simpler model = mod7a
summary(mod7a)
anova(mod7a) # investigates whether at least one of the levels within each factor is significantly different.
# Yes, at least one of the species is different (they do not all have the same relative abundances). 
emmeans(mod7c, list(pairwise ~ state + year_factor), adjust = "tukey")
emmeans(mod7c, list(pairwise ~ year_factor), adjust = "tukey")
emmeans(mod7c, list(pairwise ~ species), adjust = "tukey")
emmeans(mod7c, list(pairwise ~ insecticide), adjust = "tukey")

# Take a look at the estimates for each fixed effect. These are the estimates from summary(mod7a). You'll see that species vary a lot - and many of them are different from zero (meaning their half cover date is significantly different from zero).
plot_model(mod7a, sort.est = TRUE)
# if you want to standardize the estimates:
plot_model(mod7a, sort.est = TRUE, type="std") 
# these are the fixed predicted values: - note this is a new plot
plot_model(mod7a, type = "pred", terms = c("species", "state", "year_factor"))
# these are the random effects estimates
plot_model(mod7a, type = "re")

# including native vs. exotic
comp_kbs_spp <- within(comp_kbs_spp, origin <- relevel(factor(origin), ref = "Native")) # releveling so native is the reference
mod8 <- lmer(log(relabun) ~ state * origin + (1+factor(year_factor)|plot), comp_kbs_spp, REML=FALSE)
mod9 <- lmer(log(relabun) ~ state + origin + (1+factor(year_factor)|plot), comp_kbs_spp, REML=FALSE)
mod9a <- lmer(log(relabun) ~ state + origin + factor(year_factor) + (1|plot), comp_kbs_spp, REML=FALSE)
anova(mod8, mod9) # model 8 is a better fit to data
anova(mod8, mod9a) # mod 8
summary(mod8)
anova(mod8)
emmeans(mod8, list(pairwise ~ state * origin), adjust = "tukey")

# including growth form - first with interaction term
comp_kbs_spp <- within(comp_kbs_spp, growth_habit <- relevel(factor(growth_habit), ref = "Forb")) # releveling so forb is the reference
mod10 <- lmer(log(relabun) ~ state * growth_habit + (1+factor(year_factor)|plot), comp_kbs_spp, REML=FALSE)
mod11 <- lmer(log(relabun) ~ state + growth_habit + (1+factor(year_factor)|plot), comp_kbs_spp, REML=FALSE)
mod11a <- lmer(log(relabun) ~ state + growth_habit + factor(year_factor) + (1|plot), comp_kbs_spp, REML=FALSE)
anova(mod10, mod11) # model 10 is a better fit to data
anova(mod10, mod11a) # model 10 is still a better fit to data
summary(mod10)
anova(mod10)
emmeans(mod10, list(pairwise ~ factor(year_factor) * growth_habit), adjust = "tukey")

# New version of our model incorporating interaction term and species within year so that there is a separate intercept and slope for each species. The issue here is that there are some species that are not found each year. Easiest to remove those from another version of this dataframe before running below. Otherwise, it's not a balanced design.
# updated mod4
mod12 <- lmer(log(relabun) ~ state * factor(year_factor) + (1+factor(year_factor)|species), comp_kbs_spp)

# So another version of this model would include the interaction but not include the nesting (and thus would assume that species aren't observed ea yr)
# updated mod5
mod13 <- lmer(log(relabun) ~ state * factor(year_factor) + (1|species), comp_kbs_spp)

# All the models ran:
mod1 <- lmer(log(relabun) ~ state*year_factor + insecticide*year_factor + (1|species) + (1|plot), comp_kbs_spp, REML=FALSE)
mod2 <- lmer(log(relabun) ~ state*year_factor + insecticide*year_factor + (1|species), comp_kbs_spp, REML=FALSE)
mod3 <- lmer(log(relabun) ~ state*year_factor + (1|species), comp_kbs_spp, REML=FALSE)
mod4 <- lmer(log(relabun) ~ state*year_factor + insecticide + (1|species) + (1|plot), comp_kbs_spp, REML=FALSE)
mod5 <- lmer(log(relabun) ~ state + year_factor + (1|species), comp_kbs_spp, REML=FALSE)
mod6 <- lmer(log(relabun) ~ state + year_factor + (1|species) + (1 + year|plot), comp_kbs_spp, REML=FALSE)
mod7 <- lmer(log(relabun) ~ state + species + (1+factor(year_factor)|plot), comp_kbs_spp, REML=FALSE)
mod7a <- lmer(log(relabun) ~ state + species + factor(year_factor) + (1|plot), comp_kbs_spp, REML=FALSE)
mod7b <- lmer(log(relabun) ~ state * factor(year_factor) + species + (1|plot), comp_kbs_spp, REML=FALSE)
mod7c <- lmer(log(relabun) ~ state + species + factor(year_factor) + insecticide + (1|plot), comp_kbs_spp, REML=FALSE)
mod8 <- lmer(log(relabun) ~ state * origin + (1+factor(year_factor)|plot), comp_kbs_spp, REML=FALSE)
mod9 <- lmer(log(relabun) ~ state + origin + (1+factor(year_factor)|plot), comp_kbs_spp, REML=FALSE)
mod9a <- lmer(log(relabun) ~ state + origin + factor(year_factor) + (1|plot), comp_kbs_spp, REML=FALSE)
mod10 <- lmer(log(relabun) ~ state * growth_habit + (1+factor(year_factor)|plot), comp_kbs_spp, REML=FALSE)
mod11 <- lmer(log(relabun) ~ state + growth_habit + (1+factor(year_factor)|plot), comp_kbs_spp, REML=FALSE)
mod11a <- lmer(log(relabun) ~ state + growth_habit + factor(year_factor) + (1|plot), comp_kbs_spp, REML=FALSE)
mod12 <- lmer(log(relabun) ~ state * factor(year_factor) + (1+factor(year_factor)|species), comp_kbs_spp)
mod13 <- lmer(log(relabun) ~ state * factor(year_factor) + (1|species), comp_kbs_spp)
AICctab(mod1, mod2, mod3, mod5, mod6, mod7, mod7a, mod7b, mod7c, mod8, mod9, mod9a, mod10, mod11, mod11a, mod13, weights=T) #mod7a is the best fitting model (took out mod12 bc it wouldn't run)
summary(mod7a)
anova(mod7a)
```

# KBS Plot-level Mixed Effects Models:
```{r, error=TRUE}
mod1p <- lmer(log(plot_cover_avg) ~ state + (1|plot), comp_kbs_plot, REML=FALSE)
mod2p <- lmer(log(plot_cover_avg) ~ insecticide + (1|plot), comp_kbs_plot, REML=FALSE)
mod3p <- lmer(log(plot_cover_avg) ~ insecticide + state + (1|plot), comp_kbs_plot, REML=FALSE)
mod4p <- lmer(log(plot_cover_avg) ~ insecticide * state + (1|plot), comp_kbs_plot, REML=FALSE)
mod5p <- lmer(log(plot_cover_avg) ~ state + year_factor + (1|plot), comp_kbs_plot, REML=FALSE)
mod6p <- lmer(log(plot_cover_avg) ~ state + year_factor + insecticide + (1|plot), comp_kbs_plot, REML=FALSE)
mod7p <- lmer(log(plot_cover_avg) ~ state * year_factor + (1|plot), comp_kbs_plot, REML=FALSE)
mod8p <- lmer(log(plot_cover_avg) ~ state * year_factor + insecticide + (1|plot), comp_kbs_plot, REML=FALSE)
mod9p <- lmer(log(plot_cover_avg) ~ state * insecticide + year_factor + (1|plot), comp_kbs_plot, REML=FALSE)
mod10p <- lmer(log(plot_cover_avg) ~ state + insecticide * year_factor + (1|plot), comp_kbs_plot, REML=FALSE)
mod11p <- lmer(log(plot_cover_avg) ~ state * year_factor * insecticide + (1|plot), comp_kbs_plot, REML=FALSE)
AICctab(mod1p, mod2p, mod3p, mod4p, mod5p, mod6p, mod7p, mod8p, mod9p, mod10p, mod11p, weights=T) # model 11p and 10p the same
anova(mod7p, mod5p) #7p just barely better
summ(mod7p)
emmeans(mod7p, list(pairwise ~ state*year_factor), adjust = "tukey")

# including native vs. exotic
kbs_comp_plot_origin <- within(kbs_comp_plot_origin, origin <- relevel(factor(origin), ref = "Native")) # releveling so native is the reference
mod12p <- lmer(log(plot_cover_avg) ~ state * origin + (1+year_factor|plot), kbs_comp_plot_origin, REML=FALSE)
mod13p <- lmer(log(plot_cover_avg) ~ state + origin + (1+year_factor|plot), kbs_comp_plot_origin, REML=FALSE)
mod14p <- lmer(log(plot_cover_avg) ~ state + origin + year_factor + (1|plot), kbs_comp_plot_origin, REML=FALSE)
anova(mod12p, mod13p) # go with model 12pu 
anova(mod12p, mod14p) # mod 14pu
summ(mod14p)
anova(mod14p)
emmeans(mod14p, list(pairwise ~ state + origin + year_factor), adjust = "tukey")

# including growth form - first with interaction term
kbs_comp_plot_growthhabit <- within(kbs_comp_plot_growthhabit, growth_habit <- relevel(factor(growth_habit), ref = "Forb")) # releveling so forb is the reference
mod15p <- lmer(log(plot_cover_avg) ~ state * growth_habit + (1+year_factor|plot), kbs_comp_plot_growthhabit, REML=FALSE)
mod16p <- lmer(log(plot_cover_avg) ~ state + growth_habit + (1+year_factor|plot), kbs_comp_plot_growthhabit, REML=FALSE)
mod17p <- lmer(log(plot_cover_avg) ~ state + growth_habit + year_factor + (1|plot), kbs_comp_plot_growthhabit, REML=FALSE)
anova(mod15p, mod16p) # go with model 15pu 
anova(mod15p, mod17p) # mod 17pu
summ(mod17p)
anova(mod17p)
emmeans(mod17p, list(pairwise ~ state + growth_habit + year_factor), adjust = "tukey")
```

UMBS species level mixed effects models
```{r, error=TRUE}
# UMBS SPECIES LEVEL - Looking at RELATIVE ABUNDANCE
mod1u <- lmer(log(relabun) ~ state*year_factor + insecticide*year_factor + (1|species) + (1|plot), comp_umbs_spp, REML = FALSE)
ggplot(comp_umbs_spp,aes(x=plot,y=relabun,col=species)) + geom_jitter() + geom_boxplot(alpha=0.2) + facet_wrap(~plot)

# Check Assumptions:
# (1) Linearity: if covariates are not categorical (year isn't)
# (2) Homogeneity: Need to Check by plotting residuals vs predicted values.
par(mfrow=c(1,2))
plot(mod1u)
# Homogeneity of variance is ok here (increasing variance in resids is not increasing with fitted values)
# Check for homogeneity of variances (true if p>0.05). If the result is not significant, the assumption of equal variances (homoscedasticity) is met (no significant difference between the group variances).
# *****Levene’s Test - tests whether or not the variance among two or more groups is equal - If the p-value is less than our chosen significance level, we can reject the null hypothesis and conclude that we have enough evidence to state that the variance among the groups is not equal (which we want).

leveneTest(residuals(mod1u) ~ comp_umbs_spp$state)
# Assumption not met
leveneTest(residuals(mod1u) ~ comp_umbs_spp$species)
# Assumption met
leveneTest(residuals(mod1u) ~ comp_umbs_spp$insecticide) 
# Assumption not met - this seems weird
leveneTest(residuals(mod1u) ~ comp_umbs_spp$plot)
# Assumption met

# (3) Normality of error term: need to check by histogram, QQplot of residuals, could do Kolmogorov-Smirnov test.
# Check for normal residuals
qqPlot(resid(mod1u))
hist(residuals(mod1u))
shapiro.test(resid(mod1u)) # not normally distributed resids bc p<0.05
outlierTest(mod1u) # no outliers

# (4) Normality of random effect: Get the estimate of random effect (e.g., random intercepts), and check them as you would check the residual. 
require(lme4)
r_int<- ranef(mod1u)$plot$`(Intercept)`
qqnorm(r_int)
#qline(r_int) #doesn't work
shapiro.test(r_int) 
# Normally distributed random effect - p-value > 0.05
```

```{r, error=TRUE}
# Do we need to include plot as a random effect with the KBS models?   
mod1u <- lmer(log(relabun) ~ state*year_factor + insecticide*year_factor + (1|species) + (1|plot), comp_umbs_spp, REML = FALSE)
mod2u <- lmer(log(relabun) ~ state*year_factor + insecticide*year_factor + (1|species), comp_umbs_spp, REML=FALSE)
# Run analysis of variance on each model (see this for more explanation on how anova on a linear mixed effects model is similar to an anove on a regular linear model: https://m-clark.github.io/docs/mixedModels/anovamixed.html)
anova(mod1u)
anova(mod2u)

anova(mod1u, mod2u) # Go with model 2 since pvalue >0.05, aka more complex model does not have something in it that matters (the two models are essentially the same so go w model with fewer parameters)
summ(mod1u) 
summ(mod2u)
AICctab(mod1u, mod2u, weights=T)

# Plot the fixed effects estimates for different models
# these are the fixed effects estimates from summary(mod1)
plot_model(mod2u, sort.est = TRUE)
# these are the fixed predicted values:
plot_model(mod2u, type = "pred", terms = c("year_factor", "state", "insecticide"))
# these are the random effects estimates
plot_model(mod2u, type = "re", terms = c("species"))


# Does year need to be interactive with insecticide? - already removed insecticide
mod3 <- lmer(log(relabun) ~ state*year_factor + insecticide + (1|species) + (1|plot), comp_umbs_spp, REML=FALSE)
anova(mod2u, mod3u) # Looks like no, pvalue > 0.05, so year interactive with insecticide does not improve model fit so we will go with mod 2u

# Do we need to include insecticide? (dropping insecticide from the model)
mod4u <- lmer(log(relabun) ~ state*year_factor + (1|species), comp_umbs_spp, REML=FALSE)
anova(mod2u, mod4u) # Looks like no, pvalue > 0.05, so insecticide does not improve model fit so we will go with mod 4u

AICctab(mod2u, mod4u, weights=T)

# Does year need to be interactive with state?
mod5u <- lmer(log(relabun) ~ state + year_factor + (1|species), comp_umbs_spp, REML=FALSE)
anova(mod4u, mod5u)
# No, P>0.05 so state*year_factor doesn't strongly improve model fit so we will shift to mod5u, but AIC values are close!!

# Plot the fixed effects estimates for different models
# these are the fixed effects estimates from summary(mod5)
plot_model(mod5u, sort.est = TRUE)
# these are the fixed predicted values:
plot_model(mod5u, type = "pred", terms = c("year_factor", "state"))
# these are the random effects estimates
plot_model(mod5u, type = "re", terms = c("species"))

# If we wanted to include plots nested within year it would look like this:
mod6u <- lmer(log(relabun) ~ state + year_factor + (1|species) + (1 + year_factor|plot), comp_umbs_spp, REML=FALSE)
anova(mod5u, mod6u) # go with simpler model = mod5u
anova(mod5u)

# reminder what model 5u is:
# mod5u <- lmer(log(relabun) ~ state + year_factor + (1|species), comp_umbs_spp, REML=FALSE)

# mod5u (and mod6u) are pretty complex in terms of interpretation (they actually don't have many parameters though). We could consider an alternative model that's simpler to understand and also one that provides more insight about the species. That would be something like this:
mod7u <- lmer(log(relabun) ~ state + species + (1+factor(year_factor)|plot), comp_umbs_spp, REML=FALSE)
mod7au <- lmer(log(relabun)  ~ state + species + factor(year_factor) + (1|plot), comp_umbs_spp, REML=FALSE)
mod7bu <- lmer(log(relabun)  ~ state * factor(year_factor) + species + (1|plot), comp_umbs_spp, REML=FALSE)
mod7cu <- lmer(log(relabun)  ~ state + species + factor(year_factor) + insecticide + (1|plot), comp_umbs_spp, REML=FALSE)
anova(mod5u, mod7u) # model 7u is a better fit to data
anova(mod7u, mod7au) #mod 7au
anova(mod7au, mod7bu) #mod 7au - interaction between state and year does not improve model go with simpler model (mod7a)
anova(mod7au, mod7cu) #models are not different than one another, go with simpler model = mod7au
summary(mod7au)
anova(mod7au) # investigates whether at least one of the levels within each factor is significantly different.
# Yes, at least one of the species is different (they do not all have the same relative abundances). 
emmeans(mod7au, list(pairwise ~ state + year_factor), adjust = "tukey")
emmeans(mod7au, list(pairwise ~ year_factor), adjust = "tukey")
emmeans(mod7au, list(pairwise ~ species), adjust = "tukey")
emmeans(mod7au, list(pairwise ~ insecticide), adjust = "tukey")


# Take a look at the estimates for each fixed effect. These are the estimates from summary(mod7a). You'll see that species vary a lot - and many of them are different from zero (meaning their half cover date is significantly different from zero).
plot_model(mod7au, sort.est = TRUE)
# if you want to standardize the estimates:
plot_model(mod7au, sort.est = TRUE, type="std") 
# these are the fixed predicted values: - note this is a new plot
plot_model(mod7au, type = "pred", terms = c("species", "state", "year_factor"))
# these are the random effects estimates
plot_model(mod7au, type = "re")

# including native vs. exotic
comp_umbs_spp <- within(comp_umbs_spp, origin <- relevel(factor(origin), ref = "Native")) # releveling so native is the reference
mod8u <- lmer(log(relabun) ~ state * origin + (1+factor(year_factor)|plot), comp_umbs_spp, REML=FALSE)
mod9u <- lmer(log(relabun) ~ state + origin + (1+factor(year_factor)|plot), comp_umbs_spp, REML=FALSE)
mod9au <- lmer(log(relabun) ~ state + origin + factor(year_factor) + (1|plot), comp_umbs_spp, REML=FALSE)
anova(mod8u, mod9u) # model 8 is a better fit to data
anova(mod8u, mod9au) # mod 8
summary(mod8u)
anova(mod8u)
emmeans(mod8u, list(pairwise ~ state * origin), adjust = "tukey")

# including growth form - first with interaction term
comp_umbs_spp <- within(comp_umbs_spp, growth_habit <- relevel(factor(growth_habit), ref = "Forb")) # releveling so forb is the reference
mod10u <- lmer(log(relabun) ~ state * growth_habit + (1+factor(year_factor)|plot), comp_umbs_spp, REML=FALSE)
mod11u <- lmer(log(relabun) ~ state + growth_habit + (1+factor(year_factor)|plot), comp_umbs_spp, REML=FALSE)
mod11au <- lmer(log(relabun) ~ state + growth_habit + factor(year_factor) + (1|plot), comp_umbs_spp, REML=FALSE)
anova(mod10u, mod11u) # model 10 is a better fit to data
anova(mod10u, mod11au) # model 10 is still a better fit to data
summary(mod10u)
anova(mod10u)
emmeans(mod10u, list(pairwise ~ factor(year_factor) * growth_habit), adjust = "tukey")

# New version of our model incorporating interaction term and species within year so that there is a separate intercept and slope for each species. The issue here is that there are some species that are not found each year. Easiest to remove those from another version of this dataframe before running below. Otherwise, it's not a balanced design.
# updated mod4
mod12u <- lmer(log(relabun) ~ state * factor(year_factor) + (1+factor(year_factor)|species), comp_umbs_spp)

# So another version of this model would include the interaction but not include the nesting (and thus would assume that species aren't observed ea yr)
# updated mod5
mod13u <- lmer(log(relabun) ~ state * factor(year_factor) + (1|species), comp_umbs_spp)

# All the models ran:
mod1u <- lmer(log(relabun) ~ state*year_factor + insecticide*year_factor + (1|species) + (1|plot), comp_umbs_spp, REML=FALSE)
mod2u <- lmer(log(relabun) ~ state*year_factor + insecticide*year_factor + (1|species), comp_umbs_spp, REML=FALSE)
mod3u <- lmer(log(relabun) ~ state*year_factor + (1|species), comp_umbs_spp, REML=FALSE)
mod4u <- lmer(log(relabun) ~ state*year_factor + insecticide + (1|species) + (1|plot), comp_umbs_spp, REML=FALSE)
mod5u <- lmer(log(relabun) ~ state + year_factor + (1|species), comp_umbs_spp, REML=FALSE)
mod6u <- lmer(log(relabun) ~ state + year_factor + (1|species) + (1 + year|plot), comp_umbs_spp, REML=FALSE)
mod7u <- lmer(log(relabun) ~ state + species + (1+factor(year_factor)|plot), comp_umbs_spp, REML=FALSE)
mod7au <- lmer(log(relabun) ~ state + species + factor(year_factor) + (1|plot), comp_umbs_spp, REML=FALSE)
mod7bu <- lmer(log(relabun) ~ state * factor(year_factor) + species + (1|plot), comp_umbs_spp, REML=FALSE)
mod7cu <- lmer(log(relabun) ~ state + species + factor(year_factor) + insecticide + (1|plot), comp_umbs_spp, REML=FALSE)
mod8u <- lmer(log(relabun) ~ state * origin + (1+factor(year_factor)|plot), comp_umbs_spp, REML=FALSE)
mod9u <- lmer(log(relabun) ~ state + origin + (1+factor(year_factor)|plot), comp_umbs_spp, REML=FALSE)
mod9au <- lmer(log(relabun) ~ state + origin + factor(year_factor) + (1|plot), comp_umbs_spp, REML=FALSE)
mod10u <- lmer(log(relabun) ~ state * growth_habit + (1+factor(year_factor)|plot), comp_umbs_spp, REML=FALSE)
mod11u <- lmer(log(relabun) ~ state + growth_habit + (1+factor(year_factor)|plot), comp_umbs_spp, REML=FALSE)
mod11au <- lmer(log(relabun) ~ state + growth_habit + factor(year_factor) + (1|plot), comp_umbs_spp, REML=FALSE)
mod12u <- lmer(log(relabun) ~ state * factor(year_factor) + (1+factor(year_factor)|species), comp_umbs_spp)
mod13u <- lmer(log(relabun) ~ state * factor(year_factor) + (1|species), comp_umbs_spp)
AICctab(mod1u, mod2u, mod3u, mod5u, mod6u, mod7u, mod7au, mod7bu, mod7cu, mod8u, mod9u, mod9au, mod10u, mod11u, mod11au, mod13u, weights=T) #mod7a is the best fitting model (took out mod12 bc it wouldn't run)
summary(mod7au)
anova(mod7au)
```
# UBS Plot-level Mixed Effects Models:
```{r, error=TRUE}
mod1pu <- lmer(log(plot_cover_avg) ~ state + (1|plot), comp_umbs_plot, REML=FALSE)
mod2pu <- lmer(log(plot_cover_avg) ~ insecticide + (1|plot), comp_umbs_plot, REML=FALSE)
mod3pu <- lmer(log(plot_cover_avg) ~ insecticide + state + (1|plot), comp_umbs_plot, REML=FALSE)
mod4pu <- lmer(log(plot_cover_avg) ~ insecticide * state + (1|plot), comp_umbs_plot, REML=FALSE)
mod5pu <- lmer(log(plot_cover_avg) ~ state + year_factor + (1|plot), comp_umbs_plot, REML=FALSE)
mod6pu <- lmer(log(plot_cover_avg) ~ state + year_factor + insecticide + (1|plot), comp_umbs_plot, REML=FALSE)
mod7pu <- lmer(log(plot_cover_avg) ~ state * year_factor + (1|plot), comp_umbs_plot, REML=FALSE)
mod8pu <- lmer(log(plot_cover_avg) ~ state * year_factor + insecticide + (1|plot), comp_umbs_plot, REML=FALSE)
mod9pu <- lmer(log(plot_cover_avg) ~ state * insecticide + year_factor + (1|plot), comp_umbs_plot, REML=FALSE)
mod10pu <- lmer(log(plot_cover_avg) ~ state + insecticide * year_factor + (1|plot), comp_umbs_plot, REML=FALSE)
mod11pu <- lmer(log(plot_cover_avg) ~ state * year_factor * insecticide + (1|plot), comp_umbs_plot, REML=FALSE)
AICctab(mod1pu, mod2pu, mod3pu, mod4pu, mod5pu, mod6pu, mod7pu, mod8pu, mod9pu, mod10pu, mod11pu, weights=T) # model 11p and 10p the same
anova(mod7pu, mod8pu) #7p just barely better
summ(mod7pu)
summ(mod8pu)
emmeans(mod8pu, list(pairwise ~ state*year_factor), adjust = "tukey")

# including native vs. exotic
umbs_comp_plot_origin <- within(umbs_comp_plot_origin, origin <- relevel(factor(origin), ref = "Native")) # releveling so native is the reference
mod12pu <- lmer(log(plot_cover_avg) ~ state * origin + (1+year_factor|plot), umbs_comp_plot_origin, REML=FALSE)
mod13pu <- lmer(log(plot_cover_avg) ~ state + origin + (1+year_factor|plot), umbs_comp_plot_origin, REML=FALSE)
mod14pu <- lmer(log(plot_cover_avg) ~ state + origin + year_factor + (1|plot), umbs_comp_plot_origin, REML=FALSE)
anova(mod12pu, mod13pu) # go with model 13pu 
anova(mod13pu, mod14pu) # mod 14pu
summary(mod14pu)
summ(mod14pu)
anova(mod14pu)
emmeans(mod14pu, list(pairwise ~ state + origin + year_factor), adjust = "tukey")

# including growth form - first with interaction term
umbs_comp_plot_growthhabit <- within(umbs_comp_plot_growthhabit, growth_habit <- relevel(factor(growth_habit), ref = "Forb")) # releveling so forb is the reference
mod15pu <- lmer(log(plot_cover_avg) ~ state * growth_habit + (1+year_factor|plot), umbs_comp_plot_growthhabit, REML=FALSE)
mod16pu <- lmer(log(plot_cover_avg) ~ state + growth_habit + (1+year_factor|plot), umbs_comp_plot_growthhabit, REML=FALSE)
mod17pu <- lmer(log(plot_cover_avg) ~ state + growth_habit + year_factor + (1|plot), umbs_comp_plot_growthhabit, REML=FALSE)
anova(mod15pu, mod16pu) # go with model 16pu 
anova(mod16pu, mod17pu) # mod 17pu
summary(mod17pu)
summ(mod17pu)
anova(mod17pu)
emmeans(mod17pu, list(pairwise ~ state + growth_habit + year_factor), adjust = "tukey")
```

