---
title: "warmXtrophic Project: Plant Composition Data Wrangling"
author: "Kara Dobson, Moriah Young, Pat Bills"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=80),tidy=TRUE)
```

```{r}
# Clear all existing data
rm(list=ls())

#Load packages
library(tidyverse)
library(ggplot2)
library(plotrix)

# Set working directory
Sys.getenv("L1DIR")
L0_dir <- Sys.getenv("L0DIR")
L1_dir <- Sys.getenv("L1DIR")
L2_dir <- Sys.getenv("L2DIR")
list.files(L1_dir)

# read in plant comp data
comp <- read.csv(file.path(L1_dir, "plant_composition/final_plantcomp_species_removed_L1.csv"))

# read in temp data
GDD <- read.csv(file.path(L1_dir, "HOBO_data/GDD_L1.csv"))
temps <- read.csv(file.path(L1_dir, "HOBO_data/plant_comp_temps_L1.csv"))


# delete 2021 UMBS data from dataframe 
comp <- comp[-which(comp$year == "2021" & comp$site == "umbs"),]
```

```{r}
# adding sequential year variable starting at 1: this is because the years (e.g. 2015, 2016, etc) are large numbers compared with other values in the dataset. We can always label axes with these real years.
comp$year_factor[comp$year == 2015] <- 1
comp$year_factor[comp$year == 2016] <- 2
comp$year_factor[comp$year == 2017] <- 3
comp$year_factor[comp$year == 2018] <- 4
comp$year_factor[comp$year == 2019] <- 5
comp$year_factor[comp$year == 2020] <- 6
comp$year_factor[comp$year == 2021] <- 7

comp$year_factor <- as.factor(comp$year_factor)
str(comp)
comp$cover <- as.numeric(comp$cover)
str(comp)
```

Peak biomass Selection
```{r}
# select peak biomass dates 
# select for all June dates
june_comp <- dplyr::filter(comp, month %in% c(6)) 
june_comp$date <- as.Date(june_comp$date)
unique(june_comp$date)
str(june_comp)

# need to change "2021-06-12" to "2021-06-14" - this is because these two dates together make one data collection point (they were collected over two days instead of one)
june_comp$date[june_comp$date == "2021-06-12"] <- "2021-06-14"
unique(june_comp$date)

# select the max date in the month per year
june_comp1 <- june_comp %>% group_by(site, year, month) %>% slice_max(julian)
unique(june_comp1$date)

# select for all July dates
july_comp <- dplyr::filter(comp, month %in% c(7))
july_comp$date <- as.Date(july_comp$date)
unique(july_comp$date)
july_comp1 <- july_comp %>% group_by(site, year, month) %>% slice_max(julian)
unique(july_comp1$date)

# select for all August dates
august_comp <- dplyr::filter(comp, month %in% c(8))
august_comp$date <- as.Date(august_comp$date)
unique(august_comp$date)
unique(august_comp$julian)
# need to change "2020-08-10" and "2020-08-14" to "2020-08-15" - this is because these dates together make one data collection point (they were collected over three days instead of one)
august_comp$date[august_comp$date == "2020-08-10"] <- "2020-08-15"
august_comp$date[august_comp$date == "2020-08-14"] <- "2020-08-15"
# now do the same so that the dates match the same julian data
august_comp$julian[august_comp$julian == 223] <- 228
august_comp$julian[august_comp$julian == 227] <- 228
unique(august_comp$date) # check to see if above code did what you wanted it to do
august_comp1 <- august_comp %>% group_by(site, year, month) %>% slice_max(date)
unique(august_comp1$date)

# merge the june, july, and august data frames together
peak_comp <- full_join(august_comp1, july_comp1, by = c("site", "plot", "species", "cover", "date", "julian", "year", "month", "treatment_key", "state", "insecticide", "scientific_name", "common_name", "USDA_code", "LTER_code", "origin", "group", "family", "duration", "growth_habit", "year_factor"))

peak_comp <- full_join(peak_comp, june_comp1, by = c("site", "plot", "species", "cover", "date", "julian", "year", "month", "treatment_key", "state", "insecticide", "scientific_name", "common_name", "USDA_code", "LTER_code", "origin", "group", "family", "duration", "growth_habit", "year_factor"))
```

Calculate absolute and relative abundance at peak biomass dates
```{r}
peak_comp1 <- peak_comp %>%  
        select(c("species", "site", "year_factor", "month", "date", "year", "julian", "plot", "cover", "state", "insecticide", "treatment_key"))
peak_comp1$cover <- as.numeric(as.character(peak_comp1$cover))

# first get summed cover for all plants per plot
plot_cover_total = aggregate(cover ~ plot*year_factor*year*date*month*site*state*insecticide*treatment_key, data = peak_comp1, FUN = sum, na.rm = T)
names(plot_cover_total)[names(plot_cover_total) == "cover"] <- "plot_cover_total"
peak_comp2 <- merge(peak_comp1, plot_cover_total, by = c("plot","year_factor", "year", "date", "month", "site", "state", "treatment_key", "insecticide"))

#calculate absolute percent cover per species in each plot
peak_comp2$absolute <- peak_comp2$cover/100

#calculate relative percent cover per species in each plot
peak_comp2$relative <- peak_comp2$cover/peak_comp2$plot_cover_total
```

Origin
```{r}
origin_comp <- peak_comp %>% 
        select(species, site, year_factor, month, date, year, julian, plot, cover, state, insecticide, origin, treatment_key)
origin_comp$cover <- as.numeric(as.character(origin_comp$cover))

# first get summed cover for all plants per plot
origin_cover_total = aggregate(cover ~ plot*year_factor*year*date*month*site*state*insecticide*origin*treatment_key, data = origin_comp, FUN = sum, na.rm = T)
names(origin_cover_total)[names(origin_cover_total) == "cover"] <- "origin_cover_total"

origin_cover_total1 <- merge(origin_cover_total, plot_cover_total, by = c("plot","year_factor", "year", "date", "month", "site", "state", "insecticide", "treatment_key"))

#calculate relative percent cover
origin_cover_total1$relative <- origin_cover_total1$origin_cover_total/origin_cover_total1$plot_cover_total

#calculate absolute percent cover
origin_cover_total1$absolute <- origin_cover_total1$origin_cover_total/100

# subset data to just include forbs and graminoids
origin_cover_total1 <- subset(origin_cover_total1, origin == "Exotic" | origin == "Native")
```

Growth Habit
```{r}
growth_comp <- peak_comp %>% 
        select(species, site, year_factor, month, date, year, julian, plot, cover, state, insecticide, growth_habit, treatment_key)
growth_comp$cover <- as.numeric(as.character(growth_comp$cover))

# first get summed cover for all plants per plot
growth_cover_total = aggregate(cover ~ plot*year_factor*year*date*month*site*state*insecticide*growth_habit*treatment_key, data = growth_comp, FUN = sum, na.rm = T)
names(growth_cover_total)[names(growth_cover_total) == "cover"] <- "growth_cover_total"

growth_cover_total1 <- merge(growth_cover_total, plot_cover_total, by = c("plot","year_factor", "year", "date", "month", "site", "state", "insecticide", "treatment_key"))

#calculate relative percent cover
growth_cover_total1$relative <- growth_cover_total1$growth_cover_total/growth_cover_total1$plot_cover_total

#calculate relative percent cover
growth_cover_total1$absolute <- growth_cover_total1$growth_cover_total/100

# subset data to just include forbs and graminoids
growth_cover_total1 <- subset(growth_cover_total1, growth_habit == "Forb" | growth_habit == "Graminoid")

```

Plot June, July, and August relative abundances to see which month we should choose for each site
```{r}
comp_umbs_19 <- subset(peak_comp2, year == "2019" & site == "umbs")

avg_comp_umbs_19 <- comp_umbs_19 %>%
  group_by(site, state, month) %>%
  summarize(avg_plot_cover_total = mean(plot_cover_total, na.rm = TRUE),
            se = std.error(plot_cover_total, na.rm = TRUE))

ggplot(avg_comp_umbs_19, aes(x = month, y = avg_plot_cover_total, fill = state)) +
        #facet_wrap(.~month) +
        geom_line() +
        geom_point() +
        theme_classic()


##############################################
avg_comp <- peak_comp2 %>%
  group_by(site, state, month, year) %>%
  summarize(avg_plot_cover_total = mean(plot_cover_total, na.rm = TRUE),
            se = std.error(plot_cover_total, na.rm = TRUE))

# Order warmed and ambient so that warmed shows up first in plotting (and is default is red = warmed; blue = ambient). First make it a factor
avg_comp$state <- as.factor(avg_comp$state)
levels(avg_comp$state)
# [1] "ambient" "warmed" 
avg_comp$state <- factor(avg_comp$state, levels(avg_comp$state)[c(2,1)])
levels(avg_comp$state)
# [1] "warmed"  "ambient"

avg_comp_plot <- function(loc) { 
  avg_comp <- subset(avg_comp, site == loc)
  return(ggplot(avg_comp, aes(x = month, y = avg_plot_cover_total, fill = state)) +
        facet_grid(.~year) +
        geom_line(aes(color = state)) +
        geom_point(aes(color = state)) +
        labs(x = "Month", y = "Average Total Plot Cover", title = loc) +
        theme_classic())
}

avg_comp_plot("umbs")
avg_comp_plot("kbs")

```

Based on graphs - for calculating absolute and relative abundance of plant composition, max June date will be used for UMBS and max August date will be used for KBS for each year.
```{r}
peak_comp_umbs <- peak_comp2[which(peak_comp2$month == "7" & peak_comp2$site == "umbs"),]
peak_comp_kbs <- peak_comp2[which(peak_comp2$month == "8" & peak_comp2$site == "kbs"),]
peak_comp <- full_join(peak_comp_umbs, peak_comp_kbs, by = c("species", "site", "year_factor", "month", "date", "year", "julian", "plot", "cover", "state", "insecticide", "treatment_key", "plot_cover_total", "relative", "absolute"))
write.csv(peak_comp, file.path(L2_dir, "plant_composition/final_plant_comp_species_L2.csv"), row.names = F)

# Average relative abundance by PLOT LEVEL
relative_plot <- peak_comp %>%
  group_by(plot, year, state, site, month, insecticide, treatment_key, year_factor) %>%
  summarize(avg_relative_abundance = mean(relative, na.rm=T))

# Average absolute abundance by PLOT LEVEL
absolute_plot <- peak_comp %>%
  group_by(plot, year, state, site, month, insecticide, treatment_key, year_factor) %>%
  summarize(avg_absolute_abundance = mean(absolute, na.rm=T))

peak_comp_plot <- full_join(relative_plot, absolute_plot, by = c("site", "year_factor", "month", "year", "plot", "state", "insecticide", "treatment_key"))

# KD edits - merging in temp data
peak_comp_plot_temps <- left_join(peak_comp_plot, temps, by=c('site','year','state'))
# merge GDD - first, select the right julian date for GDD (based on HOBO_GDD_L1.R script)
GDD_plant_comp <- GDD %>%
        filter(site == "kbs" & julian == 237 |
                       site == "umbs" & julian == 209) %>%
        select(-max_temp, -min_temp, -mean_GDD_temp, -GDD_base_10, -julian)
peak_comp_plot_temps2 <- left_join(peak_comp_plot_temps, GDD_plant_comp, by=c('site','year','state'))

write.csv(peak_comp_plot_temps2, file.path(L2_dir, "plant_composition/final_plant_comp_plot_L2.csv"), row.names = F)

################ Origin ####################
origin_peak_comp_umbs <- origin_cover_total1[which(origin_cover_total1$month == "7" & origin_cover_total1$site == "umbs"),]
origin_peak_comp_kbs <- origin_cover_total1[which(origin_cover_total1$month == "8" & origin_cover_total1$site == "kbs"),]
origin_peak_comp <- full_join(origin_peak_comp_umbs, origin_peak_comp_kbs, by = c("site", "year_factor", "month", "date", "year", "plot", "state", "insecticide", "treatment_key", "origin_cover_total", "plot_cover_total", "relative", "absolute", "origin"))
write.csv(origin_peak_comp, file.path(L2_dir, "plant_composition/final_plant_comp_origin_species_L2.csv"), row.names = F)

# Average relative abundance by PLOT LEVEL
origin_relative_plot <- origin_peak_comp %>%
  group_by(plot, year, state, site, month, insecticide, treatment_key, year_factor, origin) %>%
  summarize(avg_relative_abundance = mean(relative, na.rm=T))

# Average absolute abundance by PLOT LEVEL
origin_absolute_plot <- origin_peak_comp %>%
  group_by(plot, year, state, site, month, insecticide, treatment_key, year_factor, origin) %>%
  summarize(avg_absolute_abundance = mean(absolute, na.rm=T))

origin_peak_comp_plot <- full_join(origin_relative_plot, origin_absolute_plot, by = c("site", "year_factor", "month", "year", "plot", "state", "insecticide", "treatment_key", "origin"))
write.csv(origin_peak_comp_plot, file.path(L2_dir, "plant_composition/final_plant_comp_origin_plot_L2.csv"), row.names = F)

################ Growth Habit ####################
growth_peak_comp_umbs <- growth_cover_total1[which(growth_cover_total1$month == "7" & growth_cover_total1$site == "umbs"),]
growth_peak_comp_kbs <- growth_cover_total1[which(growth_cover_total1$month == "8" & growth_cover_total1$site == "kbs"),]
growth_peak_comp <- full_join(growth_peak_comp_umbs, growth_peak_comp_kbs, by = c("site", "year_factor", "month", "date", "year", "plot", "state", "insecticide", "treatment_key", "plot_cover_total", "growth_cover_total", "relative", "absolute", "growth_habit"))
write.csv(growth_peak_comp, file.path(L2_dir, "plant_composition/final_plant_comp_growth_habit_species_L2.csv"), row.names = F)

# Average relative abundance by PLOT LEVEL
growth_relative_plot <- growth_peak_comp %>%
  group_by(plot, year, state, site, month, insecticide, treatment_key, year_factor, growth_habit) %>%
  summarize(avg_relative_abundance = mean(relative, na.rm=T))

# Average absolute abundance by PLOT LEVEL
growth_absolute_plot <- growth_peak_comp %>%
  group_by(plot, year, state, site, month, insecticide, treatment_key, year_factor, growth_habit) %>%
  summarize(avg_absolute_abundance = mean(absolute, na.rm=T))

growth_peak_comp_plot <- full_join(growth_relative_plot, growth_absolute_plot, by = c("site", "year_factor", "month", "year", "plot", "state", "insecticide", "treatment_key", "growth_habit"))
write.csv(growth_peak_comp_plot, file.path(L2_dir, "plant_composition/final_plant_comp_growth_habit_plot_L2.csv"), row.names = F)
```

