---
title: "warmXtrophic Project: Plant Composition Data Analyses"
author: "Kara Dobson, Moriah Young, Pat Bills"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=80),tidy=TRUE)
```

# Load in packages & data
```{r, message = F}
# Clear all existing data
rm(list=ls())

#Load packages
library(tidyverse)
library(ggplot2)
library(lme4)
library(lmerTest)
library(olsrr)
library(predictmeans)
library(car)
library(fitdistrplus)
library(ggpubr)
library(rstatix)
library(vegan)
library(interactions)
library(sjPlot)
library(effects)
library(glmmTMB)
library(labdsv) # used with Vegan package, the matrify() and matrify2() functions
library(agricolae) # HSD.test() function
library(emmeans)
library(scales)
library(fitdistrplus)
library(moments)# for calculating skewness of data
library(jtools) # summ() function
library(GGally) # ggpairs() function
library(bbmle) # AICtab() function

# Set working directory
Sys.getenv("L1DIR")
L0_dir <- Sys.getenv("L0DIR")
L1_dir <- Sys.getenv("L1DIR")
L2_dir <- Sys.getenv("L2DIR")
list.files(L1_dir)

# Set ggplot2 plotting
# This code for ggplot2 sets the theme to mostly black and white 
# (Arial font, and large font, base size=24)
theme_set(theme_bw(14))
theme_update(axis.text.x = element_text(size = 12, angle = 90),
             axis.text.y = element_text(size = 12))

# read in plant comp data
comp_species <- read.csv(file.path(L2_dir, "plant_composition/final_plant_comp_species_L2.csv"))
comp_plot <- read.csv(file.path(L2_dir, "plant_composition/final_plant_comp_plot_L2.csv"))

comp_species <- comp_species %>% select(-X) # get rid of "X" column that shows up
comp_plot <- comp_plot %>% select(-X) # get rid of "X" column that shows up

# Order warmed and ambient so that warmed shows up first in plotting (and is default is red = warmed; blue = ambient). First make it a factor
comp_species$state <- as.factor(comp_species$state)
levels(comp_species$state)
# [1] "ambient" "warmed" 
comp_species$state <- factor(comp_species$state, levels(comp_species$state)[c(2,1)])
levels(comp_species$state)
# [1] "warmed"  "ambient"

comp_plot$state <- as.factor(comp_plot$state)
levels(comp_plot$state)
# [1] "ambient" "warmed" 
comp_plot$state <- factor(comp_plot$state, levels(comp_plot$state)[c(2,1)])
levels(comp_plot$state)
# [1] "warmed"  "ambient"

# create dataframes for kbs and umbs - remember that these contain species within plots
comp_kbs_spp <- subset(comp_species, site == "kbs")
comp_umbs_spp <- subset(comp_species, site == "umbs")

# create dataframes for kbs and umbs
comp_kbs_plot <- subset(comp_plot, site == "kbs")
comp_umbs_plot <- subset(comp_plot, site == "umbs")

# read in meta data
meta <- read.csv(file.path(L0_dir, "plot.csv")) # dataframe above already has meta data in it
```
# Data exploration - some different ways of visualizing these data

# KBS SPECIES LEVEL - Looking at RELATIVE ABUNDANCE
```{r, warning=F}
### KBS ###
hist(comp_kbs_spp$relabun)
qqnorm(comp_kbs_spp$relabun)
shapiro.test(comp_kbs_spp$relabun) # pvalue is < 0.05 so we reject the null hypothesis that the data is normal (aka not normally distributed)

# Visualizing median Julian date for umbs at the PLOT LEVEL
ggplot(comp_kbs_spp, aes(relabun, fill = plot)) + 
        geom_histogram(binwidth = 0.5) + 
        facet_grid(year_factor ~ site, margins = TRUE, scales = "free")

ggplot(comp_kbs_spp, aes(relabun, fill = as.factor(plot))) + geom_histogram(binwidth = 0.5) + 
        facet_grid(state~year_factor, margins = TRUE, scales = "free")

ggplot(comp_kbs_spp, aes(relabun, fill = plot, color=plot)) +
        geom_density(alpha = 0.1)

#ggplot(umbs_flwr_plot, aes(julian_median, fill = plot, color=plot)) +
#        geom_density(alpha = 0.1) +
#        facet_wrap(~year)
#
#ggplot(umbs_flwr_plot, aes(julian_median, fill = plot, color=plot)) +
#        geom_density(alpha = 0.1) +
#        facet_wrap(~year + plot)

# Exploring distributions for these right-skewed data:
descdist(comp_kbs_spp$relabun, discrete = FALSE)

# Gamma distribution 
fit.gamma <- fitdist(comp_kbs_spp$relabun, "gamma")
plot(fit.gamma)

# Weibull distribution
fit.weibull  <- fitdist(comp_kbs_spp$relabun, "weibull")
plot(fit.weibull)

fit.beta <- fitdist(comp_kbs_spp$relabun, "beta")
plot(fit.beta)

# Lognormal distribution 
fit.ln <- fitdist(comp_kbs_spp$relabun, "lnorm")
plot(fit.ln)

# Exponential distribution is another option
fit.exp <- fitdist(comp_kbs_spp$relabun, "exp")
plot(fit.exp)


par(mfrow=c(2,2))
plot.legend <- c("Gamma", "Weibull", "Log Normal", "Exponential", "Beta")
denscomp(list(fit.gamma, fit.weibull, fit.ln, fit.exp, fit.beta), legendtext = plot.legend)
cdfcomp (list(fit.gamma, fit.weibull, fit.ln, fit.exp, fit.beta), legendtext = plot.legend)
qqcomp  (list(fit.gamma, fit.weibull, fit.ln, fit.exp, fit.beta), legendtext = plot.legend)
ppcomp  (list(fit.gamma, fit.weibull, fit.ln, fit.exp, fit.beta), legendtext = plot.legend)

# Goodness of fit comparisons across fits (can't include the sqrt normal bc it becomes diff response values)
gofstat(list(fit.gamma, fit.weibull, fit.ln, fit.exp, fit.beta), fitnames = c("Gamma", "Weibull", "Log Normal", "Exp", "Beta"))
# Beta distribution looks to be the best based on AIC and BIC values
```

# KBS PLOT LEVEL - Looking at PLOT AVG TOTALS 
```{r, warning=F}
### KBS ###
hist(comp_kbs_plot$plot_cover_avg)
qqnorm(comp_kbs_plot$plot_cover_avg)
shapiro.test(comp_kbs_plot$plot_cover_avg) # pvalue is < 0.05 so we reject the null hypothesis that the data is normal (aka not normally distributed)

# Visualizing median Julian date for umbs at the PLOT LEVEL
ggplot(comp_kbs_plot, aes(plot_cover_avg, fill = plot)) + 
        geom_histogram(binwidth = 0.5) + 
        facet_grid(year_factor ~ site, margins = TRUE, scales = "free")

ggplot(comp_kbs_plot, aes(plot_cover_avg, fill = as.factor(plot))) + geom_histogram(binwidth = 0.5) + 
        facet_grid(state~year_factor, margins = TRUE, scales = "free")

ggplot(comp_kbs_plot, aes(plot_cover_avg, fill = plot, color=plot)) +
        geom_density(alpha = 0.1)


#ggplot(umbs_flwr_plot, aes(julian_median, fill = plot, color=plot)) +
#        geom_density(alpha = 0.1) +
#        facet_wrap(~year)
#
#ggplot(umbs_flwr_plot, aes(julian_median, fill = plot, color=plot)) +
#        geom_density(alpha = 0.1) +
#        facet_wrap(~year + plot)

# Exploring distributions for these right-skewed data:
descdist(comp_kbs_plot$plot_cover_avg, discrete = FALSE)

# Gamma distribution 
fit.gamma <- fitdist(comp_kbs_plot$plot_cover_avg, "gamma")
plot(fit.gamma)

# Weibull distribution
fit.weibull  <- fitdist(comp_kbs_plot$plot_cover_avg, "weibull")
plot(fit.weibull)

# Lognormal distribution 
fit.ln <- fitdist(comp_kbs_plot$plot_cover_avg, "lnorm")
plot(fit.ln)

par(mfrow=c(2,2))
plot.legend <- c("Gamma", "Weibull", "Log Normal")
denscomp(list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
cdfcomp (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
qqcomp  (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
ppcomp  (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)

# Goodness of fit comparisons across fits (can't include the sqrt normal bc it becomes diff response values)
gofstat(list(fit.gamma, fit.weibull, fit.ln), fitnames = c("Gamma", "Weibull", "Log Normal"))
# Log Normal the best...
```

# Leverage plots and detecting Outliers. https://www.statmethods.net/stats/rdiagnostics.html
# These illustrate whether certain data points have more leverage (more influence), and thus could be outliers. It's a way of detecting outliers. Leverage plots can help identify whether a point has high or low influence, based on its leverage and residual and determining model fit with and without the point in question. Ultimately you decide whether the points are outliers or not, based on the knowledge of the system and how much it changes the model when included vs. excluded from the data used to fit the model. Here is a good overview of the combination of leverage and residual: scroll down to sections beginning at "13.3 Unusual Observations": https://daviddalpiaz.github.io/appliedstats/model-diagnostics.html
```{r}
# species level data
# KBS State-only model
fit_spp_state_kbs <- lm(log(relabun) ~ state, data = comp_kbs_spp)
outlierTest(fit_spp_state_kbs) # outliers - row 449
qqPlot(fit_spp_state_kbs, main="QQ Plot") 
hist(fit_spp_state_kbs$residuals)
leveragePlots(fit_spp_state_kbs)
ols_test_normality(fit_spp_state_kbs)

comp_kbs_spp_no_outliers <- comp_kbs_spp[-449,] # remove outliers
fit_spp_state_kbs1 <- lm(log(relabun) ~ state, data = comp_kbs_spp_no_outliers)
outlierTest(fit_spp_state_kbs1) # outliers - row 449
qqPlot(fit_spp_state_kbs1, main="QQ Plot") 
hist(fit_spp_state_kbs1$residuals)
leveragePlots(fit_spp_state_kbs1)
ols_test_normality(fit_spp_state_kbs1)


# KBS State and year model
fit_spp_stateyear_kbs <- lm(log(relabun) ~ state + year_factor, data = comp_kbs_spp)
outlierTest(fit_spp_stateyear_kbs) # no outliers
qqPlot(fit_spp_stateyear_kbs, main="QQ Plot") 
hist(fit_spp_stateyear_kbs$residuals)
leveragePlots(fit_spp_stateyear_kbs)
ols_test_normality(fit_spp_stateyear_kbs)

# Interaction plot (ignore for now the repeated measures with species); see: https://cran.r-project.org/web/packages/interactions/vignettes/interactions.html and: https://interactions.jacob-long.com/

fit3 <- lm(log(relabun) ~ state + year_factor + species, data = comp_kbs_spp)
interact_plot(fit3, pred = year_factor, modx = state)
fit4 <- lm(log(relabun) ~ state * year_factor + species, data = comp_kbs_spp)
interact_plot(fit4, pred = year_factor, modx = state, mod2 = species)

```
```{r}
# Plot level data
# KBS State-only model
fit_plot_state_kbs <- lm(log(plot_cover_avg) ~ state, data = comp_kbs_plot)
outlierTest(fit_plot_state_kbs) # outlier - row 252
comp_kbs_plot_no_outliers <- comp_kbs_plot[-252,] # remove outliers
qqPlot(fit_plot_state_kbs, main="QQ Plot") 
hist(fit_plot_state_kbs$residuals)
leveragePlots(fit_plot_state_kbs)
ols_test_normality(fit_plot_state_kbs)

# KBS State and year model
fit_plot_stateyear_kbs <- lm(log(plot_cover_avg) ~ state + year_factor, data = comp_kbs_plot)
outlierTest(fit_plot_stateyear_kbs) # outlier - row 190
comp_kbs_plot_no_outliers <- comp_kbs_plot[-190,] # remove outliers
qqPlot(fit_plot_stateyear_kbs, main="QQ Plot") 
hist(fit_plot_stateyear_kbs$residuals)
leveragePlots(fit_plot_stateyear_kbs) 
ols_test_normality(fit_plot_stateyear_kbs) # not all are under 0.05 but I think we can move forward with assumptions of normality here after log transformation

# Interaction plot (ignore for now the repeated measures with species); see: https://cran.r-project.org/web/packages/interactions/vignettes/interactions.html and: https://interactions.jacob-long.com/

fit3 <- lm(log(julian_median) ~ state * year_factor, data = comp_kbs_plot)
interact_plot(fit3, pred = year_factor, modx = state)
```

MIXED EFFECT MODELS

```{r}
# KBS SPECIES LEVEL - Looking at RELATIVE ABUNDANCE
mod1 <- lmer(log(relabun) ~ state*year_factor + insecticide*year_factor + (1|species) + (1|plot), comp_kbs_spp, REML = FALSE)
ggplot(comp_kbs_spp,aes(x=plot,y=relabun,col=species)) + geom_jitter() + geom_boxplot(alpha=0.2) + facet_wrap(~plot)

# Check Assumptions:
# (1) Linearity: if covariates are not categorical (year isn't)
# (2) Homogeneity: Need to Check by plotting residuals vs predicted values.
par(mfrow=c(1,2))
plot(mod1)
# Homogeneity of variance is ok here (increasing variance in resids is not increasing with fitted values)
# Check for homogeneity of variances (true if p>0.05). If the result is not significant, the assumption of equal variances (homoscedasticity) is met (no significant difference between the group variances).
# *****Levene’s Test - tests whether or not the variance among two or more groups is equal - If the p-value is less than our chosen significance level, we can reject the null hypothesis and conclude that we have enough evidence to state that the variance among the groups is not equal (which we want).

leveneTest(residuals(mod1) ~ comp_kbs_spp$state)
# Assumption met
leveneTest(residuals(mod1) ~ comp_kbs_spp$species)
# Assumption met
leveneTest(residuals(mod1) ~ comp_kbs_spp$insecticide) 
# Assumption not met - this seems weird
leveneTest(residuals(mod1) ~ comp_kbs_spp$plot)
# Assumption met

# (3) Normality of error term: need to check by histogram, QQplot of residuals, could do Kolmogorov-Smirnov test.
# Check for normal residuals
qqPlot(resid(mod1))
hist(residuals(mod1))
shapiro.test(resid(mod1)) # not normally distributed resids bc p<0.05
outlierTest(mod1) # no outliers

# (4) Normality of random effect: Get the estimate of random effect (e.g., random intercepts), and check them as you would check the residual. 
require(lme4)
r_int<- ranef(mod1)$plot$`(Intercept)`
qqnorm(r_int)
qqline(r_int)
shapiro.test(r_int) 
# Not normally distributed random effect - p-value = 0.001565
```

```{r}
# Do we need to include plot as a random effect with the UMBS models?   
mod1 <- lmer(log(relabun) ~ state*year_factor + insecticide*year_factor + (1|species) + (1|plot), comp_kbs_spp, REML = FALSE)
mod2 <- lmer(log(relabun) ~ state*year_factor + insecticide*year_factor + (1|species), comp_kbs_spp, REML=FALSE)
# Run analysis of variance on each model (see this for more explanation on how anova on a linear mixed effects model is similar to an anove on a regular linear model: https://m-clark.github.io/docs/mixedModels/anovamixed.html)
anova(mod1)
anova(mod2)

anova(mod1, mod2) # Go with model 2 since pvalue >0.05, aka more complex model does not have something in it that matters (the two models are essentially the same so go w model with fewer parameters)
summary(mod1) 
summary(mod2)
AICctab(mod1, mod2, weights=T)

# Plot the fixed effects estimates for different models
# these are the fixed effects estimates from summary(mod1)
plot_model(mod2, sort.est = TRUE)
# these are the fixed predicted values:
plot_model(mod2, type = "pred", terms = c("year_factor", "state", "insecticide"))
# these are the random effects estimates
plot_model(mod2, type = "re", terms = c("species"))

# Do we need to include insecticide? (dropping insecticide from the model)
mod3 <- lmer(log(relabun) ~ state*year_factor + (1|species), comp_kbs_spp, REML=FALSE)
anova(mod2, mod3)

AICctab(mod2, mod3, weights=T)
# Looks like no pvalue > 0.05, so insecticide does not improve model fit so we will go with mod 3

# Does year need to be interactive with insecticide? - already removed insecticide
# mod4 <- lmer(log(julian_median) ~ state*year_factor + insecticide + (1|species) + (1|plot), umbs_flwr_spp, REML=FALSE)

# Does year need to be interactive with state?
mod5 <- lmer(log(relabun) ~ state + year_factor + (1|species), comp_kbs_spp, REML=FALSE)
anova(mod3, mod5)
# No, P>0.05 so state*year_factor doesn't strongly improve model fit so we will shift to mod5, but AIC values are super close!!

# Plot the fixed effects estimates for different models
# these are the fixed effects estimates from summary(mod5)
plot_model(mod5, sort.est = TRUE)
# these are the fixed predicted values:
plot_model(mod5, type = "pred", terms = c("year_factor", "state"))
# these are the random effects estimates
plot_model(mod5, type = "re", terms = c("species"))

# If we wanted to include plots nested within year it would look like this:
mod6 <- lmer(log(relabun) ~ state + year_factor + (1|species) + (1 + year_factor|plot), comp_kbs_spp, REML=FALSE)
anova(mod5, mod6)
anova(mod5)

# mod4 (and mod5) are pretty complex in terms of interpretation (they actually don't have many parameters though). We could consider an alternative model that's simpler to understand and also one that provides more insight about the species. That would be something like this:
mod7 <- lmer(log(relabun) ~ state + species + (1+factor(year_factor)|plot), comp_kbs_spp, REML=FALSE)
mod7a <- lmer(log(relabun)  ~ state + species + factor(year_factor) + (1|plot), comp_kbs_spp, REML=FALSE)
mod7b <- lmer(log(relabun)  ~ state * factor(year_factor) + species + (1|plot), comp_kbs_spp, REML=FALSE)
mod7c <- lmer(log(relabun)  ~ state + species + factor(year_factor) + insecticide + (1|plot), comp_kbs_spp, REML=FALSE)
anova(mod5, mod7) # model 7 is a better fit to data
anova(mod7, mod7a) #mod 7a
anova(mod7a, mod7b) #mod 7a - interaction between state and year does not improve model go with simpler model (mod7a)
anova(mod7b, mod7c) #models are not different than one another, go with simpler model = mod7a
summary(mod7a)
anova(mod7a) # investigates whether at least one of the levels within each factor is significantly different.
# Yes, at least one of the species is different (they do not all have the median first flwr dates). 
emmeans(mod7c, list(pairwise ~ state + year_factor), adjust = "tukey")
emmeans(mod7c, list(pairwise ~ year_factor), adjust = "tukey")
emmeans(mod7c, list(pairwise ~ species), adjust = "tukey")
emmeans(mod7c, list(pairwise ~ insecticide), adjust = "tukey")


# Take a look at the estimates for each fixed effect. These are the estimates from summary(mod7a). You'll see that species vary a lot - and many of them are different from zero (meaning their half cover date is significantly different from zero).
plot_model(mod7a, sort.est = TRUE)
# if you want to standardize the estimates:
plot_model(mod7a, sort.est = TRUE, type="std") 
# these are the fixed predicted values: - note this is a new plot
plot_model(mod7a, type = "pred", terms = c("species", "state", "year_factor"))
# these are the random effects estimates
plot_model(mod7a, type = "re")

# including native vs. exotic
comp_kbs_spp <- within(comp_kbs_spp, origin <- relevel(factor(origin), ref = "Native")) # releveling so native is the reference
mod8 <- lmer(log(relabun) ~ state * origin + (1+factor(year_factor)|plot), comp_kbs_spp, REML=FALSE)
mod9 <- lmer(log(relabun) ~ state + origin + (1+factor(year_factor)|plot), comp_kbs_spp, REML=FALSE)
mod9a <- lmer(log(relabun) ~ state + origin + factor(year_factor) + (1|plot), comp_kbs_spp, REML=FALSE)
anova(mod8, mod9) # model 8 is a better fit to data
anova(mod8, mod9a) # mod 8
summary(mod8)
anova(mod8)
emmeans(mod8, list(pairwise ~ state * origin), adjust = "tukey")

# including growth form - first with interaction term
comp_kbs_spp <- within(comp_kbs_spp, growth_habit <- relevel(factor(growth_habit), ref = "Forb")) # releveling so forb is the reference
mod10 <- lmer(log(relabun) ~ state * growth_habit + (1+factor(year_factor)|plot), comp_kbs_spp, REML=FALSE)
mod11 <- lmer(log(relabun) ~ state + growth_habit + (1+factor(year_factor)|plot), comp_kbs_spp, REML=FALSE)
mod11a <- lmer(log(relabun) ~ state + growth_habit + factor(year_factor) + (1|plot), comp_kbs_spp, REML=FALSE)
anova(mod10, mod11) # model 10 is a better fit to data
anova(mod10, mod11a) # model 10 is still a better fit to data
summary(mod10)
anova(mod10)
emmeans(mod10, list(pairwise ~ factor(year_factor) * growth_habit), adjust = "tukey")

# New version of our model incorporating interaction term and species within year so that there is a separate intercept and slope for each species. The issue here is that there are some species that are not found each year. Easiest to remove those from another version of this dataframe before running below. Otherwise, it's not a balanced design.
# updated mod4
mod12 <- lmer(log(relabun) ~ state * factor(year_factor) + (1+factor(year_factor)|species), comp_kbs_spp)

# So another version of this model would include the interaction but not include the nesting (and thus would assume that species aren't observed ea yr)
# updated mod5
mod13 <- lmer(log(relabun) ~ state * factor(year_factor) + (1|species), comp_kbs_spp)

# All the models ran:
mod1 <- lmer(log(relabun) ~ state*year_factor + insecticide*year_factor + (1|species) + (1|plot), comp_kbs_spp, REML=FALSE)
mod2 <- lmer(log(relabun) ~ state*year_factor + insecticide*year_factor + (1|species), comp_kbs_spp, REML=FALSE)
mod3 <- lmer(log(relabun) ~ state*year_factor + (1|species), comp_kbs_spp, REML=FALSE)
mod4 <- lmer(log(relabun) ~ state*year_factor + insecticide + (1|species) + (1|plot), comp_kbs_spp, REML=FALSE)
mod5 <- lmer(log(relabun) ~ state + year_factor + (1|species), comp_kbs_spp, REML=FALSE)
mod6 <- lmer(log(relabun) ~ state + year_factor + (1|species) + (1 + year|plot), comp_kbs_spp, REML=FALSE)
mod7 <- lmer(log(relabun) ~ state + species + (1+factor(year_factor)|plot), comp_kbs_spp, REML=FALSE)
mod7a <- lmer(log(relabun) ~ state + species + factor(year_factor) + (1|plot), comp_kbs_spp, REML=FALSE)
mod7b <- lmer(log(relabun) ~ state * factor(year_factor) + species + (1|plot), comp_kbs_spp, REML=FALSE)
mod7c <- lmer(log(relabun) ~ state + species + factor(year_factor) + insecticide + (1|plot), comp_kbs_spp, REML=FALSE)
mod8 <- lmer(log(relabun) ~ state * origin + (1+factor(year_factor)|plot), comp_kbs_spp, REML=FALSE)
mod9 <- lmer(log(relabun) ~ state + origin + (1+factor(year_factor)|plot), comp_kbs_spp, REML=FALSE)
mod9a <- lmer(log(relabun) ~ state + origin + factor(year_factor) + (1|plot), comp_kbs_spp, REML=FALSE)
mod10 <- lmer(log(relabun) ~ state * growth_habit + (1+factor(year_factor)|plot), comp_kbs_spp, REML=FALSE)
mod11 <- lmer(log(relabun) ~ state + growth_habit + (1+factor(year_factor)|plot), comp_kbs_spp, REML=FALSE)
mod11a <- lmer(log(relabun) ~ state + growth_habit + factor(year_factor) + (1|plot), comp_kbs_spp, REML=FALSE)
mod12 <- lmer(log(relabun) ~ state * factor(year_factor) + (1+factor(year_factor)|species), comp_kbs_spp)
mod13 <- lmer(log(relabun) ~ state * factor(year_factor) + (1|species), comp_kbs_spp)
AICctab(mod1, mod2, mod3, mod5, mod6, mod7, mod7a, mod7b, mod7c, mod8, mod9, mod9a, mod10, mod11, mod11a, mod13, weights=T) #mod7a is the best fitting model (took out mod12 bc it wouldn't run)
summary(mod7a)
anova(mod7a)
```

# KBS Plot-level Mixed Effects Models:
```{r}
hist(comp_kbs_plot$plot_cover_avg) # skewed to the right - positive skew
qqnorm(comp_kbs_plot$plot_cover_avg) # this doesn't look good
shapiro.test(comp_kbs_plot$plot_cover_avg) # not normal

# Exploring distributions for these right-skewed data:
descdist(comp_kbs_plot$plot_cover_avg, discrete = FALSE)

mod1p <- lmer(log(plot_cover_avg) ~ state + (1|plot), comp_kbs_plot, REML=FALSE)
mod2p <- lmer(plot_cover_avg ~ insecticide + (1|plot), comp_kbs_plot, REML=FALSE)
mod3p <- lmer(plot_cover_avg ~ insecticide + state + (1|plot), comp_kbs_plot, REML=FALSE)
mod4p <- lmer(plot_cover_avg ~ insecticide * state + (1|plot), comp_kbs_plot, REML=FALSE)
mod5p <- lmer(plot_cover_avg ~ state + year_factor + (1|plot), comp_kbs_plot, REML=FALSE)
mod6p <- lmer(plot_cover_avg ~ state + year_factor + insecticide + (1|plot), comp_kbs_plot, REML=FALSE)
mod7p <- lmer(plot_cover_avg ~ state * year_factor + (1|plot), comp_kbs_plot, REML=FALSE)
mod8p <- lmer(plot_cover_avg ~ state * year_factor + insecticide + (1|plot), comp_kbs_plot, REML=FALSE)
mod9p <- lmer(plot_cover_avg ~ state * insecticide + year_factor + (1|plot), comp_kbs_plot, REML=FALSE)
mod10p <- lmer(plot_cover_avg ~ state + insecticide * year_factor + (1|plot), comp_kbs_plot, REML=FALSE)
mod11p <- lmer(plot_cover_avg ~ state * year_factor * insecticide + (1|plot), comp_kbs_plot, REML=FALSE)
AICctab(mod1p, mod2p, mod3p, mod4p, mod5p, mod6p, mod7p, mod8p, mod9p, mod10p, mod11p, weights=T) # model 11p and 10p the same
anova(mod10p, mod11p) #11p just barely better, going with 10p because is simpler

summary(mod10p)
emmeans(mod10p, list(pairwise ~ state+insecticide*year_factor), adjust = "tukey")
```








