---
title: "Plant Composition Diversity Plots"
author: "Moriah Young"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

COLLABORATORS: Phoebe Zarnetske, Mark Hammond, Pat Bills, Kara Dobson  
DATA INPUT: Plant composition diversity measures from warmX Google drive L2 folder
DATA OUTPUT: Plots of Plant composition diversity at KBS and UMBS - Rmd and PDF located in Github folder
PROJECT: warmXtrophic

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Clear all existing data
rm(list=ls())

# Load packages
library(plotrix)
library(tidyverse)
library(ggpubr)
library(hrbrthemes)

# Set working directory
#Sys.getenv("L1DIR")
L1_dir<-Sys.getenv("L1DIR")
L2_dir<-Sys.getenv("L2DIR")
#list.files(L1_dir)

# Read in data
diversity <- read.csv(file.path(L2_dir, "plant_composition/final_plant_comp_diversity_L2.csv"))

# making site names capital for cleaner plots - from Kara
change_site <- function(df){
  df$site[df$site == "umbs"] <- "UMBS"
  df$site[df$site == "kbs"] <- "KBS"
  return(df)
}

diversity <- change_site(diversity) # using function from above

# creating a vector with cleaned up insecticide labels for plotting - from Kara
insect_labels <- c("insects" = "Herbivory", "no_insects" = "Reduced Herbivory")

# averages + std error for plot level data - state only
sum_shannon <- diversity %>%
  group_by(site, state, year) %>%
  summarize(avg_shannon = mean(shannon, na.rm = TRUE),
            se = std.error(shannon, na.rm = TRUE))

sum_simpson <- diversity %>%
  group_by(site, state, year) %>%
  summarize(avg_simpson = mean(simpson, na.rm = TRUE),
            se = std.error(simpson, na.rm = TRUE))

sum_richness <- diversity %>%
  group_by(site, state, year) %>%
  summarize(avg_richness = mean(richness, na.rm = TRUE),
            se = std.error(richness, na.rm = TRUE))

# Order warmed and ambient so that warmed shows up first in plotting (and is default is red = warmed; blue = ambient). First make it a factor
sum_shannon$state <- as.factor(sum_shannon$state)
levels(sum_shannon$state)
# [1] "ambient" "warmed" 
sum_shannon$state <- factor(sum_shannon$state, levels(sum_shannon$state)[c(2,1)])
levels(sum_shannon$state)
# [1] "warmed"  "ambient"

sum_simpson$state <- as.factor(sum_simpson$state)
levels(sum_simpson$state)
# [1] "ambient" "warmed" 
sum_simpson$state <- factor(sum_simpson$state, levels(sum_simpson$state)[c(2,1)])
levels(sum_simpson$state)
# [1] "warmed"  "ambient"

sum_richness$state <- as.factor(sum_richness$state)
levels(sum_richness$state)
# [1] "ambient" "warmed" 
sum_richness$state <- factor(sum_richness$state, levels(sum_richness$state)[c(2,1)])
levels(sum_richness$state)
# [1] "warmed"  "ambient"

# Including insecticide treatment
# averages + std error for plot level data + insecticide treatment
sum_shannon_i <- diversity %>%
  group_by(site, state, insecticide, year) %>%
  summarize(avg_shannon = mean(shannon, na.rm = TRUE),
            se = std.error(shannon, na.rm = TRUE))

sum_simpson_i <- diversity %>%
  group_by(site, state, insecticide, year) %>%
  summarize(avg_simpson = mean(simpson, na.rm = TRUE),
            se = std.error(simpson, na.rm = TRUE))


sum_richness_i <- diversity %>%
  group_by(site, state, insecticide, year) %>%
  summarize(avg_richness = mean(richness, na.rm = TRUE),
            se = std.error(richness, na.rm = TRUE))

# Order warmed and ambient so that warmed shows up first in plotting (and is default is red = warmed; blue = ambient). First make it a factor
sum_shannon_i$state <- as.factor(sum_shannon_i$state)
levels(sum_shannon_i$state)
# [1] "ambient" "warmed" 
sum_shannon_i$state <- factor(sum_shannon_i$state, levels(sum_shannon_i$state)[c(2,1)])
levels(sum_shannon_i$state)
# [1] "warmed"  "ambient"

sum_simpson_i$state <- as.factor(sum_simpson_i$state)
levels(sum_simpson_i$state)
# [1] "ambient" "warmed" 
sum_simpson_i$state <- factor(sum_simpson_i$state, levels(sum_simpson_i$state)[c(2,1)])
levels(sum_simpson_i$state)
# [1] "warmed"  "ambient"

sum_richness_i$state <- as.factor(sum_richness_i$state)
levels(sum_richness_i$state)
# [1] "ambient" "warmed" 
sum_richness_i$state <- factor(sum_richness_i$state, levels(sum_richness_i$state)[c(2,1)])
levels(sum_richness_i$state)
# [1] "warmed"  "ambient"

# using function from above to change site names to uppercase
diversity <- change_site(diversity) 
sum_shannon <- change_site(sum_shannon)
sum_simpson <- change_site(sum_simpson)
sum_richness <- change_site(sum_richness)
sum_shannon_i <- change_site(sum_shannon_i)
sum_simpson_i <- change_site(sum_simpson_i)
sum_richness_i <- change_site(sum_richness_i)

# Set ggplot2 plotting
# This code for ggplot2 sets the theme to mostly black and white 
# (Arial font, and large font, base size=24)
#theme_set(theme_bw(14))
#theme_update(axis.text.x = element_text(size = 13),
#             axis.text.y = element_text(size = 16),
#             axis.title = element_text(size=16,face="bold"))
```

Averages + std error for plot level data + insecticide treatment across all years (what will be used in the paper)
```{r}
# averages + std error for plot level data - state only
sum_shannon_all <- diversity %>%
  group_by(site, state, insecticide) %>%
  summarize(avg_shannon = mean(shannon, na.rm = TRUE),
            se = std.error(shannon, na.rm = TRUE))

sum_simpson_all <- diversity %>%
  group_by(site, state, insecticide) %>%
  summarize(avg_simpson = mean(simpson, na.rm = TRUE),
            se = std.error(simpson, na.rm = TRUE))

sum_richness_all <- diversity %>%
  group_by(site, state, insecticide) %>%
  summarize(avg_richness = mean(richness, na.rm = TRUE),
            se = std.error(richness, na.rm = TRUE))
```


Richness only dot plot
```{r}
sum_richness_all$full_treat <- paste(sum_richness_all$state, sum_richness_all$insecticide, sep="_")
richness_dot_i <- function(loc) { 
       richness_plot <- subset(sum_richness_all, site == loc)
        return(ggplot(richness_plot, aes(x = state, y = avg_richness, fill=insecticide)) +
                       #geom_point(stat = "identity",position=position_dodge(0.2),size=5) +
                       geom_pointrange(aes(ymin=avg_richness-se, ymax=avg_richness+se), pch=21,size=1,position=position_dodge(0.3)) +
                       #scale_shape_manual(name="Treatment",
                       #                   values = c(1, 19),
                       #                   labels=c("Herbivory","Reduced Herbivory")) +
                       scale_fill_manual(name="Treatment",
                                         values = c("#FFB451", "#0b0055"),
                                         labels=c("Herbivory","Reduced Herbivory")) +
                               labs(x = NULL, y = NULL, title = NULL) +
                       scale_x_discrete(labels=c("ambient" = "Ambient", "warmed" = "Warmed")) +
                       coord_cartesian(ylim=c(4.0,8.5)) +
                       theme_bw(14))
}
richness_dot_i_kbs <- richness_dot_i("KBS")
richness_dot_i_kbs <- richness_dot_i_kbs + 
        theme(plot.title = element_text(size=20),
                                     axis.text.y = element_text(size=17),
                                     axis.text.x = element_text(size=17),
                                     axis.title.y = element_text(size=17),
                                     legend.text = element_text(size=17),
                                     legend.title = element_text(size=17)) +
        scale_y_continuous(breaks = c(4, 5, 6, 7, 8), 
                           labels = c("     4", "     5", "     6", "     7", "     8"))

richness_dot_i_umbs <- richness_dot_i("UMBS")
richness_dot_i_umbs <- richness_dot_i_umbs + labs(y=NULL) + 
        theme(axis.title.y=element_blank(),
              plot.title = element_text(size=20),
              axis.text.y = element_blank(),
              axis.text.x = element_text(size=17),
              legend.text = element_text(size=17),
              legend.title = element_text(size=17)) +
        scale_y_continuous(breaks = c(4, 5, 6, 7, 8), 
                           labels = c("    4", "    5", "    6", "    7", "    8"))

richness_all_dot_i <- ggpubr::ggarrange(richness_dot_i_kbs, richness_dot_i_umbs,
                                nrow = 1, ncol = 2, common.legend = T, legend="right",
                                widths = c(1, 0.9))

png("species_richness_L2_all_dot_insect.png", units="in", width=8, height=5, res=300)
annotate_figure(richness_all_dot_i,
                left = text_grob("Species Richness", color = "black", rot = 90, size=17))
dev.off()
```

Shannon and Simpsons diversity
```{r}
# Shannon diversity
sum_shannon_all$full_treat <- paste(sum_shannon_all$state, sum_shannon_all$insecticide, sep="_")
shannon_dot_i <- function(loc) { 
       shannon_plot <- subset(sum_shannon_all, site == loc)
        return(ggplot(shannon_plot, aes(x = state, y = avg_shannon, fill=insecticide)) +
                       #geom_point(stat = "identity",position=position_dodge(0.2),size=5) +
                       geom_pointrange(aes(ymin=avg_shannon-se, ymax=avg_shannon+se), pch=21,size=1,position=position_dodge(0.3)) +
                       #scale_shape_manual(name="Treatment",
                       #                   values = c(1, 19),
                       #                   labels=c("Herbivory","Reduced Herbivory")) +
                       scale_fill_manual(name="Treatment",
                                         values = c("#FFB451", "#0b0055"),
                                         labels=c("Herbivory","Reduced Herbivory")) +
                               labs(x = NULL, y = "Shannon Diversity Index", title = loc) +
                       scale_x_discrete(labels=c("ambient" = "Ambient", "warmed" = "Warmed")) +
                       coord_cartesian(ylim=c(1.00,1.355)) +
                       theme_bw(14))
}
shannon_dot_i_kbs <- shannon_dot_i("KBS")
shannon_dot_i_kbs <- shannon_dot_i_kbs + 
        annotate("text", x = 0.5, y=1.35, label = "A", size=5) +
        theme(plot.title = element_text(size=20),
              axis.text.y = element_text(size=17),
              axis.title.y = element_text(size=17),
              axis.text.x = element_blank(),
              legend.text = element_text(size=17),
              legend.title = element_text(size=17)) +
        scale_y_continuous(breaks = c(1.0, 1.1, 1.2, 1.3), 
                           labels = c("  1.0", "  1.1", "  1.2", "  1.3"))
shannon_dot_i_umbs <- shannon_dot_i("UMBS")
shannon_dot_i_umbs <- shannon_dot_i_umbs + 
        labs(y=NULL) + 
        annotate("text", x = 0.5, y=1.35, label = "B", size=5) +
        theme(axis.title.y = element_blank(),
              axis.text.x = element_blank(),
              plot.title = element_text(size=20),
              axis.text.y = element_blank(),
              legend.text = element_text(size=17),
              legend.title = element_text(size=17)) +
        scale_y_continuous(breaks = c(1.0, 1.1, 1.2, 1.3), 
                           labels = c("  1.0", "  1.1", "  1.2", "  1.3"))

# Simpson diversity
sum_simpson_all$full_treat <- paste(sum_simpson_all$state, sum_simpson_all$insecticide, sep="_")
simpson_dot_i <- function(loc) { 
       simpson_plot <- subset(sum_simpson_all, site == loc)
        return(ggplot(simpson_plot, aes(x = state, y = avg_simpson, fill=insecticide)) +
                       #geom_point(stat = "identity",position=position_dodge(0.2),size=5) +
                       geom_pointrange(aes(ymin=avg_simpson-se, ymax=avg_simpson+se), pch=21,size=1,position=position_dodge(0.3)) +
                       #scale_shape_manual(name="Treatment",
                       #                   values = c(1, 19),
                       #                   labels=c("Herbivory","Reduced Herbivory")) +
                       scale_fill_manual(name="Treatment",
                                         values = c("#FFB451", "#0b0055"),
                                         labels=c("Herbivory","Reduced Herbivory")) +
                               labs(x = NULL, y = "Simpson's Diversity Index", title = NULL) +
                       scale_x_discrete(labels=c("ambient" = "Ambient", "warmed" = "Warmed")) +
                       coord_cartesian(ylim=c(0.50,0.65)) +
                       theme_bw(14))
}
simpson_dot_i_kbs <- simpson_dot_i("KBS")
simpson_dot_i_kbs <- simpson_dot_i_kbs + 
        annotate("text", x = 0.5, y=0.65, label = "C", size=5) +
        theme(axis.text.x = element_text(size=17),
              axis.text.y = element_text(size=17),
              axis.title.y = element_text(size=17),
              legend.text = element_text(size=17),
              legend.title = element_text(size=17))

simpson_dot_i_umbs <- simpson_dot_i("UMBS")
simpson_dot_i_umbs <- simpson_dot_i_umbs + 
        labs(y=NULL) + 
        annotate("text", x = 0.5, y=0.65, label = "D", size=5) +
        theme(axis.text.x = element_text(size=17),
              axis.title.y = element_blank(),
              axis.text.y = element_blank(),
              legend.text = element_text(size=17),
              legend.title = element_text(size=17))

png("diversity_indices_L2_all_dot_insect.png", units="in", width=9, height=7, res=300)
ggpubr::ggarrange(shannon_dot_i_kbs, shannon_dot_i_umbs,
                                simpson_dot_i_kbs, simpson_dot_i_umbs,
                                nrow = 2, ncol = 2, common.legend = T, legend="right",
                  widths = c(1.1, 0.9))
dev.off()
```

# used in manuscript
Dot plot with insecticide
```{r}
# Shannon diversity
sum_shannon_all$full_treat <- paste(sum_shannon_all$state, sum_shannon_all$insecticide, sep="_")
shannon_dot_i <- function(loc) { 
       shannon_plot <- subset(sum_shannon_all, site == loc)
        return(ggplot(shannon_plot, aes(x = state, y = avg_shannon, fill=insecticide)) +
                       #geom_point(stat = "identity",position=position_dodge(0.2),size=5) +
                       geom_pointrange(aes(ymin=avg_shannon-se, ymax=avg_shannon+se), pch=21,size=1,position=position_dodge(0.3)) +
                       #scale_shape_manual(name="Treatment",
                       #                   values = c(1, 19),
                       #                   labels=c("Herbivory","Reduced Herbivory")) +
                       scale_fill_manual(name="Treatment",
                                         values = c("#FFB451", "#0b0055"),
                                         labels=c("Herbivory","Reduced Herbivory")) +
                               labs(x = NULL, y = "Shannon Diversity \n Index", title = loc) +
                       scale_x_discrete(labels=c("ambient" = "Ambient", "warmed" = "Warmed")) +
                       coord_cartesian(ylim=c(1.00,1.355)) +
                       theme_bw(14))
}
shannon_dot_i_kbs <- shannon_dot_i("KBS")
shannon_dot_i_kbs <- shannon_dot_i_kbs + 
        annotate("text", x = 0.5, y=1.35, label = "A", size=5) +
        theme(plot.title = element_text(size=20),
              axis.text.y = element_text(size=17),
              axis.title.y = element_text(size=17),
              axis.text.x = element_blank(),
              legend.text = element_text(size=17),
              legend.title = element_text(size=17)) +
        scale_y_continuous(breaks = c(1.0, 1.1, 1.2, 1.3), 
                           labels = c("  1.0", "  1.1", "  1.2", "  1.3"))
shannon_dot_i_umbs <- shannon_dot_i("UMBS")
shannon_dot_i_umbs <- shannon_dot_i_umbs + 
        labs(y=NULL) + 
        annotate("text", x = 0.5, y=1.35, label = "B", size=5) +
        theme(axis.title.y = element_blank(),
              axis.text.x = element_blank(),
              plot.title = element_text(size=20),
              axis.text.y = element_blank(),
              legend.text = element_text(size=17),
              legend.title = element_text(size=17)) +
        scale_y_continuous(breaks = c(1.0, 1.1, 1.2, 1.3), 
                           labels = c("  1.0", "  1.1", "  1.2", "  1.3"))

# Simpson diversity
sum_simpson_all$full_treat <- paste(sum_simpson_all$state, sum_simpson_all$insecticide, sep="_")
simpson_dot_i <- function(loc) { 
       simpson_plot <- subset(sum_simpson_all, site == loc)
        return(ggplot(simpson_plot, aes(x = state, y = avg_simpson, fill=insecticide)) +
                       #geom_point(stat = "identity",position=position_dodge(0.2),size=5) +
                       geom_pointrange(aes(ymin=avg_simpson-se, ymax=avg_simpson+se), pch=21,size=1,position=position_dodge(0.3)) +
                       #scale_shape_manual(name="Treatment",
                       #                   values = c(1, 19),
                       #                   labels=c("Herbivory","Reduced Herbivory")) +
                       scale_fill_manual(name="Treatment",
                                         values = c("#FFB451", "#0b0055"),
                                         labels=c("Herbivory","Reduced Herbivory")) +
                               labs(x = NULL, y = "Simpson's Diversity \n Index", title = NULL) +
                       scale_x_discrete(labels=c("ambient" = "Ambient", "warmed" = "Warmed")) +
                       coord_cartesian(ylim=c(0.50,0.65)) +
                       theme_bw(14))
}
simpson_dot_i_kbs <- simpson_dot_i("KBS")
simpson_dot_i_kbs <- simpson_dot_i_kbs + 
        annotate("text", x = 0.5, y=0.65, label = "C", size=5) +
        theme(axis.text.x = element_blank(),
              axis.text.y = element_text(size=17),
              axis.title.y = element_text(size=17),
              legend.text = element_text(size=17),
              legend.title = element_text(size=17))

simpson_dot_i_umbs <- simpson_dot_i("UMBS")
simpson_dot_i_umbs <- simpson_dot_i_umbs + 
        labs(y=NULL) + 
        annotate("text", x = 0.5, y=0.65, label = "D", size=5) +
        theme(axis.title.y = element_blank(),
              axis.text.x = element_blank(),
              axis.text.y = element_blank(),
              legend.text = element_text(size=17),
              legend.title = element_text(size=17))

# Richness
sum_richness_all$full_treat <- paste(sum_richness_all$state, sum_richness_all$insecticide, sep="_")
richness_dot_i <- function(loc) { 
       richness_plot <- subset(sum_richness_all, site == loc)
        return(ggplot(richness_plot, aes(x = state, y = avg_richness, fill=insecticide)) +
                       #geom_point(stat = "identity",position=position_dodge(0.2),size=5) +
                       geom_pointrange(aes(ymin=avg_richness-se, ymax=avg_richness+se), pch=21,size=1,position=position_dodge(0.3)) +
                       #scale_shape_manual(name="Treatment",
                       #                   values = c(1, 19),
                       #                   labels=c("Herbivory","Reduced Herbivory")) +
                       scale_fill_manual(name="Treatment",
                                         values = c("#FFB451", "#0b0055"),
                                         labels=c("Herbivory","Reduced Herbivory")) +
                               labs(x = NULL, y = "Species\nRichness", title = NULL) +
                       scale_x_discrete(labels=c("ambient" = "Ambient", "warmed" = "Warmed")) +
                       coord_cartesian(ylim=c(4.0,8.5)) +
                       theme_bw(14))
}
richness_dot_i_kbs <- richness_dot_i("KBS")
richness_dot_i_kbs <- richness_dot_i_kbs + 
        annotate("text", x = 0.5, y = 8.5, label = "E", size=5) +
        theme(axis.text.y = element_text(size=17),
              axis.text.x = element_text(size=17),
              axis.title.y = element_text(size=17),
              legend.text = element_text(size=17),
              legend.title = element_text(size=17)) +
        scale_y_continuous(breaks = c(4, 5, 6, 7, 8), 
                           labels = c("     4", "     5", "     6", "     7", "     8"))

richness_dot_i_umbs <- richness_dot_i("UMBS")
richness_dot_i_umbs <- richness_dot_i_umbs + labs(y=NULL) + 
        annotate("text", x = 0.5, y=8.5, label = "F", size=5) +
        theme(axis.title.y=element_blank(),
              axis.text.y = element_blank(),
              axis.text.x = element_text(size=17),
              legend.text = element_text(size=17),
              legend.title = element_text(size=17)) +
        scale_y_continuous(breaks = c(4, 5, 6, 7, 8), 
                           labels = c("    4", "    5", "    6", "    7", "    8"))

png("diversity_indices_L2_all_dot_insect.png", units="in", width=10, height=9, res=300)
ggpubr::ggarrange(shannon_dot_i_kbs, shannon_dot_i_umbs,
                                simpson_dot_i_kbs, simpson_dot_i_umbs,
                                richness_dot_i_kbs, richness_dot_i_umbs,
                                nrow = 3, ncol = 2, common.legend = T, legend="right",
                  widths = c(1, 0.8),
                  heights = c(1.35, 1.2, 1.3))
dev.off()
```

## used in supp
## Combining diversity indices - state only
## Line plot
```{r, echo=F}
# how to make the desired plots below:
# http://www.sthda.com/english/wiki/ggplot2-line-plot-quick-start-guide-r-software-and-data-visualization

# shannon plot
sum_shannon$year <- as.factor(sum_shannon$year)
shannon_overall_plot <- function(loc) { 
  shannon_plot <- subset(sum_shannon, site == loc)
  return(ggplot(shannon_plot, aes(x = year, y = avg_shannon, group = state)) +
                           coord_cartesian(ylim = c(0.5,1.75)) +
                 geom_errorbar(aes(ymin=avg_shannon-se, ymax=avg_shannon+se), width=.1, position=position_dodge(0.15)) +
           geom_line(aes(color=state), size = 1) +
           geom_point(aes(color=state), size = 2) +
           scale_color_manual(values = c("#a6bddb", "#fb6a4a"), labels=c("Ambient","Warmed")) +
           labs(x = NULL, y = "Shannon", title = loc, color = "Treatment"))
}

shannon_overall_kbs <- shannon_overall_plot("KBS")
shannon_overall_kbs <- shannon_overall_kbs + theme(axis.text.x=element_blank())
shannon_overall_umbs <- shannon_overall_plot("UMBS")
shannon_overall_umbs <- shannon_overall_umbs + theme(axis.text.x=element_blank(),
                                                     axis.text.y=element_blank()) + labs(y = NULL)

# simpson plot
sum_simpson$year <- as.factor(sum_simpson$year)
simpson_overall_plot <- function(loc) { 
  simpson_plot <- subset(sum_simpson, site == loc)
  return(ggplot(simpson_plot, aes(x = year, y = avg_simpson, group = state)) +
                 coord_cartesian(ylim = c(0.3, 0.8)) +
                 geom_errorbar(aes(ymin=avg_simpson-se, ymax=avg_simpson+se), width=.1, position=position_dodge(0.15)) +
           geom_line(aes(color=state), size = 1) +
           geom_point(aes(color=state), size = 2) +
           scale_color_manual(values = c("#a6bddb", "#fb6a4a"), labels=c("Ambient","Warmed")) +
           labs(x = NULL, y = "Simpson", title = loc, color = "Treatment"))
}

simpson_overall_kbs <- simpson_overall_plot("KBS")
simpson_overall_kbs <- simpson_overall_kbs + theme(axis.text.x=element_blank()) + labs(title = NULL)
simpson_overall_umbs <- simpson_overall_plot("UMBS")
simpson_overall_umbs <- simpson_overall_umbs + labs(y = NULL, title = NULL) + theme(axis.text.x=element_blank(),
                                                                                    axis.text.y=element_blank())

# richness plot
sum_richness$year <- as.factor(sum_richness$year)
richness_overall_plot <- function(loc) { 
  richness_plot <- subset(sum_richness, site == loc)
  return(ggplot(richness_plot, aes(x = year, y = avg_richness, group = state)) +
                 coord_cartesian(ylim = c(3, 10.5)) +
                 geom_errorbar(aes(ymin=avg_richness-se, ymax=avg_richness+se), width=.1, position=position_dodge(0.15)) +
           geom_line(aes(color=state), size = 1) +
           geom_point(aes(color=state), size = 2) +
           scale_color_manual(values = c("#a6bddb", "#fb6a4a"), labels=c("Ambient","Warmed")) +
           labs(x = NULL, y = "Richness", title = loc, color = "Treatment"))
}

richness_overall_kbs <- richness_overall_plot("KBS")
richness_overall_kbs <- richness_overall_kbs + labs(title = NULL)
richness_overall_umbs <- richness_overall_plot("UMBS")
richness_overall_umbs <- richness_overall_umbs + labs(y = NULL, title = NULL) + theme(axis.text.y=element_blank())

diversity_overall_line <- ggpubr::ggarrange(shannon_overall_kbs, shannon_overall_umbs,
                          simpson_overall_kbs, simpson_overall_umbs,
                          richness_overall_kbs, richness_overall_umbs,
                          nrow = 3, ncol = 2, common.legend = T, legend="bottom")
diversity_overall_line
png("plant_comp_diversity_plots_L2_all_diversity_lineplot.png", units="in", width=10, height=10, res=300)
annotate_figure(diversity_overall_line,
                bottom = text_grob("Year", color = "black"))
dev.off()
```











## Not used in manuscript or supp
## Combining diversity indices - state only
## Boxplot
```{r, echo=F}
# shannon plot
shannon_overall_plot <- function(loc) { 
  shannon_plot <- subset(diversity, site == loc)
  return(ggplot(shannon_plot, aes(x = state, y = shannon, fill=state)) +
           facet_grid(.~insecticide, labeller = as_labeller(insect_labels)) +
                 facet_grid(.~year) +
           geom_boxplot(outlier.shape=NA, alpha=0.7) +
           geom_jitter(aes(alpha=0.8, color=state, fill=state), shape=16, size=2) +
           labs(x = NULL, y = "Shannon", title = loc) +
           scale_color_manual(values = c("#a6bddb", "#fb6a4a")) +
           scale_fill_manual(values = c("#a6bddb", "#fb6a4a")) +
           scale_x_discrete(labels=c("ambient" = "A", "warmed" = "W")) +
            theme_classic())
}

shannon_overall_kbs <- shannon_overall_plot("KBS")
shannon_overall_kbs <- shannon_overall_kbs + theme(axis.text.x=element_blank())
shannon_overall_umbs <- shannon_overall_plot("UMBS")
shannon_overall_umbs <- shannon_overall_umbs + theme(axis.text.x=element_blank()) + labs(y = NULL)

# simpson plot
simpson_overall_plot <- function(loc) { 
  simpson_plot <- subset(diversity, site == loc)
  return(ggplot(simpson_plot, aes(x = state, y = simpson, fill=state)) +
           facet_grid(.~insecticide, labeller = as_labeller(insect_labels)) +
                 facet_grid(.~year) +
           geom_boxplot(outlier.shape=NA, alpha=0.7) +
           geom_jitter(aes(alpha=0.8, color=state, fill=state), shape=16, size=2) +
           labs(x = NULL, y = "Simpson", title = loc) +
           scale_color_manual(values = c("#a6bddb", "#fb6a4a")) +
           scale_fill_manual(values = c("#a6bddb", "#fb6a4a")) +
           scale_x_discrete(labels=c("ambient" = "A", "warmed" = "W")) +
           theme_classic())
}

simpson_overall_kbs <- simpson_overall_plot("KBS")
simpson_overall_kbs <- simpson_overall_kbs + theme(axis.text.x=element_blank()) + labs(title = NULL)
simpson_overall_umbs <- simpson_overall_plot("UMBS")
simpson_overall_umbs <- simpson_overall_umbs + labs(y = NULL, title = NULL) + theme(axis.text.x=element_blank())

# richness plot
richness_overall_plot <- function(loc) { 
  richness_plot <- subset(diversity, site == loc)
  return(ggplot(richness_plot, aes(x = state, y = richness, fill=state)) +
           facet_grid(.~insecticide, labeller = as_labeller(insect_labels)) +
                 facet_grid(.~year) +
           geom_boxplot(outlier.shape=NA, alpha=0.7) +
           geom_jitter(aes(alpha=0.8, color=state, fill=state), shape=16, size=2) +
           labs(x = NULL, y = "Species Richness", title = loc) +
           scale_color_manual(values = c("#a6bddb", "#fb6a4a")) +
           scale_fill_manual(values = c("#a6bddb", "#fb6a4a")) +
           scale_x_discrete(labels=c("ambient" = "A", "warmed" = "W")) +
            theme_classic())
}

richness_overall_kbs <- richness_overall_plot("KBS")
richness_overall_kbs <- richness_overall_kbs + labs(title = NULL)
richness_overall_umbs <- richness_overall_plot("UMBS")
richness_overall_umbs <- richness_overall_umbs + labs(y = NULL, title = NULL)

diversity_overall_boxplot <- ggpubr::ggarrange(shannon_overall_kbs, shannon_overall_umbs,
                          simpson_overall_kbs, simpson_overall_umbs,
                          richness_overall_kbs, richness_overall_umbs,
                          nrow = 3, ncol = 2, legend = "none")
diversity_overall_boxplot
png("plant_comp_diversity_plots_L2_all_diversity_boxplot.png", units="in", width=8, height=8, res=300)
annotate_figure(diversity_overall_boxplot,
                left = text_grob("Ecological Diversity Indices", color = "black", rot = 90),
                bottom = text_grob("Year", color = "black"))
dev.off()
```

## not in manuscript or supp
## Combining diversity indices - all treatments
## Line plot with insecticide - code from Kara
```{r}
# Shannon
sum_shannon_i$year <- as.factor(sum_shannon_i$year)
sum_shannon_i$full_treat <- paste(sum_shannon_i$state, sum_shannon_i$insecticide, sep="_")
sum_shannon_line_i <- function(loc) { 
        shannon_plot <- subset(sum_shannon_i, site == loc)
        return(ggplot(shannon_plot, aes(x = year, y = avg_shannon, group=full_treat, linetype=full_treat, color = full_treat)) +
                       geom_errorbar(aes(ymin=avg_shannon-se, ymax=avg_shannon+se), width=.1, position=position_dodge(0.15)) +
                       geom_line(size = 1) +
                       geom_point(size = 2) +
                       scale_color_manual(name="Treatment",
                                          values = c("#a6bddb", "#a6bddb", "#fb6a4a", "#fb6a4a"),
                                          labels=c("Ambient + Herbivory","Ambient + Reduced Herbivory","Warmed + Herbivory", "Warmed + Reduced Herbivory")) +
                       scale_linetype_manual(name="Treatment",
                                             values = c("solid", "dashed", "solid", "dashed"),
                                             labels=c("Ambient + Herbivory","Ambient + Reduced Herbivory","Warmed + Herbivory", "Warmed + Reduced Herbivory")) +
                       labs(x = NULL, y = "Shannon", title = loc) +
                       theme_bw(14))
}
sum_shannon_line_i_kbs <- sum_shannon_line_i("KBS")
sum_shannon_line_i_kbs <- sum_shannon_line_i_kbs + theme(axis.text.x=element_blank())
sum_shannon_line_i_umbs <- sum_shannon_line_i("UMBS")
sum_shannon_line_i_umbs <- sum_shannon_line_i_umbs + labs(y=NULL) + theme(axis.text.x=element_blank())

#Simpson
sum_simpson_i$year <- as.factor(sum_simpson_i$year)
sum_simpson_i$full_treat <- paste(sum_simpson_i$state, sum_simpson_i$insecticide, sep="_")
sum_simpson_line_i <- function(loc) { 
        simpson_plot <- subset(sum_simpson_i, site == loc)
        return(ggplot(simpson_plot, aes(x = year, y = avg_simpson, group=full_treat, linetype=full_treat, color = full_treat)) +
                       geom_errorbar(aes(ymin=avg_simpson-se, ymax=avg_simpson+se), width=.1, position=position_dodge(0.15)) +
                       geom_line(size = 1) +
                       geom_point(size = 2) +
                       scale_color_manual(name="Treatment",
                                          values = c("#a6bddb", "#a6bddb", "#fb6a4a", "#fb6a4a"),
                                          labels=c("Ambient + Herbivory","Ambient + Reduced Herbivory","Warmed + Herbivory", "Warmed + Reduced Herbivory")) +
                       scale_linetype_manual(name="Treatment",
                                             values = c("solid", "dashed", "solid", "dashed"),
                                             labels=c("Ambient + Herbivory","Ambient + Reduced Herbivory","Warmed + Herbivory", "Warmed + Reduced Herbivory")) +
                       labs(x = NULL, y = "Simpson", title = loc) +
                       theme_bw(14))
}
sum_simpson_line_i_kbs <- sum_simpson_line_i("KBS")
sum_simpson_line_i_kbs <- sum_simpson_line_i_kbs + theme(axis.text.x=element_blank()) + labs(title=NULL)
sum_simpson_line_i_umbs <- sum_simpson_line_i("UMBS")
sum_simpson_line_i_umbs <- sum_simpson_line_i_umbs + labs(y=NULL, title=NULL) + theme(axis.text.x=element_blank())

#Richness
sum_richness_i$year <- as.factor(sum_richness_i$year)
sum_richness_i$full_treat <- paste(sum_richness_i$state, sum_richness_i$insecticide, sep="_")
sum_richness_line_i <- function(loc) { 
        richness_plot <- subset(sum_richness_i, site == loc)
        return(ggplot(richness_plot, aes(x = year, y = avg_richness, group=full_treat, linetype=full_treat, color = full_treat)) +
                       geom_errorbar(aes(ymin=avg_richness-se, ymax=avg_richness+se), width=.1, position=position_dodge(0.15)) +
                       geom_line(size = 1) +
                       geom_point(size = 2) +
                       scale_color_manual(name="Treatment",
                                          values = c("#a6bddb", "#a6bddb", "#fb6a4a", "#fb6a4a"),
                                          labels=c("Ambient + Herbivory","Ambient + Reduced Herbivory","Warmed + Herbivory", "Warmed + Reduced Herbivory")) +
                       scale_linetype_manual(name="Treatment",
                                             values = c("solid", "dashed", "solid", "dashed"),
                                             labels=c("Ambient + Herbivory","Ambient + Reduced Herbivory","Warmed + Herbivory", "Warmed + Reduced Herbivory")) +
                       labs(x = NULL, y = "Species Richness", title = loc) +
                       theme_bw(14))
}
sum_richness_line_i_kbs <- sum_richness_line_i("KBS")
sum_richness_line_i_kbs <- sum_richness_line_i_kbs + labs(title=NULL)
sum_richness_line_i_umbs <- sum_richness_line_i("UMBS")
sum_richness_line_i_umbs <- sum_richness_line_i_umbs + labs(y=NULL, title=NULL)

diversity_line_i <- ggpubr::ggarrange(sum_shannon_line_i_kbs, sum_shannon_line_i_umbs,
                          sum_simpson_line_i_kbs, sum_simpson_line_i_umbs,
                          sum_richness_line_i_kbs, sum_richness_line_i_umbs,
                          nrow = 3, ncol = 2, legend = "none")
diversity_line_i
png("plant_comp_diversity_plots_L2_all_diversity_lineplot_insect.png", units="in", width=11, height=8, res=300)
annotate_figure(diversity_line_i,
                left = text_grob("Ecological Diversity Indices", color = "black", rot = 90, size=15),
                bottom = text_grob("Year", color = "black", size=15))
dev.off()
```


