---
title: "warmXtrophic Project: Plant Composition Diversity Data Analyses"
author: "Moriah Young, Pat Bills"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=80),tidy=TRUE)
```

# Load in packages & data
```{r, message = F}
# Clear all existing data
rm(list=ls())

#Load packages
library(tidyverse)
library(ggplot2)
library(lme4)
library(olsrr)
library(predictmeans)
library(car)
library(fitdistrplus)
library(ggpubr)
library(rstatix)
library(vegan)
library(interactions)
library(emmeans)
library(sjPlot)
library(effects)
library(glmmTMB)
library(labdsv) # used with Vegan package, the matrify() and matrify2() functions
library(agricolae) # HSD.test() function
library(bbmle)
library(jtools) # summ() function

# Set working directory
Sys.getenv("L1DIR")
L0_dir <- Sys.getenv("L0DIR")
L1_dir <- Sys.getenv("L1DIR")
L2_dir <- Sys.getenv("L2DIR")
list.files(L1_dir)

# read in plant comp data
comp <- read.csv(file.path(L1_dir, "plant_composition/final_plantcomp_L1.csv"))
comp <- comp %>% select(-X) # get rid of "X" column that shows up

# Remove non-plant data
comp <- comp[!(comp$species=="Bare_Ground" | 
                       comp$species=="Unknown" | 
                       comp$species=="Brown" | 
                       comp$species=="Litter" | 
                       comp$species=="Vert_Litter" | 
                       comp$species=="Animal_Disturbance"), ]

# read in meta data
meta <- read.csv(file.path(L0_dir, "plot.csv")) # dataframe above already has meta data in it
```

Function to get data into wide format in order to work in the Vegan package
```{r}
# Function to get data in wide format to work in Vegan package - taken from link below
# https://stackoverflow.com/questions/50691393/transform-community-data-into-wide-format-for-vegan-package

matrify2 <-  function(data) {
    #Data must have columns: plot, SPEC, abundance measure,Year
    if (ncol(data) != 4)
        stop("data frame must have four column format")
    plt <- factor(data[, 1])
    spc <- factor(data[, 2])
    abu <- data[, 3]
    yrs <- factor(data[, 4])
    plt.codes <- sort(levels(factor(plt)))##object with sorted plot numbers
    spc.codes <- levels(factor(spc))##object with sorted SPEC names
    yrs.codes <- sort(levels(factor(yrs)))##object with sorted sampling Years
    taxa <- matrix(0, nrow = length(plt.codes)*length(yrs.codes), ncol = length(spc.codes))##Create empty matrix with proper dimensions (unique(plotxYear) by # of SPEC)
    plt.list <- rep(plt.codes,length(yrs.codes))##Create a list of all the plot numbers (in order of input data) to add as an ID column at end of function
    yrs.list <- rep(yrs.codes,each=length(plt.codes))##Create a list of all the Year numbers (in order of input data) to add as an ID column at end of function
    col <- match(spc, spc.codes)##object that determines the alphabetical order ranking of each SPEC in the spc.code list
    row.plt <- match(plt, plt.codes)##object that determines the rank order ranking of each plot of the input data in the plt.code list
    row.yrs <- match(yrs,yrs.codes)##object that determines the rank order ranking of each Year of the input data in he #yrs.code list
    for (i in 1:length(abu)) {
        row <- (row.plt[i])+length(plt.codes)*(row.yrs[i]-1)##Determine row number by assuming each row represents a specific plot & year in an object of rep(plot,each=Year)
        if(!is.na(abu[i])) {##ONly use value if !is.na .. [ignore all is.NA values]
            taxa[row, col[i]] <- sum(taxa[row, col[i]], abu[i])##Add abundance measure of row i to the proper SPEC column and plot/Year row. Sum across all identical individuals.
        }
    }
    taxa <- data.frame(taxa)##Convert to data.frame for easier manipulation
    taxa <- cbind(yrs.list,plt.list,taxa)##Add ID columns for plot and Year to each row already representing the abundance of Each SPEC of that given plot/Year.
    names(taxa) <- c('Year','Plot',spc.codes)
    taxa
}
```

Calculating Shannon and Simpsons Diversity and Species Richness
```{r}
#diversity_by_year <- function(comp, site, div_index = "shannon"){
# subset comp data by site
comp_kbs <- subset(comp, site == "kbs") %>% dplyr::select(plot, species, cover, year)
comp_umbs <- subset(comp, site == "umbs") %>% dplyr::select(plot, species, cover, year)


# convert the abundance (cover) data to wide format for each species in columns for the vegan package
#kbs
comp_kbs$cover <- as.numeric(comp_kbs$cover) # change cover data to numeric
comp_wide_kbs <- matrify2(comp_kbs) # use matrify2 function
#umbs
comp_umbs$cover <- as.numeric(comp_umbs$cover) # change cover data to numeric
comp_wide_umbs <- matrify2(comp_umbs) # use matrify2 function

# comp_wide_data assumes to have columns Year, Plot, and columns for each species found, e.g. for Vegan

# first, split up the wide data into a list of years.  Each list item is a year of data
comp_wide_by_year_kbs <- dplyr::group_by(comp_wide_kbs, Year) %>% dplyr::group_split()
comp_wide_by_year_umbs <- dplyr::group_by(comp_wide_umbs, Year) %>% dplyr::group_split()

# we need to add plot names.  Get those Plot names by taking a column from any one of the years
# since we are assuming the Plot column is the exact same across years and IN THE SAME ORDER
# Moriah - this might be a problem bc I know at kbs some data wasn't taken for one of plots in later years due to a groundhog hole in the plot
plot_names <- comp_wide_by_year_kbs[[1]]$Plot
plot_names <- comp_wide_by_year_umbs[[1]]$Plot

# remove the plot and year columns from each item in the list so that Vegan will work.
# This assumes row order is the exact same for all years (each row a plot)
comp_wide_by_year_kbs <- lapply(comp_wide_by_year_kbs, dplyr::select, c(-Year, -Plot))
comp_wide_by_year_umbs <- lapply(comp_wide_by_year_umbs, dplyr::select, c(-Year, -Plot))

#  apply the diversity function to each year  - in this case the main index is plot, each year considered separately
shannon_by_year_list_kbs <- lapply(comp_wide_by_year_kbs,vegan::diversity, index = "shannon")
shannon_by_year_list_umbs <- lapply(comp_wide_by_year_umbs,vegan::diversity, index = "shannon")

simpson_by_year_list_kbs <- lapply(comp_wide_by_year_kbs,vegan::diversity, index = "simpson")
simpson_by_year_list_umbs <- lapply(comp_wide_by_year_umbs,vegan::diversity, index = "simpson")

richness_by_year_list_kbs <- lapply(comp_wide_by_year_kbs,vegan::specnumber) # species richness
richness_by_year_list_umbs <- lapply(comp_wide_by_year_umbs,vegan::specnumber) # species richness

# each item in the list is a year of diversity, so name those with the years we know we have
names(shannon_by_year_list_kbs) <- as.character(2015:2021)
names(shannon_by_year_list_umbs) <- as.character(2015:2021)
names(simpson_by_year_list_kbs) <- as.character(2015:2021)
names(simpson_by_year_list_umbs) <- as.character(2015:2021)
names(richness_by_year_list_kbs) <- as.character(2015:2021)
names(richness_by_year_list_umbs) <- as.character(2015:2021)

# "unlist" and create a new data frame, each year a column, each row a plot, and add a new row with the plot names
shannon_kbs <- do.call(cbind,shannon_by_year_list_kbs) %>% cbind(Plot = plot_names ) %>% as.data.frame()
shannon_umbs <- do.call(cbind,shannon_by_year_list_umbs) %>% cbind(Plot = plot_names ) %>% as.data.frame()
simpson_kbs <- do.call(cbind,simpson_by_year_list_kbs) %>% cbind(Plot = plot_names ) %>% as.data.frame()
simpson_umbs <- do.call(cbind,simpson_by_year_list_umbs) %>% cbind(Plot = plot_names ) %>% as.data.frame()
richness_kbs <- do.call(cbind,richness_by_year_list_kbs) %>% cbind(Plot = plot_names ) %>% as.data.frame()
richness_umbs <- do.call(cbind,richness_by_year_list_umbs) %>% cbind(Plot = plot_names ) %>% as.data.frame()
#  an alternative tidyverse way x<- diversity_by_year(diversity_by_year_list)

## optional step!
shannon_kbs
shannon_umbs
simpson_kbs
simpson_umbs
richness_kbs
richness_umbs

# this output has a column for each year 2015, 2016, and Plot, but if you need it narrow use 'melt' from reshape2: 
library(reshape2)

#calculate shannon diversity
shannon_by_plot_year_kbs <- reshape2::melt(shannon_kbs, id =  "Plot", variable.name = c("Year"), value.name = "shannon")
shannon_by_plot_year_kbs$site <- "kbs" # adding site column
shannon_by_plot_year_umbs <- reshape2::melt(shannon_umbs, id =  "Plot", variable.name = c("Year"), value.name = "shannon") 
shannon_by_plot_year_umbs$site <- "umbs" # adding site column

#calculate simpson diversity
simpson_by_plot_year_kbs <- reshape2::melt(simpson_kbs, id =  "Plot", variable.name = c("Year"), value.name = "simpson") 
simpson_by_plot_year_kbs$site <- "kbs" # adding site column
simpson_by_plot_year_umbs <- reshape2::melt(simpson_umbs, id =  "Plot", variable.name = c("Year"), value.name = "simpson") 
simpson_by_plot_year_umbs$site <- "umbs" # adding site column

#calculate species richness
richness_by_plot_year_kbs <- reshape2::melt(richness_kbs, id =  "Plot", variable.name = c("Year"), value.name = "richness") 
richness_by_plot_year_kbs$site <- "kbs" # adding site column
richness_by_plot_year_umbs <- reshape2::melt(richness_umbs, id =  "Plot", variable.name = c("Year"), value.name = "richness")
richness_by_plot_year_umbs$site <- "umbs" # adding site column

#combine umbs and kbs shannon diversity measures into 1 dataframe
shannon_diversity <- full_join(shannon_by_plot_year_kbs, shannon_by_plot_year_umbs, by = c("Plot", "Year", "shannon", "site"))

#combine umbs and kbs simpson diversity measures into 1 dataframe
simpson_diversity <- full_join(simpson_by_plot_year_kbs, simpson_by_plot_year_umbs, by = c("Plot", "Year", "simpson", "site"))

#combine umbs and kbs richness measures into 1 dataframe
richness <- full_join(richness_by_plot_year_kbs, richness_by_plot_year_umbs, by = c("Plot", "Year", "richness", "site"))

# combine simpson and shannon diversity data frames into 1
comp_diversity <- full_join(simpson_diversity, shannon_diversity, by = c("Plot", "Year", "site"))
# Looks like diversity and simpson diveristy measures are the same?? Need to look into this
comp_diversity <- full_join(comp_diversity, richness, by = c("Plot", "Year", "site"))

names(comp_diversity) <- tolower(names(comp_diversity)) # column names to lower case so I can combine this data frame with the meta data

# merge meta data with comp_diversity
comp_diversity <- full_join(comp_diversity, meta, by = "plot")

comp_diversity$simpson <- as.numeric(comp_diversity$simpson)
comp_diversity$shannon <- as.numeric(comp_diversity$shannon)
comp_diversity$richness <- as.numeric(comp_diversity$richness)

# adding sequential year variable starting at 1: this is because the years (e.g. 2015, 2016, etc) are large numbers compared with other values in the dataset. We can always label axes with these real years.
comp_diversity$year_factor[comp_diversity$year == 2015] <- 1
comp_diversity$year_factor[comp_diversity$year == 2016] <- 2
comp_diversity$year_factor[comp_diversity$year == 2017] <- 3
comp_diversity$year_factor[comp_diversity$year == 2018] <- 4
comp_diversity$year_factor[comp_diversity$year == 2019] <- 5
comp_diversity$year_factor[comp_diversity$year == 2020] <- 6
comp_diversity$year_factor[comp_diversity$year == 2021] <- 7

comp_diversity <- comp_diversity[,c("site", "plot", "year", "year_factor", "treatment_key", "state", "insecticide", "simpson", "shannon", "richness")] #reorder columns

# write a new csv with diversity indices and upload to the shared google drive L2 data folder
write.csv(comp_diversity, file.path(L2_dir, "plant_composition/final_plant_comp_diversity_L2.csv"))

# create separate data frames for kbs and umbs sites
kbs_diversity <- subset(comp_diversity, site == "kbs")
kbs_diversity <- kbs_diversity[-167,] # remove this row with zero values for shannon diversity and species richness - this zero is from there not being any data in plot D5 in 2021 due to a large ground hog hole
umbs_diversity <- subset(comp_diversity, site == "umbs")
```

Simpson's Index
KBS
```{r, warning=F}
### KBS ###
hist(kbs_diversity$simpson) # skewed to the left
qqnorm(kbs_diversity$simpson)
shapiro.test(kbs_diversity$simpson) # pvalue is < 0.05 so we reject the null hypothesis that the data is normal (aka not normally distributed)

# Visualizing plot average totals for kbs at the PLOT LEVEL
ggplot(kbs_diversity, aes(simpson, fill = plot)) + 
        geom_histogram(binwidth = 0.5) + 
        facet_grid(year_factor ~ site, margins = TRUE, scales = "free")

ggplot(kbs_diversity, aes(simpson, fill = as.factor(plot))) + geom_histogram(binwidth = 0.5) + 
        facet_grid(state~year_factor, margins = TRUE, scales = "free")

ggplot(kbs_diversity, aes(simpson, fill = plot, color=plot)) +
        geom_density(alpha = 0.1)

ggplot(kbs_diversity, aes(simpson, fill = plot, color=plot)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year_factor)

ggplot(kbs_diversity, aes(simpson, fill = plot, color=plot)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year_factor + plot)

# Exploring distributions for these right-skewed data:
descdist(kbs_diversity$simpson, discrete = FALSE)

# Gamma distribution 
fit.gamma <- fitdist(kbs_diversity$simpson, "gamma")
plot(fit.gamma)

# Weibull distribution
fit.weibull  <- fitdist(kbs_diversity$simpson, "weibull")
plot(fit.weibull)

# Lognormal distribution 
fit.ln <- fitdist(kbs_diversity$simpson, "lnorm")
plot(fit.ln)

# Exponential distribution is another option
fit.exp <- fitdist(kbs_diversity$simpson, "exp")
plot(fit.exp)

par(mfrow=c(2,2))
plot.legend <- c("Gamma", "Weibull", "Log Normal", "Exponential")
denscomp(list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)
cdfcomp (list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)
qqcomp  (list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)
ppcomp  (list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)

# Goodness of fit comparisons across fits
gofstat(list(fit.gamma, fit.weibull, fit.ln, fit.exp), fitnames = c("Gamma", "Weibull", "Log Normal", "Exp"))
# log normal distribution looks to be the best based on AIC and BIC values
```
UMBS
```{r, warning=F}
### UMBS ###
hist(umbs_diversity$simpson) # skewed to the left
qqnorm(umbs_diversity$simpson)
shapiro.test(umbs_diversity$simpson) # pvalue is < 0.05 so we reject the null hypothesis that the data is normal (aka not normally distributed)

# Visualizing plot average totals for umbs at the PLOT LEVEL
ggplot(umbs_diversity, aes(simpson, fill = plot)) + 
        geom_histogram(binwidth = 0.5) + 
        facet_grid(year_factor ~ site, margins = TRUE, scales = "free")

ggplot(umbs_diversity, aes(simpson, fill = as.factor(plot))) + geom_histogram(binwidth = 0.5) + 
       facet_grid(state~year_factor, margins = TRUE, scales = "free")

ggplot(umbs_diversity, aes(simpson, fill = plot, color=plot)) +
        geom_density(alpha = 0.1)

ggplot(umbs_diversity, aes(simpson, fill = plot, color=plot)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year_factor)

ggplot(umbs_diversity, aes(simpson, fill = plot, color=plot)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year_factor + plot)

# Exploring distributions for these right-skewed data:
descdist(umbs_diversity$simpson, discrete = FALSE)

# Gamma distribution 
fit.gamma <- fitdist(umbs_diversity$simpson, "gamma")
plot(fit.gamma)

# Weibull distribution
fit.weibull  <- fitdist(umbs_diversity$simpson, "weibull")
plot(fit.weibull)

# Lognormal distribution 
fit.ln <- fitdist(umbs_diversity$simpson, "lnorm")
plot(fit.ln)

# Exponential distribution is another option
fit.exp <- fitdist(umbs_diversity$simpson, "exp")
plot(fit.exp)

par(mfrow=c(2,2))
plot.legend <- c("Gamma", "Weibull", "Log Normal", "Exponential")
denscomp(list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)
cdfcomp (list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)
qqcomp  (list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)
ppcomp  (list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)

# Goodness of fit comparisons across fits
gofstat(list(fit.gamma, fit.weibull, fit.ln, fit.exp), fitnames = c("Gamma", "Weibull", "Log Normal", "Exp"))
# log normal distribution looks to be the best based on AIC and BIC values
```

Shannon Index
KBS
```{r, warning=F}
### KBS ###
hist(kbs_diversity$shannon) # skewed to the left
qqnorm(kbs_diversity$shannon)
shapiro.test(kbs_diversity$shannon) # pvalue is < 0.05 so we reject the null hypothesis that the data is normal (aka not normally distributed)

# Visualizing plot average totals for kbs at the PLOT LEVEL
ggplot(kbs_diversity, aes(shannon, fill = plot)) + 
        geom_histogram(binwidth = 0.5) + 
        facet_grid(year_factor ~ site, margins = TRUE, scales = "free")

ggplot(kbs_diversity, aes(shannon, fill = as.factor(plot))) + geom_histogram(binwidth = 0.5) + 
        facet_grid(state~year_factor, margins = TRUE, scales = "free")

ggplot(kbs_diversity, aes(shannon, fill = plot, color=plot)) +
        geom_density(alpha = 0.1)

ggplot(kbs_diversity, aes(shannon, fill = plot, color=plot)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year_factor)

ggplot(kbs_diversity, aes(shannon, fill = plot, color=plot)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year_factor + plot)

# Exploring distributions for these right-skewed data:
descdist(kbs_diversity$shannon, discrete = FALSE)

# Gamma distribution 
fit.gamma <- fitdist(kbs_diversity$shannon, "gamma")
plot(fit.gamma)

# Weibull distribution
fit.weibull  <- fitdist(kbs_diversity$shannon, "weibull")
plot(fit.weibull)

# Lognormal distribution 
fit.ln <- fitdist(kbs_diversity$shannon, "lnorm")
plot(fit.ln)

# Exponential distribution is another option
fit.exp <- fitdist(kbs_diversity$shannon, "exp")
plot(fit.exp)

par(mfrow=c(2,2))
plot.legend <- c("Gamma", "Weibull", "Log Normal", "Exponential")
denscomp(list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)
cdfcomp (list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)
qqcomp  (list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)
ppcomp  (list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)

# Goodness of fit comparisons across fits
gofstat(list(fit.gamma, fit.weibull, fit.ln, fit.exp), fitnames = c("Gamma", "Weibull", "Log Normal", "Exp"))
# weibull distribution looks to be the best based on AIC and BIC values
```

UMBS
```{r, warning=F}
### UMBS ###
hist(umbs_diversity$shannon) 
qqnorm(umbs_diversity$shannon)
shapiro.test(umbs_diversity$shannon) # pvalue is > 0.05 so we do not reject the null hypothesis that the data is normal (aka normally distributed)

# Visualizing plot average totals for umbs at the PLOT LEVEL
ggplot(umbs_diversity, aes(shannon, fill = plot)) + 
        geom_histogram(binwidth = 0.5) + 
        facet_grid(year_factor ~ site, margins = TRUE, scales = "free")

ggplot(umbs_diversity, aes(shannon, fill = as.factor(plot))) + geom_histogram(binwidth = 0.5) + 
       facet_grid(state~year_factor, margins = TRUE, scales = "free")

ggplot(umbs_diversity, aes(shannon, fill = plot, color=plot)) +
        geom_density(alpha = 0.1)

ggplot(umbs_diversity, aes(shannon, fill = plot, color=plot)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year_factor)

ggplot(umbs_diversity, aes(shannon, fill = plot, color=plot)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year_factor + plot)

# Exploring distributions for these right-skewed data:
descdist(umbs_diversity$shannon, discrete = FALSE)

# Gamma distribution 
fit.gamma <- fitdist(umbs_diversity$shannon, "gamma")
plot(fit.gamma)

# Weibull distribution
fit.weibull  <- fitdist(umbs_diversity$shannon, "weibull")
plot(fit.weibull)

# Lognormal distribution 
fit.ln <- fitdist(umbs_diversity$shannon, "lnorm")
plot(fit.ln)

# Exponential distribution is another option
fit.exp <- fitdist(umbs_diversity$shannon, "exp")
plot(fit.exp)

par(mfrow=c(2,2))
plot.legend <- c("Gamma", "Weibull", "Log Normal", "Exponential")
denscomp(list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)
cdfcomp (list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)
qqcomp  (list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)
ppcomp  (list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)

# Goodness of fit comparisons across fits
gofstat(list(fit.gamma, fit.weibull, fit.ln, fit.exp), fitnames = c("Gamma", "Weibull", "Log Normal", "Exp"))
# weibull best distributions based on AIC and BIC values
```

Species Richness
KBS
```{r, warning=F}
### KBS ###
hist(kbs_diversity$richness) # looks pretty good
qqnorm(kbs_diversity$richness)
shapiro.test(kbs_diversity$richness) # pvalue is < 0.05 so we reject the null hypothesis that the data is normal (aka not normally distributed)

# Visualizing plot average totals for kbs at the PLOT LEVEL
ggplot(kbs_diversity, aes(richness, fill = plot)) + 
        geom_histogram(binwidth = 0.5) + 
        facet_grid(year_factor ~ site, margins = TRUE, scales = "free")

ggplot(kbs_diversity, aes(richness, fill = as.factor(plot))) + geom_histogram(binwidth = 0.5) + 
        facet_grid(state~year_factor, margins = TRUE, scales = "free")

ggplot(kbs_diversity, aes(richness, fill = plot, color=plot)) +
        geom_density(alpha = 0.1)

ggplot(kbs_diversity, aes(richness, fill = plot, color=plot)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year_factor)

ggplot(kbs_diversity, aes(richness, fill = plot, color=plot)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year_factor + plot)

# Exploring distributions for these right-skewed data:
descdist(kbs_diversity$richness, discrete = FALSE)

# Gamma distribution 
fit.gamma <- fitdist(kbs_diversity$richness, "gamma")
plot(fit.gamma)

# Weibull distribution
fit.weibull  <- fitdist(kbs_diversity$richness, "weibull")
plot(fit.weibull)

# Lognormal distribution 
fit.ln <- fitdist(kbs_diversity$richness, "lnorm")
plot(fit.ln)

# Exponential distribution is another option
fit.exp <- fitdist(kbs_diversity$richness, "exp")
plot(fit.exp)

par(mfrow=c(2,2))
plot.legend <- c("Gamma", "Weibull", "Log Normal", "Exponential")
denscomp(list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)
cdfcomp (list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)
qqcomp  (list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)
ppcomp  (list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)

# Goodness of fit comparisons across fits
gofstat(list(fit.gamma, fit.weibull, fit.ln, fit.exp), fitnames = c("Gamma", "Weibull", "Log Normal", "Exp"))
# weibull distribution looks to be the best based on AIC and BIC values
```

UMBS
```{r, warning=F}
### UMBS ###
hist(umbs_diversity$richness) # skewed to the right
qqnorm(umbs_diversity$richness)
shapiro.test(umbs_diversity$richness) # pvalue is < 0.05 so we reject the null hypothesis that the data is normal (aka not normally distributed)

# Visualizing plot average totals for umbs at the PLOT LEVEL
ggplot(umbs_diversity, aes(richness, fill = plot)) + 
        geom_histogram(binwidth = 0.5) + 
        facet_grid(year_factor ~ site, margins = TRUE, scales = "free")

ggplot(umbs_diversity, aes(richness, fill = as.factor(plot))) + geom_histogram(binwidth = 0.5) + 
       facet_grid(state~year_factor, margins = TRUE, scales = "free")

ggplot(umbs_diversity, aes(richness, fill = plot, color=plot)) +
        geom_density(alpha = 0.1)

ggplot(umbs_diversity, aes(richness, fill = plot, color=plot)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year_factor)

ggplot(umbs_diversity, aes(richness, fill = plot, color=plot)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year_factor + plot)

# Exploring distributions for these right-skewed data:
descdist(umbs_diversity$richness, discrete = FALSE)

# Gamma distribution 
fit.gamma <- fitdist(umbs_diversity$richness, "gamma")
plot(fit.gamma)

# Weibull distribution
fit.weibull  <- fitdist(umbs_diversity$richness, "weibull")
plot(fit.weibull)

# Lognormal distribution 
fit.ln <- fitdist(umbs_diversity$richness, "lnorm")
plot(fit.ln)

# Exponential distribution is another option
fit.exp <- fitdist(umbs_diversity$richness, "exp")
plot(fit.exp)

par(mfrow=c(2,2))
plot.legend <- c("Gamma", "Weibull", "Log Normal", "Exponential")
denscomp(list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)
cdfcomp (list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)
qqcomp  (list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)
ppcomp  (list(fit.gamma, fit.weibull, fit.ln, fit.exp), legendtext = plot.legend)

# Goodness of fit comparisons across fits
gofstat(list(fit.gamma, fit.weibull, fit.ln, fit.exp), fitnames = c("Gamma", "Weibull", "Log Normal", "Exp"))
# log normal distribution looks to be the best based on AIC and BIC values
```

# Leverage plots and detecting Outliers. https://www.statmethods.net/stats/rdiagnostics.html
# These illustrate whether certain data points have more leverage (more influence), and thus could be outliers. It's a way of detecting outliers. Leverage plots can help identify whether a point has high or low influence, based on its leverage and residual and determining model fit with and without the point in question. Ultimately you decide whether the points are outliers or not, based on the knowledge of the system and how much it changes the model when included vs. excluded from the data used to fit the model. Here is a good overview of the combination of leverage and residual: scroll down to sections beginning at "13.3 Unusual Observations": https://daviddalpiaz.github.io/appliedstats/model-diagnostics.html

SIMPSON
```{r}
# KBS State-only model
fit_state_kbs <- lm(log(simpson) ~ state, data = kbs_diversity)
outlierTest(fit_state_kbs) # yes row 148, 145
qqPlot(fit_state_kbs, main="QQ Plot") 
hist(fit_state_kbs$residuals)
leveragePlots(fit_state_kbs)
ols_test_normality(fit_state_kbs)

# KBS State and year model
fit_stateyear_kbs <- lm(log(simpson) ~ state + year, data = kbs_diversity)
outlierTest(fit_stateyear_kbs) # yes, row 148
qqPlot(fit_stateyear_kbs, main="QQ Plot") 
hist(fit_stateyear_kbs$residuals)
leveragePlots(fit_stateyear_kbs)
ols_test_normality(fit_stateyear_kbs)

# Interaction plot (ignore for now the repeated measures with species); see: https://cran.r-project.org/web/packages/interactions/vignettes/interactions.html and: https://interactions.jacob-long.com/

fit3 <- lm(log(simpson) ~ state + year_factor, data = kbs_diversity)
interact_plot(fit3, pred = year_factor, modx = state)

```
```{r}
# UMBS State-only model
fit_state_umbs <- lm(log(simpson) ~ state, data = umbs_diversity)
outlierTest(fit_state_umbs) # yes, row 202
qqPlot(fit_state_umbs, main="QQ Plot") 
hist(fit_state_umbs$residuals)
leveragePlots(fit_state_umbs)
ols_test_normality(fit_state_umbs)

# UMBS State and year model
fit_stateyear_umbs <- lm(log(simpson) ~ state + year, data = umbs_diversity)
outlierTest(fit_stateyear_kbs) # row 48
qqPlot(fit_stateyear_umbs, main="QQ Plot") 
hist(fit_stateyear_umbs$residuals)
leveragePlots(fit_stateyear_umbs)
ols_test_normality(fit_stateyear_umbs)

# Interaction plot (ignore for now the repeated measures with species); see: https://cran.r-project.org/web/packages/interactions/vignettes/interactions.html and: https://interactions.jacob-long.com/

# I can't get these to work
fit3 <- lm(log(simpson) ~ state + year, data = umbs_diversity)
#interact_plot(fit3, pred = year_factor, modx = state)
```

SHANNON
```{r}
# KBS State-only model
fit_state_kbs_shannon <- lm(log(shannon) ~ state, data = kbs_diversity)
outlierTest(fit_state_kbs_shannon) # yes row 148
qqPlot(fit_state_kbs_shannon, main="QQ Plot") 
hist(fit_state_kbs_shannon$residuals)
leveragePlots(fit_state_kbs_shannon)
ols_test_normality(fit_state_kbs_shannon)

# KBS State and year model
fit_stateyear_kbs_shannon <- lm(log(shannon) ~ state + year, data = kbs_diversity)
outlierTest(fit_stateyear_kbs_shannon) # no outliers
qqPlot(fit_stateyear_kbs_shannon, main="QQ Plot") 
hist(fit_stateyear_kbs_shannon$residuals)
leveragePlots(fit_stateyear_kbs_shannon)
ols_test_normality(fit_stateyear_kbs_shannon)

# Interaction plot (ignore for now the repeated measures with species); see: https://cran.r-project.org/web/packages/interactions/vignettes/interactions.html and: https://interactions.jacob-long.com/

# I can't get these to work
fit3 <- lm(log(shannon) ~ state + year_factor, data = kbs_diversity)
interact_plot(fit3, pred = year_factor, modx = state)
```

```{r}
# UMBS State-only model
fit_state_umbs_shannon <- lm(shannon ~ state, data = umbs_diversity)
outlierTest(fit_state_umbs_shannon) # no outliers
qqPlot(fit_state_umbs_shannon, main="QQ Plot") 
hist(fit_state_umbs_shannon$residuals)
leveragePlots(fit_state_umbs_shannon)
ols_test_normality(fit_state_umbs_shannon)

# UMBS State and year model
fit_stateyear_umbs_shannon <- lm(shannon ~ state + year, data = umbs_diversity)
outlierTest(fit_stateyear_umbs_shannon) # no outliers
qqPlot(fit_stateyear_umbs_shannon, main="QQ Plot") 
hist(fit_stateyear_umbs_shannon$residuals)
leveragePlots(fit_stateyear_umbs_shannon)
ols_test_normality(fit_stateyear_umbs_shannon)

# Interaction plot (ignore for now the repeated measures with species); see: https://cran.r-project.org/web/packages/interactions/vignettes/interactions.html and: https://interactions.jacob-long.com/

# I can't get these to work
fit3 <- lm(shannon ~ state + year_factor, data = umbs_diversity)
interact_plot(fit3, pred = year_factor, modx = state)
```
RICHNESS
```{r}
# KBS State-only model
fit_state_kbs_richness <- lm(log(richness) ~ state, data = kbs_diversity)
outlierTest(fit_state_kbs_richness) # yes row 67
qqPlot(fit_state_kbs_richness, main="QQ Plot") 
hist(fit_state_kbs_richness$residuals)
leveragePlots(fit_state_kbs_richness)
ols_test_normality(fit_state_kbs_richness)

# KBS State and year model
fit_stateyear_kbs_richness <- lm(log(richness) ~ state + year, data = kbs_diversity)
outlierTest(fit_stateyear_kbs_richness) # no outliers
qqPlot(fit_stateyear_kbs_richness, main="QQ Plot") 
hist(fit_stateyear_kbs_richness$residuals)
leveragePlots(fit_stateyear_kbs_richness)
ols_test_normality(fit_stateyear_kbs_richness)

# Interaction plot (ignore for now the repeated measures with species); see: https://cran.r-project.org/web/packages/interactions/vignettes/interactions.html and: https://interactions.jacob-long.com/

# I can't get these to work
fit3 <- lm(log(shannon) ~ state + year_factor, data = kbs_diversity)
interact_plot(fit3, pred = year_factor, modx = state)
```

```{r}
# UMBS State-only model
fit_state_umbs_richness <- lm(log(richness) ~ state, data = umbs_diversity)
outlierTest(fit_state_umbs_richness) # no outliers
qqPlot(fit_state_umbs_richness, main="QQ Plot") 
hist(fit_state_umbs_richness$residuals)
leveragePlots(fit_state_umbs_richness)
ols_test_normality(fit_state_umbs_richness)

# UMBS State and year model
fit_stateyear_umbs_richness <- lm(log(richness) ~ state + year, data = umbs_diversity)
outlierTest(fit_stateyear_umbs_richness) # no outliers
qqPlot(fit_stateyear_umbs_richness, main="QQ Plot") 
hist(fit_stateyear_umbs_richness$residuals)
leveragePlots(fit_stateyear_umbs_richness)
ols_test_normality(fit_stateyear_umbs_richness)

# Interaction plot (ignore for now the repeated measures with species); see: https://cran.r-project.org/web/packages/interactions/vignettes/interactions.html and: https://interactions.jacob-long.com/

# I can't get these to work
fit3 <- lm(log(richness) ~ state + year_factor, data = umbs_diversity)
interact_plot(fit3, pred = year_factor, modx = state)
```

MIXED EFFECT MODELS
SIMPSON
KBS 
```{r}
mod1 <- lmer(log(simpson) ~ state*year + insecticide*year + (1|plot), kbs_diversity, REML = FALSE)

# Check Assumptions:
# (1) Linearity: if covariates are not categorical (year isn't)
# (2) Homogeneity: Need to Check by plotting residuals vs predicted values.
par(mfrow=c(1,2))
plot(mod1)
# Homogeneity of variance is ok here (increasing variance in resids is not increasing with fitted values)
# Check for homogeneity of variances (true if p>0.05). If the result is not significant, the assumption of equal variances (homoscedasticity) is met (no significant difference between the group variances).
# *****Levene’s Test - tests whether or not the variance among two or more groups is equal - If the p-value is less than our chosen significance level, we can reject the null hypothesis and conclude that we have enough evidence to state that the variance among the groups is not equal (which we want).

leveneTest(residuals(mod1) ~ kbs_diversity$state)
# Assumption not met
leveneTest(residuals(mod1) ~ kbs_diversity$insecticide) 
# Assumption met 
leveneTest(residuals(mod1) ~ kbs_diversity$plot)
# Assumption not met

# (3) Normality of error term: need to check by histogram, QQplot of residuals, could do Kolmogorov-Smirnov test.
# Check for normal residuals
qqPlot(resid(mod1))
hist(residuals(mod1))
shapiro.test(resid(mod1)) # not normally distributed resids bc p<0.05
outlierTest(mod1) # row 148

# (4) Normality of random effect: Get the estimate of random effect (e.g., random intercepts), and check them as you would check the residual. 
require(lme4)
r_int<- ranef(mod1)$plot$`(Intercept)`
qqnorm(r_int)
qqline(r_int)
shapiro.test(r_int) 
# Normally distributed random effect pvalue > 0.05
```

```{r}
# Do we need to include plot as a random effect with the KBS models?   
mod1 <- lmer(log(simpson) ~ state*year + insecticide*year + (1|plot), kbs_diversity, REML = FALSE)
mod2 <- lmer(log(simpson) ~ state*year + insecticide + year + (1|plot), kbs_diversity, REML=FALSE)
# Run analysis of variance on each model (see this for more explanation on how anova on a linear mixed effects model is similar to an anove on a regular linear model: https://m-clark.github.io/docs/mixedModels/anovamixed.html)
anova(mod1)
anova(mod2)

anova(mod1, mod2) # Go with model 1 since pvalue <0.05, aka more complex model does have something in it that matters
summary(mod1)
summary(mod2)
AICctab(mod1, mod2, weights=T) # model 1

# Plot the fixed effects estimates for different models
# these are the fixed effects estimates from summary(mod1)
plot_model(mod1, sort.est = TRUE)
# these are the fixed predicted values:
plot_model(mod1, type = "pred", terms = c("year", "state", "insecticide"))
# these are the random effects estimates
plot_model(mod1, type = "re", terms = c("species"))

# Does year need to be interactive with state?
mod3 <- lmer(log(simpson) ~ state + year + insecticide*year + (1|plot), kbs_diversity, REML = FALSE)
anova(mod2, mod3)

AICctab(mod1, mod3, weights=T) # going with mod3

# Do we need to include insecticide? (dropping insecticide from the model)
mod5 <- lmer(log(simpson) ~ state + year + (1|plot), kbs_diversity, REML = FALSE)
anova(mod3, mod5)
# Yes, p<0.05 so insecticide*year does strongly improve model fit so we will stick with the more complex mod3

# Plot the fixed effects estimates for different models
# these are the fixed effects estimates from summary(mod5)
plot_model(mod3, sort.est = TRUE)
# these are the fixed predicted values:
plot_model(mod3, type = "pred", terms = c("year", "state"))
# these are the random effects estimates
plot_model(mod3, type = "re", terms = c("species"))

# the best model fit appears to be = 
# mod3 <- lmer(log(simpson) ~ state + year + insecticide*year + (1|plot), kbs_diversity, REML = FALSE)
summ(mod3)
emmeans(mod3, list(pairwise ~ state + year + insecticide*year), adjust = "tukey")
```

UMBS
```{r}
mod1u <- lmer(log(simpson) ~ state*year + insecticide*year + (1|plot), umbs_diversity, REML = FALSE)

# Check Assumptions:
# (1) Linearity: if covariates are not categorical (year isn't)
# (2) Homogeneity: Need to Check by plotting residuals vs predicted values.
par(mfrow=c(1,2))
plot(mod1u)
# Homogeneity of variance is ok here (increasing variance in resids is not increasing with fitted values)
# Check for homogeneity of variances (true if p>0.05). If the result is not significant, the assumption of equal variances (homoscedasticity) is met (no significant difference between the group variances).
# *****Levene’s Test - tests whether or not the variance among two or more groups is equal - If the p-value is less than our chosen significance level, we can reject the null hypothesis and conclude that we have enough evidence to state that the variance among the groups is not equal (which we want).

leveneTest(residuals(mod1u) ~ umbs_diversity$state)
# Assumption met
leveneTest(residuals(mod1u) ~ umbs_diversity$insecticide) 
# Assumption not met 
leveneTest(residuals(mod1u) ~ umbs_diversity$plot)
# Assumption not met

# (3) Normality of error term: need to check by histogram, QQplot of residuals, could do Kolmogorov-Smirnov test.
# Check for normal residuals
qqPlot(resid(mod1u))
hist(residuals(mod1u))
shapiro.test(resid(mod1u)) # not normally distributed resids bc p<0.05
outlierTest(mod1u) # row 202

# (4) Normality of random effect: Get the estimate of random effect (e.g., random intercepts), and check them as you would check the residual. 
require(lme4)
r_int_u <- ranef(mod1u)$plot$`(Intercept)`
qqnorm(r_int_u)
qqline(r_int_u)
shapiro.test(r_int_u) 
# Normally distributed random effect pvalue > 0.05
```
UMBS
```{r}
# Do we need to include plot as a random effect with the UMBS models?   
mod1u <- lmer(log(simpson) ~ state*year + insecticide*year + (1|plot), umbs_diversity, REML = FALSE)
mod2u <- lmer(log(simpson) ~ state*year + insecticide + year + (1|plot), umbs_diversity, REML=FALSE)
# Run analysis of variance on each model (see this for more explanation on how anova on a linear mixed effects model is similar to an anove on a regular linear model: https://m-clark.github.io/docs/mixedModels/anovamixed.html)
anova(mod1u)
anova(mod2u)

anova(mod1u, mod2u) # Go with model 2u since pvalue >0.05, aka more complex model does not have something in it that matters
summary(mod1u)
summary(mod2u)
AICctab(mod1u, mod2u, weights=T) # model 2u

# Plot the fixed effects estimates for different models
# these are the fixed effects estimates from summary(mod1)
plot_model(mod2u, sort.est = TRUE)
# these are the fixed predicted values:
plot_model(mod2u, type = "pred", terms = c("year", "state", "insecticide"))
# these are the random effects estimates
plot_model(mod2u, type = "re", terms = c("species"))

# Does year need to be interactive with state?
mod3u <- lmer(log(simpson) ~ state + insecticide + year + (1|plot), umbs_diversity, REML = FALSE)
anova(mod2u, mod3u)
AICctab(mod1u, mod3u, weights=T) # going with mod3u

# Do we need to include insecticide? (dropping insecticide from the model)
mod5u <- lmer(log(simpson) ~ state + year + (1|plot), umbs_diversity, REML = FALSE)
anova(mod3u, mod5u)
# No, p>0.05 so insecticide*year doesn't strongly improve model fit so we will go with the more simple model mod5u

# Plot the fixed effects estimates for different models
# these are the fixed effects estimates from summary(mod5u)
plot_model(mod5u, sort.est = TRUE)
# these are the fixed predicted values:
plot_model(mod5u, type = "pred", terms = c("year", "state"))
# these are the random effects estimates
plot_model(mod5u, type = "re", terms = c("species"))

# If we wanted to include plots nested within year it would look like this:
# mod6 <- lmer(log(simpson) ~ state + year + insecticide*year + (1 + year|plot), kbs_diversity, REML=FALSE)
# anova(mod5, mod6)
# anova(mod5)
# cant get mod6 to work

# the best model fit appears to be = 
# mod5u <- lmer(log(simpson) ~ state + year + (1|plot), umbs_diversity, REML = FALSE)
summ(mod5u)
emmeans(mod5u, list(pairwise ~ state + year), adjust = "tukey")
```

SHANNON
KBS 
```{r}
# Do we need to include plot as a random effect with the KBS models?   
mod1ks <- lmer(log(shannon) ~ state*year + insecticide*year + (1|plot), kbs_diversity, REML = FALSE)
mod2ks <- lmer(log(shannon) ~ state*year + insecticide + year + (1|plot), kbs_diversity, REML=FALSE)
# Run analysis of variance on each model (see this for more explanation on how anova on a linear mixed effects model is similar to an anova on a regular linear model: https://m-clark.github.io/docs/mixedModels/anovamixed.html)
anova(mod1ks)
anova(mod2ks)

anova(mod1ks, mod2ks) # Go with model 1 since pvalue <0.05, aka more complex model does have something in it that matters
summary(mod1ks)
summary(mod2ks)
AICctab(mod1ks, mod2ks, weights=T) # model 1

# Plot the fixed effects estimates for different models
# these are the fixed effects estimates from summary(mod1)
plot_model(mod1ks, sort.est = TRUE)
# these are the fixed predicted values:
plot_model(mod1ks, type = "pred", terms = c("year", "state", "insecticide"))
# these are the random effects estimates
plot_model(mod1ks, type = "re", terms = c("species"))

# Does year need to be interactive with state?
mod3ks <- lmer(log(shannon) ~ state + year + insecticide*year + (1|plot), kbs_diversity, REML = FALSE)
anova(mod2ks, mod3ks)

AICctab(mod1ks, mod3ks, weights=T) # going with mod3

# Do we need to include insecticide? (dropping insecticide from the model)
mod5ks <- lmer(log(shannon) ~ state + year + (1|plot), kbs_diversity, REML = FALSE)
anova(mod3ks, mod5ks)
# Yes, p<0.05 so insecticide*year does strongly improve model fit so we will stick with the more complex mod3

# Plot the fixed effects estimates for different models
# these are the fixed effects estimates from summary(mod5)
plot_model(mod3ks, sort.est = TRUE)
# these are the fixed predicted values:
plot_model(mod3ks, type = "pred", terms = c("year", "state"))
# these are the random effects estimates
plot_model(mod3ks, type = "re", terms = c("species"))

# If we wanted to include plots nested within year it would look like this:
# mod6ks <- lmer(log(shannon) ~ state + year + insecticide*year + (1 + year|plot), kbs_diversity, REML=FALSE)
# anova(mod5ks, mod6ks)
# anova(mod5ks)
# cant get mod6 to work

# the best model fit appears to be = 
# mod3ks <- lmer(log(shannon) ~ state + year + insecticide*year + (1|plot), kbs_diversity, REML = FALSE)
summ(mod3ks)
emmeans(mod3ks, list(pairwise ~ state + year + insecticide*year), adjust = "tukey")
```

UMBS
```{r}
# Do we need to include plot as a random effect with the UMBS models?   
mod1us <- lmer(shannon ~ state*year + insecticide*year + (1|plot), umbs_diversity, REML = FALSE)
mod2us <- lmer(shannon ~ state*year + insecticide + year + (1|plot), umbs_diversity, REML=FALSE)
# Run analysis of variance on each model (see this for more explanation on how anova on a linear mixed effects model is similar to an anove on a regular linear model: https://m-clark.github.io/docs/mixedModels/anovamixed.html)
anova(mod1us)
anova(mod2us)
anova(mod1us, mod2us) # Go with model 2 since pvalue >0.05, aka more complex model does not have something in it that matters
summary(mod1us)
summary(mod2us)
AICctab(mod1us, mod2us, weights=T) # model 1

# Plot the fixed effects estimates for different models
# these are the fixed effects estimates from summary(mod1)
plot_model(mod2us, sort.est = TRUE)
# these are the fixed predicted values:
plot_model(mod2us, type = "pred", terms = c("year", "state", "insecticide"))
# these are the random effects estimates
plot_model(mod2us, type = "re", terms = c("species"))

# Does year need to be interactive with state?
mod3us <- lmer(shannon ~ state + year + insecticide + (1|plot), umbs_diversity, REML = FALSE)
anova(mod2us, mod3us)

AICctab(mod1us, mod3us, weights=T) # going with mod3

# Do we need to include insecticide? (dropping insecticide from the model)
mod5us <- lmer(shannon ~ state + year + (1|plot), umbs_diversity, REML = FALSE)
anova(mod3us, mod5us)
# No p>0.05 so insecticide does not strongly improve model fit so we will go with model 5

# Plot the fixed effects estimates for different models
# these are the fixed effects estimates from summary(mod5)
plot_model(mod5us, sort.est = TRUE)
# these are the fixed predicted values:
plot_model(mod5us, type = "pred", terms = c("year", "state"))
# these are the random effects estimates
plot_model(mod5us, type = "re", terms = c("species"))

# If we wanted to include plots nested within year it would look like this:
# mod6us <- lmer(log(shannon) ~ state + year + insecticide*year + (1 + year|plot), umbs_diversity, REML=FALSE)
# anova(mod5us, mod6us)
# anova(mod5us)
# cant get mod6 to work

# the best model fit appears to be = 
# mod5us <- lmer(shannon ~ state + year + (1|plot), umbs_diversity, REML = FALSE)
summ(mod5us)
emmeans(mod5us, list(pairwise ~ state + year), adjust = "tukey")
```

RICHNESS
KBS 
```{r}
# Do we need to include plot as a random effect with the UMBS models?   
mod1kr <- lmer(log(richness) ~ state*year + insecticide*year + (1|plot), kbs_diversity, REML = FALSE)
mod2kr <- lmer(log(richness) ~ state*year + insecticide + year + (1|plot), kbs_diversity, REML=FALSE)
# Run analysis of variance on each model (see this for more explanation on how anova on a linear mixed effects model is similar to an anove on a regular linear model: https://m-clark.github.io/docs/mixedModels/anovamixed.html)
anova(mod1kr)
anova(mod2kr)
anova(mod1kr, mod2kr) # Go with model 1 since pvalue <0.05, aka more complex model does have something in it that matters
summary(mod1kr)
summary(mod2kr)
AICctab(mod1kr, mod2kr, weights=T) # model 1

# Plot the fixed effects estimates for different models
# these are the fixed effects estimates from summary(mod1)
plot_model(mod1kr, sort.est = TRUE)
# these are the fixed predicted values:
plot_model(mod1kr, type = "pred", terms = c("year", "state", "insecticide"))
# these are the random effects estimates
plot_model(mod1kr, type = "re", terms = c("species"))

# Does year need to be interactive with state?
mod3kr <- lmer(log(richness) ~ state + year + insecticide*year + (1|plot), kbs_diversity, REML = FALSE)
anova(mod1kr, mod3kr) # go with less complex model
AICctab(mod1kr, mod3kr, weights=T) # going with mod3

# Do we need to include insecticide? (dropping insecticide from the model)
mod5kr <- lmer(log(richness) ~ state + year + (1|plot), kbs_diversity, REML = FALSE)
anova(mod3kr, mod5kr)
# Yes, p<0.05 so insecticide*year does strongly improve model fit so we will stick with the more complex mod3

# Plot the fixed effects estimates for different models
# these are the fixed effects estimates from summary(mod5)
plot_model(mod3kr, sort.est = TRUE)
# these are the fixed predicted values:
plot_model(mod3kr, type = "pred", terms = c("year", "state"))
# these are the random effects estimates
plot_model(mod3kr, type = "re", terms = c("species"))

# If we wanted to include plots nested within year it would look like this:
# mod6ks <- lmer(log(richness) ~ state + year + insecticide*year + (1 + year|plot), kbs_diversity, REML=FALSE)
# anova(mod5kr, mod6kr)
# anova(mod5kr)
# cant get mod6 to work

# the best model fit appears to be = 
# mod3kr <- lmer(log(richness) ~ state + year + insecticide*year + (1|plot), kbs_diversity, REML = FALSE)
summ(mod3kr)
emmeans(mod3kr, list(pairwise ~ state + year + insecticide*year), adjust = "tukey")
```

UMBS 
```{r}
# Do we need to include plot as a random effect with the UMBS models?   
mod1ur <- lmer(log(richness) ~ state*year + insecticide*year + (1|plot), umbs_diversity, REML = FALSE)
mod2ur <- lmer(log(richness) ~ state*year + insecticide + year + (1|plot), umbs_diversity, REML=FALSE)
# Run analysis of variance on each model (see this for more explanation on how anova on a linear mixed effects model is similar to an anove on a regular linear model: https://m-clark.github.io/docs/mixedModels/anovamixed.html)
anova(mod1ur)
anova(mod2ur)
anova(mod1ur, mod2ur) # Go with model 2 since pvalue >0.05, aka more complex model does not have something in it that matters
summary(mod1ur)
summary(mod2ur)
AICctab(mod1ur, mod2ur, weights=T) # model 2

# Plot the fixed effects estimates for different models
# these are the fixed effects estimates from summary(mod1)
plot_model(mod2ur, sort.est = TRUE)
# these are the fixed predicted values:
plot_model(mod2ur, type = "pred", terms = c("year", "state", "insecticide"))
# these are the random effects estimates
plot_model(mod2ur, type = "re", terms = c("species"))

# Does year need to be interactive with state?
mod3ur <- lmer(log(richness) ~ state + year + insecticide*year + (1|plot), umbs_diversity, REML = FALSE)
anova(mod2ur, mod3ur)
AICctab(mod1ur, mod3ur, weights=T) # going with mod3

# Do we need to include insecticide? (dropping insecticide from the model)
mod5ur <- lmer(log(richness) ~ state + year + (1|plot), umbs_diversity, REML = FALSE)
anova(mod3ur, mod5ur)
# p>0.05 so insecticide*year does not strongly improve model fit so we will go with mod5

# Plot the fixed effects estimates for different models
# these are the fixed effects estimates from summary(mod5)
plot_model(mod5ur, sort.est = TRUE)
# these are the fixed predicted values:
plot_model(mod5ur, type = "pred", terms = c("year", "state"))
# these are the random effects estimates
plot_model(mod5ur, type = "re", terms = c("species"))

# If we wanted to include plots nested within year it would look like this:
# mod6us <- lmer(log(richness) ~ state + year + insecticide*year + (1 + year|plot), umbs_diversity, REML=FALSE)
# anova(mod5ur, mod6ur)
# anova(mod5ur)
# cant get mod6 to work

# the best model fit appears to be = 
# mod5ur <- lmer(log(richness) ~ state + year + (1|plot), umbs_diversity, REML = FALSE)
summ(mod5ur)
emmeans(mod5ur, list(pairwise ~ state + year), adjust = "tukey")
```





Code below is a function written by Pat but unsuccessfully subsets sites so you get the same values for both kbs and umbs - above is a clumsy fix by Moriah (no function)
```{r, eval=FALSE}
#' function to calculate annual diversity index for a specific site
#'
#' after reading a comp file, this function should do all that's needed to prep it and
#' run the diversity function on for each year.  diversity indexes are for the year only,
#' the diversity indexes use total abundances for a year,  do not sum/count/pool abundances in other years
#'
#' @param comp plant composition data as read from project folder
#' @param site one of kbs or umbs as coded in the comp data
#' @param div_index is the same as 'index' for vegan::diversity function "shannon", "simpson" or "invsimpson".
#'
#' @returns a matrix (data frame) of diversity indices for one site with years in the columns, and plot in rows
#'
diversity_by_year <- function(comp, site, div_index = "shannon"){
    comp_site <- subset(comp, site == site) %>% dplyr::select(plot, species, cover, year)

    # remove non-species using "not in"
    #obs_to_exclude = c("Bare_Ground", "Unknown","Brown","Litter", "Vert_Litter", "Animal_Disturbance")
    #comp_site <-dplyr::filter(comp_site, !(species %in% obs_to_exclude))

    # convert the abundance data to abundance for each species in columns for the vegan package
    comp_wide <- matrify2(comp_site)

    # comp_wide_data is assumes to have columns Year, Plot, and columns for each species found, e.g. for Vegan

    # first, split up the wide data into a list of years.  Each list item is a year of data
    comp_wide_by_year <- dplyr::group_by(comp_wide, Year) %>% dplyr::group_split()

    # we need to add plot names.  Get those Plot names by taking a column from any one of the years
    # since we are assuming the Plot column is the exact same across years and IN THE SAME ORDER
    plot_names <- comp_wide_by_year[[1]]$Plot

    # remove the plot and year columns from each item in the list so that Vegan will work.
    # This assumes row order is the exact same for all years (each row a plot)
    comp_wide_by_year<- lapply(comp_wide_by_year, dplyr::select, c(-Year, -Plot))

    #  apply  the diversity function to each year  - in this case the main index is plot, each year considered separately
    diversity_by_year_list<- lapply(comp_wide_by_year,vegan::diversity, index = div_index)

    # each item in the list is a year of diversity, so name those with the years we know we have
    names(diversity_by_year_list) <- as.character(2015:2021)

    # "unlist" and create a new data frame, each year a column, each row a plot, and add a new row with the plot names
    x<- do.call(cbind,diversity_by_year_list) %>% cbind(Plot = plot_names ) %>% as.data.frame()
    #  an alternative tidyverse way x<- diversity_by_year(diversity_by_year_list)

    ## optional step!
    return(x)
}

comp$cover <- as.numeric(comp$cover)

# use the one function above to both matrify and calculate Shannon diversity index per year
diversity_by_year_kbs  <- diversity_by_year(comp, site = "kbs", div_index = "shannon")
diversity_by_year_umbs <- diversity_by_year(comp, site = "umbs", div_index = "shannon")

# this output has a column for each year 2015, 2016, and Plot, but if you need it narrow use 'melt' from reshape2: 
library(reshape2)
diversity_by_plot_year_kbs <- reshape2::melt(diversity_by_year_kbs, id =  "Plot", variable.name = c("Year"), value.name = "shannon") 
diversity_by_plot_year_umbs <- reshape2::melt(diversity_by_year_umbs, id =  "Plot", variable.name = c("Year"), value.name = "shannon") 

# To do just August (peak_comp):

peak_comp <- dplyr::filter(comp, month == 8)

peak_shannon_by_year_kbs  <- diversity_by_year(peak_comp, site = "kbs", div_index = "shannon")
peak_shannon_by_year_umbs  <- diversity_by_year(peak_comp, site = "umbs", div_index = "shannon")

peak_simpson_by_year_kbs <- diversity_by_year(peak_comp, site = "kbs", div_index = "simpson")
peak_simpson_by_year_umbs <- diversity_by_year(peak_comp, site = "umbs", div_index = "simpson")

# this output has a column for each year 2015, 2016, and Plot, but if you need it narrow use 'melt' from reshape2: 
peak_shannon_by_plot_year_kbs <- reshape2::melt(peak_shannon_by_year_kbs, id =  "Plot", variable.name = c("Year"), value.name = "shannon") 
peak_shannon_by_plot_year_umbs <- reshape2::melt(peak_shannon_by_year_umbs, id =  "Plot", variable.name = c("Year"), value.name = "shannon") 

# this output has a column for each year 2015, 2016, and Plot, but if you need it narrow use 'melt' from reshape2: 
peak_simpson_by_plot_year_kbs <- reshape2::melt(peak_simpson_by_year_kbs, id =  "Plot", variable.name = c("Year"), value.name = "simpson") 
peak_simpson_by_plot_year_umbs <- reshape2::melt(peak_simpson_by_year_umbs, id =  "Plot", variable.name = c("Year"), value.name = "simpson") 

diversity_kbs <- left_join(peak_shannon_by_plot_year_kbs, peak_simpson_by_plot_year_kbs)
diversity_kbs$site <- "kbs" #add site column

diversity_umbs <- left_join(peak_shannon_by_plot_year_umbs, peak_simpson_by_plot_year_umbs)
diversity_umbs$site <- "umbs" #add site column
```
