---
title: "Plant Comp Plots"
author: "Moriah Young"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

COLLABORATORS: Kara Dobson, Phoebe Zarnetske, Mark Hammond, Pat Bills 
DATA INPUT: Clean & plot plant comp csv from the shared Google drive  
DATA OUTPUT: Code and Rmd are in the scripts folder in Github  
PROJECT: warmXtrophic 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=F}
# Clear all existing data
rm(list=ls())

# Load packages
library(tidyverse)
library(plotrix)
library(ggpubr)

# Set working directory
Sys.getenv("L0DIR")
L0_dir <- Sys.getenv("L0DIR")
L1_dir <- Sys.getenv("L1DIR")
L2_dir <- Sys.getenv("L2DIR")

# Read in plant comp data and meta-data - plot level
comp <- read.csv(file.path(L2_dir, "plant_composition/final_plant_comp_plot_L2.csv"))
growth <- read.csv(file.path(L2_dir, "plant_composition/final_plant_comp_growth_habit_plot_L2.csv"))
origin <- read.csv(file.path(L2_dir, "plant_composition/final_plant_comp_origin_plot_L2.csv"))

# species level
comp_spp <- read.csv(file.path(L2_dir, "plant_composition/final_plant_comp_species_L2.csv"))
```

Set ggplot2 plotting
```{r}
# This code for ggplot2 sets the theme to mostly black and white 
# (Arial font, and large font, base size=24)
#theme_set(theme_bw(14))
#theme_update(axis.text.x = element_text(size = 12, angle = 90),
#             axis.text.y = element_text(size = 12))

# making site names capital for cleaner plots
change_site <- function(df){
        df$site[df$site == "umbs"] <- "UMBS"
        df$site[df$site == "kbs"] <- "KBS"
        return(df)
}
comp <- change_site(comp)
growth <- change_site(growth)
origin <- change_site(origin)
comp_spp <- change_site(comp_spp)

# creating a vector with cleaned up insecticide labels for plotting
insect_labels <- c("insects" = "Herbivory", "no_insects" = "Reduced Herbivory")
```

Figures in main manuscript
```{r}
# filter the data to contain the mean and std error for percent cover of each site
pc_avg <- comp %>%
  group_by(site, state, insecticide, year) %>%
  summarize(avg_comp = mean(avg_pc, na.rm = TRUE),
            se = std.error(avg_pc, na.rm = TRUE))

# filter the data to contain the mean and std error for percent cover of each site for origin
origin_pc_avg <- origin %>%
  group_by(site, state, insecticide, origin, year) %>%
  summarize(avg_pc = mean(avg_origin_pc, na.rm = TRUE),
            se = std.error(avg_origin_pc, na.rm = TRUE))

# filter the data to contain the mean and std error for percent cover of each site for growth habit
growth_pc_avg <- growth %>%
  group_by(site, state, insecticide, growth_habit, year) %>%
  summarize(avg_pc = mean(avg_growth_pc, na.rm = TRUE),
            se = std.error(avg_growth_pc, na.rm = TRUE))

# adding empty values for 2021 to umbs so the x-axis matches the other plots
pc_avg$year <- as.factor(pc_avg$year)
pc_avg$full_treat <- paste(pc_avg$state, pc_avg$insecticide, sep="_")
#pc_avg2 <- pc_avg
#de<-data.frame("UMBS", NA, NA, "2021", NA, NA, NA)
#names(de)<-c("site","state", "insecticide", "year", "avg_comp", "se", "full_treat")
#de$year <- as.factor(de$year)
#pc_avg <- rbind(pc_avg2, de)

# origin
origin_pc_avg$year <- as.factor(origin_pc_avg$year)
origin_pc_avg$full_treat <- paste(origin_pc_avg$state, origin_pc_avg$insecticide, sep="_")
#origin_pc_avg2 <- origin_pc_avg
#de<-data.frame("UMBS", NA, NA, "Native", "2021", NA, NA, NA)
#de2<-data.frame("UMBS", NA, NA, "Exotic", "2021", NA, NA, NA)
#names(de)<-c("site","state", "insecticide", "origin", "year", "avg_pc", "se", "full_treat")
#names(de2)<-c("site","state", "insecticide", "origin", "year", "avg_pc", "se", "full_treat")
#de$year <- as.factor(de$year)
#de2$year <- as.factor(de2$year)
#origin_pc_avg <- rbind(origin_pc_avg2, de, de2)

# growth habit
growth_pc_avg$year <- as.factor(growth_pc_avg$year)
growth_pc_avg$full_treat <- paste(growth_pc_avg$state, growth_pc_avg$insecticide, sep="_")
#growth_pc_avg2 <- growth_pc_avg
#de<-data.frame("UMBS", NA, NA, "Graminoid", "2021", NA, NA, NA)
#de2<-data.frame("UMBS", NA, NA, "Forb", "2021", NA, NA, NA)
#names(de)<-c("site","state", "insecticide", "growth_habit", "year", "avg_pc", "se", "full_treat")
#names(de2)<-c("site","state", "insecticide", "growth_habit", "year", "avg_pc", "se", "full_treat")
#de$year <- as.factor(de$year)
#de2$year <- as.factor(de2$year)
#growth_pc_avg <- rbind(growth_pc_avg2, de, de2)
```

General Plot Level - all years
```{r}
plot_line <- function(loc) { 
  all_plot <- subset(pc_avg, site == loc)
  return(ggplot(all_plot, aes(x = year, y = avg_comp, group=full_treat, linetype=full_treat, fill=full_treat,color = full_treat, shape=full_treat)) +
                 geom_line(size = 1) +
                 geom_pointrange(aes(ymin=avg_comp-se, ymax=avg_comp+se), linetype="solid",pch=21,size=0.6) +
                 scale_color_manual(name="Treatment",
                           values = c("#a6bddb", "#a6bddb", "#AE1F00", "#AE1F00"),
                           labels=c("Ambient + Herbivory","Ambient + Reduced Herbivory","Warmed + Herbivory", "Warmed + Reduced Herbivory")) +
                 scale_fill_manual(name="Treatment",
                                   values = c("#a6bddb", "#a6bddb", "#AE1F00", "#AE1F00"),
                                   labels=c("Ambient + Herbivory","Ambient + Reduced Herbivory","Warmed + Herbivory", "Warmed + Reduced Herbivory")) +
                 scale_linetype_manual(name="Treatment",
                                       values = c("solid", "dotted", "solid", "dotted"),
                                       labels=c("Ambient + Herbivory","Ambient + Reduced Herbivory","Warmed + Herbivory", "Warmed + Reduced Herbivory")) +
                 labs(x = NULL, y = "Cover (%)", title=loc, color="Treatment") +
                 #ylim(0, 30) +
                 theme_bw())
}

plot_line_kbs <- plot_line("KBS")
plot_line_kbs <- plot_line_kbs + labs(title="KBS", y=NULL) + theme(axis.text.x=element_blank())
plot_line_umbs <- plot_line("UMBS")
plot_line_umbs <- plot_line_umbs + labs(title="UMBS", y=NULL)

plot_line <- ggpubr::ggarrange(plot_line_kbs, plot_line_umbs,
                          nrow = 2, ncol = 1, common.legend = T, legend="right")
png("avg_percent_cover_plots_by_year_L2.png", units="in", width=9, height=8, res=300)
annotate_figure(plot_line,
                left = text_grob("Cover (%)", color = "black", rot = 90, size=15),
                bottom = text_grob("Year", color = "black", size=15))
dev.off()
```

Supplemental Figures

Origin Plot Level - all years
```{r}
native_pc_avg <- origin_pc_avg %>% filter(origin == "Native")

plot_line_n <- function(loc) { 
  all_plot <- subset(native_pc_avg, site == loc)
  return(ggplot(all_plot, aes(x = year, y = avg_pc, group=full_treat, linetype=full_treat, fill=full_treat,color = full_treat, shape=full_treat)) +
                 geom_line(size = 1) +
                 geom_pointrange(aes(ymin=avg_pc-se, ymax=avg_pc+se), linetype="solid",pch=21,size=0.6) +
                 scale_color_manual(name="Treatment",
                           values = c("#a6bddb", "#a6bddb", "#AE1F00", "#AE1F00"),
                           labels=c("Ambient + Herbivory","Ambient + Reduced Herbivory","Warmed + Herbivory", "Warmed + Reduced Herbivory")) +
                 scale_fill_manual(name="Treatment",
                                   values = c("#a6bddb", "#a6bddb", "#AE1F00", "#AE1F00"),
                                   labels=c("Ambient + Herbivory","Ambient + Reduced Herbivory","Warmed + Herbivory", "Warmed + Reduced Herbivory")) +
                 scale_linetype_manual(name="Treatment",
                                       values = c("solid", "dotted", "solid", "dotted"),
                                       labels=c("Ambient + Herbivory","Ambient + Reduced Herbivory","Warmed + Herbivory", "Warmed + Reduced Herbivory")) +
                 labs(x = "Year", y = "Cover (%)", title=loc, color="Treatment") +
                 #ylim(0, 30) +
                 theme_bw(14))
}

plot_line_kbs_n <- plot_line_n("KBS")
plot_line_kbs_n <- plot_line_kbs_n + labs(title="KBS", y=NULL, subtitle = "Native") +
        theme(axis.text.x=element_blank()) +
        annotate("text", x = 0.7, y=111, label = "A", size=5)
plot_line_umbs_n <- plot_line_n("UMBS")
plot_line_umbs_n <- plot_line_umbs_n + labs(title="UMBS", y=NULL, subtitle = "Native") +
        theme(axis.text.x=element_blank()) +
        annotate("text", x = 0.7, y=49, label = "B", size=5)

exotic_pc_avg <- origin_pc_avg %>% filter(origin == "Exotic")

plot_line_e <- function(loc) { 
  all_plot <- subset(exotic_pc_avg, site == loc)
  return(ggplot(all_plot, aes(x = year, y = avg_pc, group=full_treat, linetype=full_treat, fill=full_treat,color = full_treat, shape=full_treat)) +
                 geom_line(size = 1) +
                 geom_pointrange(aes(ymin=avg_pc-se, ymax=avg_pc+se), linetype="solid",pch=21,size=0.6) +
                 scale_color_manual(name="Treatment",
                           values = c("#a6bddb", "#a6bddb", "#AE1F00", "#AE1F00"),
                           labels=c("Ambient + Herbivory","Ambient + Reduced Herbivory","Warmed + Herbivory", "Warmed + Reduced Herbivory")) +
                 scale_fill_manual(name="Treatment",
                                   values = c("#a6bddb", "#a6bddb", "#AE1F00", "#AE1F00"),
                                   labels=c("Ambient + Herbivory","Ambient + Reduced Herbivory","Warmed + Herbivory", "Warmed + Reduced Herbivory")) +
                 scale_linetype_manual(name="Treatment",
                                       values = c("solid", "dotted", "solid", "dotted"),
                                       labels=c("Ambient + Herbivory","Ambient + Reduced Herbivory","Warmed + Herbivory", "Warmed + Reduced Herbivory")) +
                 labs(x = "Year", y = "Cover (%)", title=loc, color="Treatment") +
                 #ylim(0, 30) +
                 theme_bw(14))
}

plot_line_kbs_e <- plot_line_e("KBS")
plot_line_kbs_e <- plot_line_kbs_e + labs(title=NULL, y=NULL, subtitle = "Exotic") +
        annotate("text", x = 0.7, y=36, label = "C", size=5)
plot_line_umbs_e <- plot_line_e("UMBS")
plot_line_umbs_e <- plot_line_umbs_e + labs(title=NULL, y=NULL, subtitle = "Exotic") +
        annotate("text", x = 0.7, y=55, label = "D", size=5)

plot_line_origin <- ggpubr::ggarrange(plot_line_kbs_n, plot_line_umbs_n,
                                      plot_line_kbs_e, plot_line_umbs_e,
                                      nrow = 2, ncol = 2, common.legend = T, legend="right")
plot_line_origin

png("origin_avg_percent_cover_plots_by_year_L2.png", units="in", width=10, height=8, res=300)
annotate_figure(plot_line_origin,
                left = text_grob("Cover (%)", color = "black", rot = 90, size=15))
dev.off()
```

Growth habit plot level - all years
```{r}
forb_pc_avg <- growth_pc_avg %>% filter(growth_habit == "Forb")

plot_line_f <- function(loc) { 
  all_plot <- subset(forb_pc_avg, site == loc)
  return(ggplot(all_plot, aes(x = year, y = avg_pc, group=full_treat, linetype=full_treat, fill=full_treat,color = full_treat, shape=full_treat)) +
                 geom_line(size = 1) +
                 geom_pointrange(aes(ymin=avg_pc-se, ymax=avg_pc+se), linetype="solid",pch=21,size=0.6) +
                 scale_color_manual(name="Treatment",
                           values = c("#a6bddb", "#a6bddb", "#AE1F00", "#AE1F00"),
                           labels=c("Ambient + Herbivory","Ambient + Reduced Herbivory","Warmed + Herbivory", "Warmed + Reduced Herbivory")) +
                 scale_fill_manual(name="Treatment",
                                   values = c("#a6bddb", "#a6bddb", "#AE1F00", "#AE1F00"),
                                   labels=c("Ambient + Herbivory","Ambient + Reduced Herbivory","Warmed + Herbivory", "Warmed + Reduced Herbivory")) +
                 scale_linetype_manual(name="Treatment",
                                       values = c("solid", "dotted", "solid", "dotted"),
                                       labels=c("Ambient + Herbivory","Ambient + Reduced Herbivory","Warmed + Herbivory", "Warmed + Reduced Herbivory")) +
                 labs(x = "Year", y = "Cover (%)", title=loc, color="Treatment") +
                 #ylim(0, 30) +
                 theme_bw(14))
}

plot_line_kbs_f <- plot_line_f("KBS")
plot_line_kbs_f <- plot_line_kbs_f + labs(title="KBS", y=NULL, subtitle = "Forb") + theme(axis.text.x=element_blank()) +
        annotate("text", x = 0.7, y=91, label = "A", size=5)
plot_line_umbs_f <- plot_line_f("UMBS")
plot_line_umbs_f <- plot_line_umbs_f + labs(title="UMBS", y=NULL, subtitle = "Forb") + theme(axis.text.x=element_blank()) +
        annotate("text", x = 0.7, y=38, label = "B", size=5)

graminoid_pc_avg <- growth_pc_avg %>% filter(growth_habit == "Graminoid")

plot_line_g <- function(loc) { 
  all_plot <- subset(graminoid_pc_avg, site == loc)
  return(ggplot(all_plot, aes(x = year, y = avg_pc, group=full_treat, linetype=full_treat, fill=full_treat,color = full_treat, shape=full_treat)) +
                 geom_line(size = 1) +
                 geom_pointrange(aes(ymin=avg_pc-se, ymax=avg_pc+se), linetype="solid",pch=21,size=0.6) +
                 scale_color_manual(name="Treatment",
                           values = c("#a6bddb", "#a6bddb", "#AE1F00", "#AE1F00"),
                           labels=c("Ambient + Herbivory","Ambient + Reduced Herbivory","Warmed + Herbivory", "Warmed + Reduced Herbivory")) +
                 scale_fill_manual(name="Treatment",
                                   values = c("#a6bddb", "#a6bddb", "#AE1F00", "#AE1F00"),
                                   labels=c("Ambient + Herbivory","Ambient + Reduced Herbivory","Warmed + Herbivory", "Warmed + Reduced Herbivory")) +
                 scale_linetype_manual(name="Treatment",
                                       values = c("solid", "dotted", "solid", "dotted"),
                                       labels=c("Ambient + Herbivory","Ambient + Reduced Herbivory","Warmed + Herbivory", "Warmed + Reduced Herbivory")) +
                 labs(x = "Year", y = "Cover (%)", title=loc, color="Treatment") +
                 #ylim(0, 30) +
                 theme_bw(14))
}

plot_line_kbs_g <- plot_line_g("KBS")
plot_line_kbs_g <- plot_line_kbs_g + labs(title=NULL, y=NULL, subtitle = "Graminoid") +
        annotate("text", x = 0.7, y=50, label = "C", size=5)
plot_line_umbs_g <- plot_line_g("UMBS")
plot_line_umbs_g <- plot_line_umbs_g + labs(title=NULL, y=NULL, subtitle = "Graminoid") +
        annotate("text", x = 0.7, y=50, label = "D", size=5)

plot_line_gh <- ggpubr::ggarrange(plot_line_kbs_f, plot_line_umbs_f,
                                      plot_line_kbs_g, plot_line_umbs_g,
                          nrow = 2, ncol = 2, common.legend = T, legend="right")
plot_line_gh

png("growth_habit_avg_percent_cover_plots_by_year_L2.png", units="in", width=10, height=8, res=300)
annotate_figure(plot_line_gh,
                left = text_grob("Cover (%)", color = "black", rot = 90, size=15))
dev.off()
```

Averages + std error for species level data + insecticide treatment across all years - supplement
```{r}
# filter data to contain the averages and std error for each site & species
sum_comp_spp <- comp_spp %>%
        group_by(site, state, species) %>%
        summarize(avg_pc = mean(cover, na.rm = TRUE),
                  se = std.error(cover, na.rm = TRUE))
#sum_comp_spp_i <- comp_spp %>%
#        group_by(site, state, insecticide, species) %>%
#        summarize(avg_pc = mean(cover, na.rm = TRUE),
#                  se = std.error(cover, na.rm = TRUE))

## only keep species that were recorded in both warmed and ambient plots
#sum_flwr_dur_spp_i2 <- sum_flwr_dur_spp_i %>%
#        group_by(site, species) %>% 
#        filter(all(c('warmed', 'ambient') %in% state))
sum_comp_spp_i <- comp_spp %>%
        group_by(site, year, species) %>%
        filter(length(plot) >= 12) %>% # only keeping species present in at least 12 plots each year
        filter(all(c('warmed', 'ambient') %in% state)) %>% # only keep species that were recorded in both warmed and ambient plots
        group_by(site, state, insecticide, species) %>%
        summarize(avg_pc = mean(cover, na.rm = TRUE),
                  se = std.error(cover, na.rm = TRUE),
                  count=n()) # sample size
```

Species level plots for supplemental doc
```{r}
####### percent cover species level ########
# only keep species that were recorded in both warmed and ambient plots
#sum_comp_spp_i2 <- sum_comp_spp_i %>%
#        group_by(site, species) %>% 
#        filter(all(c('warmed', 'ambient') %in% state))
comp_plot_overall <- function(loc) { 
        comp_spp <- subset(sum_comp_spp_i, site == loc)
        return(ggplot(comp_spp, aes(x = state, y = avg_pc, fill = insecticide)) +
                       facet_wrap(~species, ncol=4,scales="free") +
                       geom_pointrange(aes(ymin=avg_pc-se, ymax=avg_pc+se), pch=21,size=0.6,position=position_dodge(0.3)) +
                       scale_fill_manual(name="Treatment",
                                         values = c("#FFB451", "#0b0055"),
                                         labels=c("Herbivory","Reduced Herbivory")) +
                       #geom_bar(position = "identity", stat = "identity", color = "black") +
                       #geom_errorbar(aes(ymin = avg_julian - se, ymax = avg_julian + se), width = 0.2,
                       #              position = "identity") +
                       labs(x = NULL, y = NULL, title=loc) +
                       scale_x_discrete(labels=c("ambient" = "Ambient  ", "warmed" = "  Warmed")) +
                       #scale_fill_manual(name="Treatment",
                       #                  labels=c("Ambient, herbivory","Ambient, reduced herbivory",
                       #                           "Warmed, herbivory","Warmed, reduced herbivory"),
                       #                  values = c("#a6bddb", "#687689", "#fb6a4a", "#9D422E")) +
                       theme_bw() +
                       theme(axis.text.x = element_text(size=14),
                             axis.text.y = element_text(size=14),
                             legend.text = element_text(size=15),
                             legend.title = element_text(size=15),
                             title = element_text(size=17),
                             strip.text = element_text(size=15)))
}
kbs_comp_spp <- comp_plot_overall("KBS")
umbs_comp_spp <- comp_plot_overall("UMBS")
comp_overall_merge <- ggpubr::ggarrange(kbs_comp_spp, umbs_comp_spp,
                                         ncol = 2, common.legend = T,legend="right")
png("comp_species.png", units="in", width=21, height=10, res=300)
annotate_figure(comp_overall_merge,
                left = text_grob("Cover (%)", color = "black", rot = 90, size=17))
dev.off()
```

mean temp plots
```{r}
comp_kbsp <- comp %>%
        filter(site == "KBS" & state == "ambient")
comp_umbsp <- comp %>%
        filter(site == "UMBS" & state == "ambient")

kbs_comp_mean <- ggplot(comp_kbsp, aes(x = mean_temp, y = avg_pc, color = state))+
        geom_point() +
        geom_smooth(method = 'lm', aes(fill = state)) + 
        labs(x = "Mean temperature (°C)",y=NULL) +
        scale_color_manual(values = c("#0b0055", "#fb6a4a"),
                           labels = c("Ambient","Warmed"),
                           name="Treatment") +
        scale_fill_manual(values = c("#a6bddb", "#fb6a4a"),
                          labels = c("Ambient","Warmed"),
                          name = "Treatment") +
        theme_classic() +
        theme(legend.title=element_text(size=12), 
              legend.text=element_text(size=12))
umbs_comp_mean <- ggplot(comp_umbsp, aes(x = mean_temp, y = avg_pc, color = state))+
        geom_point() +
        geom_smooth(method = 'lm', aes(fill = state)) + 
        labs(x = "Mean temperature (°C)",y=NULL) +
        scale_color_manual(values = c("#0b0055", "#fb6a4a"),
                           labels = c("Ambient","Warmed"),
                           name="Treatment") +
        scale_fill_manual(values = c("#a6bddb", "#fb6a4a"),
                          labels = c("Ambient","Warmed"),
                          name = "Treatment") +
        theme_classic() +
        theme(legend.title=element_text(size=12), 
              legend.text=element_text(size=12))

gdd_comp_merge <- ggpubr::ggarrange(kbs_comp_mean,umbs_comp_mean,
                                    ncol = 2, common.legend=T,legend="none")
png("comp_gdd_temp.png", units="in", width=7, height=5, res=300)
annotate_figure(gdd_comp_merge,
                left = text_grob("Average percent cover (%)", color = "black", rot = 90, size=15),
                top = text_grob("KBS                                                      UMBS", color = "black", size=15))
dev.off()
```








Not used in manuscript or supp

Averages + std error for plot level data + insecticide treatment averaged across all years
```{r}
sum_comp_plot_i <- comp %>%
        group_by(site, state, insecticide) %>%
        summarize(avg_pc_plot = mean(avg_pc, na.rm = TRUE),
                  se = std.error(avg_pc, na.rm = TRUE))

sum_origin_plot_i <- origin %>%
        group_by(site, state, insecticide, origin) %>%
        summarize(avg_pc = mean(avg_origin_pc, na.rm = TRUE),
                  se = std.error(avg_origin_pc, na.rm = TRUE))

sum_growth_plot_i <- growth %>%
        group_by(site, state, insecticide, growth_habit) %>%
        summarize(avg_pc = mean(avg_growth_pc, na.rm = TRUE),
                  se = std.error(avg_growth_pc, na.rm = TRUE))
```

Dot plot with insecticide
This plot has been updated with percent cover rather than relative abundance after Jen's feedback 8/1/23
```{r}
# plant composition
sum_comp_plot_i$full_treat <- paste(sum_comp_plot_i$state, sum_comp_plot_i$insecticide, sep="_")
comp_dot_i <- function(loc) { 
       comp_plot <- subset(sum_comp_plot_i, site == loc)
        return(ggplot(comp_plot, aes(x = state, y = avg_pc_plot, fill=insecticide)) +
                       #geom_point(stat = "identity",position=position_dodge(0.2),size=5) +
                       geom_pointrange(aes(ymin=avg_pc_plot-se, ymax=avg_pc_plot+se), pch=21,size=1,position=position_dodge(0.3)) +
                       #scale_shape_manual(name="Treatment",
                       #                   values = c(1, 19),
                       #                   labels=c("Herbivory","Reduced Herbivory")) +
                       scale_fill_manual(name="Treatment",
                                         values = c("#FFB451", "#0b0055"),
                                         labels=c("Herbivory","Reduced Herbivory")) +
                               labs(x = NULL, y = NULL, title = loc) +
                       scale_x_discrete(labels=c("ambient" = "Ambient", "warmed" = "Warmed")) +
                       coord_cartesian(ylim=c(14,24)) +
                       theme_bw(14))
}
comp_dot_i_kbs <- comp_dot_i("KBS")
comp_dot_i_kbs <- comp_dot_i_kbs + theme(plot.title = element_text(size=20),
                                     axis.text.y = element_text(size=17),
                                     axis.text.x = element_text(size=17),
                                     axis.title.y = element_text(size=17),
                                     legend.text = element_text(size=17),
                                     legend.title = element_text(size=17))
comp_dot_i_umbs <- comp_dot_i("UMBS")
comp_dot_i_umbs <- comp_dot_i_umbs + labs(y=NULL) + theme(axis.title.y=element_blank(),
                                                      plot.title = element_text(size=20),
                                                      axis.text.y = element_blank(),
                                                      axis.text.x = element_text(size=17),
                                                      legend.text = element_text(size=17),
                                                      legend.title = element_text(size=17))

comp_all_dot_i <- ggpubr::ggarrange(comp_dot_i_kbs, comp_dot_i_umbs,
                                nrow = 1, ncol = 2, common.legend = T, legend="right",
                                widths = c(1, 0.9))
png("percent_cover_plots_L2_all_plant_comp_dot_insect.png", units="in", width=8, height=5, res=300)
annotate_figure(comp_all_dot_i,
                left = text_grob("Cover (%)", color = "black", rot = 90, size=17))
dev.off()
```

Origin & Growth Habit
```{r}
# origin
sum_origin_plot_i$full_treat <- paste(sum_origin_plot_i$state, sum_origin_plot_i$insecticide, sep="_")
origin_dot_i <- function(loc) { 
       origin_plot <- subset(sum_origin_plot_i, site == loc)
        return(ggplot(origin_plot, aes(x = state, y = avg_pc, color=insecticide, shape=origin)) +
                       #geom_point(stat = "identity",position=position_dodge(0.2),size=5) +
                       geom_pointrange(aes(ymin=avg_pc-se, ymax=avg_pc+se),size=1,position=position_dodge(0.3)) +
                       scale_shape_manual(name="Origin",
                                          values = c(1, 19),
                                          labels=c("Exotic","Native")) +
                       scale_color_manual(name="Treatment",
                                         values = c("#FFB451", "#0b0055"),
                                         labels=c("Herbivory","Reduced Herbivory")) +
                               labs(x = NULL, y = "Origin", title = loc) +
                       scale_x_discrete(labels=c("ambient" = "Ambient", "warmed" = "Warmed")) +
                       #coord_cartesian(ylim=c(0.1,0.7)) +
                       theme_bw(14))
}
origin_dot_i_kbs <- origin_dot_i("KBS")
origin_dot_i_kbs
origin_dot_i_kbs <- origin_dot_i_kbs + 
        annotate("text", x = 0.5, y=51, label = "A", size=5) +
        theme(plot.title = element_text(size=20),
              axis.text.y = element_text(size=17),
              axis.text.x = element_blank(),
              axis.title.y = element_text(size=17),
              legend.position = "none")
origin_dot_i_umbs <- origin_dot_i("UMBS")
origin_dot_i_umbs <- origin_dot_i_umbs + 
        labs(y=NULL) + 
        annotate("text", x = 0.5, y=51, label = "B", size=5) +
        theme(axis.title.y=element_blank(),
              plot.title = element_text(size=20),
              axis.text.x = element_blank(),
              axis.text.y = element_blank(),
              legend.text = element_text(size=17),
              legend.title = element_text(size=17))

# growth habit
sum_growth_plot_i$full_treat <- paste(sum_growth_plot_i$state, sum_growth_plot_i$insecticide, sep="_")
growth_dot_i <- function(loc) { 
       growth_plot <- subset(sum_growth_plot_i, site == loc)
        return(ggplot(growth_plot, aes(x = state, y = avg_pc, color=insecticide, shape=growth_habit)) +
                       #geom_point(stat = "identity",position=position_dodge(0.2),size=5) +
                       geom_pointrange(aes(ymin=avg_pc-se, ymax=avg_pc+se),size=1,position=position_dodge(0.3)) +
                       scale_shape_manual(name="Growth Habit",
                                          values = c(1, 19),
                                          labels=c("Forb","Graminoid")) +
                       scale_color_manual(name="Treatment",
                                         values = c("#FFB451", "#0b0055"),
                                         labels=c("Herbivory","Reduced Herbivory")) +
                               labs(x = NULL, y = "Growth Habit", title = NULL) +
                       scale_x_discrete(labels=c("ambient" = "Ambient", "warmed" = "Warmed")) +
                       #coord_cartesian(ylim=c(10,42)) +
                       theme_bw(14))
}
growth_dot_i_kbs <- growth_dot_i("KBS")
growth_dot_i_kbs
growth_dot_i_kbs <- growth_dot_i_kbs + 
        annotate("text", x = 0.5, y=41, label = "C", size=5) +
        theme(plot.title = element_text(size=20),
              axis.text.y = element_text(size=17),
              axis.text.x = element_text(size=17),
              axis.title.y = element_text(size=17),
              legend.position = "none")
growth_dot_i_umbs <- growth_dot_i("UMBS")
growth_dot_i_umbs <- growth_dot_i_umbs + 
        labs(y=NULL) + 
        annotate("text", x = 0.5, y=41, label = "D", size=5) +
        theme(axis.title.y=element_blank(),
              plot.title = element_text(size=20),
              axis.text.y = element_blank(),
              axis.text.x = element_text(size=17),
              legend.text = element_text(size=17),
              legend.title = element_text(size=17))

legend_1 <- get_legend(origin_dot_i_umbs)
legend_2 <- get_legend(growth_dot_i_umbs)
legends <- ggarrange(legend_1, legend_2, nrow = 2)

origin_growth_dot_i <- ggpubr::ggarrange(origin_dot_i_kbs, origin_dot_i_umbs,
                  growth_dot_i_kbs, growth_dot_i_umbs,
                  nrow = 2, ncol = 2,
                  widths = c(1, 1.52))

png("percent_cover_plots_L2_growth_origin_dot_insect.png", units="in", width=10, height=8, res=300)
annotate_figure(origin_growth_dot_i,
                left = text_grob("Cover (%)", color = "black", rot = 90, size=17))
dev.off()
```

#GDD and temp comparisons
```{r}

comp_kbsp <- comp %>%
        filter(site == "KBS" & state == "ambient")
comp_umbsp <- comp %>%
        filter(site == "UMBS" & state == "ambient")

# looking to see which temp variable best correlates with flowring
kbs_comp_gdd <- ggplot(comp_kbsp, aes(x = GDD_cumulative, y = avg_pc, color = state))+
        geom_point() +
        geom_smooth(method = 'lm', aes(fill = state)) + 
        labs(x = "Cumulative growing degree days (GDD)",y=NULL) +
        scale_color_manual(values = c("#0b0055", "#fb6a4a"),
                           labels = c("Ambient","Warmed"),
                           name="Treatment") +
        scale_fill_manual(values = c("#a6bddb", "#fb6a4a"),
                          labels = c("Ambient","Warmed"),
                          name = "Treatment") +
        theme_classic() +
        theme(legend.title=element_text(size=12), 
              legend.text=element_text(size=12))
umbs_comp_gdd <- ggplot(comp_umbsp, aes(x = GDD_cumulative, y = avg_pc, color = state))+
        geom_point() +
        geom_smooth(method = 'lm', aes(fill = state)) + 
        labs(x = "Cumulative growing degree days (GDD)",y=NULL) +
        scale_color_manual(values = c("#0b0055", "#fb6a4a"),
                           labels = c("Ambient","Warmed"),
                           name="Treatment") +
        scale_fill_manual(values = c("#a6bddb", "#fb6a4a"),
                          labels = c("Ambient","Warmed"),
                          name = "Treatment") +
        theme_classic() +
        theme(legend.title=element_text(size=12), 
              legend.text=element_text(size=12))

kbs_comp_med <- ggplot(comp_kbsp, aes(x = median_temp, y = avg_pc, color = state))+
        geom_point() +
        geom_smooth(method = 'lm', aes(fill = state)) + 
        labs(x = "Median temperature (°C)",y=NULL) +
        scale_color_manual(values = c("#0b0055", "#fb6a4a"),
                           labels = c("Ambient","Warmed"),
                           name="Treatment") +
        scale_fill_manual(values = c("#a6bddb", "#fb6a4a"),
                          labels = c("Ambient","Warmed"),
                          name = "Treatment") +
        theme_classic() +
        theme(legend.title=element_text(size=12), 
              legend.text=element_text(size=12))
umbs_comp_med <- ggplot(comp_umbsp, aes(x = median_temp, y = avg_pc, color = state))+
        geom_point() +
        geom_smooth(method = 'lm', aes(fill = state)) + 
        labs(x = "Median temperature (°C)",y=NULL) +
        scale_color_manual(values = c("#0b0055", "#fb6a4a"),
                           labels = c("Ambient","Warmed"),
                           name="Treatment") +
        scale_fill_manual(values = c("#a6bddb", "#fb6a4a"),
                          labels = c("Ambient","Warmed"),
                          name = "Treatment") +
        theme_classic() +
        theme(legend.title=element_text(size=12), 
              legend.text=element_text(size=12))
kbs_comp_max <- ggplot(comp_kbsp, aes(x = max_temp, y = avg_pc, color = state))+
        geom_point() +
        geom_smooth(method = 'lm', aes(fill = state)) + 
        labs(x = "Max temperature (°C)",y=NULL) +
        scale_color_manual(values = c("#0b0055", "#fb6a4a"),
                           labels = c("Ambient","Warmed"),
                           name="Treatment") +
        scale_fill_manual(values = c("#a6bddb", "#fb6a4a"),
                          labels = c("Ambient","Warmed"),
                          name = "Treatment") +
        theme_classic() +
        theme(legend.title=element_text(size=12), 
              legend.text=element_text(size=12))
umbs_comp_max <- ggplot(comp_umbsp, aes(x = max_temp, y = avg_pc, color = state))+
        geom_point() +
        geom_smooth(method = 'lm', aes(fill = state)) + 
        labs(x = "Max temperature (°C)",y=NULL) +
        scale_color_manual(values = c("#0b0055", "#fb6a4a"),
                           labels = c("Ambient","Warmed"),
                           name="Treatment") +
        scale_fill_manual(values = c("#a6bddb", "#fb6a4a"),
                          labels = c("Ambient","Warmed"),
                          name = "Treatment") +
        theme_classic() +
        theme(legend.title=element_text(size=12), 
              legend.text=element_text(size=12))

```

Origin - Native vs. Exotic
```{r}
native <- origin_pc_avg[which(origin_pc_avg$origin == "Native"),]
exotic <- origin_pc_avg[which(origin_pc_avg$origin == "Exotic"),]

native_line <- function(loc) { 
  native_plot <- subset(native, site == loc)
  return(ggplot(native_plot, aes(x = year, y = avg_pc, group = state)) +
                 geom_errorbar(aes(ymin=avg_pc-se, ymax=avg_pc+se), position=position_dodge(0.1), width=.25) +
           geom_line(aes(color=state), size = 1) +
           geom_point(aes(color=state), size = 2) +
           scale_color_manual(values = c("#a6bddb", "#AE1F00"), labels=c("Ambient","Warmed")) +
           labs(x = NULL, y = "Cover (%)", title=loc, color="Treatment") +
                 theme_bw())
}

native_line_kbs <- native_line("KBS")
native_line_kbs <- native_line_kbs + labs(title="KBS", subtitle= "Native", y=NULL) + theme(axis.text.x=element_blank())
native_line_umbs <- native_line("UMBS")
native_line_umbs <- native_line_umbs + labs(title="UMBS", subtitle= "Native", y=NULL)

native_line <- ggpubr::ggarrange(native_line_kbs, native_line_umbs,
                          nrow = 2, ncol = 1, common.legend = T, legend="bottom")
#png("native_plots_L2.png", units="in", width=9, height=8, res=300)
#annotate_figure(native_line,
#                left = text_grob("Cover (%)", color = "black", rot = 90, size=15),
#                bottom = text_grob("Year", color = "black", size=15))
#dev.off()

exotic_line <- function(loc) { 
  exotic_plot <- subset(exotic, site == loc)
  return(ggplot(exotic_plot, aes(x = year, y = avg_pc, group = state)) +
                 geom_errorbar(aes(ymin=avg_pc-se, ymax=avg_pc+se), position=position_dodge(0.1), width=.25) +
           geom_line(aes(color=state), size = 1) +
           geom_point(aes(color=state), size = 2) +
           scale_color_manual(values = c("#a6bddb", "#AE1F00"), labels=c("Ambient","Warmed")) +
           labs(x = NULL, y = "Cover (%)", title=loc, color="Treatment") +
                 theme_bw())
}

exotic_line_kbs <- exotic_line("KBS")
exotic_line_kbs <- exotic_line_kbs + labs(title="KBS", subtitle="Exotic", y=NULL) + theme(axis.text.x=element_blank())
exotic_line_umbs <- exotic_line("UMBS")
exotic_line_umbs <- exotic_line_umbs + labs(title="UMBS", subtitle="Exotic", y=NULL)

origin_line <- ggpubr::ggarrange(native_line_kbs, exotic_line_kbs,
                                 native_line_umbs, exotic_line_umbs,
                          nrow = 2, ncol = 2, common.legend = T, legend="bottom")
png("origin_percent_cover_plots_L2.png", units="in", width=9, height=8, res=300)
annotate_figure(origin_line,
                left = text_grob("Cover (%)", color = "black", rot = 90, size=15),
                bottom = text_grob("Year", color = "black", size=15))
dev.off()
```

Growth Habit - Forbs vs. Graminoids
```{r}
forb <- growth_pc_avg[which(growth_pc_avg$growth_habit == "Forb"),]
graminoid <- growth_pc_avg[which(growth_pc_avg$growth_habit == "Graminoid"),]

forb_line <- function(loc) { 
  forb_plot <- subset(forb, site == loc)
  return(ggplot(forb_plot, aes(x = year, y = avg_pc, group = state)) +
                 geom_errorbar(aes(ymin=avg_pc-se, ymax=avg_pc+se), position=position_dodge(0.1), width=.25) +
           geom_line(aes(color=state), size = 1) +
           geom_point(aes(color=state), size = 2) +
           scale_color_manual(values = c("#a6bddb", "#AE1F00"), labels=c("Ambient","Warmed")) +
           labs(x = NULL, y = "Cover (%)", title=loc, color="Treatment") +
                 theme_bw())
}

forb_line_kbs <- forb_line("KBS")
forb_line_kbs <- forb_line_kbs + labs(title="KBS", subtitle= "Forb", y=NULL) + theme(axis.text.x=element_blank())
forb_line_umbs <- forb_line("UMBS")
forb_line_umbs <- forb_line_umbs + labs(title="UMBS", subtitle= "Forb", y=NULL)

#forb_line <- ggpubr::ggarrange(forb_line_kbs, forb_line_umbs,
#                          nrow = 2, ncol = 1, common.legend = T, legend="bottom")
#png("forb_plots_L2.png", units="in", width=9, height=8, res=300)
#annotate_figure(forb_line,
#                left = text_grob("Average Relative Abundance", color = "black", rot = 90, size=15),
#                bottom = text_grob("Year", color = "black", size=15))
#dev.off()

graminoid_line <- function(loc) { 
  graminoid_plot <- subset(graminoid, site == loc)
  return(ggplot(graminoid_plot, aes(x = year, y = avg_pc, group = state)) +
                 geom_errorbar(aes(ymin=avg_pc-se, ymax=avg_pc+se), position=position_dodge(0.1), width=.25) +
           geom_line(aes(color=state), size = 1) +
           geom_point(aes(color=state), size = 2) +
           scale_color_manual(values = c("#a6bddb", "#AE1F00"), labels=c("Ambient","Warmed")) +
           labs(x = NULL, y = "Cover (%)", title=loc, color="Treatment") +
                 theme_bw())
}

graminoid_line_kbs <- graminoid_line("KBS")
graminoid_line_kbs <- graminoid_line_kbs + labs(title="KBS", subtitle="Graminoid", y=NULL) + theme(axis.text.x=element_blank())
graminoid_line_umbs <- graminoid_line("UMBS")
graminoid_line_umbs <- graminoid_line_umbs + labs(title="UMBS", subtitle="Graminoid", y=NULL)

growthhabit_line <- ggpubr::ggarrange(forb_line_kbs, graminoid_line_kbs,
                                    forb_line_umbs, graminoid_line_umbs,
                          nrow = 2, ncol = 2, common.legend = T, legend="bottom")
png("growthhabit_percent_cover_plots_L2.png", units="in", width=9, height=8, res=300)
annotate_figure(growthhabit_line,
                left = text_grob("Cover (%)", color = "black", rot = 90, size=15),
                bottom = text_grob("Year", color = "black", size=15))
dev.off()
```

