---
title: "warmXtrophic Project: Date of First Seed Set Analyses"
author: "Moriah Young"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load in packages & data
```{r}
# clear all existing data
rm(list=ls())

#Load packages
library(tidyverse)
library(ggplot2)
library(lme4)
library(lmerTest)
library(emmeans)
library(vegan)
library(car)
library(rstatix)
library(scales)
library(fitdistrplus)
library(moments)# for calculating skewness of data
library(ggpubr)
library(jtools) # summ() function
library(predictmeans)
library(olsrr)
library(car)
library(fitdistrplus)
library(ggpubr)
library(interactions)
library(sjPlot)
library(effects)
library(glmmTMB)
library(GGally) # ggpairs() function
library(bbmle) # AICtab() function
library(knitr)

# Set working directory
Sys.getenv("L1DIR")
L1_dir <- Sys.getenv("L1DIR")
L2_dir <- Sys.getenv("L2DIR")

# Read in data
sd_plot <- read.csv(file.path(L2_dir, "phenology/final_sd_plot_L2.csv")) # plot level data
sd_plot_origin <- read.csv(file.path(L2_dir, "phenology/final_sd_plot_origin_L2.csv")) # plot level data for origin
sd_plot_growthhabit <- read.csv(file.path(L2_dir, "phenology/final_sd_plot_growthhabit_L2.csv")) # plot level data for growth habit

# species level seed set
# note: the species data frame is only used to demonstrate that seed set differs between species (for the supplement)
sd_spp <- read.csv(file.path(L2_dir, "phenology/final_sd_species_L2.csv")) 

sd_plot$julian_min <- as.numeric(as.integer(sd_plot$julian_min))
sd_spp$julian_min <- as.numeric(as.integer(sd_spp$julian_min))

# delete 2021 data from dataframe -doesn't make sense to have bc we can't get min flower/sd, median flower, or 
# flower duration from 2021 UMBS since data was not collected at the same frequency as previous years
#sd_plot <- sd_plot[-which(sd_plot$year == "2021" & sd_plot$site == "umbs"),]

# Order warmed and ambient so that warmed shows up first in plotting (and is default is red = warmed; blue = ambient). First make it a factor
sd_plot$state <- as.factor(sd_plot$state)
levels(sd_plot$state)
# [1] "ambient" "warmed" 
sd_plot$state <- factor(sd_plot$state, levels(sd_plot$state)[c(2,1)])
levels(sd_plot$state)
# [1] "warmed"  "ambient"

# make the year_factor a factor
sd_plot$year_factor <- as.factor(sd_plot$year_factor)
sd_spp$year_factor <- as.factor(sd_spp$year_factor)

umbs_sd_plot <- subset(sd_plot, site == "umbs") # pull out umbs only data at plot level
kbs_sd_plot <- subset(sd_plot, site == "kbs") # pull out kbs only data at plot level
kbs_sd_plot_origin <- subset(sd_plot_origin, site == "kbs")
kbs_sd_plot_growthhabit <- subset(sd_plot_growthhabit, site == "kbs")
umbs_sd_plot_origin <- subset(sd_plot_origin, site == "umbs")
umbs_sd_plot_growthhabit <- subset(sd_plot_growthhabit, site == "umbs")

# species specific dataframes for kbs and umbs
sd_kbs_spp <- subset(sd_spp, site == "kbs")
sd_umbs_spp <- subset(sd_spp, site == "umbs")
```

# KBS PLOT LEVEL - Looking at MINIMUM JULIAN DAY (aka date of observed first seed set)
```{r, warning=F, error=TRUE}
### KBS ###
hist(kbs_sd_plot$julian_min)
qqnorm(kbs_sd_plot$julian_min)
shapiro.test(kbs_sd_plot$julian_min) # pvalue is < 0.05 so we reject the null hypothesis that the data is normal (aka not normally distributed)

# Exploring distributions for these right-skewed data:
descdist(kbs_sd_plot$julian_min, discrete = FALSE)

# Gamma distribution 
fit.gamma <- fitdist(kbs_sd_plot$julian_min, "gamma")
plot(fit.gamma)

# Weibull distribution
fit.weibull  <- fitdist(kbs_sd_plot$julian_min, "weibull")
plot(fit.weibull)

# Lognormal distribution 
fit.ln <- fitdist(kbs_sd_plot$julian_min, "lnorm")
plot(fit.ln)

par(mfrow=c(2,2))
plot.legend <- c("Gamma", "Weibull", "Log Normal")
denscomp(list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
cdfcomp (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
qqcomp  (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
ppcomp  (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)

gofstat(list(fit.gamma, fit.weibull, fit.ln), fitnames = c("Gamma", "Weibull", "Log Normal"))
# Lognormal is better - going with a lognormal transformation for kbs seed set at the plot level
```

# Leverage plots and detecting Outliers. https://www.statmethods.net/stats/rdiagnostics.html
# These illustrate whether certain data points have more leverage (more influence), and thus could be outliers. It's a way of detecting outliers. Leverage plots can help identify whether a point has high or low influence, based on its leverage and residual and determining model fit with and without the point in question. Ultimately you decide whether the points are outliers or not, based on the knowledge of the system and how much it changes the model when included vs. excluded from the data used to fit the model. Here is a good overview of the combination of leverage and residual: scroll down to sections beginning at "13.3 Unusual Observations": https://daviddalpiaz.github.io/appliedstats/model-diagnostics.html
```{r, error=TRUE}
# KBS Plot level data
# KBS State-only model
fit_plot_state_kbs <- lm(log(julian_min) ~ state, data = kbs_sd_plot)
outlierTest(fit_plot_state_kbs) # outliers - 144, 217
qqPlot(fit_plot_state_kbs, main="QQ Plot") 
hist(fit_plot_state_kbs$residuals)
leveragePlots(fit_plot_state_kbs)
ols_test_normality(fit_plot_state_kbs)

# KBS State and year model
fit_plot_stateyear_kbs <- lm(log(julian_min) ~ state + year_factor, data = kbs_sd_plot)
outlierTest(fit_plot_stateyear_kbs) # no outliers
qqPlot(fit_plot_stateyear_kbs, main="QQ Plot") 
hist(fit_plot_stateyear_kbs$residuals)
leveragePlots(fit_plot_stateyear_kbs)
ols_test_normality(fit_plot_stateyear_kbs)

# Interaction plot (ignore for now the repeated measures with species); see: https://cran.r-project.org/web/packages/interactions/vignettes/interactions.html and: https://interactions.jacob-long.com/

fit3 <- lm(log(julian_min) ~ state * year, data = kbs_sd_plot)
interact_plot(fit3, pred = year, modx = state) # this doesn't work
```

KBS Plot-level Mixed Effects Models
```{r, error=TRUE}
mod1p <- lmer(log(julian_min) ~ state + (1|plot), kbs_sd_plot, REML=FALSE)
mod2p <- lmer(log(julian_min) ~ insecticide + (1|plot), kbs_sd_plot, REML=FALSE)
mod3p <- lmer(log(julian_min) ~ insecticide + state + (1|plot), kbs_sd_plot, REML=FALSE)
mod4p <- lmer(log(julian_min) ~ insecticide * state + (1|plot), kbs_sd_plot, REML=FALSE)
mod5p <- lmer(log(julian_min) ~ state + year_factor + (1|plot), kbs_sd_plot, REML=FALSE)
mod6p <- lmer(log(julian_min) ~ state + year_factor + insecticide + (1|plot), kbs_sd_plot, REML=FALSE)
mod7p <- lmer(log(julian_min) ~ state * year_factor + (1|plot), kbs_sd_plot, REML=FALSE)
mod8p <- lmer(log(julian_min) ~ state * year_factor + insecticide + (1|plot), kbs_sd_plot, REML=FALSE)
mod9p <- lmer(log(julian_min) ~ state * insecticide + year_factor + (1|plot), kbs_sd_plot, REML=FALSE)
mod10p <- lmer(log(julian_min) ~ state + insecticide * year_factor + (1|plot), kbs_sd_plot, REML=FALSE)
mod11p <- lmer(log(julian_min) ~ state * year_factor * insecticide + (1|plot), kbs_sd_plot, REML=FALSE)
AICctab(mod1p, mod2p, mod3p, mod4p, mod5p, mod6p, mod7p, mod8p, mod9p, mod10p, mod11p, weights=T) 
anova(mod7p, mod8p) 
AICctab(mod7p, mod8p, weights=T)
summary(mod7p)
anova(mod7p)
emmeans(mod7p, list(pairwise ~ state*year_factor), adjust = "tukey")

# calculating effect size - accounting for log transformation
exp(5.46150 + (-0.11720*0)) # 235.4503 warmed
exp(5.46150 + (-0.11720*1)) # 209.4112 ambient
# effect:
209.4112 - 235.4503 # -26.0391 --> ambient plots advanced date of first seed set by 26.0391 days

# including herb reduction as fixed effects in model
anova(mod8p)
# including herb reduction as interactive term
anova(mod9p)

#Our hypothesized model is  mod9p
summary(mod9p)
anova(mod9p)

# making a table
kable(anova(mod9p)) %>% kableExtra::kable_styling()

# comparisons
emmeans(mod9p, list(pairwise ~ state*insecticide), adjust = "tukey")
emmeans(mod9p, list(pairwise ~ state), adjust = "tukey")
mod9p.emm <- emmeans(mod9p, ~ state*insecticide)
contrast(mod9p.emm, "consec", simple = "each", combine = F, adjust = "mvt")
emmip(mod9p, insecticide~state)

# calculating effect size - accounting for log transformation
exp((5.40746) +  (-0.0342)*0) # 223.0643 
exp((5.40746) +  (-0.0342)*1) # 215.5645
# effect:
215.5645 - 223.0643 # = -7.4998  
# warming plots delayed seed set by ~7.5 days 

# using effect size from emmeans - emmeans(mod9p, list(pairwise ~ state), adjust = "tukey")
# calculating effect size - accounting for log transformation
exp((5.40746) +  (-0.025)*0) # 223.0643 
exp((5.40746) +  (-0.025)*1) # 217.5568
# effect:
217.5568 - 223.0643 # = -5.5075  
# warming plots delayed seed set by ~5.5 days 

# adding in our temp data into some models
# note: including state, year, and temp data into a model leads to rank deficiency
# so below, we test for date of first seed set as a function of just temp to see how real temp data affects date of first seed set
modtest1 <- lmer(log(julian_min) ~ mean_temp + (1|plot), kbs_sd_plot, REML=FALSE)
modtest2 <- lmer(log(julian_min) ~ GDD_cumulative + (1|plot), kbs_sd_plot, REML=FALSE)

AICtab(modtest1, modtest2, weights=T)

anova(modtest1)
summary(modtest1)
# calculating effect size - accounting for log transformation
exp((4.984e+00) +  (-2.009e-02)*0) # 146.0574 
exp((4.984e+00) +  (-2.009e-02)*1) # 143.1524
# effect:
143.1524 - 146.0574 # = -2.905  
# warming plots delayed seed set by ~2.91 days
```

UMBS PLOT LEVEL
# Looking at MINIMUM JULIAN DAY (aka date of observed first seed set)
```{r, warning=F, error=TRUE}
### UMBS ###
hist(umbs_sd_plot$julian_min)
qqnorm(umbs_sd_plot$julian_min)
shapiro.test(umbs_sd_plot$julian_min) # pvalue is >0.05 so we accept the null hypothesis that the data is normal (aka not normally distributed)

# Exploring distributions for these right-skewed data:
descdist(umbs_sd_plot$julian_min, discrete = FALSE) # normal

# Gamma distribution 
fit.gamma <- fitdist(umbs_sd_plot$julian_min, "gamma")
plot(fit.gamma)

# Weibull distribution
fit.weibull  <- fitdist(umbs_sd_plot$julian_min, "weibull")
plot(fit.weibull)

# Lognormal distribution 
fit.ln <- fitdist(umbs_sd_plot$julian_min, "lnorm")
plot(fit.ln)

par(mfrow=c(2,2))
plot.legend <- c("Gamma", "Weibull", "Log Normal")
denscomp(list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
cdfcomp (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
qqcomp  (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
ppcomp  (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)

gofstat(list(fit.gamma, fit.weibull, fit.ln), fitnames = c("Gamma", "Weibull", "Log Normal"))
# log normal is slightly better here but going to go with a normal transformation for umbs seed set at the plot level based on graphs and shapiro test above

```

```{r, error=TRUE}
# Plot level data
# UMBS State-only model
fit_plot_state_umbs <- lm(julian_min ~ state, data = umbs_sd_plot)
outlierTest(fit_plot_state_umbs) # no outliers
qqPlot(fit_plot_state_umbs, main="QQ Plot") 
hist(fit_plot_state_umbs$residuals)
leveragePlots(fit_plot_state_umbs)
ols_test_normality(fit_plot_state_umbs)

# UMBS State and year model
fit_plot_stateyear_umbs <- lm(julian_min ~ state + year_factor, data = umbs_sd_plot)
outlierTest(fit_plot_stateyear_umbs) 
qqPlot(fit_plot_stateyear_umbs, main="QQ Plot") 
hist(fit_plot_stateyear_umbs$residuals)
leveragePlots(fit_plot_stateyear_umbs)
ols_test_normality(fit_plot_stateyear_umbs)

# Interaction plot (ignore for now the repeated measures with species); see: https://cran.r-project.org/web/packages/interactions/vignettes/interactions.html and: https://interactions.jacob-long.com/

fit3_plot_umbs <- lm(julian_min ~ state * year, data = umbs_sd_plot)
interact_plot(fit3_plot_umbs, pred = year, modx = state)
```

UMBS Plot-level Mixed Effects Models
```{r}
mod1pu <- lmer(julian_min ~ state + (1|plot), umbs_sd_plot, REML=FALSE)
mod2pu <- lmer(julian_min ~ insecticide + (1|plot), umbs_sd_plot, REML=FALSE)
mod3pu <- lmer(julian_min ~ insecticide + state + (1|plot), umbs_sd_plot, REML=FALSE)
mod4pu <- lmer(julian_min ~ insecticide * state + (1|plot), umbs_sd_plot, REML=FALSE)
mod5pu <- lmer(julian_min ~ state + year_factor + (1|plot), umbs_sd_plot, REML=FALSE)
mod6pu <- lmer(julian_min ~ state + year_factor + insecticide + (1|plot), umbs_sd_plot, REML=FALSE)
mod7pu <- lmer(julian_min ~ state * year_factor + (1|plot), umbs_sd_plot, REML=FALSE)
mod8pu <- lmer(julian_min ~ state * year_factor + insecticide + (1|plot), umbs_sd_plot, REML=FALSE)
mod9pu <- lmer(julian_min ~ state * insecticide + year_factor + (1|plot), umbs_sd_plot, REML=FALSE)
mod10pu <- lmer(julian_min ~ state + insecticide * year_factor + (1|plot), umbs_sd_plot, REML=FALSE)
mod11pu <- lmer(julian_min ~ state * year_factor * insecticide + (1|plot), umbs_sd_plot, REML=FALSE)
AICctab(mod1pu, mod2pu, mod3pu, mod4pu, mod5pu, mod6pu, mod7pu, mod8pu, mod9pu, mod10pu, mod11pu, weights=T) # model 11p and 10p the same
anova(mod5pu, mod6pu) #go with mod5pu
AICctab(mod5pu, mod6pu, weights=T)
summary(mod5pu)
anova(mod5pu)
emmeans(mod5pu, list(pairwise ~ state+year_factor), adjust = "tukey")

# including herb reduction in model
anova(mod6pu)
# including herb reduction as interactive term
anova(mod9pu)

#Our hypothesized model is  mod9p
summary(mod9pu)
anova(mod9pu)

# making a table
kable(anova(mod9pu)) %>% kableExtra::kable_styling()

# comparisons
emmeans(mod9pu, list(pairwise ~ state*insecticide), adjust = "tukey")
mod9pu.emm <- emmeans(mod9pu, ~ state*insecticide)
contrast(mod9pu.emm, "consec", simple = "each", combine = F, adjust = "mvt")
emmip(mod9pu, insecticide~state)

# adding in our temp data into some models
# note: including state, year, and temp data into a model leads to rank deficiency
# so below, we test for date of first seed set as a function of just temp to see how real temp data affects date of first seed set
modtest1 <- lmer(julian_min ~ mean_temp + (1|plot), umbs_sd_plot, REML=FALSE)
modtest2 <- lmer(julian_min ~ GDD_cumulative + (1|plot), umbs_sd_plot, REML=FALSE)

AICtab(modtest1, modtest2, weights=T)

anova(modtest1)
summary(modtest1)

```

```{r}
# KBS
# species model for supp
mod_spp_k <- lmer(julian_min ~ state * insecticide + species + as.factor(year_factor) + (1|plot), sd_kbs_spp, REML=FALSE)
anova(mod_spp_k)
# making a table
kable(anova(mod_spp_k)) %>% kableExtra::kable_styling()

# UMBS
# species model for supp
mod_spp_u <- lmer(julian_min ~ state * insecticide + species + as.factor(year_factor) + (1|plot), sd_umbs_spp, REML=FALSE)
anova(mod_spp_u)
# making a table
kable(anova(mod_spp_u)) %>% kableExtra::kable_styling()
```

