---
title: "warmXtrophic Project: Date of First Seed Set Analyses"
author: "Moriah Young"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load in packages & data
```{r}
# clear all existing data
rm(list=ls())

#Load packages
library(tidyverse)
library(ggplot2)
library(lme4)
library(lmerTest)
library(emmeans)
library(vegan)
library(car)
library(rstatix)
library(scales)
library(fitdistrplus)
library(moments)# for calculating skewness of data
library(ggpubr)
library(jtools) # summ() function
library(predictmeans)
library(olsrr)
library(car)
library(fitdistrplus)
library(ggpubr)
library(interactions)
library(sjPlot)
library(effects)
library(glmmTMB)
library(GGally) # ggpairs() function
library(bbmle) # AICtab() function
library(knitr)
library(MuMIn)

# Set working directory
Sys.getenv("L1DIR")
L1_dir <- Sys.getenv("L1DIR")
L2_dir <- Sys.getenv("L2DIR")

# Read in data
sd_plot <- read.csv(file.path(L2_dir, "phenology/final_sd_plot_L2.csv")) # plot level data
sd_plot_origin <- read.csv(file.path(L2_dir, "phenology/final_sd_plot_origin_L2.csv")) # plot level data for origin
sd_plot_growthhabit <- read.csv(file.path(L2_dir, "phenology/final_sd_plot_growthhabit_L2.csv")) # plot level data for growth habit

# species level seed set
# note: the species data frame is only used to demonstrate that seed set differs between species (for the supplement)
sd_spp <- read.csv(file.path(L2_dir, "phenology/final_sd_species_L2.csv")) 

sd_plot$julian_min <- as.numeric(as.integer(sd_plot$julian_min))
sd_spp$julian_min <- as.numeric(as.integer(sd_spp$julian_min))

# delete 2021 data from dataframe -doesn't make sense to have bc we can't get min flower/sd, median flower, or 
# flower duration from 2021 UMBS since data was not collected at the same frequency as previous years
#sd_plot <- sd_plot[-which(sd_plot$year == "2021" & sd_plot$site == "umbs"),]

# make the year_factor a factor
sd_plot$year_factor <- as.factor(sd_plot$year_factor)
sd_spp$year_factor <- as.factor(sd_spp$year_factor)

umbs_sd_plot <- subset(sd_plot, site == "umbs") # pull out umbs only data at plot level
kbs_sd_plot <- subset(sd_plot, site == "kbs") # pull out kbs only data at plot level
kbs_sd_plot_origin <- subset(sd_plot_origin, site == "kbs")
kbs_sd_plot_growthhabit <- subset(sd_plot_growthhabit, site == "kbs")
umbs_sd_plot_origin <- subset(sd_plot_origin, site == "umbs")
umbs_sd_plot_growthhabit <- subset(sd_plot_growthhabit, site == "umbs")

# species specific dataframes for kbs and umbs
sd_kbs_spp <- subset(sd_spp, site == "kbs")
sd_umbs_spp <- subset(sd_spp, site == "umbs")
```

# KBS PLOT LEVEL - Looking at MINIMUM JULIAN DAY (aka date of observed first seed set)
```{r, warning=F, error=TRUE}
### KBS ###
hist(kbs_sd_plot$julian_min)
qqnorm(kbs_sd_plot$julian_min)
shapiro.test(kbs_sd_plot$julian_min) # pvalue is < 0.05 so we reject the null hypothesis that the data is normal (aka not normally distributed)

# Exploring distributions for these right-skewed data:
descdist(kbs_sd_plot$julian_min, discrete = FALSE)

# Gamma distribution 
fit.gamma <- fitdist(kbs_sd_plot$julian_min, "gamma")
plot(fit.gamma)

# Weibull distribution
fit.weibull  <- fitdist(kbs_sd_plot$julian_min, "weibull")
plot(fit.weibull)

# Lognormal distribution 
fit.ln <- fitdist(kbs_sd_plot$julian_min, "lnorm")
plot(fit.ln)

par(mfrow=c(2,2))
plot.legend <- c("Gamma", "Weibull", "Log Normal")
denscomp(list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
cdfcomp (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
qqcomp  (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
ppcomp  (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)

gofstat(list(fit.gamma, fit.weibull, fit.ln), fitnames = c("Gamma", "Weibull", "Log Normal"))
```

# Leverage plots and detecting Outliers. https://www.statmethods.net/stats/rdiagnostics.html
# These illustrate whether certain data points have more leverage (more influence), and thus could be outliers. It's a way of detecting outliers. Leverage plots can help identify whether a point has high or low influence, based on its leverage and residual and determining model fit with and without the point in question. Ultimately you decide whether the points are outliers or not, based on the knowledge of the system and how much it changes the model when included vs. excluded from the data used to fit the model. Here is a good overview of the combination of leverage and residual: scroll down to sections beginning at "13.3 Unusual Observations": https://daviddalpiaz.github.io/appliedstats/model-diagnostics.html
```{r}
# KBS Plot level data
# KBS State-only model
fit_plot_state_kbs <- lm(julian_min ~ state, data = kbs_sd_plot)
outlierTest(fit_plot_state_kbs) # outliers - 144, 204
qqPlot(fit_plot_state_kbs, main="QQ Plot") 
hist(fit_plot_state_kbs$residuals)
leveragePlots(fit_plot_state_kbs)
ols_test_normality(fit_plot_state_kbs)

# KBS State and year model
fit_plot_stateyear_kbs <- lm(julian_min ~ state + year_factor, data = kbs_sd_plot)
outlierTest(fit_plot_stateyear_kbs) # outliers - 144, 204
qqPlot(fit_plot_stateyear_kbs, main="QQ Plot") 
hist(fit_plot_stateyear_kbs$residuals)
leveragePlots(fit_plot_stateyear_kbs)
ols_test_normality(fit_plot_stateyear_kbs)

# Interaction plot (ignore for now the repeated measures with species); see: https://cran.r-project.org/web/packages/interactions/vignettes/interactions.html and: https://interactions.jacob-long.com/

fit3 <- lm(julian_min ~ state * year, data = kbs_sd_plot)
interact_plot(fit3, pred = year, modx = state) # this doesn't work

# remove outliers, lines 144 and 204
kbs_sd_plot_nooutlier <- kbs_sd_plot[!rownames(kbs_sd_plot) == '144', ]
kbs_sd_plot_nooutlier <- kbs_sd_plot_nooutlier[!rownames(kbs_sd_plot_nooutlier) == '204', ]
# KBS State-only model
fit_plot_state_kbs <- lm(julian_min ~ state, data = kbs_sd_plot_nooutlier)
outlierTest(fit_plot_state_kbs) # no outliers
qqPlot(fit_plot_state_kbs, main="QQ Plot") 
hist(fit_plot_state_kbs$residuals)
leveragePlots(fit_plot_state_kbs)
ols_test_normality(fit_plot_state_kbs)
```

Mixed Effects Model
```{r}
# KBS PLOT LEVEL - Looking at FIRST date of flowering
mod1 <- lmer(julian_min ~ state * insecticide * year_factor + (1|plot), kbs_sd_plot_nooutlier, REML=FALSE)

# Check Assumptions:
# (1) Linearity: if covariates are not categorical (year isn't)
# (2) Homogeneity: Need to Check by plotting residuals vs predicted values.
par(mfrow=c(1,2))
plot(mod1) # the ones in the data are making this look weird - idk this doesn't look great
# Check for homogeneity of variances (true if p>0.05). If the result is not significant, the assumption of equal variances (homoscedasticity) is met (no significant difference between the group variances).
# *****Levene’s Test - tests whether or not the variance among two or more groups is equal - If the p-value is less than our chosen significance level, we can reject the null hypothesis and conclude that we have enough evidence to state that the variance among the groups is not equal (which we want).

leveneTest(residuals(mod1) ~ kbs_sd_plot_nooutlier$state)
# Assumption met
leveneTest(residuals(mod1) ~ kbs_sd_plot_nooutlier$insecticide) 
# Assumption not met
leveneTest(residuals(mod1) ~ kbs_sd_plot_nooutlier$plot)
# Assumption met

# (3) Normality of error term: need to check by histogram, QQplot of residuals, could do Kolmogorov-Smirnov test.
# Check for normal residuals
qqPlot(resid(mod1))
hist(residuals(mod1))
shapiro.test(resid(mod1)) # normally distributed resids bc p>0.05
outlierTest(mod1) # 192

kbs_sd_plot_nooutlier <- kbs_sd_plot_nooutlier[!rownames(kbs_sd_plot_nooutlier) == '192', ]
mod1 <- lmer(julian_min ~ state * insecticide * year_factor + (1|plot), kbs_sd_plot_nooutlier, REML=FALSE)
qqPlot(resid(mod1))
hist(residuals(mod1))
shapiro.test(resid(mod1)) # normally distributed resids bc p>0.05
outlierTest(mod1)
```

# KBS Plot-level model - ALL YEARS
```{r}
#Our hypothesized model is:
mod_k_all <- lmer(julian_min ~ state * insecticide * year_factor + (1|plot), kbs_sd_plot_nooutlier, REML=FALSE)
summary(mod_k_all)
anova(mod_k_all)

kable(anova(mod_k_all)) %>% kableExtra::kable_styling()

# comparisons
contrast.k <- contrast(emmeans(mod_k_all, ~state*year_factor), "pairwise", simple = "each", combine = F, adjust = "mvt")

result = as.data.frame(contrast.k)

result <- result %>%
mutate_if(is.numeric, round, digits=2)

# table
kable(result) %>% kableExtra::kable_styling()

# adding in our temp data into some models
# note: including state, year, and temp data into a model leads to rank deficiency
# so below, we test for date of first seed set as a function of just temp to see how real temp data affects date of first seed set
kbs_sd_plot_amb <- kbs_sd_plot %>%
        filter(state == "ambient")
modtest1 <- lmer(julian_min ~ mean_temp + (1|plot), kbs_sd_plot_nooutlier, REML=FALSE)
modtest2 <- lmer(julian_min ~ GDD_cumulative + (1|plot), kbs_sd_plot_nooutlier, REML=FALSE)

AICtab(modtest1, modtest2, weights=T)

anova(modtest1)
summary(modtest1)

r.squaredGLMM(modtest1)

# making a table
kable(anova(modtest1)) %>% kableExtra::kable_styling()
```

UMBS PLOT LEVEL
# Looking at MINIMUM JULIAN DAY (aka date of observed first seed set)
```{r, warning=F, error=TRUE}
### UMBS ###
hist(umbs_sd_plot$julian_min)
qqnorm(umbs_sd_plot$julian_min)
shapiro.test(umbs_sd_plot$julian_min) # pvalue is >0.05 so we accept the null hypothesis that the data is normal (aka not normally distributed)

# Exploring distributions for these right-skewed data:
descdist(umbs_sd_plot$julian_min, discrete = FALSE) # normal

# Gamma distribution 
fit.gamma <- fitdist(umbs_sd_plot$julian_min, "gamma")
plot(fit.gamma)

# Weibull distribution
fit.weibull  <- fitdist(umbs_sd_plot$julian_min, "weibull")
plot(fit.weibull)

# Lognormal distribution 
fit.ln <- fitdist(umbs_sd_plot$julian_min, "lnorm")
plot(fit.ln)

par(mfrow=c(2,2))
plot.legend <- c("Gamma", "Weibull", "Log Normal")
denscomp(list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
cdfcomp (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
qqcomp  (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)
ppcomp  (list(fit.gamma, fit.weibull, fit.ln), legendtext = plot.legend)

gofstat(list(fit.gamma, fit.weibull, fit.ln), fitnames = c("Gamma", "Weibull", "Log Normal"))
# log normal is slightly better here but going to go with a normal transformation for umbs seed set at the plot level based on graphs and shapiro test above

```

```{r}
# Plot level data
# UMBS State-only model
fit_plot_state_umbs <- lm(julian_min ~ state, data = umbs_sd_plot)
outlierTest(fit_plot_state_umbs) # no outliers
qqPlot(fit_plot_state_umbs, main="QQ Plot") 
hist(fit_plot_state_umbs$residuals)
leveragePlots(fit_plot_state_umbs)
ols_test_normality(fit_plot_state_umbs)

# UMBS State and year model
fit_plot_stateyear_umbs <- lm(julian_min ~ state + year_factor, data = umbs_sd_plot)
outlierTest(fit_plot_stateyear_umbs) 
qqPlot(fit_plot_stateyear_umbs, main="QQ Plot") 
hist(fit_plot_stateyear_umbs$residuals)
leveragePlots(fit_plot_stateyear_umbs)
ols_test_normality(fit_plot_stateyear_umbs)

# Interaction plot (ignore for now the repeated measures with species); see: https://cran.r-project.org/web/packages/interactions/vignettes/interactions.html and: https://interactions.jacob-long.com/

fit3_plot_umbs <- lm(julian_min ~ state * year, data = umbs_sd_plot)
interact_plot(fit3_plot_umbs, pred = year, modx = state)
```

# UMBS Plot-level model - ALL YEARS
```{r}
#Our hypothesized model is  mod9p
mod_u_all <- lmer(julian_min ~ state * insecticide * year_factor + (1|plot), umbs_sd_plot, REML=FALSE)
summary(mod_u_all)
anova(mod_u_all)

# comparisons
contrast.u <- contrast(emmeans(mod_u_all, ~state*insecticide*year_factor), "pairwise", simple = "each", combine = F, adjust = "mvt")

result = as.data.frame(contrast.u)

result <- result %>%
mutate_if(is.numeric, round, digits=2)

# table
kable(result) %>% kableExtra::kable_styling()

# adding in our temp data into some models
# note: including state, year, and temp data into a model leads to rank deficiency
# so below, we test for date of first seed set as a function of just temp to see how real temp data affects date of first seed set
umbs_sd_plot_amb <- umbs_sd_plot %>%
        filter(state == "ambient")
modtest1 <- lmer(julian_min ~ mean_temp + (1|plot), umbs_sd_plot, REML=FALSE)
modtest2 <- lmer(julian_min ~ GDD_cumulative + (1|plot), umbs_sd_plot, REML=FALSE)

AICtab(modtest1, modtest2, weights=T)

anova(modtest1)
summary(modtest1)

r.squaredGLMM(modtest1)

# making a table
kable(anova(modtest1), digits = 3) %>% kableExtra::kable_styling()
```

Supplement

Species models
```{r}
# KBS
# species model for supp
mod_spp_k <- lmer(julian_min ~ state * insecticide * as.factor(year_factor) + species + (1|plot), sd_kbs_spp, REML=FALSE)
anova(mod_spp_k)
# making a table
kable(anova(mod_spp_k), digits = 2) %>% kableExtra::kable_styling()

# UMBS
# species model for supp
mod_spp_u <- lmer(julian_min ~ state * insecticide * as.factor(year_factor) + species + (1|plot), sd_umbs_spp, REML=FALSE)
anova(mod_spp_u)
# making a table
kable(anova(mod_spp_u), digits = 2) %>% kableExtra::kable_styling()
```






KBS and UMBS models not used in manuscript

KBS Plot-level Mixed Effects Models
```{r}
#Our hypothesized model is  mod9p
mod9p <- lmer(log(julian_min) ~ state * insecticide + year_factor + (1|plot), kbs_sd_plot, REML=FALSE)
summary(mod9p)
anova(mod9p)

# making a table
kable(anova(mod9p)) %>% kableExtra::kable_styling()

# calculating effect size - accounting for log transformation
exp((5.392e+00) +  (1.588e-02)*0) # 219.6422 
exp((5.392e+00) +  (1.588e-02)*1) # 223.158
# effect:
223.158 - 219.6422 # = 3.5158  
# warming plots delayed seed set by ~3.5 days 
```

UMBS Plot-level Mixed Effects Models
```{r}
#Our hypothesized model is  mod9pu
mod9pu <- lmer(julian_min ~ state * insecticide + year_factor + (1|plot), umbs_sd_plot, REML=FALSE)
summary(mod9pu)
anova(mod9pu)

# making a table
kable(anova(mod9pu)) %>% kableExtra::kable_styling()
```

```{r}
# KBS
mod1p <- lmer(log(julian_min) ~ state + (1|plot), kbs_sd_plot, REML=FALSE)
mod2p <- lmer(log(julian_min) ~ insecticide + (1|plot), kbs_sd_plot, REML=FALSE)
mod3p <- lmer(log(julian_min) ~ insecticide + state + (1|plot), kbs_sd_plot, REML=FALSE)
mod4p <- lmer(log(julian_min) ~ insecticide * state + (1|plot), kbs_sd_plot, REML=FALSE)
mod5p <- lmer(log(julian_min) ~ state + year_factor + (1|plot), kbs_sd_plot, REML=FALSE)
mod6p <- lmer(log(julian_min) ~ state + year_factor + insecticide + (1|plot), kbs_sd_plot, REML=FALSE)
mod7p <- lmer(log(julian_min) ~ state * year_factor + (1|plot), kbs_sd_plot, REML=FALSE)
mod8p <- lmer(log(julian_min) ~ state * year_factor + insecticide + (1|plot), kbs_sd_plot, REML=FALSE)
mod9p <- lmer(log(julian_min) ~ state * insecticide + year_factor + (1|plot), kbs_sd_plot, REML=FALSE)
mod10p <- lmer(log(julian_min) ~ state + insecticide * year_factor + (1|plot), kbs_sd_plot, REML=FALSE)
mod11p <- lmer(log(julian_min) ~ state * year_factor * insecticide + (1|plot), kbs_sd_plot, REML=FALSE)
AICctab(mod1p, mod2p, mod3p, mod4p, mod5p, mod6p, mod7p, mod8p, mod9p, mod10p, mod11p, weights=T) 
anova(mod7p, mod8p) 
AICctab(mod7p, mod8p, weights=T)
summary(mod7p)
anova(mod7p)
emmeans(mod7p, list(pairwise ~ state*year_factor), adjust = "tukey")

# comparisons
emmeans(mod9p, list(pairwise ~ state*insecticide), adjust = "tukey")
emmeans(mod9p, list(pairwise ~ state), adjust = "tukey")
mod9p.emm <- emmeans(mod9p, ~ state*insecticide)
contrast(mod9p.emm, "consec", simple = "each", combine = F, adjust = "mvt")
emmip(mod9p, insecticide~state)

# UMBS
mod1pu <- lmer(julian_min ~ state + (1|plot), umbs_sd_plot, REML=FALSE)
mod2pu <- lmer(julian_min ~ insecticide + (1|plot), umbs_sd_plot, REML=FALSE)
mod3pu <- lmer(julian_min ~ insecticide + state + (1|plot), umbs_sd_plot, REML=FALSE)
mod4pu <- lmer(julian_min ~ insecticide * state + (1|plot), umbs_sd_plot, REML=FALSE)
mod5pu <- lmer(julian_min ~ state + year_factor + (1|plot), umbs_sd_plot, REML=FALSE)
mod6pu <- lmer(julian_min ~ state + year_factor + insecticide + (1|plot), umbs_sd_plot, REML=FALSE)
mod7pu <- lmer(julian_min ~ state * year_factor + (1|plot), umbs_sd_plot, REML=FALSE)
mod8pu <- lmer(julian_min ~ state * year_factor + insecticide + (1|plot), umbs_sd_plot, REML=FALSE)
mod9pu <- lmer(julian_min ~ state * insecticide + year_factor + (1|plot), umbs_sd_plot, REML=FALSE)
mod10pu <- lmer(julian_min ~ state + insecticide * year_factor + (1|plot), umbs_sd_plot, REML=FALSE)
mod11pu <- lmer(julian_min ~ state * year_factor * insecticide + (1|plot), umbs_sd_plot, REML=FALSE)
AICctab(mod1pu, mod2pu, mod3pu, mod4pu, mod5pu, mod6pu, mod7pu, mod8pu, mod9pu, mod10pu, mod11pu, weights=T) # model 11p and 10p the same
anova(mod5pu, mod6pu) #go with mod5pu
AICctab(mod5pu, mod6pu, weights=T)
summary(mod5pu)
anova(mod5pu)
emmeans(mod5pu, list(pairwise ~ state+year_factor), adjust = "tukey")

# including herb reduction in model
anova(mod6pu)
# including herb reduction as interactive term
anova(mod9pu)

# comparisons
emmeans(mod9pu, list(pairwise ~ state*insecticide), adjust = "tukey")
mod9pu.emm <- emmeans(mod9pu, ~ state*insecticide)
contrast(mod9pu.emm, "consec", simple = "each", combine = F, adjust = "mvt")
emmip(mod9pu, insecticide~state)
```

