---
title: "warmXtrophic Project: Specific Leaf Area (SLA) Analyses"
author: "Phoebe Zarnetske"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=80),tidy=TRUE)
```

# Load in packages & data
```{r, message = F}
# Clear all existing data
rm(list=ls())

#Load packages
library(tidyverse)
library(bbmle)
library(lmerTest)
library(fitdistrplus)
library(lme4)
library(car)


# Set ggplot2 plotting
# This code for ggplot2 sets the theme to mostly black and white 
# (Arial font, and large font, base size=24)
theme_set(theme_bw(14))
theme_update(axis.text.x = element_text(size = 12, angle = 90),
             axis.text.y = element_text(size = 12))

# Set working directory to Google Drive
setwd("/Volumes/GoogleDrive/Shared drives/SpaCE_Lab_warmXtrophic/data/")

# Read in data
sla <- read.csv("./L1/SLA/SLA_L1.csv")
```

# Check what species are measured at each site
```{r}
unique(sla$species)
with(sla,table(sla$site,sla$species))
```
## Each species is only at one site, so for now don't make separate dataframes (just use species as fixed effect)

# Data exploration - some different ways of visualizing these data
```{r}
ggplot(sla, aes(area_cm2, fill = species)) + geom_histogram(binwidth = 0.5) + 
        facet_grid(year ~ site, margins = TRUE, scales = "free")

ggplot(sla, aes(area_cm2, fill = as.factor(species))) + geom_histogram(binwidth = 0.5) + 
        facet_grid(species~year, margins = TRUE, scales = "free")

ggplot(sla, aes(area_cm2, fill = species, color=species)) +
        geom_density(alpha = 0.1)

ggplot(sla, aes(area_cm2, fill = species, color=species)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year)

ggplot(sla, aes(area_cm2, fill = species, color=species)) +
        geom_density(alpha = 0.1) +
        facet_wrap(~year + site)

# Exploring distributions for these right-skewed data:
# See: http://www.di.fc.ul.pt/~jpn/r/distributions/fitting.html
descdist(sla$area_cm2, discrete = FALSE)

# Gamma distribution is an option
fit.gamma <- fitdist(sla$area_cm2, "gamma")
plot(fit.gamma)

# Exponential distribution is another option
fit.exp <- fitdist(sla$area_cm2, "exp")
plot(fit.exp)

# Lognormal distribution is another option
fit.ln <- fitdist(sla$area_cm2, "lnorm")
plot(fit.ln)

# Square root transformed normal distribution 
fit.sr <- fitdist(sqrt(sla$area_cm2), "norm")
plot(fit.sr)

par(mfrow=c(2,2))
plot.legend <- c("Gamma", "Exponential", "Log Normal")
denscomp(list(fit.gamma, fit.exp, fit.ln), legendtext = plot.legend)
cdfcomp (list(fit.gamma, fit.exp, fit.ln), legendtext = plot.legend)
qqcomp  (list(fit.gamma, fit.exp, fit.ln), legendtext = plot.legend)
ppcomp  (list(fit.gamma, fit.exp, fit.ln), legendtext = plot.legend)

# Goodness of fit comparisons across fits (can't include the sqrt normal bc it becomes diff response values)
gofstat(list(fit.gamma, fit.exp, fit.ln), fitnames = c("Gamma", "Exponential", "Log Normal"))
# Lognormal is a winner so far
```


# Determining appropriate distribution in mixed effects model
```{r}
# See Ben Bolker's site for details on fitting glmms: https://bbolker.github.io/mixedmodels-misc/glmmFAQ.html
#       Fixed effects: warming, year, insecticide
#       Random effect = plot (bc observations are nested within plot)
# Fit with ML bc we are interested in estimating fixed effects more than random effects
# See https://uoftcoders.github.io/rcourse/lec08-linear-mixed-effects-models.html for more details 
m1 <- lmer(log(area_cm2) ~ state + species + year + insecticide + (1|plot), data = sla, REML=FALSE)

# Check Assumptions:
# (1) Linearity: if covariates are not categorical (year isn't)
# (2) Homogeneity: Need to Check by plotting residuals vs predicted values.
par(mfrow=c(1,2))
plot(m1, main = "Log Area (cm2)")
# Homogeneity of variance is ok here (increasing variance in resids is not increasing with fitted values)
# Check for homogeneity of variances (true if p>0.05). If the result is not significant, the assumption of equal variances (homoscedasticity) is met (no significant difference between the group variances).
leveneTest(residuals(m1) ~ sla$state)
# Assumption not met
leveneTest(residuals(m1) ~ sla$species)
# Assumption not met - not surprising given the differences among species
leveneTest(residuals(m1) ~ sla$insecticide)
# Assumption met
leveneTest(residuals(m1) ~ sla$plot)
# Assumption met

# (3) Normality of error term: need to check by histogram, QQplot of residuals, could do Kolmogorov-Smirnov test.
# Check for normal residuals
qqPlot(resid(m1), main = "Area (cm2)")
hist(residuals(m1), main = "Area (cm2)")
shapiro.test(resid(m1)) # not normally distributed resids bc p>0.05
# Outliers
sla[377,]
sla[801,]
# Outlier test - yes, these are outliers bc p<0.05 
outlierTest(m1)

# Remove these 2 points
sla1<-sla[-c(377, 801),]
m1.nooutlier <- lmer(log(area_cm2) ~ state + species + year + insecticide + (1|plot), data = sla1, REML=FALSE)
shapiro.test(resid(m1.nooutlier)) # normally distributed resids - good!

# (4) Normality of random effect: Get the estimate of random effect (e.g., random intercepts), and check them as you would check the residual. [***need to do for final model]
```

Model exploration determined log normal distribution for SLA and dropping 2 outliers. Proceed with re-checking the distribution and assumptions as above with the outliers removed.
```{r}
# Fixed effects: warming, year, insecticide
# Random effect = plot (bc observations are nested within plot)
# Fit with ML bc we are interested in estimating fixed effects more than random effects
# See https://uoftcoders.github.io/rcourse/lec08-linear-mixed-effects-models.html for more details 
m2 <- lmer(log(area_cm2) ~ state + species + year + insecticide + (1|plot), data = sla1, REML=FALSE)

# Check Assumptions:
# (1) Linearity: if covariates are not categorical (year isn't)
# (2) Homogeneity: Need to Check by plotting residuals vs predicted values.
par(mfrow=c(1,2))
plot(m2, main = "Log Area (cm2)")
# Homogeneity of variance is ok here (increasing variance in resids is not increasing with fitted values)
# Check for homogeneity of variances (true if p>0.05). If the result is not significant, the assumption of equal variances (homoscedasticity) is met (no significant difference between the group variances).
leveneTest(residuals(m2) ~ sla1$state)
# Assumption not met
leveneTest(residuals(m2) ~ sla1$species)
# Assumption not met - not surprising given the differences among species
leveneTest(residuals(m2) ~ sla1$insecticide)
# Assumption met
leveneTest(residuals(m2) ~ sla1$plot)
# Assumption met

# (3) Normality of error term: need to check by histogram, QQplot of residuals, could do Kolmogorov-Smirnov test.
# Check for normal residuals
qqPlot(resid(m2), main = "Area (cm2)")
hist(residuals(m2), main = "Area (cm2)")
shapiro.test(resid(m2)) # normally distributed resids bc p>0.05
# Outlier test - no, these are no outliers bc p>0.05 
outlierTest(m2)

# (4) Normality of random effect: Get the estimate of random effect (e.g., random intercepts), and check them as you would check the residual. 
require(lme4)
r_int<- ranef(m2)$plot$`(Intercept)`
qqnorm(r_int)
qqline(r_int)
shapiro.test(r_int) 
# Yes, normally distributed random effect.
```
According to the results above, the outlier removal helps the model meed the assumptions. Proceed with model comparison. Start with investigating the model summary to determine which coviarates to drop.
```{r}
summary(m2)
# Looks like year and warmed aren't significant - I'm leaving it in for now bc the data will change. With fuller data we'd begin model comparison here (will pick up here when those data are available). Then run the code below for diff models:

#AICctab(m2, m3, weights = T)
```

