---
title: "warmXtrophic Project: Specific Leaf Area (spla) Analyses"
author: "Phoebe Zarnetske"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=80),tidy=TRUE)
```

# Load in packages & data
```{r, message = F}
# Clear all existing data
rm(list=ls())

#Load packages
library(tidyverse)
library(bbmle)
library(lmerTest)
library(fitdistrplus) 
library(lme4)
library(car)
library(emmeans)


# Set ggplot2 plotting
# This code for ggplot2 sets the theme to mostly black and white 
# (Arial font, and large font, base size=24)
theme_set(theme_bw(14))
theme_update(axis.text.x = element_text(size = 12, angle = 90),
             axis.text.y = element_text(size = 12))

# Set working directory to Google Drive
#setwd("/Volumes/GoogleDrive/Shared drives/SpaCE_Lab_warmXtrophic/data/")
L1_dir<-Sys.getenv("L1DIR")
spla <- read.csv(file.path(L1_dir,"SLA/SLA_L1.csv"))

# Read in data
#spla <- read.csv("./L1/SLA/SLA_L1.csv")
```
# Methods for SLA: 
For a subset of species, per site, use scissors to collect 4-5 mature leaves with little to no obvious insect damage or disease off of 3-5 plants of the same species. PICK one representative leaf from the samples of leaves for SLA and place it in a separate glassine envelope or ziploc (for wet surface area and for dry weight later) - place this in a cooler ASAP. This equates to multiple SLA values per plant. As such, "plant_number" should be used as a random effect. If possible, also use "plot" as a random effect (plant_number within plot).
# Check what species are measured at each site
```{r}
unique(spla$species)
with(spla,table(spla$site,spla$species))
with(spla,table(spla$plot,spla$species))
with(spla,table(spla$year,spla$plant_number))
```
Almost all species are only at one site, except Popr. Use species as fixed effect, and site as fixed effect. Plant_number and plot as random effects account for the pseudoreplication of observations within a plant and within a plot.

# Data exploration - some different ways of visualizing these data
```{r}
ggplot(spla, aes(sla, fill = species)) + geom_histogram(binwidth = 50) + 
        facet_grid(year ~ site, margins = TRUE, scales = "free")

ggplot(spla, aes(sla, fill = as.factor(species))) + geom_histogram(binwidth = 50) + 
        facet_grid(species~year, margins = TRUE, scales = "free")

ggplot(spla, aes(sla, fill = species, color=species)) +
        geom_density(alpha = 25)

ggplot(spla, aes(sla, fill = species, color=species)) +
        geom_density(alpha = 25) +
        facet_wrap(~year)

ggplot(spla, aes(sla, fill = species, color=species)) +
        geom_density(alpha = 25) +
        facet_wrap(~year + site)


# exploring transformation for right skewed data:
hist(sqrt(spla$sla))
hist((spla$sla)^(1/3))
hist(log(spla$sla))

sqrt.test <- lmer(sqrt(sla) ~ state + species + site + (1 | plot/plant_number), data=spla,REML=F)
qqPlot(resid(sqrt.test), main = "Specific Leaf Area (cm2/g)")
hist(residuals(sqrt.test), main = "Specific Leaf Area (cm2/g)")
outlierTest(sqrt.test) # some outliers
shapiro.test(resid(sqrt.test)) # not normally distributed

cubed.test <- lmer((sla^(1/3)) ~ state + species + site + (1 | plot/plant_number), data=spla,REML=F)
qqPlot(resid(cubed.test), main = "Specific Leaf Area (cm2/g)")
hist(residuals(cubed.test), main = "Specific Leaf Area (cm2/g)")
outlierTest(cubed.test) # some outliers, but will remove these further down
shapiro.test(resid(cubed.test)) # not normally distributed
# this looks best

log.test <- lmer(log(sla) ~ state + species + site + (1 | plot/plant_number), data=spla,REML=F)
qqPlot(resid(log.test), main = "Specific Leaf Area (cm2/g)")
hist(residuals(log.test), main = "Specific Leaf Area (cm2/g)")
outlierTest(log.test) # no outliers
shapiro.test(resid(log.test)) # not normally distributed


# Exploring distributions for these right-skewed data:
# See: http://www.di.fc.ul.pt/~jpn/r/distributions/fitting.html
descdist(spla$sla, discrete = FALSE)

# Gamma distribution is an option
fit.gamma <- fitdist(spla$sla, "gamma")
plot(fit.gamma)

# Exponential distribution is another option; not as good but ok
fit.exp <- fitdist(spla$sla, "exp")
plot(fit.exp)

# Lognormal distribution is another option; not as good w Q-Q plot
fit.ln <- fitdist(spla$sla, "lnorm")
plot(fit.ln)

# Square root transformed normal distribution - not as good
fit.sr <- fitdist(sqrt(spla$sla), "norm")
plot(fit.sr)

par(mfrow=c(2,2))
plot.legend <- c("Gamma", "Exponential", "Log Normal")
denscomp(list(fit.gamma, fit.exp, fit.ln), legendtext = plot.legend)
cdfcomp (list(fit.gamma, fit.exp, fit.ln), legendtext = plot.legend)
qqcomp  (list(fit.gamma, fit.exp, fit.ln), legendtext = plot.legend)
ppcomp  (list(fit.gamma, fit.exp, fit.ln), legendtext = plot.legend)

# Goodness of fit comparisons across fits (can't include the sqrt normal bc it becomes diff response values)
gofstat(list(fit.gamma, fit.exp, fit.ln), fitnames = c("Gamma", "Exponential", "Log Normal"))
# Log normal is a winner so far (lowest AIC, BIC), but going to start with cubed transformation
```


# Determining appropriate distribution in mixed effects model.
See Ben Bolker's site for details on fitting glmms: https://bbolker.github.io/mixedmodels-misc/glmmFAQ.html

Fixed effects: warming, year, insecticide
Random effect = plant_number, plot (bc observations are nested within plant and plot)

Fit with ML (REML = FALSE) bc we are interested in estimating fixed effects more than random effects. See https://uoftcoders.github.io/rcourse/lec08-linear-mixed-effects-models.html for more details.
```{r}
m1 <- lmer(sla^(1/3) ~ state + species + year + insecticide + (1 | plot/plant_number), data = spla, REML=FALSE)
ggplot(spla,aes(x=plot,y=(sla^(1/3)),col=plant_number)) + geom_jitter() + geom_boxplot(alpha=0.2) + facet_wrap(~plant_number)


# Check Assumptions:
# (1) Linearity: if covariates are not categorical (year isn't)
# (2) Homogeneity: Need to Check by plotting residuals vs predicted values.
par(mfrow=c(1,2))
plot(m1, main = "Log Specific Leaf Area (cm2/g)")
# Homogeneity of variance is ok here (increasing variance in resids is not increasing with fitted values)
# Check for homogeneity of variances (true if p>0.05). If the result is not significant, the assumption of equal variances (homoscedasticity) is met (no significant difference between the group variances).
leveneTest(residuals(m1) ~ spla$state)
# Assumption not met
leveneTest(residuals(m1) ~ spla$species)
# Assumption not met - not surprising given the differences among species
leveneTest(residuals(m1) ~ spla$insecticide)
# Assumption met
leveneTest(residuals(m1) ~ spla$plot)
# Assumption met

# (3) Normality of error term: need to check by histogram, QQplot of residuals, could do Kolmogorov-Smirnov test.
# Check for normal residuals
qqPlot(resid(m1), main = "Specific Leaf Area (cm2/g)")
hist(residuals(m1), main = "Specific Leaf Area (cm2/g)")
shapiro.test(resid(m1)) # not normally distributed because p < 0.05

# check for outliers
outlierTest(m1) # some outliers
# Remove these points
spla1<-spla[-c(1406,1375,835,557,590,558),]

# re-run model
m1.1 <- lmer(sla^(1/3) ~ state + species + year + insecticide + (1|plot/plant_number), data = spla1, REML=FALSE)
qqPlot(resid(m1.1), main = "Specific Leaf Area (cm2/g)")
hist(residuals(m1.1), main = "Specific Leaf Area (cm2/g)")
shapiro.test(resid(m1.1)) # not normally distributed
outlierTest(m1.1) # still outliers
# remove these points
spla1[635,] # location of "638" in the dataframe
spla1[266,]
spla2<-spla1[-c(266,635),]

# re-run model
m1.2 <- lmer(sla^(1/3) ~ state + species + year + insecticide + (1|plot/plant_number), data = spla2, REML=FALSE)
qqPlot(resid(m1.2), main = "Specific Leaf Area (cm2/g)")
hist(residuals(m1.2), main = "Specific Leaf Area (cm2/g)")
shapiro.test(resid(m1.2)) # not normally distributed
outlierTest(m1.2) # no outliers

# Check homogeneity of variances again
leveneTest(residuals(m1.2) ~ spla2$state)
# Assumption not met
leveneTest(residuals(m1.2) ~ spla2$species)
# Assumption not met - not surprising given the differences among species
leveneTest(residuals(m1.2) ~ spla2$insecticide)
# Assumption met bc p>0.05
leveneTest(residuals(m1.2) ~ spla2$plot)
# Assumption met bc p>0.05

# (4) Normality of random effect: Get the estimate of random effect (e.g., random intercepts), and check them as you would check the residual. 
require(lme4)
r_int<- ranef(m1.2)$plot$`(Intercept)`
qqnorm(r_int)
qqline(r_int)
shapiro.test(r_int) 
#KD edit: this last chunk doesn't work for me
# Yes, normally distributed random effect.

```

*Export* the "no outliers" SLA L1 dataframe as a CSV to ensure it is used for subsequent analyses. This is now the updated csv with the cleaned and merged SLA L0 data, lacking outliers, uploaded to the shared google drive L1 folder:
```{r}
write.csv(spla2, file.path(L1_dir, "SLA/SLA_L1_nooutliers.csv"), row.names=F)
```

# Moving on with model exploration
```{r}
# based on the code above, we know cubed root transformation doesn't work great (based on shapiro-wilk tests)
# however, the residuals histograms still look decently normal (even if shapiro is <0.05)
# I'm going to test different cubed root-transformed models and see how they look before trying a different distribution

# full non-outlier model
m1.2 <- lmer(sla^(1/3) ~ state + species + year + insecticide + (1|plot/plant_number), data = spla2, REML=FALSE)
qqPlot(resid(m1.2), main = "Specific Leaf Area (cm2/g)")
hist(residuals(m1.2), main = "Specific Leaf Area (cm2/g)")
outlierTest(m1.2) # no outliers (they are all below +-2 resids)
shapiro.test(resid(m1.2)) # not normally distributed, but histogram looks good

m1.3 <- lmer(sla^(1/3) ~ state + species + year + (1|plot/plant_number), data = spla2, REML=FALSE)
qqPlot(resid(m1.3), main = "Specific Leaf Area (cm2/g)")
hist(residuals(m1.3), main = "Specific Leaf Area (cm2/g)")
outlierTest(m1.3) # no outliers (they are all below +-2 resids)
shapiro.test(resid(m1.3)) # not normally distributed, but histogram looks good
anova(m1.2,m1.3)
AICctab(m1.2,m1.3) # m1.3 slightly better
summary(m1.3)

m1.4 <- lmer(sla^(1/3) ~ state + species + year + site + (1|plot/plant_number), data = spla2, REML=FALSE)
qqPlot(resid(m1.4), main = "Specific Leaf Area (cm2/g)")
hist(residuals(m1.4), main = "Specific Leaf Area (cm2/g)")
outlierTest(m1.4) # no outliers (they are all below +-2 resids)
shapiro.test(resid(m1.4)) # not normally distributed, but histogram looks good
anova(m1.3,m1.4)
AICctab(m1.3,m1.4) # m1.4 better
summary(m1.4)

m1.5 <- lmer(sla^(1/3) ~ state * species + year + site + (1|plot/plant_number), data = spla2, REML=FALSE)
qqPlot(resid(m1.5), main = "Specific Leaf Area (cm2/g)")
hist(residuals(m1.5), main = "Specific Leaf Area (cm2/g)")
outlierTest(m1.5) # some outliers
shapiro.test(resid(m1.5)) # not normal, but histogram looks good
anova(m1.4,m1.5)
AICctab(m1.4,m1.5) # m1.5 better
summary(m1.5)

m1.6 <- lmer(sla^(1/3) ~ state * species + site + (1|plot/plant_number) + (1|year), data = spla2, REML=FALSE)
qqPlot(resid(m1.6), main = "Specific Leaf Area (cm2/g)")
hist(residuals(m1.6), main = "Specific Leaf Area (cm2/g)")
outlierTest(m1.6) # no outliers (they are all below +-2 resids)
shapiro.test(resid(m1.6)) # not normal, but histogram looks good
anova(m1.5,m1.6)
AICctab(m1.5,m1.6) # m1.6 better


# based on the above comparisons, I'm going with m1.6
# shapiro-wilk is < 0.05, but the histogram looks normal so I'm continuing with cubed root transformation in this model
anova(m1.6)
summary(m1.6)
emmeans(m1.6, list(pairwise ~ state*species), adjust = "tukey")
mod1.6.emm <- emmeans(m1.6, ~ state*species)
contrast(mod1.6.emm, "consec", simple = "each", combine = F, adjust = "mvt")
emmip(m1.6, species~state)
```

Testing models for each site separately
This might fit the format of our manuscript better, since all other models are ran separately for each site
(However, species are mostly unique to each site, so one combined model could work - using these separate models in the paper for now)
```{r}
sla_kbs <- spla2 %>%
        filter(site == "kbs")
sla_umbs <- spla2 %>%
        filter(site == "umbs")


### KBS ###
# hypothesized model - species random
m.hyp.k <- lmer(sla^(1/3) ~ state * insecticide + year + (1|plot/species/plant_number), data = sla_kbs, REML=FALSE)
anova(m.hyp.k)
# hypothesized model - with species
m.hyp.k2 <- lmer(sla^(1/3) ~ state * insecticide + species + year + (1|plot/species/plant_number), data = sla_kbs, REML=FALSE)
anova(m.hyp.k2)
mod.emm2 <- emmeans(m.hyp.k2, ~ state * insecticide)
contrast(mod.emm2, "pairwise", simple = "each", combine = F, adjust = "mvt")

# old model tests
# full non-outlier model
m1.2k <- lmer(sla^(1/3) ~ state + species + year + insecticide + (1|plot/plant_number), data = sla_kbs, REML=FALSE)
qqPlot(resid(m1.2k), main = "Specific Leaf Area (cm2/g)")
hist(residuals(m1.2k), main = "Specific Leaf Area (cm2/g)")
outlierTest(m1.2k) # one outlier
shapiro.test(resid(m1.2k)) # not normally distributed, but histogram looks decent

m1.3k <- lmer(sla^(1/3) ~ state + species + year + (1|plot/plant_number), data = sla_kbs, REML=FALSE)
qqPlot(resid(m1.3k), main = "Specific Leaf Area (cm2/g)")
hist(residuals(m1.3k), main = "Specific Leaf Area (cm2/g)")
outlierTest(m1.3k) # same outliers as above
shapiro.test(resid(m1.3k)) # not normally distributed, but histogram looks good
anova(m1.2k,m1.3k)
AICctab(m1.2k,m1.3k) # almost equivalent - going with simpler model (m1.3)
summary(m1.3k)

m1.5k <- lmer(sla^(1/3) ~ state * species + year + (1|plot/plant_number), data = sla_kbs, REML=FALSE)
qqPlot(resid(m1.5k), main = "Specific Leaf Area (cm2/g)")
hist(residuals(m1.5k), main = "Specific Leaf Area (cm2/g)")
outlierTest(m1.5k) # same as above
shapiro.test(resid(m1.5k)) # not normal, but histogram looks good
anova(m1.3k,m1.5k)
AICctab(m1.3k,m1.5k) # m1.5 better
summary(m1.5k)

m1.6k <- lmer(sla^(1/3) ~ state * species + (1|plot/plant_number), data = sla_kbs, REML=FALSE)
qqPlot(resid(m1.6k), main = "Specific Leaf Area (cm2/g)")
hist(residuals(m1.6k), main = "Specific Leaf Area (cm2/g)")
outlierTest(m1.6k) # none
shapiro.test(resid(m1.6k)) # pretty close
anova(m1.5k,m1.6k)
AICctab(m1.5k,m1.6k) # m1.5 better

m1.7k <- lmer(sla^(1/3) ~ state * species + year + (1|plot/plant_number), data = sla_kbs, REML=FALSE) # changed to true here and above for comparing random effects
qqPlot(resid(m1.7k), main = "Specific Leaf Area (cm2/g)")
hist(residuals(m1.7k), main = "Specific Leaf Area (cm2/g)")
outlierTest(m1.7k) # one
shapiro.test(resid(m1.7k)) # pretty close
anova(m1.7k,m1.5k)
AICctab(m1.7k,m1.5k) # m1.7 better

# going with model 1.7k for KBS
anova(m1.7k)
summary(m1.7k)
emmeans(m1.7k, list(pairwise ~ state*species), adjust = "tukey")
mod1.7k.emm <- emmeans(m1.7k, ~ state*species)
contrast(mod1.7k.emm, "consec", simple = "each", combine = F, adjust = "mvt")
emmip(m1.7k, species~state)



### UMBS ###
# hypothesized model - species random
m.hyp.u <- lmer(sla^(1/3) ~ state * insecticide + year + (1|plot/species/plant_number), data = sla_umbs, REML=FALSE)
anova(m.hyp.u)
# hypothesized model - with species
m.hyp.u2 <- lmer(sla^(1/3) ~ state * insecticide + species + year + (1|plot/species/plant_number), data = sla_umbs, REML=FALSE)
anova(m.hyp.u2)
mod.emm2 <- emmeans(m.hyp.u2, ~ state * insecticide)
contrast(mod.emm2, "pairwise", simple = "each", combine = F, adjust = "mvt")

# full non-outlier model
m1.2u <- lmer(sla^(1/3) ~ state + species + year + insecticide + (1|plot/plant_number), data = sla_umbs, REML=FALSE)
qqPlot(resid(m1.2u), main = "Specific Leaf Area (cm2/g)")
hist(residuals(m1.2u), main = "Specific Leaf Area (cm2/g)")
outlierTest(m1.2u) # none
shapiro.test(resid(m1.2u)) # no but histogram looks good

m1.3u <- lmer(sla^(1/3) ~ state + species + year + (1|plot/plant_number), data = sla_umbs, REML=FALSE)
qqPlot(resid(m1.3u), main = "Specific Leaf Area (cm2/g)")
hist(residuals(m1.3u), main = "Specific Leaf Area (cm2/g)")
outlierTest(m1.3u) # none
shapiro.test(resid(m1.3u)) # no but histogram looks good
anova(m1.2u,m1.3u)
AICctab(m1.2u,m1.3u) # m1.3 slightly better
summary(m1.3u)

m1.5u <- lmer(sla^(1/3) ~ state * species + year + (1|plot/plant_number), data = sla_umbs, REML=FALSE)
qqPlot(resid(m1.5u), main = "Specific Leaf Area (cm2/g)")
hist(residuals(m1.5u), main = "Specific Leaf Area (cm2/g)")
outlierTest(m1.5u) # one
shapiro.test(resid(m1.5u)) # no but histogram looks good
anova(m1.3u,m1.5u)
AICctab(m1.3u,m1.5u) # m1.5 better
summary(m1.5u)

m1.6u <- lmer(sla^(1/3) ~ state * species + (1|plot/plant_number), data = sla_umbs, REML=FALSE)
qqPlot(resid(m1.6u), main = "Specific Leaf Area (cm2/g)")
hist(residuals(m1.6u), main = "Specific Leaf Area (cm2/g)")
outlierTest(m1.6u) # none
shapiro.test(resid(m1.6u)) # no but histogram looks good
anova(m1.5u,m1.6u)
AICctab(m1.5u,m1.6u) # m1.5 better

m1.7u <- lmer(sla^(1/3) ~ state * species + (1|plot/plant_number) + (1|year), data = sla_umbs, REML=FALSE) # changed to true here and above for comparing random effects
qqPlot(resid(m1.7u), main = "Specific Leaf Area (cm2/g)")
hist(residuals(m1.7u), main = "Specific Leaf Area (cm2/g)")
outlierTest(m1.7u) # one
shapiro.test(resid(m1.7u)) # pretty close
anova(m1.7u,m1.6u)
anova(m1.7u,m1.5u)
AICctab(m1.7u,m1.6u) # m1.7 better
AICctab(m1.7u,m1.5u) # m1.5 better

# going with model 1.5u for UMBS
anova(m1.5u)
summary(m1.5u)
emmeans(m1.5u, list(pairwise ~ state*species), adjust = "tukey")
mod1.5u.emm <- emmeans(m1.5u, ~ state*species)
contrast(mod1.5u.emm, "consec", simple = "each", combine = F, adjust = "mvt")
emmip(m1.5u, species~state)
```
