---
title: "HOBO - Pendant Data"
author: "Kara Dobson"
output: pdf_document
date: "`r format(Sys.time(), '%B %d, %Y')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

# HOBO pendant data clean-up
## Kara's way of walking herself through cleaning up the script

This sheet is focused on cleaning up the HOBO_pendant_2017_2018.R file in the archived scripts. 
In this markdown, I'll be going through the code and writing down ways to better streamline and clean up the script. Eventually, my comments here could be implemented into code in a cleaned script, called "HOBO_pendant_cleanup" in my kara file (which could be moved to the scripts file eventually). For now, this is just my way of walking myself through the script clean-up.

The script begins:
```{r}
# **** This script makes the following datasets:
# pendk:        KBS HOBO pendant data from 2015-2018
# pendu:        UMBS HOBO pendant data from 2015-2018
# pend1518:     Combined HOBO pendant data from KBS and UMBS from 2015-2018

# Clear all existing data
rm(list=ls())

# Close graphics devices
graphics.off()
```

This could stay the same, except update the data set names and descriptions:
```{r}
# **** This script makes the following datasets:
# pendk:        KBS HOBO pendant data from 2015-2019
# pendu:        UMBS HOBO pendant data from 2015-2019
# pend1519:     Combined HOBO pendant data from KBS and UMBS from 2015-2019

# Clear all existing data
rm(list=ls())

# Close graphics devices
graphics.off()
```

Next is setting the working directory and loading packages:
```{r}
# set working directory (if you're not PLZ, change this to the correct path for your
# Google Drive directory. It should point to where we have /final_data
setwd("/Volumes/GoogleDrive/My Drive/MIWarmHerb_FieldExperiment/data/")
#setwd("/Volumes/GoogleDrive/My Drive/Research/TrophicWarming/TrophicWarming_Experiment/MIWarmHerb_FieldExperiment/data/") # PLZ's location

## Edit below for any packages you'll be using
for (package in c("ggplot2","plyr", "dplyr", "tidyr", "tidyverse","lubridate")) {
  if (!require(package, character.only=T, quietly=T)) {
    install.packages("package")
    library(package, character.only=T)
  }
}
```
The working directory portion of the script will need to be updated, especially if the data is transferred to GitHub. For now, I'll edit this in the new script as though the data exists in a github folder.
As for the packages, only tidyverse may need to be opened, as it contains the other listed packages within it:
```{r}
# Set working directory - links to the data folder in warmXtrophic repository
setwd("/warmXtrophic/data/")

## Edit below for any packages you'll be using
for (package in c("tidyverse")) {
  if (!require(package, character.only=T, quietly=T)) {
    install.packages("package")
    library(package, character.only=T)
  }
}
```


