---
title: "Practice with warmXtrophic data and RMarkdown"
author: "Moriah Young"
date: "6/3/2020"
output:
  pdf_document: default
  html_document:
    theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this document I will be practicing using RMarkdown while attempting to make a simple graph from the warmXtrophic data

First, I will make sure that the packages I need are in my working environment
```{r, message = FALSE}
for (package in c('tidyverse', 'googledrive', 'googlesheets4')) {
        if (!require(package, character.only=T, quietly=T)) {
                install.packages(package)
                library(package, character.only=T)
        }
}
```

Next, I need to download the data file that I wish to work with in R.
```{r}
# Data import and set working directory ==============================================
# I want to look at plant composition data from 2019 at UMBS
PlantComp <- read.csv("/Users/moriahyoung/Downloads/umbs_plantcomp_2019 - Sheet1.csv")
```

I want to look at the composition of spotted knapweed (Cest) in plot B2 at UMBS in 2019. To do this, I'll need to subset from the PlantComp dataframe I created above.
```{r}
# Subset the data ==========================================
# I want to look at the percent cover of the species Cest over one growing season and plot its composition over time

CestB2 <- subset(PlantComp, Species == "Cest" & Plot == "B2")
```

Now that I've subsetted my data, I'll plot the percent cover of Cest in B2 over the growing season in 2019
```{r}
# Create a plot =========================================
#I want to make a plot that shows the percent cover of Cest in plot B2 over one growing season

ggplot(data = CestB2, aes(x = Date, y = Cover)) +
        geom_bar(stat = "identity", fill = "purple") +
        labs(title = "Cest Composition in Plot B2 at UMBS",
             subtitle = "2019",
             x = "Date", y = "Percent Cover")
```

